### 2012-04-25 docs/processing.tex
reshaped directory, clean up, add presentation

\documentclass[a4paper,12pt]{article}
\begin{document}

\section{Iterative Solvers}

To minimize the error produced in the grid search for improperly chosen
$tps$ we find optimal $\alpha$, $\beta$ and crustal thickness $h$ which
define the minimal residual curve. 
\begin{equation}
Tps = h \left( \sqrt{ \frac{1}{\beta^2} - p^2} - \sqrt{ \frac{1}{\alpha^2} - p^2} \right)
\end{equation}

\subsection{Newton's Method}
To find the minimum we seek to minize the sum of the squares of the residual $r=(t_i - f(\mathbf{x}))$ 
where $f(\mathbf{x}) = Tps$ as in \emph{1)} and $\mathbf{x} = \lbrack \alpha,\beta,h \rbrack^T$.
Thus we seek 
\begin{equation}
\mbox{min}\; \vert r^Tr \vert = \mbox{min}\; \vert \phi \vert
\end{equation}
which is true when $\frac{d\phi}{d\mathbf{x}} = g(\mathbf{x}^*) = 0 $. The solution 
$\mathbf{x}^*$ is the initial guess $\mathbf{x}_0 +$ the vector $\mathbf{m} $ which 
finds the minimum $g(\mathbf{x}^*) = 0$. Taking the Taylor approximation gives,
\begin{equation}
g(\mathbf{x}+\mathbf{m}) = g(\mathbf{x}) + \mathbf{Hm} + O(\|mathbf{m}\|)
\end{equation}
The hessian $\mathbf{H}$ is equal to $\nabla^2\phi$ which is given by
$\sum_{i,j=1}^{n} \frac{\partial^2 \phi}{\partial x_i \partial x_j}$. As 
$g(\mathbf{x}+\mathbf{m})=0$ we can rewrite \emph{3)} as
\begin{equation}
\mathbf{m}=-\mathbf{H}^-1g(mathbf{x})
\end{equation}
and iterate until convergence with
\begin{equation}
\mathbf{x}_{k+1} = \mathbf{x}_k + \mathbf{m}
\end{equation}

\end{document}

### 2012-04-25 ProcessTraces.m

@@ -115,12 +115,13 @@ fLow = 0.04;
 fHigh = 1;
 numPoles = 2;
 
-brec = fbpfilt(rec,dt,fLow,fHigh,numPoles,0);
+%brec = fbpfilt(rec,dt,fLow,fHigh,numPoles,0);
 
 %load Rec_50_it
 %load Rec_100_it_splines
-%brec = REC;
-%brec = fbpfilt(brec,dt,fLow,fHigh,numPoles,0);
+load Rec_100_it_bsplines
+brec = REC;
+brec = fbpfilt(brec,dt,fLow,fHigh,numPoles,0);
 
 for ii=1:size(rec,1);
     %brtrace(ii,:)=filter(h2,rtrace(ii,:));
@@ -147,6 +148,7 @@ damp = 0.2;
 
 %% 8) Grid and Line Search
 [ results ] = GridSearch(brec,Tps',dt,pslow);
+
 %[ results ] = GsearchKanamori(brec,dt,pslow);
 %% Viewers
 %{

### 2012-04-25 dataDriver.m

@@ -32,7 +32,7 @@ remove = false;    % Removes all entries associated with particular station
 %% Run ProcessTraces
 % Run ProcessTraces then collect results into structure
 try
-    %ProcessTraces
+    ProcessTraces
     % For a description of data see DataDescription.m
     dbn.station = station;    
     dbn.processnotes = [];
@@ -122,7 +122,4 @@ end
 %    end     
 %end
 %}
-
-
-
 

### 2012-04-25 python/loopApplyDRIVER.py

@@ -10,7 +10,8 @@
 ###########################################################################
 import os, re, time, shutil
 from preprocessor import calculate_slowness, detrend_taper_rotate, NoSlownessError
-
+import subprocess
+sh = subprocess.Popen
 ###########################################################################
 # HELPER FUNCTIONS
 ###########################################################################
@@ -91,8 +92,10 @@ for network in networks:
 
             # Run Processing function
             try:
-                calculate_slowness(eventdir, sacfiles)
-                detrend_taper_rotate(eventdir, sacfiles)
+                # calculate_slowness(eventdir, sacfiles)
+                # detrend_taper_rotate(eventdir, sacfiles)
+                
+                # sh("rm {}/*STACK*".format(eventdir), shell=True, executable = "/bin/bash")
             except IOError:
                 print "IOERROR in event:", eventdir
                 logfile.write("Error Processing: " + eventdir + " IOError\n")

### 2012-04-25 python/preprocessor.py

@@ -53,6 +53,18 @@ def calculate_slowness(eventdir, sacfiles):
         if i == 0:
             evdp = st[0].stats.sac['evdp']
             gcarc = st[0].stats.sac['gcarc']
+            # Get slowness P
+            process = sh("/home/bpostlet/bin/Get_tt/get_tt -z {} -d {} -p P".format(evdp,gcarc),
+               shell=True, executable = "/bin/bash", stdout = pipe )
+            results =  process.communicate()[0].rstrip().split('\n')
+            for result in results:
+                result = result.split()
+                if result[1] == 'P':
+                    slowness = float(result[3])
+                    break
+                else:
+                    raise NoSlownessError('No_slowness_GARC_is:{}'.format(gcarc))
+            # Get PP time and set in headers
             process = sh("/home/bpostlet/bin/Get_tt/get_tt -z {} -d {} -p P".format(evdp,gcarc),
                shell=True, executable = "/bin/bash", stdout = pipe )
             results =  process.communicate()[0].rstrip().split('\n')
@@ -77,6 +89,7 @@ def detrend_taper_rotate(eventdir, sacfiles):
     """preprocess performs the demean,detrend,taper and rotation into radial and
     transverse components. It saves these at STACK_R.sac and STACK_T.sac"""
     
+    N = 16384  #Max length of seismic array, truncate for speed
     ev = []
     
     # READ 3 Component SAC files into object array.
@@ -86,11 +99,13 @@ def detrend_taper_rotate(eventdir, sacfiles):
         ev.append(st[0]) 
     
     dt = ev[1].stats.delta
+    pslow = ev[1].stats.sac['user0']
+
     # window, detrend, taper all three components
     for i in range(3):
-        ev[i].stats.sac['b'] = ev[i].stats.sac['t1'] - 70
-        n1 = round(ev[i].stats.sac['b']/dt)
-        ev[i].data = ev[i].data[n1:n1+16384] # window 70s before pick, extend to a pwr of 2
+        ev[i].stats.sac['b'] = ev[i].stats.sac['t0'] - 70 # start seismogram 70 seconds before P arrival, adjust beginning time in header.
[…]

### 2012-04-27 ProcessTraces.m
Curvelet denoising bitches\!

@@ -2,6 +2,7 @@
 % Script to load up sac files, extract out some info, p-value etc
 % Rotate traces, deconvolve traces -> then off to be stacked.
 
+loadtools;
 
 %% 1) Filter Event Directories
 %
@@ -29,54 +30,26 @@ pIndex = pIndex(:,any(pIndex)); % Strip out indices with no traces
 nbins = length(pslow); % Number of bins we now have.
 %}
 %% 4) Normalize
-dt = header{1,1}.DELTA;
-n1 = round(100/dt);
-%rms = zeros(length(ptrace),2);
+dt = header{1}.DELTA;
+ext = round(20/dt);
 %{
 for ii = 1:size(ptrace,1)
-    rms(ii,1) = norm(ptrace(ii,1:n1))/sqrt(n1);
-    %rms(ii,2) = norm(ptrace{ii}(1,end-n1-1:end))/sqrt(n1);
+    T1 = round((header{1}.T1 - header{1}.B)/dt);
+    normnoise = norm(ptrace(ii, 1 : ext));
+    normP = norm(ptrace(ii, T1 : T1 + ext ));
+    nr = (normP/normnoise);
+    P(ii,:) = detrend(ptrace(ii,:)) ; %.* nr;
+    S(ii,:) = detrend(strace(ii,:)) ; %.* nr;
 end
-sfact = 0.5; % Rescale spread between magnitudes between sfact->1
-range = 1-sfact;
-scale = [];
-for ii = 1:nbins
-    rmsN = rms(pIndex(:,ii));
-    rmsR = rmsN/max(rmsN);
-    sc = rmsR*range/max(rmsR) + sfact; % Project into new range
-    scale = [scale; max(rmsN)*(sc./rmsN) ]; % Find multiplier that will rescale traces
-end
-
-for ii = 1:size(ptrace,1)
-    ptrace(ii,:) = ptrace(ii,:) * scale(ii);
-    strace(ii,:) = strace(ii,:) * scale(ii);
-end
+ptrace = P;
+strace = S;
 %}
-
-%rmsR = rms(:,1)./rms(:,2);
-%figure(2222)
-%plot(rms)
-%figure(1111)
-%plot(rmsR)
-
 %% 5)  Window with Taper and fourier transform signal.
 %
 viewtaper  = 0;
[…]

### 2012-04-27 dataDriver.m

@@ -11,7 +11,7 @@ addpath functions
 
 %% Variables
 user = getenv('USER');
-sacfolder = '/media/TerraS/X5';
+sacfolder = '/media/TerraS/CNSN';
 %sacfolder = '/media/TerraS/CNSN';
 datadir = ['/home/',user,'/programming/matlab/thesis/data'];
 databasedir = [datadir,'/database'];
@@ -19,7 +19,7 @@ rfile = 'STACK_R.sac';
 zfile = 'STACK_Z.sac';
 %%  Select Station to Process and load station data.
 %station = 'VTIN';
-station = 'CRLN';
+station = 'GAC';
 workingdir = fullfile(sacfolder,station);
 %load(fullfile(databasedir,[station,'.mat']))
 %workingdir = fullfile(['/home/',user,'/Programming/data/'],station);
@@ -32,7 +32,7 @@ remove = false;    % Removes all entries associated with particular station
 %% Run ProcessTraces
 % Run ProcessTraces then collect results into structure
 try
-    ProcessTraces
+    % ProcessTraces
     % For a description of data see DataDescription.m
     dbn.station = station;    
     dbn.processnotes = [];
@@ -74,9 +74,15 @@ catch e
         dbn.failmessage)
 end
 %% Plot the results if we completed the processing
+close all
 if dbn.scanstatus
     plotStack(dbn);
-    %plotStack(db(2));
+    err = results.smax - results.stderr1;
+    %number of indices inside contour range for Vp
+    errNs = sum(any(results.stackvr > err,2)); 
+    dv = results.vRange(2) - results.vRange(1);
+    errV = 0.5 * dv * errNs;
+    fprintf('Vp is +/- %1.2f km/s\n',errV)
 end
 %% Enter Processing Notes:
 notes = inputdlg('Enter Notes','Processing Notes',[3 80]);

### 2012-04-27 functions/newtonFit.m

@@ -32,13 +32,13 @@ while (iter < round(itermax)) && (deltaTps > tol);
     H = [dfaa, dfab, dfah
          dfab, dfbb, dfbh
          dfah, dfbh, dfhh]; % Hessian
-    if iter == 1
-        fprintf('----prerun-----\n')
-        fprintf('J Matrix\n')
-        full(J)
-        fprintf('Eig(H)\n')
-        eig(H)
-    end
+    %if iter == 1
+    %    fprintf('----prerun-----\n')
+    %    fprintf('J Matrix\n')
+    %    full(J)
+    %    fprintf('Eig(H)\n')
+    %    eig(H)
+    %end
         
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     delm = (IRLSsolver(-H,J,40,0.001*tol)); % Linear reweighted solution with 10x tolerance of Newton
@@ -63,11 +63,11 @@ while (iter < round(itermax)) && (deltaTps > tol);
 
 end
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-        fprintf('----postrun-----\n')
-        fprintf('J Matrix\n')
-        full(J)
-        fprintf('Eig(H)\n')
-        eig(H)
+        %fprintf('----postrun-----\n')
+        %fprintf('J Matrix\n')
+        %full(J)
+        %fprintf('Eig(H)\n')
+        %eig(H)
 
 
 % Check if iter hit itermax, if so issue warning %%%%%%%%

### 2012-04-27 normalizetraces.m

@@ -1,0 +1,24 @@
+% Normalize data
+
+close all
+dt = header{1}.DELTA;
+ext = round(20/dt);
+  
+for ii = 1:size(ptrace,1)
+    %figure(1)
+    T1 = round((header{1}.T1 - header{1}.B)/dt);
+    varnoise = norm(ptrace(ii, 1 : ext));
+    varP = norm(ptrace(ii, T1 : T1 + ext ));
+    %subplot(2,1,1)
+    %    plot(ptrace(ii, T1 : T1 + round(10/dt)))
+    P(ii,:) = ptrace(ii,:)./max(ptrace(ii,:)) * varP/varnoise;
+    %subplot(2,1,2)
+    %    plot(ptrace(ii, 1 : round(10/dt)))
+    %    title('Noise')
+    %plot(ptrace(ii,1: T1 + round(40/dt)))
+    %title(sprintf(['norm of noise: %1.2e -- norm of Pcoda: %1.2e\n Ratio'...
+    %        ' of norm is: %1.2e\n\n'],varnoise,varP,varP/varnoise))
+    %line([T1, T1 + ext],[0,0],'Color',[1 0 0],'LineWidth',2)
+    %line([1, ext],[0,0],'Color',[0 1 1],'LineWidth',2)
+    scale(ii) = varP/varnoise;
+end

### 2012-04-27 python/codecounter.py

@@ -1,0 +1,31 @@
+#!/usr/bin/python2
+
+import os, sys
+
+def countLinesOfCode(directory):
+    """
+    Procedure to count total lines of code in each file
+    """
+    total = 0
+    listing = os.listdir(directory)
+    for fname in listing:
+        subtotal = 0
+        Fname = os.path.join(directory,fname)
+        if fname[-2:] == "py" or fname[-1:] == "m":
+            with open(Fname) as f:
+                for i, l in enumerate(f):
+                    pass
+                subtotal = i + 1
+            print fname, "line num:", subtotal
+            total += subtotal
+    print total
+
+if __name__ == "__main__":
+    """
+    TEST ALL THE THINGS
+    """
+    if (len(sys.argv) < 2):
+        print "Please supply a directory"
+        exit()
+    
+    countLinesOfCode(sys.argv[1])

### 2012-07-09 matlab/L1/Splines/FFTfractsplinefilters.m
added L1 solver into repo

@@ -1,0 +1,123 @@
+function [FFTanalysisfilters,FFTsynthesisfilters]=FFTfractsplinefilters(M,alpha,type)
+% FFTFRACSPLINEFILTERS Filters for the various orthogonal fractional 
+% 	spline wavelet transforms (orthonormal or semi-orthonormal).  
+% 	[FFTanalysisfilters,FFTsynthesisfilters]=FFTfractsplinefilters(M,alpha,type) 
+% 	Computation of the frequency response of the low- and high-pass 
+% 	filters that generate the orthonormal or semi-orthonormal (B-spline  
+%	or dual) fractional splines of degree alpha, and of type '+' (causal),
+% 	 '-' (anticausal) or '*' (symmetric).
+% 	
+% 	Input:
+% 	M : size of the input signal = length of FFTfilters = 2^N
+% 	alpha>-0.5 : degree of the fractional splines 
+% 	type	: type of B-splines
+%		= '+ortho' (causal orthonormal, default), '-ortho' (anticausal  
+%		orthonormal) or '*ortho' (symmetric orthonormal);
+% 		= '+bspline' (causal B-spline), '-bspline' (anticausal 
+%		B-spline) or '*bspline' (symmetric B-spline). 
+% 		= '+dual' (causal dual), '-dual' (anticausal dual) or '*dual' 
+% 		(symmetric dual). The last option is the flipped version of the
+%		B-spline one.
+% 	Output:
+% 	FFTanalysisfilters	= [lowpassfilter;highpassfilter]	: FFT filter arrays
+% 	FFTsynthesisfilters	= [lowpassfilter;highpassfilter] 	: FFT filter arrays
+% 
+% 	See also FFTWAVELETANALYSIS, FFTWAVELETSYNTHESIS.
+% 	Uses FRACTSPLINEAUTOCORR
+% 	
+% 	Author: Thierry Blu, October 1999
+% 	Biomedical Imaging Group, EPFL, Lausanne, Switzerland.
+% 	This software is downloadable at http://bigwww.epfl.ch/
+% 	
+% 	References:
+% 	[1] M. Unser and T. Blu, "Fractional splines and wavelets," 
+% 	SIAM Review, in press.
[…]

### 2012-07-09 matlab/L1/Splines/opSplineWavelet.m

@@ -1,0 +1,51 @@
+function op = opSplineWavelet(m,n, filt, smooth, levels, type)
+% OPSPINEWAVELET  Wavelet operator (modified for spline wavelets)
+%
+%    OPWAVELET(M,N,FAMILY,FILTER,LEVELS,TYPE) creates a wavelet
+%    operator of given FAMILY, for M by N matrices. The wavelet
+%    transformation is computed using the Rice Wavelet Toolbox.
+%
+%    The remaining three parameters are optional. FILTER = 8
+%    specifies the filter length and must be even. LEVELS = 5 gives
+%    the number of levels in the transformation. Both M and N must
+%    be divisible by 2^LEVELS. TYPE = 'min' indictates what type of
+%    solution is desired; 'min' for minimum phase, 'max' for
+%    maximum phase, and 'mid' for mid-phase solutions. 
+
+%   Copyright 2007, Rayan Saab, Ewout van den Berg and Michael P. Friedlander
+%   http://www.cs.ubc.ca/labs/scl/sparco
+%   $Id: opWavelet.m 577 2007-09-17 01:48:28Z mpf $
+
+if (nargin < 4), smooth = 3;   end;
+if (nargin < 3), filt   =     128;        end;
+if (nargin < 5), levels = 5;            end;
+if (nargin < 6), type   = '*ortho';     end;
+
+
+h  =  MakeFracSplinesFilters(filt,smooth,type);
+fh = @(x,mode) opWavelet_intrnl(m,n,filt,smooth,levels,type,h,x,mode);
+op = opFunction(n*m, n*m, fh); 
+
+
+function y = opWavelet_intrnl(m,n,filt,smooth,levels,type,h,x,mode)
+if mode == 0
+   y = {n*m,n*m,[0,1,0,1],{'SplineWavelet',filt,smooth,levels,type}};
+elseif mode == 1
+   if isreal(x)
+     [y,l] = midwt(reshape(x,[m,n]), h, levels);
+   else
+     [y1,l] = midwt(reshape(real(x),[m,n]), h, levels);
+     [y2,l] = midwt(reshape(imag(x),[m,n]), h, levels);
+     y = y1 + sqrt(-1) * y2;    
+   end
+   y = reshape(y,[m*n,1]);
+else
+   if isreal(x)
+      [y,l] = mdwt(reshape(x,[m,n]), h, levels);
+   else
[…]

### 2012-07-09 matlab/L1/finalproj_server.m

@@ -1,0 +1,139 @@
+% Final Project for Wavelet class
+
+clear all; close all
+%userdir = getenv('HOME');
+%f = fullfile(userdir, 'programming','matlab'); %Set base path
+%getd = @(p)path(p,path);  %easy path function
+%getd([f,'/spotbox-v1.0/']); %Path to spot box
+addpath('/home/gmatharu/EOSC513/packages/spotbox-v1.0');
+snr = @(x,y)20*log10(norm(x(:))/norm(x(:)-y(:)));
+%addpath(genpath([f,'/CurveLab-2.1.3'])) %Path to curvelet toolbox
+%addpath(genpath([f,'/spotbox-v1.0/+spot/+rwt'])) %Path to rice toolbox
+%loadtools;  
+load('source.mat'); % Load p-coda array
+%s = s(:,1); % Trim to 1 seismic source for testing
+n = size(s,1);
+N = length(s(:));
+% Create reciever function
+rec = zeros(n,1);
+rec(round(n/16)) = 1;
+rec(round(n*(7/16))) = -0.25;
+rec(round(n*(8/16))) = 0.4;
+
+% Noise Gaussian
+delta = .1;
+noise = randn(N,1)*delta;
+sigma=norm(noise);
+
+% Noise Jitter, replace diracs with gaussians in rec function
+%{
+ind = find(rec);
+gx = zeros(n,1);
+tx = (1:n)';
+for ii = 1:length(ind)
+    a = rec(ind(ii)); % Amplitude of first spike
+    b = ind(ii);      % Location of first spike
+    c = 3;
+    if ii ~= 1
+        c = 3*c;
+    end
+    gx = gx + a*exp(- ((tx - b).^2)/(2*c^2));
+end
+%}
+%rec = gx; % Use jittered receiver function
+
+% Convolve using spot operators to generate noisy signal y
+% A = Fadj*diag(F*s)*F   y = Ax + gaussian noise
+%si = sin([1:n]*pi*0.01)'; % Practise signal
+
+dim = size(s,2);
+e = opOnes(dim,1);
+f = opDFT(n);
+% Extend DFT operator for multiple source / traces
+F = kron(opEye(dim),f); % 
+F2 = kron(ones(size(s,2),1),f);
+Ft = kron(opEye(dim),f'); % Transpose scaled operator
+Fs = opDiag(F*s(:));
+C = Ft*Fs*F2;
[…]

### 2012-07-09 matlab/L1/simdecf.m

@@ -1,0 +1,106 @@
+function [rft,xft,betax] = simdecf(wft,vft,betan,pflag)
+
+% SIMDECF Simultaneous deconvolution of multiple seismograms in
+% the frequency domain. Inputs are wavelet estimates WFT, data
+% VFT (both in frequency domain and of dimension M X N where 
+% M is number of seismograms and N is number of frequencies), 
+% and regularization parameter BETAN. If BETAN < 0 then 
+% an optimum parameter BETAX is sought using Generalized Cross 
+% Validation and used to produce impulse response RFT, and model
+% resolution kernel XFT. Optional argument PFLAG = 1 plots
+% GCV and LCURVE functions. 
+
+% Check arguments.
+if nargin < 4
+  pflag=-1;
+end
+
+% Calculate dimensions of wft, and assume that minimum dimension is 
+% number of traces, maximum size is number of frequencies.
+nm=min(size(wft));
+nn=max(size(wft));
+
+% Compute denominators.
+if nm == 1
+  wwft=wft.*conj(wft);
+  vwft=vft.*conj(wft);
+else
+  wwft=sum(wft.*conj(wft));
+  vwft=sum(vft.*conj(wft));
+end
+
+% If nonzero betan provided by user, use it for deconvolution.
+% Otherwise compute best beta using Generalized Cross Validation.
+if betan > 0 
+  betax=betan;
+else
+  beta=exp([-40:0.5:50]);
+  for ib=1:length(beta);
+
+% Define operator W W* / (W W* + B) and deconvolve to get impulse response in 
+% frequency domain.
+    wwft2=wwft+beta(ib);
+    rft=vwft./wwft2;
+    xft=wwft./wwft2;
+
+% Compute model norm.
+    modnorm(ib)=norm(rft)^2;
+
+% Compute data misfit. Note misfit is numerator of GCV function.
+% Note also earlier mistake where norm(nft)^2 was norm(nft).
+    if nm == 1
+      nft=vft-wft.*rft;
+      misfit(ib)=norm(nft)^2;
+    else
+      misfit(ib)=0.0;
+      for im=1:nm
+        nft=vft(im,:)-wft(im,:).*rft;
+        misfit(ib)=misfit(ib)+norm(nft)^2;
+      end  
+    end       
+
+% Compute denominator and GCV function. 
+    den=nn*nm-real(sum(xft));
[…]

### 2012-07-10 matlab/L1/report/Paper_EOS513.tex

% EOS 513 REPORT
\documentclass[jgrga]{agutex}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}

\newcommand{\bt}[1]{\textbf{#1}}
\newcommand{\bc}[1]{\mathcal{\textbf{#1}}}
\newcommand{\ub}[1]{\underbrace{#1}}
\newcommand{\ob}[1]{\overbrace{#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OPTIONAL:
% To produce a two-columned version:
% \documentclass[jgrga]{AGUTeX}

% Two-columned format can be used to estimate the number of pages
% for the final published PDF.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OPTIONAL:
% To create numbered lines:

% If you don't already have lineno.sty, you can download it from
% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
% Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno}
% and \linenumbers*[1]command, below:

% \usepackage{lineno}
% \linenumbers*[1]

%  To add line numbers to lines with equations:

%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
% When submitting articles through the GEMS system:
% COMMENT OUT ANY COMMANDS THAT INCLUDE GRAPHICS.
% (See FIGURES section near the end of the file.)
%
% DO NOT USE \psfrag or \subfigure commands.
%
%  Figures and tables should be placed AT THE END OF THE ARTICLE,
%  after the references.
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
% \usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
%  \setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
\authorrunninghead{MATHARU GIAN, BEN POSTLETHWAITE}

% Shorter version of title entered in capital letters:
\titlerunninghead{DECONVOLUTION L2 / L1 COMPARISON}

% Author mailing address: please repeat this command for
% each author and alphabetize authors:

%\authoraddr{R. C. Bales,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(roger@hwr.arizona.edu)}

%\authoraddr{J. R. McConnell, Division of Hydrologic
%Sciences, 123 Main Street, Desert Research Institute, Reno, NV
%89512, USA.}

%\authoraddr{E. Mosley-Thompson, Department of Geography,
%Ohio State University, 123 Orange Boulevard, Columbus, OH 43210,
%USA.}

%\authoraddr{R. Williams, Department of Space Sciences, University of
%Michigan, 123 Brown Avenue, Ann Arbor, MI 48109, USA.}

%\authoraddr{Francesco Visconti, Dipartimento di Idraulica,
%Trasporti ed Infrastrutture Civili, Politecnico di Torino,
%Corso Duca degli Abruzzi 24, I-10129, Torino, Italy.

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

\title{Sparsity Promoting L1 Deconvolution: Application for a Global Seismology Parameter Search}
%
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

%Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote;
% matching \altaffiltext will appear at bottom of page.

\authors{G. Matharu, B. Postlethwaite \altaffilmark{1}}
% E. Mosley-Thompson,\altaffilmark{2} R. Williams,\altaffilmark{3}
% J. R. McConnell\altaffilmark{4}, and Francesco Visconti\altaffilmark{5}}

\altaffiltext{1}{Department of Earth and Ocean Sciences,
University of British Columbia, Vancouver, B.C. Canada.}

%\altaffiltext{2}{Department of Geography, Ohio State University,
%Columbus, Ohio, USA.}

%\altaffiltext{3}{Department of Space Sciences, University of
%Michigan, Ann Arbor, Michigan, USA.}

%\altaffiltext{4}{Division of Hydrologic Sciences, Desert Research
%Institute, Reno, Nevada, USA.}

%\altaffiltext{5}{Dipartimento di Idraulica, Trasporti ed
%Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%

% >> Do NOT include any \begin...\end commands within
% >> the body of the abstract.

\begin{abstract}
This report will outline an attempt to utilize the inherent sparsity in seismic receiver functions in order to provide better deconvolution results than the often utilized damped-L2 methods. By promoting sparsity with the wavelet transform the use of sparse solvers such as SPGL1 becomes possible. Synthetic tests are performed using the damped L2 convolution as well as the sparse wavelet transform L1 minimization using several wavelet basis. Real data is employed and the deconvolved results from the competing methods are used for an actual parameter grid search. The final results of the grid search are then compared. Further work on the sparse-in-wavelets L1 algorthim will be required if it is to beat the damped-L2 algorthim currently being employed.
\end{abstract}
%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%

% The body of the article must start with a \begin{article} command
%
% \end{article} must follow the references section, before the figures
%  and tables.

\begin{article}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%
\section{Introduction}

Extracting material properties from seismic data spans industry and the research community. In the global seismic community material properties of the crust have been sought to validate existing geological models or provide clues for new ones. A good example of this is Poisson's ratio, which is the ratio between compression along one axis to extension on another when a material is stretched (see figure 1). Poisson's ratio is given by:
\begin{equation}
 \sigma = \frac{1}{2} \frac{ \left( \frac{ V_p }{ V_s } \right)^2 - 2 }{ \left( \frac{ V_p }{ V_s } \right)^2 - 1 }
\end{equation}

\begin{figure}
\noindent\includegraphics[width=18pc]{PoissonRatio.eps}
\caption{Poissons Ratio is the ratio of compression along one axis to extension along another}
\end{figure}

where the seismic velocities $V_p$ and $V_s$ determine the property entirely. Because of the direct link between seismic properties and Poisson's ratio this material property has been sought in numerous seismic studies such as Kanamori (2000) and Snyder (2010). All of these studies have performed a parameter search for $R = \frac{V_p}{V_s}$ and provided a background estimate for $V_p$. This allows for fast computation through a grid search and avoids the problem of the low-dependency of $V_p$ in the grid search equations. Although Poisson's ratio can be estimated from $R$ better geological constraints would be possible if $V_p$ and $V_s$ were resolved uniquely.

\section{The Problem}

In order to differentiate between $V_p$ and $V_s$ the travel times from seperate arrivals of energy must be compared. The first arrival a seismic station will pick up after the onset of an earthquake is from the higher velocity P-wave. The impulsive character and high amplitude of the P-wave arrival allow it to be windowed and filtered and used as an approximation for the source. There are three more arrivals of interest. The Ps arrival, which is P-energy converted to S-wave energy at the Moho boundary between the crust and the mantle. Two further arrivals follow which are the reflections from the free surface which have reflected again off the Moho and recorded at the station, figure (2). These reflected arrivals are usually of lower amplitude and often less well defined in the seismogram. If the three S-wave arrivals - the Ps, PpPs and PsPs - can be located in time this provides the data in which a model may be fit. Owing to the fact that the S-wave phases all travel within the same crustal thickess $H$, dividing the S-wave relected phases by the initial Ps phase will remove the dependence on crustal thickness. The traveltime ratios are then dependent on $V_p$ and $V_s$ only (Bostock, 2010). With this information a 2D grid search over the travel time equations with respect to $V_p$ and $V_s$ can be performed. The parameter $V_p$ is not very sensitive to the traveltime differences between the S-wave phases, thus a very impulsive deconvolution result is needed to steepen the gradient in $V_p$ space.

\begin{figure}
\noindent\includegraphics[width=18pc]{reflectRecs.eps}
\caption{Top: Shows direct phases Pp and Ps from a seismic event hitting receiver. Also the reflected phases PpPs and PsPs are shown. The phases ending with 's' are the phases hitting the station as S-waves. These S-wave phases are used as the primary data, the direct Pp phase is used as the source, and the two are deconvolved to attain the receiver functions shown at Bottom: The receiver functions attained from deconvolution. Depending on the quality of the data and the sharpness of the deconvolution, the results may look like those given for the Rae Domain -better- or the Hearne Domain -worse.}
\end{figure}

\section{Deconvolution - L2}
Currently simultaneous deconvolution with generalized cross validation is being employed as it performs well and is very fast. Since division in the fourier basis is deconvolution in time we can write:
\[ \bc{R} = \bc{F}^{-1}[\bc{G}(\omega)]  \]
where
\begin{equation}
 \bc{G}(\omega) = \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*}{\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right].
\end{equation}
Here $\bc{F}$ is the Fourier transform, $\bc{G}$ is the model, $\bt{S}$ is a matrix of S-wave seismograms, $\bt{P}$ is a matrix of P-wave source coda, $*$ is the complex conjugate and $\delta$ is the regularization parameter. $\delta$ is chosen by minimizing the Generalized Cross Validation function with respect to $\delta$:
\begin{equation}
\text{min} \left( \frac{ \sum_n^N \sum_m^M 
 \left( \ob{ \bt{S}_n(\omega_m) - \bt{P}(\omega_m) \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*} {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right]}^{\text{\textcolor{red}{ Misfit: more accurate with $\delta\rightarrow0$. Less Stable }}} \right)^2}  { \left( 
\ub{  NM - \sum_m^M  \frac {\sum_n^N \bt{P}_n\bt{P}_n^*}  {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  }_{\text{\textcolor{red}{ {As $\delta$ increases denominator increases, promotes stability.}}}} \right)^2  }  \right)
\end{equation}

Since the sharpness of the deconvolution helps determine the confidence level contours on the final parameter estimation the prospect of a better deconvolution methods is very appealing. As the earth model being used in this parameter search is a simple layered model and the data is piecewise smooth the wavelet transform could be effectively used to promote sparsity. If we could leverage this fact and use an L1 technique which promotes sparse solutions we may be able to increase the accuracy of the deconvolved receiver functions. 
%Deconvolution is a common yet difficult step in processing and analysis in many fields of science and technology. In the field of seismology deconvolution arises more often than most and due to the structural complexity and ambiguity of the earth poses a particularly challenging problem.

%%% End of body of article:

\section{Method}

We now outline the framework under which the deconvolution problem was performed. The general convolution problem in seismology can be written as

\begin{equation}
	u(t) = s(t)\ast g(t) + n
\end{equation}

where $u(t)$ is the ground displacement recorded at seismograms, $s(t)$ is the source function, $n$ is noise and $g(t)$ is the receiver function which we aim to obtain through a deconvolution of the source function from recorded seismograms. In this problem recorded seismograms were arranged into bins separated by slowness. Recorded seismograms with similar slownesses should provide receiver functions which are approximately the same, therefore binning them allows for the development of the deconvolution problem as an optimization problem. In an ideal case the receiver function would be represented by some type of spike train, however in the real Earth receiver functions will be noisier and blurred. As a result the receiver function can be considered to be piecewise smooth, this motivates the use of wavelets as a sparsifying transform.

Having identified the sparsifying transform to be used we represent the forward problem as a linear system.
\begin{equation}
\begin{aligned}
\textbf{y}& =\textbf{A}\textbf{x}\\
\textbf{y}& =[\mathcal{F}^{T}\mathcal{F}(s)\mathcal{F}(g)]\mathcal{W}^{T}\textbf{x}
\end{aligned}
\end{equation}
where $\textbf{y}$ is a vector which holds $n$ seismograms which are contained in any given $p$ bin. $\textbf{A}$ is a matrix which is defined in Eq. 5. and $\textbf{x}$ is a vector 
which represents the wavelet coefficients of the receiver function $g(t)$. The action of the $\textbf{A}$ matrix is to first perform the inverse wavelet transform on $\textbf{x}$, and then perform a convolution between $s(t)$ and $g(t)$ in the Fourier domain. The final result would be the recorded seismograms $\textbf{y}$ in the time domain. Rather than forming explicit matrices spot operators are used and combined using kronecker products.

The convex optimization problem is constructed as a basis persuit denoise problem,

\begin{equation}
\begin{aligned}
& \underset{x}{\text{minimize}}
& & \lVert\textbf{x}\rVert_{1} \\
& \text{subject to}
& & \lVert\textbf{Ax}-\textbf{y}\rVert_{2} \leq \sigma
\end{aligned}
\end{equation}

where $\sigma$ is an estimate of the required data misfit and is taken to be the norm of the noise in the case. We solve an optimization problem which promotes a sparse representation of the receiver function in the wavelet domain.

\section{Synthetic Trials}

Before conducting the deconvolution problem on a complete data set we created a synthetic example and tested our L1 based algorithm. The problem used windowed $P$ wave arrivals from 5 seismograms as the source functions which were convolved with a synthetic receiver function to produce synthetic seismograms. 10\% gaussian noise was added to the signal. All L1 optimization problems in this paper were solved using SPGL1, for the synthetic cases the solver was allowed to run to completion. We tested a range of different wavelet types during the trials namely Haar and Daubechies wavelets, for which shift invariant versions were used. Orthonormal and B spline wavelets were also used.

Ideally our synthetic receiver function should have been a piecewise smooth function rather than the current spike representation which is sparse in the dirac basis, and not necessarily sparse in the wavelet domain.

\begin{figure}
\noindent\includegraphics[width=20pc]{SNRwavelets.eps}
\caption{Synthetic Results: Actual receiver shown in first plot followed by the solution recovered by solving a wavelet based L1 optimization with Haar and Daubechies wavelets. Final plot is solution from an L2 based simulataneous deconvolution.}
\end{figure}

We used SNR $= -20 log \frac{\lVert \textbf{y} - \textit{w}^{T}\textbf{x} \rVert}{\lVert \textbf{y} \rVert}$ as an initial measure of quality for the recovered receiver function, however this penalized solutions which were less sparse and can be seen in the low SNR for the L2 deconvolution. Whilst having a sparse receiver function was important, a more important requirement was that the recovered peaks be as sharp as possible, a property which allows for better constrained results when performing the grid search for $V_{P}/V_{S}$ and depth to the moho at later stages. We did a simple comparison of the peak recovery for all wavelet families to determine how the performance was in this scenario.

\begin{figure}
\noindent\includegraphics[width=20pc]{SNRsplines.eps}
\caption{ Synthetic Results II: Same as above but 2nd and 3rd plots are solutions recovered using orthonormal and B spline wavelets respectively.}
\end{figure}

All methods provided a good recovery of the receiver function, the two shift invariant wavelet transforms provided the sparsest recovery for the receiver function, however they also recovered broad peaks. Both spline wavelets recoverd narrower peaks but had a less sparse recovery.

\begin{figure}
\noindent\includegraphics[width=20pc]{peak_1.eps}
\caption{Comparison of peak recovery. Haar and Daubechies wavelets recover broader peaks, a less desirable result. B-Spline wavelets appear to provide the best recovery by recovering a narrower, better located peak when compared to ortho-spline wavelets.}
\end{figure}

Through a comparison of peak recovery, shown in Fig. 5, it was deemed that B-spline wavelets provided the most desirable properties in that the solution was sparse whilst recovering a narrow and well located peak. 
\section{Results}

The data consisted of seismograms split into 36 $p$ bins with the number of seismograms per bin varying between 1-8. For each $p$ bin the L1 optimization problem defined in Eq. 6 was solved using 100 iterations of SPGL1. The initial guess provided to SPGL1 was the wavelet transformed solution obtained from the L2 simulataneous deconvolution. Each solution from SPGL1 provided the wavelet coefficients of a receiver function for the corresponding $p$ bin from which the receiver function in the time domain was obtained. 3 different wavelet types were tested on the entire data set, these consisted of shift invariant Daubechies wavelets along with orthonormal and B spline wavelets. Once all 36 receiver functions were collected for a single wavelet type, the traces were collected by bin and a grid search was conducted to determine $V_{P}$ ans $V_{S}$ values and a line search was conducted to determine an estimate for the depth to the moho.

Fig. 6 shows the receiver function section obtained through the L2 based deconvolution method and acts as a benchmark for our comparisons. The results obtained through L2 deconvolution provide relatively sharp spikes in the recovered receiver functions. The error contours on $V_{P}$ v $V_{P}/V_{S}$ are also quite tightly constrained with a small standard error.

\begin{figure}
\noindent\includegraphics[width=20pc]{Rec_section_L2decon.eps}
\caption{Receiver function section for L2 deconvolution and error contours for $V_{P}$ v $V_{P}/V_{S}$. Receiver functions traces are plotted for all 36 bins with the result showm here. Also shown is a plot of $V_{P}$ v $R$ where $R=V_{P}$ v $V_{P}/V_{S}$. Line search results for depth are also plotted shown with a depth estimate obtained at H = 36.93km.}
\end{figure}

The first wavelet based L1 trial that we ran used shift invariant Daubechies wavelets. The receiver function section shows results consistent with our earlier synthetic trials showing a continuation of two properties, the first being that the recovered receiver functions are noticeabley sparser than the L2 deconvolution. However the peak recovery is not as sharp, just as in the synthetic trials, and this leads to poorly constrained $V_{P}$ as shown by the larger error contours in Fig. 7. The recovered depth to the moho, $H(km)$, was significantly different from that obtained via the L2 method.

\begin{figure}
\noindent\includegraphics[width=20pc]{Rec_section_daub.eps}
\caption{Receiver function section for L1 deconvolution using Daubechies wavelets. Receiver function section is noticeabley sparser than the L2 section, however the recovered peaks are also broader which leads to poorly constrained $V_{P}$ v $V_{P}/V_{S}$ in comparison. H = 34.30km}
\end{figure}

Ortho splines produced results similar to B-splines but with broader error contours for $V_{P}$ v $V_{P}/V_{S}$ and also provided a depth estimate less than that from L2 deconvolution. Both ortho and B splines (Fig. 8) improved upon Daubechies wavelets by providing sharper peak recoveries in the receiver functions and better constraints on the $V_{P}$ v $V_{P}/V_{S}$ error contours. Both also provided depth estimates closer to the L2 depth estimate. The B-spline depth estimate was the closest to the L2 estimate at H = 38.82km.

\begin{figure}
\noindent\includegraphics[width=20pc]{Rec_section_bspline.eps}
\caption{Receiver function section for L1 deconvolution using B-spline wavelets. Receiver function section provides narrower peak recovery and improves the error contour constraints. The depth estimate is also closer to that from L2 at H = 38.82km }
\end{figure}

At present time the current wavelet framework has failed to produce results which are as tightly constrained as the L2 deconvolution method, however early results show promise and can likely be improved with some more sophisticated input.

\section{Conclusion}

We aimed to obtain a receiver function by deconvolving multiple source functions from multiple recorded seismograms which were grouped based on their slowness $p$. We developed a deconvolution framework based on L1 optimization and compared results to previous work using an L2 based method. We performed deconvolution by performing an inverse problem with an L1 regularization, promoting sparsity of the receiver function in the wavelet domain. SPGL1 was used to solve the convex optimization problem and was conducted for a range of different wavelet types including Daubechies, ortho-spline and B-spline wavelets. Receiver function solutions were plotted by bin and used to obtain $V_{P}$ and $V_{s}$ along with a depth estimate for the moho. In general Daubechies wavelets provided a very sparse representation of the receiver functions but failed to recover a sharp peak which lead to large error contours in $V_{P}$ v $V_{P}/V_{S}$. Spline wavelets improved by producing narrower recovered peaks and improved error estimates however at present, all L1 wavelet methods failed to produce $V_{P}$ v $V_{P}/V_{S}$ constraints which were as tightly bound as the results derived using the L2 method.

\section{Future Work}
Even with a fairly preliminary L1 minimization setup the results are encouraging. Further work could focus on the 2D receiver function plots and usage of curvelets to interpolate between poor data. This would significantly improve the estimates made by the grid search. While redundant wavelet dictionaries were employed in the non-spline cases these authors are currently unsure if the B-spline methods were shift-invariant or not. This will require further investigation. Also during a recent presentation of the work contained in this report Felix Herrmann suggested looking at Radon Transformations. Both authors are unsure of the utility of these transformations and this will also require further research. However, with a sophisticated tool-chain in place we believe that L1 minimization and the utilization of sparsity inherent in this problem will result in better deconvolution than the current L2 method being employed. Since there is a large body of global seismological work which has been built on results from L2 deconvolution there is a lot of potential to apply these methods to other problems. Along the same lines, once sparsity is being utilized, the larger compressive sensing framework should surely find a place in the global seismic research community. There is vast potential in this arena and likely many exciting avenues for phd research. Both the authors of this report are interested in exploring this new paradigm further in the context of global seismology.

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing  \appendix
%
% \section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% Or,
%
% If you use BiBTeX for your references, please produce your .bbl
% file and copy the contents into your paper here.
%
% Follow these steps:
% 1. Run LaTeX on your LaTeX file.
%
% 2. Run BiBTeX on your LaTeX file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Comment out the old \bibliographystyle and \bibliography commands.
%
% 5. Run LaTeX on your new file before submitting.
%
% AGU DOES NOT WANT a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

\begin{thebibliography}{}

\bibitem[{\textit{Herrmann}(2012)}]{herrmann}
Herrmann, F. J., (2012), Class material from EOS 513: Imaging and Estimation with Wavelets, {\it University of British Columbia}

\bibitem[{\textit{Bostock}(2010)}]{bostock}
Bostock, M. G., Kumar, M. R., (2010), Bias in seismic estimates of crustal properties, {\it Geophysical Journal International,} \textit{182}, 403--407.

\bibitem[{\textit{Thompson et al}(2010)}]{thompson}
Thompson, D. A., Bastow, I.D., Helffrich, G., Kendall, J-M., Wookey, J., Snyder, D.B., Eaton, D.W., (2010), Precambrian crustal evolution: Seismic constraints from the Canadian Shield, {\it Earth and Planetary Science Letters,} \textit{297}, 655--666.

\bibitem[{\textit{Kanamori et al.}(2000)}]{kanamori}
Kanamori, H. (2000), Moho depth variation in southern California from teleseismic receiver functions, {\it Geophysical Research,} \textit{105}, 2969--2980.

\end{thebibliography}%Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%
% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

%% ------------------------------------------------------------------------ %%
%
%  END ARTICLE
%
%% ------------------------------------------------------------------------ %%

\end{article}

%% Enter Figures and Tables here:

% When submitting articles through the GEMS system:
% COMMENT OUT ANY COMMANDS THAT INCLUDE GRAPHICS.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information
%  should be placed in footnotes below the table.

% DRAFT figure/table, including eps graphics
%
%
% \begin{table}
% \caption{}
% \end{table}
%
% ---------------
% TWO-COLUMN figure/table
%
% \begin{figure*}
% \noindent\includegraphics[width=39pc]{samplefigure.eps}
% \caption{Caption text here}
% \end{figure*}
%
% \begin{table*}
% \caption{Caption text here}
% \end{table*}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% See below for how to make landscape/sideways figures or tables.

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  LANDSCAPE/SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For figures, add \usepackage{lscape} to the file and the landscape.sty style file
% to the paper folder.
%
% \begin{figure*}[p]
% \begin{landscapefigure*}
% Illustration here.
% \caption{caption here}
% \end{landscapefigure*}
% \end{figure*}
%
% For tables, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
%
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}
%
%

### 2012-07-10 matlab/L1/report/benreport.tex

--- 2012-07-10 matlab/L1/report/Paper_EOS513.tex
@@
 
-\authors{G. Matharu, B. Postlethwaite \altaffilmark{1}}
+% \authors{R. C. Bales,\altaffilmark{1}
 % E. Mosley-Thompson,\altaffilmark{2} R. Williams,\altaffilmark{3}
@@
 
-\altaffiltext{1}{Department of Earth and Ocean Sciences,
-University of British Columbia, Vancouver, B.C. Canada.}
+%\altaffiltext{1}{Department of Hydrology and Water Resources,
+%University of Arizona, Tucson, Arizona, USA.}
 
@@
 \end{abstract}
+
 %% ------------------------------------------------------------------------ %%
@@
 %% ------------------------------------------------------------------------ %%
+
 \section{Introduction}
@@
 Poisson's ratio is given by:
-\begin{equation}
- \sigma = \frac{1}{2} \frac{ \left( \frac{ V_p }{ V_s } \right)^2 - 2 }{ \left( \frac{ V_p }{ V_s } \right)^2 - 1 }
-\end{equation}
+\[ \sigma = \frac{1}{2} \frac{ \left( \frac{ V_p }{ V_s } \right)^2 - 2 }{ \left( \frac{ V_p }{ V_s } \right)^2 - 1 } \]
 
 \begin{figure}
-\noindent\includegraphics[width=18pc]{PoissonRatio.eps}
+\noindent\includegraphics[width=18pc]{PoissonRatio.png}
 \caption{Poissons Ratio is the ratio of compression along one axis to extension along another}
@@
 \begin{figure}
-\noindent\includegraphics[width=18pc]{reflectRecs.eps}
+\noindent\includegraphics[width=18pc]{reflectRecs.jpg}
 \caption{Top: Shows direct phases Pp and Ps from a seismic event hitting receiver.
@@
 where
-\begin{equation}
- \bc{G}(\omega) = \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*}{\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right].
-\end{equation}
+\[ \bc{G}(\omega) = \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*}{\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right] \].
 Here $\bc{F}$ is the Fourier transform, $\bc{G}$ is the model, $\bt{S}$ is a matrix of S-wave seismograms, $\bt{P}$ is a matrix of P-wave source coda, $*$ is the complex conjugate and $\delta$ is the regularization parameter.
 $\delta$ is chosen by minimizing the Generalized Cross Validation function with respect to $\delta$:
-\begin{equation}
-\text{min} \left( \frac{ \sum_n^N \sum_m^M 
+\[ \text{min} \left( \frac{ \sum_n^N \sum_m^M 
  \left( \ob{ \bt{S}_n(\omega_m) - \bt{P}(\omega_m) \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*} {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right]}^{\text{\textcolor{red}{ Misfit: more accurate with $\delta\rightarrow0$.
 Less Stable }}} \right)^2}  { \left(
-\ub{  NM - \sum_m^M  \frac {\sum_n^N \bt{P}_n\bt{P}_n^*}  {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  }_{\text{\textcolor{red}{ {As $\delta$ increases denominator increases, promotes stability.}}}} \right)^2  }  \right)
-\end{equation}
+\ub{  NM - \sum_m^M  \frac {\sum_n^N \bt{P}_n\bt{P}_n^*}  {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  }_{\text{\textcolor{red}{ {As $\delta$ increases denominator increases, promotes stability.}}}} \right)^2  }  \right)  \]
 
@@
 If we could leverage this fact and use an L1 technique which promotes sparse solutions we may be able to increase the accuracy of the deconvolved receiver functions.
-%Deconvolution is a common yet difficult step in processing and analysis in many fields of science and technology.
-In the field of seismology deconvolution arises more often than most and due to the structural complexity and ambiguity of the earth poses a particularly challenging problem.
-
-%%% End of body of article:
-
-\section{Method}
-
-We now outline the framework under which the deconvolution problem was performed. The general convolution problem in seismology can be written as
-
-\begin{equation}
-	u(t) = s(t)\ast g(t) + n
-\end{equation}
-
-where $u(t)$ is the ground displacement recorded at seismograms, $s(t)$ is the source function, $n$ is noise and $g(t)$ is the receiver function which we aim to obtain through a deconvolution of the source function from recorded seismograms.
-In this problem recorded seismograms were arranged into bins separated by slowness.
-Recorded seismograms with similar slownesses should provide receiver functions which are approximately the same, therefore binning them allows for the development of the deconvolution problem as an optimization problem.
-In an ideal case the receiver function would be represented by some type of spike train, however in the real Earth receiver functions will be noisier and blurred.
-As a result the receiver function can be considered to be piecewise smooth, this motivates the use of wavelets as a sparsifying transform.
-
-Having identified the sparsifying transform to be used we represent the forward problem as a linear system.
-\begin{equation}
-\begin{aligned}
-\textbf{y}& =\textbf{A}\textbf{x}\\
-\textbf{y}& =[\mathcal{F}^{T}\mathcal{F}(s)\mathcal{F}(g)]\mathcal{W}^{T}\textbf{x}
-\end{aligned}
-\end{equation}
-where $\textbf{y}$ is a vector which holds $n$ seismograms which are contained in any given $p$ bin. $\textbf{A}$ is a matrix which is defined in Eq. 5. and $\textbf{x}$ is a vector 
-which represents the wavelet coefficients of the receiver function $g(t)$.
-The action of the $\textbf{A}$ matrix is to first perform the inverse wavelet transform on $\textbf{x}$, and then perform a convolution between $s(t)$ and $g(t)$ in the Fourier domain.
-The final result would be the recorded seismograms $\textbf{y}$ in the time domain.
-Rather than forming explicit matrices spot operators are used and combined using kronecker products.
-
-The convex optimization problem is constructed as a basis persuit denoise problem,
-
-\begin{equation}
-\begin{aligned}
-& \underset{x}{\text{minimize}}
-& & \lVert\textbf{x}\rVert_{1} \\
-& \text{subject to}
-& & \lVert\textbf{Ax}-\textbf{y}\rVert_{2} \leq \sigma
-\end{aligned}
-\end{equation}
-
-where $\sigma$ is an estimate of the required data misfit and is taken to be the norm of the noise in the case.
-We solve an optimization problem which promotes a sparse representation of the receiver function in the wavelet domain.
-
-\section{Synthetic Trials}
-
-Before conducting the deconvolution problem on a complete data set we created a synthetic example and tested our L1 based algorithm.
-The problem used windowed $P$ wave arrivals from 5 seismograms as the source functions which were convolved with a synthetic receiver function to produce synthetic seismograms.
-10\% gaussian noise was added to the signal.
-All L1 optimization problems in this paper were solved using SPGL1, for the synthetic cases the solver was allowed to run to completion.
-We tested a range of different wavelet types during the trials namely Haar and Daubechies wavelets, for which shift invariant versions were used.
-Orthonormal and B spline wavelets were also used.
-
-Ideally our synthetic receiver function should have been a piecewise smooth function rather than the current spike representation which is sparse in the dirac basis, and not necessarily sparse in the wavelet domain.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{SNRwavelets.eps}
-\caption{Synthetic Results: Actual receiver shown in first plot followed by the solution recovered by solving a wavelet based L1 optimization with Haar and Daubechies wavelets.
-Final plot is solution from an L2 based simulataneous deconvolution.}
-\end{figure}
-
-We used SNR $= -20 log \frac{\lVert \textbf{y} - \textit{w}^{T}\textbf{x} \rVert}{\lVert \textbf{y} \rVert}$ as an initial measure of quality for the recovered receiver function, however this penalized solutions which were less sparse and can be seen in the low SNR for the L2 deconvolution.
-Whilst having a sparse receiver function was important, a more important requirement was that the recovered peaks be as sharp as possible, a property which allows for better constrained results when performing the grid search for $V_{P}/V_{S}$ and depth to the moho at later stages.
-We did a simple comparison of the peak recovery for all wavelet families to determine how the performance was in this scenario.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{SNRsplines.eps}
-\caption{ Synthetic Results II: Same as above but 2nd and 3rd plots are solutions recovered using orthonormal and B spline wavelets respectively.}
-\end{figure}
-
-All methods provided a good recovery of the receiver function, the two shift invariant wavelet transforms provided the sparsest recovery for the receiver function, however they also recovered broad peaks.
-Both spline wavelets recoverd narrower peaks but had a less sparse recovery.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{peak_1.eps}
-\caption{Comparison of peak recovery.
-Haar and Daubechies wavelets recover broader peaks, a less desirable result.
-B-Spline wavelets appear to provide the best recovery by recovering a narrower, better located peak when compared to ortho-spline wavelets.}
-\end{figure}
-
-Through a comparison of peak recovery, shown in Fig.
-5, it was deemed that B-spline wavelets provided the most desirable properties in that the solution was sparse whilst recovering a narrow and well located peak.
-\section{Results}
-
-The data consisted of seismograms split into 36 $p$ bins with the number of seismograms per bin varying between 1-8.
-For each $p$ bin the L1 optimization problem defined in Eq.
-6 was solved using 100 iterations of SPGL1.
-The initial guess provided to SPGL1 was the wavelet transformed solution obtained from the L2 simulataneous deconvolution.
-Each solution from SPGL1 provided the wavelet coefficients of a receiver function for the corresponding $p$ bin from which the receiver function in the time domain was obtained.
-3 different wavelet types were tested on the entire data set, these consisted of shift invariant Daubechies wavelets along with orthonormal and B spline wavelets.
-Once all 36 receiver functions were collected for a single wavelet type, the traces were collected by bin and a grid search was conducted to determine $V_{P}$ ans $V_{S}$ values and a line search was conducted to determine an estimate for the depth to the moho.
-
-Fig.
-6 shows the receiver function section obtained through the L2 based deconvolution method and acts as a benchmark for our comparisons.
-The results obtained through L2 deconvolution provide relatively sharp spikes in the recovered receiver functions.
-The error contours on $V_{P}$ v $V_{P}/V_{S}$ are also quite tightly constrained with a small standard error.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_L2decon.eps}
-\caption{Receiver function section for L2 deconvolution and error contours for $V_{P}$ v $V_{P}/V_{S}$.
-Receiver functions traces are plotted for all 36 bins with the result showm here.
-Also shown is a plot of $V_{P}$ v $R$ where $R=V_{P}$ v $V_{P}/V_{S}$.
-Line search results for depth are also plotted shown with a depth estimate obtained at H = 36.93km.}
-\end{figure}
-
-The first wavelet based L1 trial that we ran used shift invariant Daubechies wavelets.
-The receiver function section shows results consistent with our earlier synthetic trials showing a continuation of two properties, the first being that the recovered receiver functions are noticeabley sparser than the L2 deconvolution.
-However the peak recovery is not as sharp, just as in the synthetic trials, and this leads to poorly constrained $V_{P}$ as shown by the larger error contours in Fig.
-7.
-The recovered depth to the moho, $H(km)$, was significantly different from that obtained via the L2 method.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_daub.eps}
-\caption{Receiver function section for L1 deconvolution using Daubechies wavelets.
-Receiver function section is noticeabley sparser than the L2 section, however the recovered peaks are also broader which leads to poorly constrained $V_{P}$ v $V_{P}/V_{S}$ in comparison.
-H = 34.30km}
-\end{figure}
-
-Ortho splines produced results similar to B-splines but with broader error contours for $V_{P}$ v $V_{P}/V_{S}$ and also provided a depth estimate less than that from L2 deconvolution.
-Both ortho and B splines (Fig.
-8) improved upon Daubechies wavelets by providing sharper peak recoveries in the receiver functions and better constraints on the $V_{P}$ v $V_{P}/V_{S}$ error contours.
-Both also provided depth estimates closer to the L2 depth estimate.
-The B-spline depth estimate was the closest to the L2 estimate at H = 38.82km.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_bspline.eps}
-\caption{Receiver function section for L1 deconvolution using B-spline wavelets.
-Receiver function section provides narrower peak recovery and improves the error contour constraints.
-The depth estimate is also closer to that from L2 at H = 38.82km }
-\end{figure}
-
-At present time the current wavelet framework has failed to produce results which are as tightly constrained as the L2 deconvolution method, however early results show promise and can likely be improved with some more sophisticated input.
-
-\section{Conclusion}
-
-We aimed to obtain a receiver function by deconvolving multiple source functions from multiple recorded seismograms which were grouped based on their slowness $p$.
-We developed a deconvolution framework based on L1 optimization and compared results to previous work using an L2 based method.
-We performed deconvolution by performing an inverse problem with an L1 regularization, promoting sparsity of the receiver function in the wavelet domain.
-SPGL1 was used to solve the convex optimization problem and was conducted for a range of different wavelet types including Daubechies, ortho-spline and B-spline wavelets.
-Receiver function solutions were plotted by bin and used to obtain $V_{P}$ and $V_{s}$ along with a depth estimate for the moho.
-In general Daubechies wavelets provided a very sparse representation of the receiver functions but failed to recover a sharp peak which lead to large error contours in $V_{P}$ v $V_{P}/V_{S}$.
-Spline wavelets improved by producing narrower recovered peaks and improved error estimates however at present, all L1 wavelet methods failed to produce $V_{P}$ v $V_{P}/V_{S}$ constraints which were as tightly bound as the results derived using the L2 method.
 
@@
 
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Deconvolution is a common yet difficult step in processing and analysis in many fields of science and technology.
+In the field of seismology deconvolution arises more often than most and due to the structural complexity and ambiguity of the earth poses a particularly challenging problem.
+
+%%% End of body of article:
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Optional Appendix goes here
@@
 %  ACKNOWLEDGMENTS
+
+\begin{acknowledgments}
+(Text here)
+\end{acknowledgments}
 
@@
 
-\end{thebibliography}%Reference citation examples:
+%\bibitem[{\textit{Kilby et al.}(2008)}]{jskilbye}
+%Kilby, J. S., S. Smith, and R. Jones (2008), Invention of the
+%integrated circuit, {\it IEEE Trans. Electron Devices,} \textit{23},
+%648--650.
+
+\end{thebibliography}
+
+%Reference citation examples:
 
@@
 %% Enter Figures and Tables here:
+%\begin{figure*}
+%\noindent\includegraphics[width=10pc]{PoissonRatio.png}
+%\caption{Poissons Ratio is the ratio of compression along one axis to extension along another}
+%\end{figure*}
 
@@
 
+\end{document}
+
 % DRAFT figure/table, including eps graphics
 %
+% \begin{figure}
+% \noindent\includegraphics[width=20pc]{samplefigure.eps}
+% \caption{Caption text here}
+% \end{figure}
+% \end{document}
 %
@@
 
-More Information and Advice:
+%More Information and Advice:
 
@@
 
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
+%Math coded inside display math mode \[ ...\]
+% will not be numbered, e.g.,:
+% \[ x^2=y^2 + z^2\]
 
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
+% Math coded inside \begin{equation} and \end{equation} will
+% be automatically numbered, e.g.,:
+% \begin{equation}
+% x^2=y^2 + z^2
+% \end{equation}
 
@@
 % as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
+%\begin{eqnarray}
+%  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+%        && + (y - y_{0}) \sin \Theta  \nonumber \\
+%  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+%        && + (y - y_{0}) \cos \Theta.
+%\end{eqnarray}
 

### 2012-08-08 python/dbutils.py
updates, more picks blah blah

@@ -11,7 +11,7 @@
 ###########################################################################
 # IMPORTS
 ###########################################################################
-import json, os, argparse, sys, time
+import json, os, argparse, sys, time, re
 from preprocessor import is_number
 from collections import defaultdict
 import shapefile
@@ -281,9 +281,10 @@ if __name__== '__main__' :
     # Append piped data if there is any
     # If we pipe a bunch of stations to program, query only these stations
     if not sys.stdin.isatty():
-        args.pipedData = sys.stdin.read().split('\n')
+        # trick to seperate a newline or space seperated list
+        args.pipedData =  re.findall(r'\w+', sys.stdin.read() )
     else:
-        args.pipedData = None
+        args.pipedData = False
 
     if args.update:
         updateStats(stdict, args)

### 2013-01-10 paper/draftBP.tex
paper and source stacking

--- 2012-07-10 matlab/L1/report/Paper_EOS513.tex
@@
-% EOS 513 REPORT
-\documentclass[jgrga]{agutex}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{color}
+%% ------------------------------------------------------------------------ %%
+\documentclass[draft]{article}
 
-\newcommand{\bt}[1]{\textbf{#1}}
-\newcommand{\bc}[1]{\mathcal{\textbf{#1}}}
-\newcommand{\ub}[1]{\underbrace{#1}}
-\newcommand{\ob}[1]{\overbrace{#1}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% OPTIONAL:
-% To produce a two-columned version:
-% \documentclass[jgrga]{AGUTeX}
+\title{Sesimic parameter estimation and the Canadian crust}
+\author{Ben Postlethwaite}
 
-% Two-columned format can be used to estimate the number of pages
-% for the final published PDF.
-
-% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% OPTIONAL:
-% To create numbered lines:
-
-% If you don't already have lineno.sty, you can download it from
-% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
-% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
-% Take care that you always use the latest version.
-
-% To activate the commands, uncomment \usepackage{lineno}
-% and \linenumbers*[1]command, below:
-
-% \usepackage{lineno}
-% \linenumbers*[1]
-
-%  To add line numbers to lines with equations:
-
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figures and Tables
-%
-% When submitting articles through the GEMS system:
-% COMMENT OUT ANY COMMANDS THAT INCLUDE GRAPHICS.
-% (See FIGURES section near the end of the file.)
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-%  Figures and tables should be placed AT THE END OF THE ARTICLE,
-%  after the references.
-%
-%  Uncomment the following command to include .eps files
-%  (comment out this line for draft format):
-% \usepackage[dvips]{graphicx}
-%
-%  Uncomment the following command to allow illustrations to print
-%   when using Draft:
-%  \setkeys{Gin}{draft=false}
-%
-% Substitute one of the following for [dvips] above
-% if you are using a different driver program and want to
-% proof your illustrations on your machine:
-%
-% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
-% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
-% [oztex], [textures]
-%
-% See how to enter figures and tables at the end of the article, after
-% references.
-%
 %% ------------------------------------------------------------------------ %%
-%
-%  ENTER PREAMBLE
-%
-%% ------------------------------------------------------------------------ %%
-
-% Author names in capital letters:
-\authorrunninghead{MATHARU GIAN, BEN POSTLETHWAITE}
-
-% Shorter version of title entered in capital letters:
-\titlerunninghead{DECONVOLUTION L2 / L1 COMPARISON}
-
-% Author mailing address: please repeat this command for
-% each author and alphabetize authors:
-
-%\authoraddr{R. C. Bales,
-%Department of Hydrology and Water Resources, University of
-%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
-%(roger@hwr.arizona.edu)}
-
-%\authoraddr{J. R. McConnell, Division of Hydrologic
-%Sciences, 123 Main Street, Desert Research Institute, Reno, NV
-%89512, USA.}
-
-%\authoraddr{E. Mosley-Thompson, Department of Geography,
-%Ohio State University, 123 Orange Boulevard, Columbus, OH 43210,
-%USA.}
-
-%\authoraddr{R. Williams, Department of Space Sciences, University of
-%Michigan, 123 Brown Avenue, Ann Arbor, MI 48109, USA.}
-
-%\authoraddr{Francesco Visconti, Dipartimento di Idraulica,
-%Trasporti ed Infrastrutture Civili, Politecnico di Torino,
-%Corso Duca degli Abruzzi 24, I-10129, Torino, Italy.
-
 \begin{document}
@@
 %% ------------------------------------------------------------------------ %%
-%
-%  TITLE
-%
+\begin{abstract}
+(Type abstract here)
+\end{abstract}
+
 %% ------------------------------------------------------------------------ %%
 
-\title{Sparsity Promoting L1 Deconvolution: Application for a Global Seismology Parameter Search}
-%
-% e.g., \title{Terrestrial ring current:
-% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
-%
+%% ------------------------------------------------------------------------ %%
+%% -----------------------------------------------%%
+\section{Introduction}
+(Article text here)
+\subsection{Overview}
+(Overview here)
+\subsection{Geological Background}
+(Geological background here)
+
+%% -----------------------------------------------%%
+\section{Data and method}
+(Data explanation here)
+\subsection{Vp methods}
+(Vp method here)
+\subsection{Other methods}
+(Other methods here)
+
+%% -----------------------------------------------%%
+\section{Results}
+{Results go here}
+
+%% -----------------------------------------------%%
+\section{Discussion}
+\subsection{Canada}
+(Discussion on larger - Canada wide - regions)
+\subsection{Slave Province}
+(Discussion on Slave Province)
+
+%% -----------------------------------------------%%
+\section{Conclusions}
+(Conclusions here)
 
 %% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
 
-%Use \author{\altaffilmark{}} and \altaffiltext{}
-
-% \altaffilmark will produce footnote;
-% matching \altaffiltext will appear at bottom of page.
-
-\authors{G. Matharu, B. Postlethwaite \altaffilmark{1}}
-% E. Mosley-Thompson,\altaffilmark{2} R. Williams,\altaffilmark{3}
-% J. R. McConnell\altaffilmark{4}, and Francesco Visconti\altaffilmark{5}}
-
-\altaffiltext{1}{Department of Earth and Ocean Sciences,
-University of British Columbia, Vancouver, B.C. Canada.}
-
-%\altaffiltext{2}{Department of Geography, Ohio State University,
-%Columbus, Ohio, USA.}
-
-%\altaffiltext{3}{Department of Space Sciences, University of
-%Michigan, Ann Arbor, Michigan, USA.}
-
-%\altaffiltext{4}{Division of Hydrologic Sciences, Desert Research
-%Institute, Reno, Nevada, USA.}
-
-%\altaffiltext{5}{Dipartimento di Idraulica, Trasporti ed
-%Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-%% ------------------------------------------------------------------------ %%
-
-% >> Do NOT include any \begin...\end commands within
-% >> the body of the abstract.
-
-\begin{abstract}
-This report will outline an attempt to utilize the inherent sparsity in seismic receiver functions in order to provide better deconvolution results than the often utilized damped-L2 methods.
-By promoting sparsity with the wavelet transform the use of sparse solvers such as SPGL1 becomes possible.
-Synthetic tests are performed using the damped L2 convolution as well as the sparse wavelet transform L1 minimization using several wavelet basis.
-Real data is employed and the deconvolved results from the competing methods are used for an actual parameter grid search.
-The final results of the grid search are then compared.
-Further work on the sparse-in-wavelets L1 algorthim will be required if it is to beat the damped-L2 algorthim currently being employed.
-\end{abstract}
-%% ------------------------------------------------------------------------ %%
-%
-%  BEGIN ARTICLE
-%
-%% ------------------------------------------------------------------------ %%
-
-% The body of the article must start with a \begin{article} command
-%
-% \end{article} must follow the references section, before the figures
-%  and tables.
-
-\begin{article}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-\section{Introduction}
-
-Extracting material properties from seismic data spans industry and the research community.
-In the global seismic community material properties of the crust have been sought to validate existing geological models or provide clues for new ones.
-A good example of this is Poisson's ratio, which is the ratio between compression along one axis to extension on another when a material is stretched (see figure 1).
-Poisson's ratio is given by:
-\begin{equation}
- \sigma = \frac{1}{2} \frac{ \left( \frac{ V_p }{ V_s } \right)^2 - 2 }{ \left( \frac{ V_p }{ V_s } \right)^2 - 1 }
-\end{equation}
-
-\begin{figure}
-\noindent\includegraphics[width=18pc]{PoissonRatio.eps}
-\caption{Poissons Ratio is the ratio of compression along one axis to extension along another}
-\end{figure}
-
-where the seismic velocities $V_p$ and $V_s$ determine the property entirely.
-Because of the direct link between seismic properties and Poisson's ratio this material property has been sought in numerous seismic studies such as Kanamori (2000) and Snyder (2010).
-All of these studies have performed a parameter search for $R = \frac{V_p}{V_s}$ and provided a background estimate for $V_p$.
-This allows for fast computation through a grid search and avoids the problem of the low-dependency of $V_p$ in the grid search equations.
-Although Poisson's ratio can be estimated from $R$ better geological constraints would be possible if $V_p$ and $V_s$ were resolved uniquely.
-
-\section{The Problem}
-
-In order to differentiate between $V_p$ and $V_s$ the travel times from seperate arrivals of energy must be compared.
-The first arrival a seismic station will pick up after the onset of an earthquake is from the higher velocity P-wave.
-The impulsive character and high amplitude of the P-wave arrival allow it to be windowed and filtered and used as an approximation for the source.
-There are three more arrivals of interest.
-The Ps arrival, which is P-energy converted to S-wave energy at the Moho boundary between the crust and the mantle.
-Two further arrivals follow which are the reflections from the free surface which have reflected again off the Moho and recorded at the station, figure (2).
-These reflected arrivals are usually of lower amplitude and often less well defined in the seismogram.
-If the three S-wave arrivals - the Ps, PpPs and PsPs - can be located in time this provides the data in which a model may be fit.
-Owing to the fact that the S-wave phases all travel within the same crustal thickess $H$, dividing the S-wave relected phases by the initial Ps phase will remove the dependence on crustal thickness.
-The traveltime ratios are then dependent on $V_p$ and $V_s$ only (Bostock, 2010).
-With this information a 2D grid search over the travel time equations with respect to $V_p$ and $V_s$ can be performed.
-The parameter $V_p$ is not very sensitive to the traveltime differences between the S-wave phases, thus a very impulsive deconvolution result is needed to steepen the gradient in $V_p$ space.
-
-\begin{figure}
-\noindent\includegraphics[width=18pc]{reflectRecs.eps}
-\caption{Top: Shows direct phases Pp and Ps from a seismic event hitting receiver.
-Also the reflected phases PpPs and PsPs are shown.
-The phases ending with 's' are the phases hitting the station as S-waves.
-These S-wave phases are used as the primary data, the direct Pp phase is used as the source, and the two are deconvolved to attain the receiver functions shown at Bottom: The receiver functions attained from deconvolution.
-Depending on the quality of the data and the sharpness of the deconvolution, the results may look like those given for the Rae Domain -better- or the Hearne Domain -worse.}
-\end{figure}
-
-\section{Deconvolution - L2}
-Currently simultaneous deconvolution with generalized cross validation is being employed as it performs well and is very fast.
-Since division in the fourier basis is deconvolution in time we can write:
-\[ \bc{R} = \bc{F}^{-1}[\bc{G}(\omega)]  \]
-where
-\begin{equation}
- \bc{G}(\omega) = \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*}{\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right].
-\end{equation}
-Here $\bc{F}$ is the Fourier transform, $\bc{G}$ is the model, $\bt{S}$ is a matrix of S-wave seismograms, $\bt{P}$ is a matrix of P-wave source coda, $*$ is the complex conjugate and $\delta$ is the regularization parameter.
-$\delta$ is chosen by minimizing the Generalized Cross Validation function with respect to $\delta$:
-\begin{equation}
-\text{min} \left( \frac{ \sum_n^N \sum_m^M 
- \left( \ob{ \bt{S}_n(\omega_m) - \bt{P}(\omega_m) \left[ \frac{\sum_n^N \bt{S}_n\bt{P}_n^*} {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  \right]}^{\text{\textcolor{red}{ Misfit: more accurate with $\delta\rightarrow0$.
-Less Stable }}} \right)^2}  { \left(
-\ub{  NM - \sum_m^M  \frac {\sum_n^N \bt{P}_n\bt{P}_n^*}  {\sum_n^N \bt{P}_n\bt{P}_n^* + \delta}  }_{\text{\textcolor{red}{ {As $\delta$ increases denominator increases, promotes stability.}}}} \right)^2  }  \right)
-\end{equation}
-
-Since the sharpness of the deconvolution helps determine the confidence level contours on the final parameter estimation the prospect of a better deconvolution methods is very appealing.
-As the earth model being used in this parameter search is a simple layered model and the data is piecewise smooth the wavelet transform could be effectively used to promote sparsity.
-If we could leverage this fact and use an L1 technique which promotes sparse solutions we may be able to increase the accuracy of the deconvolved receiver functions.
-%Deconvolution is a common yet difficult step in processing and analysis in many fields of science and technology.
-In the field of seismology deconvolution arises more often than most and due to the structural complexity and ambiguity of the earth poses a particularly challenging problem.
-
-%%% End of body of article:
-
-\section{Method}
-
-We now outline the framework under which the deconvolution problem was performed. The general convolution problem in seismology can be written as
-
-\begin{equation}
-	u(t) = s(t)\ast g(t) + n
-\end{equation}
-
-where $u(t)$ is the ground displacement recorded at seismograms, $s(t)$ is the source function, $n$ is noise and $g(t)$ is the receiver function which we aim to obtain through a deconvolution of the source function from recorded seismograms.
-In this problem recorded seismograms were arranged into bins separated by slowness.
-Recorded seismograms with similar slownesses should provide receiver functions which are approximately the same, therefore binning them allows for the development of the deconvolution problem as an optimization problem.
-In an ideal case the receiver function would be represented by some type of spike train, however in the real Earth receiver functions will be noisier and blurred.
-As a result the receiver function can be considered to be piecewise smooth, this motivates the use of wavelets as a sparsifying transform.
-
-Having identified the sparsifying transform to be used we represent the forward problem as a linear system.
-\begin{equation}
-\begin{aligned}
-\textbf{y}& =\textbf{A}\textbf{x}\\
-\textbf{y}& =[\mathcal{F}^{T}\mathcal{F}(s)\mathcal{F}(g)]\mathcal{W}^{T}\textbf{x}
-\end{aligned}
-\end{equation}
-where $\textbf{y}$ is a vector which holds $n$ seismograms which are contained in any given $p$ bin. $\textbf{A}$ is a matrix which is defined in Eq. 5. and $\textbf{x}$ is a vector 
-which represents the wavelet coefficients of the receiver function $g(t)$.
-The action of the $\textbf{A}$ matrix is to first perform the inverse wavelet transform on $\textbf{x}$, and then perform a convolution between $s(t)$ and $g(t)$ in the Fourier domain.
-The final result would be the recorded seismograms $\textbf{y}$ in the time domain.
-Rather than forming explicit matrices spot operators are used and combined using kronecker products.
-
-The convex optimization problem is constructed as a basis persuit denoise problem,
-
-\begin{equation}
-\begin{aligned}
-& \underset{x}{\text{minimize}}
-& & \lVert\textbf{x}\rVert_{1} \\
-& \text{subject to}
-& & \lVert\textbf{Ax}-\textbf{y}\rVert_{2} \leq \sigma
-\end{aligned}
-\end{equation}
-
-where $\sigma$ is an estimate of the required data misfit and is taken to be the norm of the noise in the case.
-We solve an optimization problem which promotes a sparse representation of the receiver function in the wavelet domain.
-
-\section{Synthetic Trials}
-
-Before conducting the deconvolution problem on a complete data set we created a synthetic example and tested our L1 based algorithm.
-The problem used windowed $P$ wave arrivals from 5 seismograms as the source functions which were convolved with a synthetic receiver function to produce synthetic seismograms.
-10\% gaussian noise was added to the signal.
-All L1 optimization problems in this paper were solved using SPGL1, for the synthetic cases the solver was allowed to run to completion.
-We tested a range of different wavelet types during the trials namely Haar and Daubechies wavelets, for which shift invariant versions were used.
-Orthonormal and B spline wavelets were also used.
-
-Ideally our synthetic receiver function should have been a piecewise smooth function rather than the current spike representation which is sparse in the dirac basis, and not necessarily sparse in the wavelet domain.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{SNRwavelets.eps}
-\caption{Synthetic Results: Actual receiver shown in first plot followed by the solution recovered by solving a wavelet based L1 optimization with Haar and Daubechies wavelets.
-Final plot is solution from an L2 based simulataneous deconvolution.}
-\end{figure}
-
-We used SNR $= -20 log \frac{\lVert \textbf{y} - \textit{w}^{T}\textbf{x} \rVert}{\lVert \textbf{y} \rVert}$ as an initial measure of quality for the recovered receiver function, however this penalized solutions which were less sparse and can be seen in the low SNR for the L2 deconvolution.
-Whilst having a sparse receiver function was important, a more important requirement was that the recovered peaks be as sharp as possible, a property which allows for better constrained results when performing the grid search for $V_{P}/V_{S}$ and depth to the moho at later stages.
-We did a simple comparison of the peak recovery for all wavelet families to determine how the performance was in this scenario.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{SNRsplines.eps}
-\caption{ Synthetic Results II: Same as above but 2nd and 3rd plots are solutions recovered using orthonormal and B spline wavelets respectively.}
-\end{figure}
-
-All methods provided a good recovery of the receiver function, the two shift invariant wavelet transforms provided the sparsest recovery for the receiver function, however they also recovered broad peaks.
-Both spline wavelets recoverd narrower peaks but had a less sparse recovery.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{peak_1.eps}
-\caption{Comparison of peak recovery.
-Haar and Daubechies wavelets recover broader peaks, a less desirable result.
-B-Spline wavelets appear to provide the best recovery by recovering a narrower, better located peak when compared to ortho-spline wavelets.}
-\end{figure}
-
-Through a comparison of peak recovery, shown in Fig.
-5, it was deemed that B-spline wavelets provided the most desirable properties in that the solution was sparse whilst recovering a narrow and well located peak.
-\section{Results}
-
-The data consisted of seismograms split into 36 $p$ bins with the number of seismograms per bin varying between 1-8.
-For each $p$ bin the L1 optimization problem defined in Eq.
-6 was solved using 100 iterations of SPGL1.
-The initial guess provided to SPGL1 was the wavelet transformed solution obtained from the L2 simulataneous deconvolution.
-Each solution from SPGL1 provided the wavelet coefficients of a receiver function for the corresponding $p$ bin from which the receiver function in the time domain was obtained.
-3 different wavelet types were tested on the entire data set, these consisted of shift invariant Daubechies wavelets along with orthonormal and B spline wavelets.
-Once all 36 receiver functions were collected for a single wavelet type, the traces were collected by bin and a grid search was conducted to determine $V_{P}$ ans $V_{S}$ values and a line search was conducted to determine an estimate for the depth to the moho.
-
-Fig.
-6 shows the receiver function section obtained through the L2 based deconvolution method and acts as a benchmark for our comparisons.
-The results obtained through L2 deconvolution provide relatively sharp spikes in the recovered receiver functions.
-The error contours on $V_{P}$ v $V_{P}/V_{S}$ are also quite tightly constrained with a small standard error.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_L2decon.eps}
-\caption{Receiver function section for L2 deconvolution and error contours for $V_{P}$ v $V_{P}/V_{S}$.
-Receiver functions traces are plotted for all 36 bins with the result showm here.
-Also shown is a plot of $V_{P}$ v $R$ where $R=V_{P}$ v $V_{P}/V_{S}$.
-Line search results for depth are also plotted shown with a depth estimate obtained at H = 36.93km.}
-\end{figure}
-
-The first wavelet based L1 trial that we ran used shift invariant Daubechies wavelets.
-The receiver function section shows results consistent with our earlier synthetic trials showing a continuation of two properties, the first being that the recovered receiver functions are noticeabley sparser than the L2 deconvolution.
-However the peak recovery is not as sharp, just as in the synthetic trials, and this leads to poorly constrained $V_{P}$ as shown by the larger error contours in Fig.
-7.
-The recovered depth to the moho, $H(km)$, was significantly different from that obtained via the L2 method.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_daub.eps}
-\caption{Receiver function section for L1 deconvolution using Daubechies wavelets.
-Receiver function section is noticeabley sparser than the L2 section, however the recovered peaks are also broader which leads to poorly constrained $V_{P}$ v $V_{P}/V_{S}$ in comparison.
-H = 34.30km}
-\end{figure}
-
-Ortho splines produced results similar to B-splines but with broader error contours for $V_{P}$ v $V_{P}/V_{S}$ and also provided a depth estimate less than that from L2 deconvolution.
-Both ortho and B splines (Fig.
-8) improved upon Daubechies wavelets by providing sharper peak recoveries in the receiver functions and better constraints on the $V_{P}$ v $V_{P}/V_{S}$ error contours.
-Both also provided depth estimates closer to the L2 depth estimate.
-The B-spline depth estimate was the closest to the L2 estimate at H = 38.82km.
-
-\begin{figure}
-\noindent\includegraphics[width=20pc]{Rec_section_bspline.eps}
-\caption{Receiver function section for L1 deconvolution using B-spline wavelets.
-Receiver function section provides narrower peak recovery and improves the error contour constraints.
-The depth estimate is also closer to that from L2 at H = 38.82km }
-\end{figure}
-
-At present time the current wavelet framework has failed to produce results which are as tightly constrained as the L2 deconvolution method, however early results show promise and can likely be improved with some more sophisticated input.
-
-\section{Conclusion}
-
-We aimed to obtain a receiver function by deconvolving multiple source functions from multiple recorded seismograms which were grouped based on their slowness $p$.
-We developed a deconvolution framework based on L1 optimization and compared results to previous work using an L2 based method.
-We performed deconvolution by performing an inverse problem with an L1 regularization, promoting sparsity of the receiver function in the wavelet domain.
-SPGL1 was used to solve the convex optimization problem and was conducted for a range of different wavelet types including Daubechies, ortho-spline and B-spline wavelets.
-Receiver function solutions were plotted by bin and used to obtain $V_{P}$ and $V_{s}$ along with a depth estimate for the moho.
-In general Daubechies wavelets provided a very sparse representation of the receiver functions but failed to recover a sharp peak which lead to large error contours in $V_{P}$ v $V_{P}/V_{S}$.
-Spline wavelets improved by producing narrower recovered peaks and improved error estimates however at present, all L1 wavelet methods failed to produce $V_{P}$ v $V_{P}/V_{S}$ constraints which were as tightly bound as the results derived using the L2 method.
-
-\section{Future Work}
-Even with a fairly preliminary L1 minimization setup the results are encouraging.
-Further work could focus on the 2D receiver function plots and usage of curvelets to interpolate between poor data.
-This would significantly improve the estimates made by the grid search.
-While redundant wavelet dictionaries were employed in the non-spline cases these authors are currently unsure if the B-spline methods were shift-invariant or not.
-This will require further investigation.
-Also during a recent presentation of the work contained in this report Felix Herrmann suggested looking at Radon Transformations.
-Both authors are unsure of the utility of these transformations and this will also require further research.
-However, with a sophisticated tool-chain in place we believe that L1 minimization and the utilization of sparsity inherent in this problem will result in better deconvolution than the current L2 method being employed.
-Since there is a large body of global seismological work which has been built on results from L2 deconvolution there is a lot of potential to apply these methods to other problems.
-Along the same lines, once sparsity is being utilized, the larger compressive sensing framework should surely find a place in the global seismic research community.
-There is vast potential in this arena and likely many exciting avenues for phd research.
-Both the authors of this report are interested in exploring this new paradigm further in the context of global seismology.
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% \appendix resets counters and redefines section heads
-% but doesn't print anything.
-% After typing  \appendix
-%
-% \section{Here Is Appendix Title}
-% will show
-% Appendix A: Here Is Appendix Title
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary or Notation section, goes here
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-% \section*{Glossary}
-% \paragraph{Term}
-% Term Definition here
-%
-%%%%%%%%%%%%%%
-% Notation -- End each entry with a period.
-% \begin{notation}
-% Term & definition.\\
-% Second term & second definition.\\
-% \end{notation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-
-%% ------------------------------------------------------------------------ %%
-%%  REFERENCE LIST AND TEXT CITATIONS
-%
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% Or,
-%
-% If you use BiBTeX for your references, please produce your .bbl
-% file and copy the contents into your paper here.
-%
-% Follow these steps:
-% 1. Run LaTeX on your LaTeX file.
-%
-% 2. Run BiBTeX on your LaTeX file.
-%
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-%
-% 4. Comment out the old \bibliographystyle and \bibliography commands.
-%
-% 5. Run LaTeX on your new file before submitting.
-%
-% AGU DOES NOT WANT a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
+% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
 
@@
 
-\bibitem[{\textit{Herrmann}(2012)}]{herrmann}
-Herrmann, F. J., (2012), Class material from EOS 513: Imaging and Estimation with Wavelets, {\it University of British Columbia}
+\end{thebibliography}
 
-\bibitem[{\textit{Bostock}(2010)}]{bostock}
-Bostock, M. G., Kumar, M. R., (2010), Bias in seismic estimates of crustal properties, {\it Geophysical Journal International,} \textit{182}, 403--407.
-
-\bibitem[{\textit{Thompson et al}(2010)}]{thompson}
-Thompson, D.
-A., Bastow, I.D., Helffrich, G., Kendall, J-M., Wookey, J., Snyder, D.B., Eaton, D.W., (2010), Precambrian crustal evolution: Seismic constraints from the Canadian Shield, {\it Earth and Planetary Science Letters,} \textit{297}, 655--666.
-
-\bibitem[{\textit{Kanamori et al.}(2000)}]{kanamori}
-Kanamori, H. (2000), Moho depth variation in southern California from teleseismic receiver functions, {\it Geophysical Research,} \textit{105}, 2969--2980.
-
-\end{thebibliography}%Reference citation examples:
+%Reference citation examples:
 
@@
 %...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
+%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
 
@@
 %...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
+%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.
 %
@@
 %% ------------------------------------------------------------------------ %%
-%
-%  END ARTICLE
-%
-%% ------------------------------------------------------------------------ %%
-
-\end{article}
-
-%% Enter Figures and Tables here:
-
-% When submitting articles through the GEMS system:
-% COMMENT OUT ANY COMMANDS THAT INCLUDE GRAPHICS.
-%
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables; all other caption information
-%  should be placed in footnotes below the table.
-
-% DRAFT figure/table, including eps graphics
-%
-%
-% \begin{table}
-% \caption{}
-% \end{table}
-%
-% ---------------
-% TWO-COLUMN figure/table
-%
-% \begin{figure*}
-% \noindent\includegraphics[width=39pc]{samplefigure.eps}
-% \caption{Caption text here}
-% \end{figure*}
-%
-% \begin{table*}
-% \caption{Caption text here}
-% \end{table*}
-%
-% ---------------
-% EXAMPLE TABLE
-%
-%\begin{table}
-%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
-%\centering
-%\begin{tabular}{l c}
-%\hline
-% Run  & Time (min)  \\
-%\hline
-%  $l1$  & 260   \\
-%  $l2$  & 300   \\
-%  $l3$  & 340   \\
-%  $h1$  & 270   \\
-%  $h2$  & 250   \\
-%  $h3$  & 380   \\
-%  $r1$  & 370   \\
-%  $r2$  & 390   \\
-%\hline
-%\end{tabular}
-%\tablenotetext{a}{Footnote text here.}
-%\end{table}
-
-% See below for how to make landscape/sideways figures or tables.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ------------------------------------------------------------------------ %%
 
@@
 
-%% ------------------------------------------------------------------------ %%
 %

### 2013-01-10 matlab/sourceStack/lagshift.m

@@ -1,0 +1,13 @@
+function shifted = lagshift(seis, lags, dt)
+
+% LAGSHIFT shift all traces by lag times given in array lags.
+% It is assumed that length(lags) == size(seis, 1)
+
+lag = round(lags/dt);
+shifted = zeros(size(seis));
+
+for ii = 1 : size(seis, 1)
+    shifted(ii, :) = circshift( seis(ii, :), [0, lag(ii)] );
+end
+
+end

### 2013-01-10 matlab/sourceStack/mccc.m

@@ -1,0 +1,84 @@
+function [tdel,rmean,sigr] = mccc(seis,dt,twin)
+
+% FUNCTION [TDEL] = MCCC(SEIS,DT,TWIN);
+% Function MCCC determines optimum relative delay times for a
+% set of seismograms based on the VanDecar & Crosson multi-channel
+% cross-correlation algorithm. SEIS is the set of seismograms. It
+% is assumed that this set includes the window of interest and
+% nothing more since we calculate the correlation functions in the
+% Fourier domain. DT is the sample interval and TWIN is the window
+% about zero in which the maximum search is performed (if TWIN is
+% not specified, the search is performed over the entire correlation
+% interval).
+
+% Set nt to twice length of seismogram section to avoid
+% spectral contamination/overlap. Note we assume that
+% columns enumerate time samples, and rows enumerate stations.
+nt = size(seis, 2) * 2;
+ns = size(seis, 1);
+tcc = zeros(ns);
+
+% Set width of window around 0 time to search for maximum.
+mask=ones(1,nt);
+if nargin == 3
+  itw = fix(twin/(2*dt));
+  mask = zeros(1,nt);
+  mask(1:itw) = 1.0;
+  mask(nt - itw-1:nt) = 1.0;
+end
+
+% Determine relative delay times between all pairs of
+% traces.
+for is = 1 : ns
+  ffis = fft(seis(is,:), nt);
+  acf = real(ifft([ffis .* conj(ffis)], nt));
+  sigt(is) = sqrt( max(acf) );
+end
+r = zeros(ns,ns);
+for is = 1 : ns-1
+  ffis = conj( fft(seis(is, :), nt) );
+  sigi = sqrt( norm(seis(is, :), 2) );
+  for js = is + 1 : ns
+    ffjs = fft( seis(js ,:), nt );
+    ccf = real( ifft([ffis .* ffjs], nt) ) .* mask;
+    [cmax, tcc(is, js)] = max(ccf);
+
+% Compute estimate of cross correlation coefficient.
+    r(is,js) = cmax / ( sigt(is) * sigt(js) );
+  end
+end
+
+% Fisher's transform of cross-correlation coefficients to produce
+% normally distributed quantity on which Gaussian statistics
[…]

### 2013-01-10 matlab/sourceStack/sourceStacker.m

@@ -1,0 +1,79 @@
+%% Source Stacker
+% Testing P-wave impulse alignment, stacking.
+clear all
+close all
+
+addpath ../../sac
+addpath ../functions
+rootdir = '/media/TerraS/CN/';
+% data attained using python program eventDistribution.py
+event = '0611122127';
+stns = {'MLON','ILKN','BOXN','COWN','KNDN','GBLN','YKW3',...
+        'JERN','SNPN','YKW2','GLWN','MCKN'};
+
+
+%% Extract SAC data
+for ii = 1 : length(stns)
+    files{ii} = fullfile(rootdir, stns{ii}, event, 'stack_P.sac');  %#ok<*SAGROW>
+end
+
+[trace, header] = getTrace(files);
+
+%% Strip out stations with unique delta
+dt = header{1}.DELTA;
+for ii = 1: length(header)
+    if dt ~= header{ii}.DELTA
+        fprintf('Error, station %s has a unique dt\n', header{ii}.KSTNM)
+        dtDiff(ii) = 1;
+    end
+end
+
+dtDiff = dtDiff == 1;
+trace(dtDiff, :) = [];
+header(dtDiff) = [];
+%% Normalize
+trace = (diag(1./max( abs(trace), [], 2)) ) * trace;
+
+%% Collect array of Pick times
+for ii = 1 : length(header)
+    picktimes(ii) = header{ii}.T1 - header{ii}.B;
+    endtimes(ii) = header{ii}.T3 - header{ii}.B; 
+end
+
+%% Loop through different windows
+endwin = 60: -1 : -20;
+
+for ii = 1:length(endwin)
+%% Set a window
+w1 = round((min(picktimes) + 0) / dt);
+w2 = round((max(endtimes) - endwin(ii) ) / dt);
+
+%% normalize traces
+wtrace = (diag(1./max( abs(trace(:, w1 : w2)), [], 2)) ) * trace(:, w1 : w2 );
+
+%% Get lags
+[tdel, rmean, sigr] = mccc(wtrace, dt);
+%% Shift unwindowed traces
+strace = lagshift(trace, tdel, dt);
+
+%% PLOT
+%figure(2)
+%plot(lagtrace(:, 200:600)')
+Y(ii) = norm(var(strace, 0, 1));
+
+numt = size(strace,1);
+figure(3)
+subplot(3,1,1)
+    plot(trace(1,:))
[…]

### 2013-01-10 python/eventDistribution.py

@@ -1,0 +1,70 @@
+#!/usr/bin/python2
+#
+# python-mode indent C-c < or C-c >
+# python-mode comment/uncomment region M-;
+
+###########################################################################
+# IMPORTS
+###########################################################################
+import sys, os, re
+from collections import defaultdict
+from dbutils import is_number
+import matplotlib.pyplot as plt
+import numpy as np
+
+stationdir = "/media/TerraS/CN/"
+
+class CrossEvent( object ):
+    def __init__(self, station):
+        self.stationSet = set()
+        self.stationSet.add(station)
+
+    def addStation(self, station):
+        self.stationSet.add(station)
+
+    def numStations(self):
+        return len(self.stationSet)
+
+
+if __name__== '__main__' :
+
+    if not sys.stdin.isatty():
+        stations =  re.findall(r'\w+', sys.stdin.read() )
+    else:
+        print "You need to pipe in stations yo"
+        exit()
+
+    d = {}
+    q = defaultdict(int)
+
+    for station in stations:
+        stdir = os.path.join(stationdir, station)
+        events = os.listdir(stdir)
+        events = filter(is_number, events)
+        for event in events:
+            if event in d:
+                d[event].addStation(station)
+            else:
+                d[event] = CrossEvent(station)
+
+
+    # singleEventStations = [list(d[event].stationSet)[0] for event in d.keys() if d[event].numStations() == 1]
+
+    # for ses in singleEventStations:
+    #     q[ses] += 1
+
+    data = np.array([ float( d[event].numStations() ) for event in d.keys() ]) #/ float(len(stations))
+
+    evs = [ ev for ev in d.keys() if d[ev].numStations() == 12]
+
+    ev = sorted(evs)[-1]
+
+    print "The following event is found in listed stations"
+    sys.stdout.write(ev + ": ")
+    sys.stdout.write("{")
+    for stn in list(d[ev].stationSet):
+        sys.stdout.write( "'" + stn + "'" + ",")
+    sys.stdout.write("\b")
+    sys.stdout.write("}\n")
+    # plt.plot( sorted(data, reverse = True) )
+    # plt.show()

### 2013-01-28 paper/draftBP.tex
distance matrix bounds stn filter function

@@
 \begin{abstract}
-(Type abstract here)
+It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
+Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
+Durrheim, W.
+Mooney, 1991).
+Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
+We apply a new stacking method (Bostock and Kumar, 2010) to retrieve estimates of bulk crustal velocities (Vp, Vs) and thickness H from these data under the assumption of 1-D structure.
+Bootstrap error analysis is performed for each station dataset and results are compared to the results produced from alternate methods (Zhu and Kanamori, 2000).
+Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
+These results will be used to evaluate competing models of early crustal formation.
 \end{abstract}

### 2013-01-28 matlab/sourceStack/distantStations.m

@@ -1,0 +1,31 @@
+% StationDistance
+function stations = distantStations(station, mindist, maxdist)
+% DISTANTSTATIONS returns all stns within mindist and maxdist from station
+
+load stndist
+n = length(stns);
+radius = 6371;
+e = ones(n, 1);
+
+%% Get input stations lat and lon
+ix = strcmp(station, stns);
+stnlat = lats(ix);
+stnlon = lons(ix);
+
+if isempty(stnlat)
+    error('Station %s not found in "stndist.mat"', station)
+end
+
+if length(stnlat) > 1
+    error('Multiple entries found for station %s in "stndist.mat"', station)
+end
+
+%% Compute distance matrix
+arclen = distance([stnlat * e, stnlon * e], [lats', lons'], radius);
+%% Sort ascending
+[arclen, I] = sort(arclen);
+stns = stns(I);
+%% pinch
+stations = stns( arclen >= mindist & arclen <= maxdist );
+
+end

### 2013-01-31 paper/draftBP.tex
paper: I started

@@
 
-\title{Sesimic parameter estimation and the Canadian crust}
+\title{Seismic parameter estimation and the Canadian crust}
 \author{Ben Postlethwaite}
@@
 \section{Data and method}
-(Data explanation here)
+   Seismic data utilized in this study comes from 343 stations across Canada from all available networks.
+Of these stations 146 were used for analysis.
+All stations are broadband with data being used between the years 2000 and 2012.
+Receiver function analysis has been widely employed over many years to investigate the earths structure.
+   Two derivations have been used in this study, one well tested stacking approach (Zhu and Kanamori, 2000) and the other a recent method that has not been employed at scale (Bostock, 2010).
+Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts.
+Deconvolution of the P-wave energy arrival, used as an estimate for the source function, from the S-wave signal produces a Green's function with energy peaks at the times of the main and reflected arrivals.
+For better separation of P and S wave energy all components are rotated into radial and transverse dimensions and transformed into P, SV and SH components with a wave field decomposition transfer matrix (Bostock, 1998).
+   Deconvolution is performed by minimizing the general cross validation function $GCV(\delta)$ where $\delta$ is used the regularizer (Bostock, 1998; Golub et al., 1979).
+This method is an $L_2$ frequency domain deconvolution which performs quickly and makes no assumptions on the noise in the data.
+All deconvolved signals are filtered between 0.04hz and 3.0hz and are stacked with weights $w_1=0.5$, $w_1=0.3$, $w_1=0.2$ for the Ps, PpPs and PpSs phases respectively.
+
+   <Error Calculations>
+
 \subsection{Vp methods}
-(Vp method here)
-\subsection{Other methods}
-(Other methods here)
+   The stacking approach outlined above provides estimates for the parameters $R=\frac{V_P}{V_S}$ as well as crustal thickness $H$ - which gives the depth of the Moho.
+A method which estimates $V_P$ and $V_S$ separately (Bostock, 2010) is also employed for select stations.
+This method makes use of the fact the dependence on $H$ in the travel time equations can be removed if we divide the reflected phases by the main arrival.
+$$ t_{Pps}(p_i) = \frac{ \sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
+$$ t_{Pss}(p_i) = \frac{2\sqrt{R^2-p_i^2V_P^2}} {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
+This has the advantage that no assumptions on $V_P$ are necessary to perform the gridsearch and stack.
+A similar stacking method as the Zhu and Kanamori approach may be employed for $t_{Pps}$ and $t_{Pss}$ as long as an estimate for $t_{Ps}$ exists.
+With estimates for $R$ and $V_P$ a simple line search along $H$ can be made using all three undivided travel time equations.
+
+Several methods for choosing $t_{Ps}$ are available.
+Direct arrival $t_{Ps}$ should contain the largest fraction of energy out of the other phases and the moveout, depending on Moho depth, should be in the order of 3 to 6 seconds.
+Given this it is trivial to define a window and select the time corripsonding to maximum amplitude.
+Another approach is to use max amplitude estimates as data in a non-linear optimization to find the $R$, $V_P$, and $H$ which minimizes the residual between $t_{Ps}$ and the data.
+The travel time equations are twice-differentiable so the quadratically convergent Newton's method may be employed.
+This approach has the advantage that noise leading to poor maximum amplitude picks are effectively collapsed onto the curve corrisponding to the travel time function.
+A third approach which offers picks along a travel time curve but better stability than the non-linear method is to perform a gridsearch with all three travel time functions, a full Kanamori stack.
+With the best estimates for $R$ and $H$ the $t_{Ps}$ function can be found and used as input into the Bostock method.
+The trade-off is that the requirement for an initial $V_P$ estimate for the initial stack introduces secondary $V_P$ dependence into the system.
 
@@
 \section{Results}
-{Results go here}
+<Results go here>
 
@@
 \subsection{Canada}
-(Discussion on larger - Canada wide - regions)
+<Discussion on larger - Canada wide - regions>
 \subsection{Slave Province}
-(Discussion on Slave Province)
+<Discussion on Slave Province>
 
@@
 \section{Conclusions}
-(Conclusions here)
+<Conclusions here>
 

### 2013-01-31 matlab/functions/newtonFit.m

@@ -8,11 +8,11 @@ q = figure(fh);
 p2 = p.^2;
 %x = 1:length(p);
 while (iter < round(itermax)) && (deltaTps > tol);
-    
+
     % Partials %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     sqrtA = sqrt(1/a^2 - p2);
     sqrtB = sqrt(1/b^2 - p2);
-    
+
     % Jacobian %%%%%%
     dfda = (-2*h* (h*(sqrtA - sqrtB) + t)' * (1./( a^3 * sqrtA )));
     dfdb =  (2*h* (h*(sqrtA - sqrtB) + t)' * (1./( b^3 * sqrtB )));
@@ -22,12 +22,12 @@ while (iter < round(itermax)) && (deltaTps > tol);
     dfah = -2* ( 2*h* (sqrtA - sqrtB) + t)' * (1./(a^3 * sqrtA)) ;
     dfbh =  2* ( 2*h* (sqrtA - sqrtB) + t)' * (1./(b^3 * sqrtB)) ;
     dfaa = 2*h *( h*(-3*a^2*p2.*sqrtB + 3*a^2*p2.*sqrtA - 3*sqrtA + 2*sqrtB) + ...
-        t .* (3*a^2 *p2 - 2) )' * (1./(a^4 * sqrtA .* (a^2* p2 - 1))); 
+        t .* (3*a^2 *p2 - 2) )' * (1./(a^4 * sqrtA .* (a^2* p2 - 1)));
     dfbb = 2*h *( h*(-2*b^2 *(sqrtA.*sqrtB + 3*p2) + ...
         3*b^4*(p2 .* sqrtA.*sqrtB + p.^4) + 3) + ...
-        b^2.*t.*sqrtB.*(3*b^2 .*p2 - 2))' * (1./(b^4 * (b^2*p2 - 1).^2));   
+        b^2.*t.*sqrtB.*(3*b^2 .*p2 - 2))' * (1./(b^4 * (b^2*p2 - 1).^2));
     dfab = sum( -2*h^2 ./ ( a^2 * b^2 * sqrt(1-a^2*p2).*sqrt(1-b^2*p2)));
-      
+
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     J = [dfda; dfdb; dfdh]; % Jacobian
     H = [dfaa, dfab, dfah
[…]

### 2013-02-01 paper/draftBP.tex
test wavelet thresholding on RFs write a little more

@@
 Receiver function analysis has been widely employed over many years to investigate the earths structure.
-   Two derivations have been used in this study, one well tested stacking approach (Zhu and Kanamori, 2000) and the other a recent method that has not been employed at scale (Bostock, 2010).
+   Two derivations have been used in this study, one well tested stacking approach (Zhu and Kanamori, 2000) and the other a recent method that has not yet been employed at scale (Bostock, 2010).
 Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts.
@@
 All deconvolved signals are filtered between 0.04hz and 3.0hz and are stacked with weights $w_1=0.5$, $w_1=0.3$, $w_1=0.2$ for the Ps, PpPs and PpSs phases respectively.
-
-   <Error Calculations>
+Error calculations are performed with bootstrap resampling by rerunning the processing with randomly chosen receiver functions - allowing multiples - 1024 times.
+The error is obtained by taking the standard deviation of the results.
 
@@
 \section{Results}
-<Results go here>
+Canada data
+
+Data across geological provinces
+
+Archean versus Proterozoic
 

### 2013-02-01 matlab/ProcessStack.m

@@ -185,7 +185,7 @@ end
 % Pslow indices.
 clear wft vft
 
-rec = recStack;
+%rec = recStack;
 
 %% Weed out poor rf results
 % Cut out traces where no betax was found during simdecf

### 2013-02-01 matlab/dataDriver.m

@@ -18,7 +18,7 @@ sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'ACKN';
+station = 'ARTN'; %
 
 %{
 Need to find which stations go with which events.
@@ -80,7 +80,8 @@ end
 %     dlist{end+1} = fullfile(workingdir, events{ii});
 % end
 % db = process(db, dlist, station, workingdir, method, vp, 0);
-ProcessStack
+%ProcessStack
+ProcessTraces
 %% Assign Data
 
 [ db ] = assigndb( db, method, station, brec(:,1:round(45/dt)), ...

### 2013-02-01 matlab/functions/simdecf.m

@@ -28,7 +28,7 @@ end
 if betan > 0
   betax = betan;
 else
-  beta = exp(-16 : 0.2 : 15);
+  beta = exp(-18 : 0.2 : 15);
   gcvf = zeros(1, length(beta));
   for ib = 1:length(beta);
 

### 2013-02-01 python/emailresponse.py

@@ -12,31 +12,64 @@ def extract_body(payload):
     else:
         return '\n'.join([extract_body(part.get_payload()) for part in payload])
 
+
+def dataReady(subj, lines):
+""" Finds myfile.seed and writes it out
+to the terminal on stdout """
+    for line in lines:
+        fields = line.split()
+        for field in fields:
+            if 'seed' in field:
+                sys.stdout.write(field + "\n")
+                sys.stdout.flush()
+                return
+
+def dataError(subj, lines):
+    """ Sends error message with subject event name
+    down standard error """
+    for line in lines:
+        if 'ERROR_LOG' in line:
+            sys.stderr.write('error with ' + subj[6:] + "\n")
+            sys.stderr.flush()
+
+
+
+
+EMAIL = 'auto.iris.response@gmail.com'
+PASSWD = 'auto iris key'
+MAILBOX = 'inbox'
+
 conn = imaplib.IMAP4_SSL("imap.gmail.com", 993)
 conn.login(EMAIL, PASSWD)
-conn.select('seisftp')
+conn.select(MAILBOX)
 
 while True:
-    typ, data = conn.search(None, 'FROM', '"autoDRM"', 'UNDELETED')
+    # Get all emails in MAILBOX FROM autoDRM
+    typ, data = conn.search(None, 'FROM', '"autoDRM"', 'unseen')
     # ITERATE THRU ALL MATCHED ITEMS
     for num in data[0].split():
         typ, msg_data = conn.fetch(num, '(RFC822)')
         # GET MSG DATA
         for response_part in msg_data:
             if isinstance(response_part, tuple):
+                # Message extraction... see python module email
                 msg = email.message_from_string(response_part[1])
-                payload=msg.get_payload()
-                body=extract_body(payload)
+                subject = msg['subject']
+                payload = msg.get_payload()
+                body = extract_body(payload)
                 lines = body.split('\n')
                 for line in lines:
-                    if "TITLE EVENT" in line:
-                        fields = line.split()
-                        sys.stdout.write(fields[3] + "\n")
-                        sys.stdout.flush()
-        # DELETE MSG
-        typ, response = conn.store(num, '+FLAGS', r'\Deleted')
-    conn.expunge()
-    sleep(20)
+                    if "FTP_FILE" in line:
+                        dataReady(subject, lines)
+                        break
+                    if 'ERROR_LOG' in line:
+                        dataError(subject, lines)
[…]

### 2013-02-01 shellscripts/evmail.sh

@@ -51,7 +51,7 @@ while read event; do
 
   echo  "Processing event" $event
   echo "BEGIN" > $event"_request"
-  echo "EMAIL post.ben.here@gmail.com" >> $event"_request"
+  echo "EMAIL auto.iris.response@gmail.com" >> $event"_request"
   echo "OUT_FILE " $event".seed" >> $event"_request"
   echo "TITLE EVENT "$event  >> $event"_request"
 # THESE ARE CNSN BH STATIONS.

### 2013-02-02 paper/draftBP.tex
tiny addition to draft

@@
 \section{Results}
-Canada data
+Ideas:
+Reliability of methods:
+  Compare bulk properties garnered from MB to Kan Approach.
+  Compare Kan approach values to others in literature
+  Canada wide -> Regional
+
+The bulk crustal composition of Canada shows a $frac{V_P}{V_S}$ ratio of $Z$.
+When compared to the granite - tonalitic - mafic gneiss spectrum this value in very close to the tonalitic gneiss value of $Y$.
+Compared to other bulk crustal averages this value appears to be X.
+[Find estimates of bulk averages].
 

### 2013-02-05 paper/draftBP.tex
paper

--- 2013-02-01 paper/draftBP.tex
@@
 \section{Results}
-Canada data
+Ideas:
+Reliability of methods:
+  Compare bulk properties garnered from MB to Kan Approach.
+  Compare Kan approach values to others in literature
+  Canada wide -> Regional
 
-Data across geological provinces
+  The preliminary interrogation of the data set yields the observation that the bulk Canadian crust is more felsic than originally anticipated.
+Previous experiments show crustal averages of 0.265 (Christensen and Mooney, 1995) or have averages of at least 0.27 for shields, platforms and paleozoic units (Zandt and Ammon, 1995).
+The weighted average for this data set shows a Poisson' ratio of 0.258 with the weighting for each station given by the calculation of the stations inverse Voronoi cell area.
+At the regional scale we have values of 0.251.
+0.254, 0.260, 0.275 for the Churchill, Superior, Slave and Grenville Provinces Poisson's ratios respectively.
+Most of the Canadian Shield stations also appear to have low Poisson values with the exception of the Grenville Province which has a high Poisson Ratio of 0.275.
 
-Archean versus Proterozoic
+  The data bears out earlier results showing Proterozoic crust to have a higher Poisson's ratio than Archean crust, likely the result of a more mafic composition.
+The increased crustal thickness of Proterozoic crust is clearly seen in the active source data while it not visible in data from processed seismic stations.
+The most likely reason for not seeing the signal of thicker crust in the station data is owing to poor station coverage in Proterozoic regions.
+
+  Further work on the Bostock-Kumar stacking approach is warranted from a look at the cleanest stations.
+Before it can be employed in large scale analysis additional denoising methods or alternative deconvolution techniques will need to be investigated to reduce the noise in the data.
 
@@
 %% ------------------------------------------------------------------------ %%
+
+M. G. Bostock, M. R. Kumar, 2010. Bias in seismic estimates of crustal properties, Geophys. J. Int., 182, 403-407.
+N. I. Christensen, 1996. Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156,
+R. J. Durrheim, W. D. Mooney, 1991. Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
+W. D. Mooney. 2012. Personal communication. Compiled GSC active source data for the Canada.
+D.A.
+Thompson, I.D.
+Bastow, G.
+Helffrich, J-M.
+Kendall, J.
+Wookey, D.B.
+Snyder, D.W.
+Eaton, 2010.
+Precambrian crustal evolution: Seismic constraints from the Canadian Shield, Earth and Planetary Science Letters, Volume 297, Issues 3–4, 1 September 2010, Pages 655–666
+L. Zhu, H. Kanamori, 2000. Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys Res, 105, 2969-2980.
+G. Zandt, C. J. Ammon, 1995. Continental crust composition constrained by measurements of crustal Poisson's ratio, Nature, 374, 152-154.
 

### 2013-02-15 paper/draftBP.tex
paper

@@
 %% ------------------------------------------------------------------------ %%
-\documentclass[draft]{article}
+\documentclass[draft, 12pt]{article}
+\usepackage[english]{babel}
+\usepackage[final]{graphicx}
+\graphicspath{ {figures/} }
+\DeclareGraphicsExtensions{.png}
 
@@
 \section{Data and method}
-   Seismic data utilized in this study comes from 343 stations across Canada from all available networks.
-Of these stations 146 were used for analysis.
-All stations are broadband with data being used between the years 2000 and 2012.
-Receiver function analysis has been widely employed over many years to investigate the earths structure.
-   Two derivations have been used in this study, one well tested stacking approach (Zhu and Kanamori, 2000) and the other a recent method that has not yet been employed at scale (Bostock, 2010).
+   Seismic data utilized in this study come from 343 stations across Canada from all available networks.
+Of these stations 146 are of sufficient quality to be useful for study.
+All stations are broadband with data selected between the years 2000 and 2012, for a total of more than 700 earthquake sources.
+All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing.
+Data is selected for reasonable signal to noise ratioand sufficient impulsiveness that arrival times can be calculated.
+All data is analyzed using receiver function methods, an approach which has been widely employed over many years to investigate the earths structure.
+
+   Two derivations of receiver function analysis have been used in this study, one well tested stacking approach and the other a recent method that has not yet been employed at scale.
 Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts.
@@
 
-\subsection{Vp methods}
+\subsection{Vp/Vs method}
+   A well tested approach for calculating $R=\frac{V_P}{V_S}$ and $H$ was employed by Zhu and Kanamori (2000).
+After deconvolution the signals are stacked and a gridsearch is performed using the travel time equations for different values of $R$ and $H$ and a spectrum of slowness.
+For each parameter combination the travel time functions output estimates for the differential arrival time of the S-component main and reflected phases.
+These times are used to pick the energy in each receiver function - with the large amplitudes at arrivals stacking to provide an estimate for the optimal parameter estimation.
+
+\subsection{Vp method}
    The stacking approach outlined above provides estimates for the parameters $R=\frac{V_P}{V_S}$ as well as crustal thickness $H$ - which gives the depth of the Moho.
@@
 
+\subsection{Vp database}
+   Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources.
+The data was compiled by Mooney (personal communication, 2012).
+
 %% -----------------------------------------------%%
@@
   Canada wide -> Regional
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{thompsonComparison}
+  \caption{A picture of a gull.}
+\end{figure}
 

### 2013-02-15 python/thompson_synder2010paper.py

@@ -72,23 +72,45 @@ stdRsnyd = stdRsnyd[~ind]
 stns = stns[~ind]
 
 corr = pearsonr(R, Rsnyd)
-print corr
+print "Correlation between datasets is {0:.2f}".format(corr[0])
 
 t = np.arange(len(R))
-plt.plot(t, R, '-ob', lw = 4, ms = 12, label = "Vp/Vs estimate -  current data set")
-plt.plot(t, Rsnyd, '-og', lw = 4, ms = 12, label = "Vp/Vs estimate Thompson et al.")
-plt.plot(t, Rjapan, '*',  ms = 13, color = 'orange',  label = "Vp/Vs estimate Japan Source Region")
-plt.plot(t, Rchili, '*',  ms = 13, color = 'yellow', label = "Vp/Vs estimate Chili Source Region")
+
+#######################
+# Plotting formatter
+#######################
+ratio = 1.5
+lw = 4 / ratio# line width
+ms = 12 / ratio# marker size
+caplen = 7 / ratio
+capwid = 2 / ratio
+elw = 2 / ratio
+ticks = 16 / ratio
+label = 16 / ratio
+title = 18 / ratio
+leg = 16 / ratio
+
+ax = plt.subplot(111)
+plt.plot(t, R, '-ob', lw = lw, ms = ms, label = "Vp/Vs estimate -  current data set")
+plt.plot(t, Rsnyd, '-og', lw = lw, ms = ms, label = "Vp/Vs estimate Thompson et al.")
+plt.plot(t, Rjapan, '*',  ms = ms, color = 'orange',  label = "Vp/Vs estimate Japan Source Region")
+plt.plot(t, Rchili, '*',  ms = ms, color = 'yellow', label = "Vp/Vs estimate Chili Source Region")
 plt.errorbar(t, R, yerr=stdR, xerr=None, fmt=None, ecolor = 'blue',
-             elinewidth = 2, capsize = 7, mew = 2, label = "2 std dev Bootstrap")
+             elinewidth = elw, capsize = caplen, mew = capwid, label = "2 std error")
 plt.errorbar(t, Rsnyd, yerr=stdRsnyd, xerr=None, fmt=None, ecolor = 'green',
-             elinewidth = 2, capsize = 7, mew = 2, label = "1 std error Eaton et. al. 2006")
-plt.legend()
-plt.title("Comparison for given Canadian Shield Stations - " +
-          "Thompson et al (2010) data.\n" +
[…]

### 2013-02-16 paper/draftBP.tex
paper

@@
 \section{Results}
-Ideas:
-Reliability of methods:
-  Compare bulk properties garnered from MB to Kan Approach.
-  Compare Kan approach values to others in literature
-  Canada wide -> Regional
+As regional studies exist which utilize similar methods and have corrisponding parameter estimates for some of the stations used in this study it is possilbe to directly compare values to those previously published.
+A comprehensive study in the Hudson Bay region of the Canadian Shield (Thomson et.
+al., 2010) uses an approach similar to the Zhu and Kanamori method provides estimates for $H$ and $R=/frac{V_P}{V_S}$ for 35 stations.
+5 stations with $R$ estimates above 1.8 are removed due to the uncertainty surrounding these high values.
+Both datasets are plotted in Figure~\ref{fig:thompsonComp}.
+The Pearson correlation coefficient between these datasets is 0.49.
+[Compare this value to uncertainty / deviation in the data?
+].
+
+Directly comparing $V_P$ estimates processed with the MB algorithm is not possible as this method has not been employed in publication.
+However, comparisons can be made to active source records for experiments within close proximity to a given seismic station (Figure~ \ref{fig:activeComp}.
+The correlation between these datasets is 0.235.
+[This is low, explain here?].
+Another method of determining the reliability of estimates is to check the $/frac{V_P}{V_S}$ ratio taken from the Kanamori approach to the $/frac{V_P}{V_S}$ value computed from the MB approach.
+These values, working with the same data, preprocessed using the same methods and tools should be equal.
+Selecting only those stations who have an error of less than +/-0.05 $R$, 132 stations, we get a correlation of 0.57.
+Reducing the error tolerance and selecting stations with less than +/-0.01 $R$, leaving 31 of the cleanest stations, we get a correlation coefficient of 0.95.
 
@@
     \includegraphics[width=\textwidth]{thompsonComparison}
-  \caption{A picture of a gull.}
+  \caption{Comparison of $\frac{V_P}{V_R}$ data processed for this study with data from Thomspson et. al. (2010).}
+  \label{fig:thompsonComp}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{activeSourceComparison}
+  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longtitude.}
+  \label{fig:activeComp}
 \end{figure}

### 2013-02-16 python/closestShots.py

@@ -48,7 +48,7 @@ if __name__  == "__main__":
 # Filter dictionary to stations of this dist or less
 # And also provide a wieghted average Vp and H
 
-    ddeg = 0.5
+    ddeg = 1.0
     mdict = {}
 
     for stn in distd:

### 2013-02-19 matlab/ProcessStack.m
presentation GEM2

@@ -10,7 +10,7 @@ fLow = 0.04; % Lower frequency cutoff
 fHigh = 3.0; % Upper frequency cutoff
 snrlim = 0.995;
 %% Get list of stations within bounds
-mindist = 250;
+mindist = 100;
 maxdist = 1200;
 stns = distantStations(station, mindist, maxdist);
 clear mindist maxdist

### 2013-02-19 matlab/sourceStack/getTrace.m

@@ -5,7 +5,6 @@ for ii = 1:length(files)
 
     % TRY I/O: Read info from sac files
     S  = readsac(files{ii});
-   
     % Truncate if longer
     if S.NPTS > N
         S.DATA1(N+1:end) = [];

### 2013-02-19 python/crust2shape.py

@@ -182,9 +182,9 @@ def mooney2shapefile(mooney):
 
     w.field("mcode", 'C', '3')
     w.field("Geotype", 'C', '45')
-    w.field("mVp", 'C', '6')
-    w.field("mVs", 'C', '6')
-    w.field("mH", 'C', '6')
+    w.field("mVp", 'N', 7, 4)
+    w.field("mVs", 'N', 7, 4)
+    w.field("mH", 'N', 7, 4)
 
     for (lon, lat), key, txt, (vp, vs, ht) in mooney:
         # Set lon & lat
@@ -198,9 +198,9 @@ def mooney2shapefile(mooney):
 
         w.record( key,
                   txt,
-                  "{0:2.3f}".format(vp),
-                  "{0:2.3f}".format(vs),
-                  "{0:2.3f}".format(ht))
+                  "{0:2.4f}".format(vp),
+                  "{0:2.4f}".format(vs),
+                  "{0:2.4f}".format(ht))
 
 
     w.save(shpfile)

### 2013-02-19 python/dbutils.py

@@ -21,7 +21,6 @@ import scipy.io as sio
 databasedir = '/media/TerraS/database'
 netdir = '/media/TerraS/CN'
 dbfile = os.environ['HOME'] + '/thesis/data/stations.json'
-shpfile = os.environ['HOME'] + '/thesis/mapping/stations'
 stationlist = os.environ['HOME'] + '/thesis/shellscripts/cnsn_stn.list'
 updtime = os.environ['HOME'] + '/thesis/data/updtime.data'
 
@@ -116,10 +115,14 @@ def buildStations(stdict, cnsnlist):
 
     open(dbfile,'w').write( json.dumps(stdict, sort_keys = True, indent = 2 ))
 
-def json2shapefile(stdict, sc):
+def json2shapefile(stdict, fout, sc):
     ''' Converts the stations.json data into a shapefile for usage with
     GIS programs such as QGIS'''
 
+    if fout == None:
+        fout = os.environ['HOME'] + '/thesis/mapping/stations'
+    else:
+        fout = os.environ['HOME'] + '/thesis/mapping/' + fout
     w = shapefile.Writer( shapeType = 1 )
     # Set fields for attribute table
 
@@ -129,14 +132,14 @@ def json2shapefile(stdict, sc):
     w.field("status", 'C', '16')
     #w.field("Vp", "C", "5")
     for field in fields:
-        w.field(field, 'C', '6')
+        w.field(field, 'N', 7, 4)
 
     for key in stdict.keys():
         # Set lon & lat
         w.point( stdict[key]["lon"], stdict[key]["lat"] )
         values = []
         for f in fields:
-            values.append( '{0:2.3f}'.format(sc.flattendict(stdict[key])[f]) if f in sc.flattendict(stdict[key]) else "None ")
+            values.append( '{0:2.4f}'.format(sc.flattendict(stdict[key])[f]) if f in sc.flattendict(stdict[key]) else "None   ")
         w.record( key,
                   stdict[key]["network"],
                   stdict[key]["status"],
@@ -147,7 +150,7 @@ def json2shapefile(stdict, sc):
                   values[4],
                   values[5])
 
-    w.save(shpfile)
+    w.save(fout)
 
 def missingComps(s):
     if "MissingComponents" in s:
@@ -409,8 +412,8 @@ if __name__== '__main__' :
     group.add_argument('-u','--update', action = 'store_true',
                         help = "updates database with statistics from data files")
 
[…]

### 2013-02-19 python/histplots.py

@@ -22,7 +22,7 @@ stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
 moonfile = os.environ['HOME'] + '/thesis/data/moonvpGeology.json'
 vfile = os.environ['HOME'] + '/thesis/data/voronoi.data'
 
-parameterType = 'velocity'
+parameterType = 'thickness'
 
 def poisson(R, reverse = False):
     ''' Function to go from Vp/Vs -> Poisson's ratio '''

### 2013-02-19 python/shapefile.py

@@ -737,6 +737,8 @@ class Writer:
                     value = str(value)[0].upper()
                 else:
                     value = str(value)[:size].ljust(size)
+                if len(value) != size:
+                    print "value", value, "is not equal to size", size
                 assert len(value) == size
                 value = b(value)
                 f.write(value)

### 2013-02-27 paper/draftBP.tex
paper / slides / stuff

@@
 \begin{abstract}
-   It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
-Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
-Durrheim, W.
-Mooney, 1991).
-Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
-We apply a new stacking method (Bostock and Kumar, 2010) to retrieve estimates of bulk crustal velocities (Vp, Vs) and thickness H from these data under the assumption of 1-D structure.
-Bootstrap error analysis is performed for each station dataset and results are compared to the results produced from alternate methods (Zhu and Kanamori, 2000).
-Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
-These results will be used to evaluate competing models of early crustal formation.
 \end{abstract}
@@
 All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing.
-Data is selected for reasonable signal to noise ratioand sufficient impulsiveness that arrival times can be calculated.
+Data is selected for reasonable signal to noise ratio and sufficient impulsiveness that arrival times can be calculated.
 All data is analyzed using receiver function methods, an approach which has been widely employed over many years to investigate the earths structure.
@@
 For better separation of P and S wave energy all components are rotated into radial and transverse dimensions and transformed into P, SV and SH components with a wave field decomposition transfer matrix (Bostock, 1998).
+
    Deconvolution is performed by minimizing the general cross validation function $GCV(\delta)$ where $\delta$ is used the regularizer (Bostock, 1998; Golub et al., 1979).
@@
 
+(MOST LIKELY UNNECESSARY INFO)
 Several methods for choosing $t_{Ps}$ are available.
 Direct arrival $t_{Ps}$ should contain the largest fraction of energy out of the other phases and the moveout, depending on Moho depth, should be in the order of 3 to 6 seconds.
-Given this it is trivial to define a window and select the time corripsonding to maximum amplitude.
+Given this it is trivial to define a window and select the time corresponding to maximum amplitude.
 Another approach is to use max amplitude estimates as data in a non-linear optimization to find the $R$, $V_P$, and $H$ which minimizes the residual between $t_{Ps}$ and the data.
 The travel time equations are twice-differentiable so the quadratically convergent Newton's method may be employed.
-This approach has the advantage that noise leading to poor maximum amplitude picks are effectively collapsed onto the curve corrisponding to the travel time function.
+This approach has the advantage that noise leading to poor maximum amplitude picks are effectively collapsed onto the curve corresponding to the travel time function.
 A third approach which offers picks along a travel time curve but better stability than the non-linear method is to perform a gridsearch with all three travel time functions, a full Kanamori stack.
@@
 \section{Results}
-As regional studies exist which utilize similar methods and have corrisponding parameter estimates for some of the stations used in this study it is possilbe to directly compare values to those previously published.
+
+\subsection{Comparisons}
+As regional studies exist which utilize similar methods and have corresponding parameter estimates for some of the stations used in this study it is possible to directly compare values to those previously published.
 A comprehensive study in the Hudson Bay region of the Canadian Shield (Thomson et.
-al., 2010) uses an approach similar to the Zhu and Kanamori method provides estimates for $H$ and $R=/frac{V_P}{V_S}$ for 35 stations.
+al., 2010) uses an approach similar to the Zhu and Kanamori method published estimates for $H$ and $R=\frac{V_P}{V_S}$ for 35 stations.
 5 stations with $R$ estimates above 1.8 are removed due to the uncertainty surrounding these high values.
-Both datasets are plotted in Figure~\ref{fig:thompsonComp}.
-The Pearson correlation coefficient between these datasets is 0.49.
+There is strong correlation of 0.95 between crustal thickness values for both datasets (Figure~\ref{fig:thompsonCompH}).
+The velocity ratio data has a lower correlation of 0.5 (Figure~\ref{fig:thompsonCompR}).
 [Compare this value to uncertainty / deviation in the data?
@@
 Directly comparing $V_P$ estimates processed with the MB algorithm is not possible as this method has not been employed in publication.
-However, comparisons can be made to active source records for experiments within close proximity to a given seismic station (Figure~ \ref{fig:activeComp}.
+However, comparisons can be made to active source records for experiments within close proximity to a given seismic station (Figure~\ref{fig:activeComp}.
 The correlation between these datasets is 0.235.
 [This is low, explain here?].
-Another method of determining the reliability of estimates is to check the $/frac{V_P}{V_S}$ ratio taken from the Kanamori approach to the $/frac{V_P}{V_S}$ value computed from the MB approach.
+Another method of determining the reliability of estimates is to check the $\frac{V_P}{V_S}$ ratio taken from the Kanamori approach to the $\frac{V_P}{V_S}$ value computed from the MB approach.
 These values, working with the same data, preprocessed using the same methods and tools should be equal.
-Selecting only those stations who have an error of less than +/-0.05 $R$, 132 stations, we get a correlation of 0.57.
+Selecting only those stations with an error of less than +/-0.05 $R$, 132 stations, we get a correlation of 0.57.
 Reducing the error tolerance and selecting stations with less than +/-0.01 $R$, leaving 31 of the cleanest stations, we get a correlation coefficient of 0.95.
@@
   \centering
-    \includegraphics[width=\textwidth]{thompsonComparison}
-  \caption{Comparison of $\frac{V_P}{V_R}$ data processed for this study with data from Thomspson et. al. (2010).}
-  \label{fig:thompsonComp}
+    \includegraphics[width=\textwidth]{thompsonComparisonH}
+  \caption{Comparison of crustal thickness $H$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.95}
+  \label{fig:thompsonCompH}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{thompsonComparisonR}
+  \caption{Comparison of $\frac{V_P}{V_R}$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.5}
+  \label{fig:thompsonCompR}
 \end{figure}
@@
 
-  The preliminary interrogation of the data set yields the observation that the bulk Canadian crust is more felsic than originally anticipated.
-Previous experiments show crustal averages of 0.265 (Christensen and Mooney, 1995) or have averages of at least 0.27 for shields, platforms and paleozoic units (Zandt and Ammon, 1995).
-The weighted average for this data set shows a Poisson' ratio of 0.258 with the weighting for each station given by the calculation of the stations inverse Voronoi cell area.
-At the regional scale we have values of 0.251.
-0.254, 0.260, 0.275 for the Churchill, Superior, Slave and Grenville Provinces Poisson's ratios respectively.
-Most of the Canadian Shield stations also appear to have low Poisson values with the exception of the Grenville Province which has a high Poisson Ratio of 0.275.
+\subsection{Crustal Thickness}
+ (MAY NOT INCLUDE)
+\begin{itemize}
+  \item Jump in crustal thickness over the fault as you go from the Slave to the Rae or Churchill Province.
+  \item Trend in Southern Ontario as we move north and up the St Lawrence Seaway.
+  \item Appears to be secular variation in the North Eastern Churchill. As we move North and East we have thickening of continental crust.
+\end{itemize}
 
-  The data bears out earlier results showing Proterozoic crust to have a higher Poisson's ratio than Archean crust, likely the result of a more mafic composition.
+\subsection{Vp/Vs}
+\begin{itemize}
+  \item reference Figure~\ref{fig:VpVsMap}
+  \item For perspective, the value for Quartzite is 1.65 and Basalt is 1.85.
+  \item North East of the Slave province, has cluster of low Vp/Vs values in a semi circle of high values.
+  \item Appears to be an shift throughout the Superior province from high Vp/Vs in the South to low Vp/Vs in the North East.
+  \item Compare to Crust 2.0 overlay (Figure~\ref{fig:VpVsMapCrust}).  Average value for Churchill province Vp/Vs is 1.73 while average value computed from Crust 2.0 is 1.77.
+\end{itemize}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{VpVsMap}
+  \caption{}
+  \label{fig:VpVsMap}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{VpVsMapCrust}
+  \caption{}
+  \label{fig:VpVsMapCrust}
+\end{figure}
+
+%% -----------------------------------------------%%
+\section{Discussion}
+\subsection{Canada}
+
+  The preliminary interrogation of the data set yields the observation that the bulk Canadian Shield has lower Vp/Vs than anticipated.
+Previous experiments show crustal averages of 1.77 (Christensen and Mooney, 1995) and 1.78 (Zandt and Ammon, 1995).
+This compares to computed values of 1.73, 1.74, 1.795 for the Churchill, Superior and Grenville Provinces Vp/Vs ratios respectively.
+
+(MIGHT NOT INCLUDE)  The data bears out earlier results showing Proterozoic crust to have a higher seismic velocity ratio than Archean crust.
+This follows previous studies on crustal formation that have noted this trend in other regions (Durrheim and Mooney, 1991) .
 The increased crustal thickness of Proterozoic crust is clearly seen in the active source data while it not visible in data from processed seismic stations.
-The most likely reason for not seeing the signal of thicker crust in the station data is owing to poor station coverage in Proterozoic regions.
 
@@
 
-%% -----------------------------------------------%%
-\section{Discussion}
-\subsection{Canada}
-<Discussion on larger - Canada wide - regions>
 \subsection{Slave Province}
@@
 
-More Information and Advice:
-
 %
@@
 %% ------------------------------------------------------------------------ %%
+
+ABSTRACT
+   It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
+Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
+Durrheim, W.
+Mooney, 1991).
+Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
+We apply a new stacking method (Bostock and Kumar, 2010) to retrieve estimates of bulk crustal velocities (Vp, Vs) and thickness H from these data under the assumption of 1-D structure.
+Bootstrap error analysis is performed for each station dataset and results are compared to the results produced from alternate methods (Zhu and Kanamori, 2000).
+Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
+These results will be used to evaluate competing models of early crustal formation.
 

### 2013-02-27 python/mooneyraw2shape.py

@@ -91,15 +91,15 @@ def mooney2shapefile(moon):
     # Set fields for attribute table
 
     w.field("mcode", 'C', '5')
-    w.field("m2Vp", 'C', '6')
-    w.field("m2H", 'C', '6')
+    w.field("mVp", 'N', 7, 3)
+    w.field("mH", 'N', 7, 3)
 
     for k in moon.keys():
         # Set lon & lat
         w.point( moon[k]["lon"], moon[k]["lat"] )
         w.record( k,
-                  "{0:2.3f}".format(moon[k]["Vp"]),
-                  "{0:2.3f}".format(moon[k]["H"]))
+                  "{0:3.3f}".format(moon[k]["Vp"]),
+                  "{0:3.3f}".format(moon[k]["H"]))
 
     w.save(shpfile)
 

### 2013-02-27 python/poisson.py

@@ -1,0 +1,30 @@
+#!/usr/bin/python2
+#
+# python-mode indent C-c < or C-c >
+# python-mode comment/uncomment region M-;
+
+###########################################################################
+# IMPORTS
+###########################################################################
+
+import re, sys
+from histplots import poisson
+
+
+
+if __name__== '__main__' :
+
+    reverse = False
+
+    if len(sys.argv) > 1:
+        if sys.argv[1] == '-r':
+            reverse = True
+
+    numbers = sys.stdin.read().split()
+
+    numbers = map(float, numbers)
+
+    for n in numbers:
+        print poisson(n, reverse)
+
+

### 2013-03-01 matlab/dataDriver.m
prepare for initial manuscript

@@ -20,7 +20,7 @@ sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'ULM'; %
+station = 'KNDN'; %
 
 %{
 Need to find which stations go with which events.
@@ -82,8 +82,8 @@ end
 %     dlist{end+1} = fullfile(workingdir, events{ii});
 % end
 % db = process(db, dlist, station, workingdir, method, vp, 0);
-%ProcessStack
-ProcessTraces
+ProcessStack
+%ProcessTraces
 %% Assign Data
 
 [ db ] = assigndb( db, method, station, brec(:,1:round(45/dt)), ...

### 2013-03-06 paper/draftBP.tex
prepare for initial manuscript
outline

@@
 \section{Data and method}
-   Seismic data utilized in this study come from 343 stations across Canada from all available networks.
-Of these stations 146 are of sufficient quality to be useful for study.
-All stations are broadband with data selected between the years 2000 and 2012, for a total of more than 700 earthquake sources.
+    The data utilized in this study are teleseismic P-wave seismograms from 343 broadband stations across Canada.
+These stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
+Seismograms have been selected between the years 2000 and 2012 which employ more 1000 earthquake sources.
 All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing.

### 2013-03-06 node/nodecleaner.js

@@ -6,9 +6,8 @@ var rmdir = require('rimraf')
 var stations = ["DPQ","GAC","YKW3","OTT"]
 var stn = "DPQ"
 
+
 var finder = f.find("/media/TerraS/CN/" + stn)
-
-
 
 finder.on('directory', function (dir, stat) {
 

### 2013-03-06 node/tallyData.js

@@ -1,0 +1,15 @@
+var f = require('findit')
+var fs = require('fs')
+var path = require('path')
+
+var finder = f.find("/media/TerraS/CN/" )
+var count = 0
+
+finder.on('file', function (file) {
+  if (path.basename(file) === 'stack_P.sac')
+    count++
+})
+
+finder.on('end', function() {
+  console.log("total files = " + String(count))
+})

### 2013-05-10 mapping/web/server.js
finish comparisons

@@ -8,7 +8,7 @@ var st = require("st")
   , fs = require("fs")
 
 // Listen on port
-var PORT = 8082
+var PORT = 8080
   , STATIC = 'public'
   , mount = st(
     { path: './' + STATIC

### 2013-05-10 matlab/autogrid.m

@@ -7,10 +7,10 @@ addpath([userdir,'/programming/matlab/jsonlab']);
 databasedir = '/media/TerraS/database';
 
 
-json = loadjson([userdir,'/thesis/data/thompsonPaper.json']);
+json = loadjson([userdir,'/thesis/data/darbyshirePaper.json']);
 stns = fieldnames(json);
 
-fhout = fopen([userdir,'/thesis/data/thomsponProcessed.json'], 'w');
+fhout = fopen([userdir,'/thesis/data/darbyshireProcessed.json'], 'w');
             
 %Setup parallel toolbox
 if ~matlabpool('size')
@@ -22,7 +22,8 @@ for stn = stns'
 
     station = stn{1};
     load(fullfile(databasedir, station));
-    vp = 6.39;
+    vp = 6.39; %Eaton & DARBYshire
+    %vp = 6.5; % Thompson
     
     method = 'kanamori';
     [ results ] = gridsearchKan(db.rec, db.dt, db.pslow, vp);

### 2013-05-10 matlab/dataDriver.m

@@ -20,7 +20,7 @@ sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'GBLN'; %
+station = 'PKLO'; %
 % KGNO
 %{
 

### 2013-05-10 python/compareResults.py

@@ -25,8 +25,8 @@ data = ['/thesis/data/thompsonPaper.json',
         '/thesis/data/darbyshireProcessed.json']
 
 leglabel = ["Thompson et. al.",
-         "Eaton et. al.",
-         "Darbyshire et. al."]
+            "Eaton et. al.",
+            "Darbyshire et. al."]
 
 #######################
 # Plotting formatter
@@ -45,7 +45,7 @@ title = 18 / ratio
 leg = 16 / ratio
 
 
-for i in range(2):
+for i in range(3):
 
     pro = os.environ['HOME'] + data[2*i + 1]
     pub = os.environ['HOME'] + data[2*i]
@@ -58,9 +58,9 @@ for i in range(2):
     p1.sync(p2)
 
     Rcorr = pearsonr(p1.R, p2.R)
-    print "Correlation between Vp/Vs datasets is {0:.2f}".format(Rcorr[0])
+    print "Correlation between", leglabel[i], "Vp/Vs datasets is {0:.2f}".format(Rcorr[0])
     Hcorr = pearsonr(p1.H, p2.H)
-    print "Correlation between H datasets is {0:.2f}".format(Hcorr[0])
+    print "Correlation between", leglabel[i], "H datasets is {0:.2f}".format(Hcorr[0])
 
 
     t = np.arange(len(p1.R))
@@ -109,4 +109,4 @@ for i in range(2):
     plt.axis("tight")
 
 
-plt.show()
+#plt.show()

### 2013-06-12 paper/draftBP.tex
finish comparisons
intro and paper

--- 2013-02-27 paper/draftBP.tex
@@
-%% ------------------------------------------------------------------------ %%
 \documentclass[draft, 12pt]{article}
@@
 
-%% ------------------------------------------------------------------------ %%
 \begin{abstract}
+
 \end{abstract}
 
-%% ------------------------------------------------------------------------ %%
-
-%% ------------------------------------------------------------------------ %%
 %% -----------------------------------------------%%
 \section{Introduction}
-(Article text here)
-\subsection{Overview}
-(Overview here)
-\subsection{Geological Background}
-(Geological background here)
+
+  Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
+The primary reason for this is the poor resolution afforded by the seismic networks currently and historically deployed across Canada.
+The Canadian continental landmass is composed of at least fifteen large geological provinces as recognized by the Geological Survey of Canada (GSC).
+Each of these regions is itself complex and heterogeneous and often larger than most European nations.
+On top of this there is poor seismic coverage, roughly one seismic station per 25,000 km^2.
+Many of these stations are clustered near areas of geologic interest such as the Cascadia Subduction zone or population and research centres such as Southern Ontario or areas of significant resource interest like the diamondiferous Great Slave Lake area.
+This leaves vast areas of the Canadian landmass completely unsampled.
+Despite these limitations, a low resolution comprehensive and comparative study is still feasible for some of the geological regions comprising the Canadian continental crust.
+Direct comparison between the bulk average seismic properties of geological regions and between aggregated regions and global averages provide some new insight into the crustal composition of the Canadian landmass.
+The accumulated dataset also affords investigations into variations of bulk crustal parameters such as Poisson's Ratio and crustal thickness with age and tectonic environment as well as some basic statistical data on interesting crustal features.
+
+  This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
+It begins with a discussion of the raw data itself followed by a review of the receiver function method and a more detailed explanation of the inversion algorithms used to produce the dataset.
+Three previous publications utilizing similar processing schemes provide unique subsets of data to compare with the values computed in this survey for quality assurance.
+...
+
+\subsection{Geological and tectonic summary}
+
+\subsection{Previous geophysical studies}
 
 %% -----------------------------------------------%%
-\section{Data and method}
-   Seismic data utilized in this study come from 343 stations across Canada from all available networks.
-Of these stations 146 are of sufficient quality to be useful for study.
-All stations are broadband with data selected between the years 2000 and 2012, for a total of more than 700 earthquake sources.
+\section{Data and methods}
+  Analysis of bulk Canadian continental crust requires estimates of crustal properties such as seismic-wave velocity and crustal thickness.
+This study draws on three sources for these crustal properties.
+The primary source of estimates are data computed by first processing raw seismograms into receiver functions and then inverting the RF's for parameter estimates.
+These primary data are supplemented and qualified by two existing datasets, the statistically compiled Crust 2.0 dataset and a compilation of pre-processed active source data.
+
+\subsection{Teleseismic Data Set}
+
+  The primary data utilized in this study are computed from teleseismic P-wave seismograms representing discrete seismic events.
+These events are comprised by the seismic records of more than 700 earthquake sources occuring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
+Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
+Events are filtered by the epicentral distance from source to receiver with only the events within a 30 to 100 degree window being included.
+Seismic events are further filtered by hand picking those with reasonable signal to noise ratio and of sufficient impulsiveness that arrival times can be accurately measured.
+After selection and filtering more than 80,000 events are available for further processing.
+
+Seismograms are selected between the years 2000 and 2012 which employ more 1000 earthquake sources.
 All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing.
@@
 
-(MIGHT NOT INCLUDE)  The data bears out earlier results showing Proterozoic crust to have a higher seismic velocity ratio than Archean crust.
+(MIGHT NOT INCLUDE)  The data bares out earlier results showing Proterozoic crust to have a higher seismic velocity ratio than Archean crust.
 This follows previous studies on crustal formation that have noted this trend in other regions (Durrheim and Mooney, 1991) .

### 2013-06-12 matlab/junkcode.m

@@ -64,5 +64,47 @@ end
 %
 %thresh = 0.3;
 %crec = performCurveletDenoise(brec,dt,thresh);
+clear all
+tic;
+m = 30;
+n = 160;
 
+% m = 5;
+% n = 15;
+nwin = n - (m-1);
+
+I = speye(m);
+
+
+tic
+for ii = 1:nwin
+    ind = (ii-1)*m + 1;
+    A( ind : ind + m - 1, ii: m+ii-1) = I;
+end
+toc;
+
+
+
+% A = sparse(A);
+% 
+% 
+% x1 = randn(n,1);
+% x2 = x1 + randn(n,1);
+% 
+% 
+% b1 = A*x1;
+% b2 = A*x2;
+% 
+% B1 = reshape(b1, m, nwin);
+% B2 = reshape(b2, m, nwin);
+% 
+% B1 = B1 ./ kron( ones(m,1), sqrt(sum(B1.^2, 1)));
+% B2 = B2 ./ kron( ones(m,1), sqrt(sum(B2.^2, 1)));
+% 
+% corr = spdiags(B2(:), 0, nwin*m, nwin*m) * B1(:);
+% 
+% C = reshape(corr, m, nwin);
+% 
+% corr = sum(C, 1);
+% toc
 

### 2013-06-12 matlab/runAutoProcessing.m

@@ -40,19 +40,15 @@ for ii = 1 : length(s)
         func(db, dbfile);
 %         if isfield(db, 'conrad')
 %             if isfield(db.conrad, 'hdisc')
-%                 hcount = hcount + 1;
-%                 H = db.conrad.H;
 %                 for jj = 1:length(db.conrad.hdisc)
-%                     h(indh) = H(db.conrad.hdisc(jj));
+%                     h(indh) = db.conrad.hdisc(jj);
 %                     indh = indh + 1;
 %                 end
 %                 
 %             end
 %             if isfield(db.conrad, 'hdiscp')
-%                 hpcount = hpcount + 1;
-%                 H = db.conrad.H;
 %                 for jj = 1:length(db.conrad.hdiscp)
-%                     hpicked(indhp) = H(db.conrad.hdiscp(jj));
+%                     hpicked(indhp) = db.conrad.hdiscp(jj);
 %                     indhp = indhp + 1;
 %                 end
 %             end

### 2013-06-12 python/dbutils.py

@@ -151,7 +151,7 @@ def json2shapefile(stdict, fout, sc):
         for f in conradfields:
             d = None
             if f in sc.flattendict(stdict[key]) and sc.flattendict(stdict[key])[f]:
-                disc = [disc for disc in sc.flattendict(stdict[key])[f] if (13.0 < disc < 19.0)]
+                disc = [disc for disc in sc.flattendict(stdict[key])[f] if (13.0 < disc < 21.0)]
                 if disc:
                     d = disc[0]
             values.append('{0:2.4f}'.format(d) if d else 0.0)

### 2013-06-14 paper/draftBP.tex
draftBP

@@
 Each of these regions is itself complex and heterogeneous and often larger than most European nations.
-On top of this there is poor seismic coverage, roughly one seismic station per 25,000 km^2.
+On top of this there is poor seismic coverage, roughly one seismic station per 25,000 $km^2$.
 Many of these stations are clustered near areas of geologic interest such as the Cascadia Subduction zone or population and research centres such as Southern Ontario or areas of significant resource interest like the diamondiferous Great Slave Lake area.
@@
 This study draws on three sources for these crustal properties.
-The primary source of estimates are data computed by first processing raw seismograms into receiver functions and then inverting the RF's for parameter estimates.
-These primary data are supplemented and qualified by two existing datasets, the statistically compiled Crust 2.0 dataset and a compilation of pre-processed active source data.
+The first and primary source for these estimates are data computed by processing raw seismograms into receiver functions and then inverting these RF's for parameter estimates.
+The other two datasets, the statistically compiled Crust 2.0 dataset and a compilation of pre-processed active source data, are used to qualify and compare with the primary dataset.
 
@@
   The primary data utilized in this study are computed from teleseismic P-wave seismograms representing discrete seismic events.
-These events are comprised by the seismic records of more than 700 earthquake sources occuring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
+These events are comprised of seismic records representing more than 700 earthquake sources occuring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
 Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
@@
 
-Seismograms are selected between the years 2000 and 2012 which employ more 1000 earthquake sources.
-All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing.
-Data is selected for reasonable signal to noise ratio and sufficient impulsiveness that arrival times can be calculated.
-All data is analyzed using receiver function methods, an approach which has been widely employed over many years to investigate the earths structure.
+  The first stage of processing requires the transformation of teleseismic data into receiver functions.
+In its generic form this transformation involves deconvolving an approximation of the source signal from channels representing ground motion.
+The resulting waveform contains discrete pulses which corrispond to arrivals of S-wave energy scattered from subsurface discontinuities.
+The resolution of these pulses has first-order effectsTo gain better resolution of these peaks we follow a wave field decomposition approach to To better focus this energy the North and East channels are rotated into radial and transverse components.
+Then the radial and vertical components are then decomposed into P, SV and SH components with a wave field decomposition transfer matrix [Bostock, 1998].
 
@@
     \includegraphics[width=\textwidth]{thompsonComparisonR}
-  \caption{Comparison of $\frac{V_P}{V_R}$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.5}
+  \caption{Comparison of $V_P / V_R$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.5}
   \label{fig:thompsonCompR}

### 2013-06-14 paper/draftBP.tex
paper

@@
 \usepackage[final]{graphicx}
+\usepackage{color}
 \graphicspath{ {figures/} }
 \DeclareGraphicsExtensions{.png}
+
+\newcommand{\cc}[1]{\textcolor{red}{#1}}
 
@@
   The first stage of processing requires the transformation of teleseismic data into receiver functions.
-In its generic form this transformation involves deconvolving an approximation of the source signal from channels representing ground motion.
+In its generic form this transformation involves deconvolving an approximation of the earthquake source from channels representing ground motion.
 The resulting waveform contains discrete pulses which corrispond to arrivals of S-wave energy scattered from subsurface discontinuities.
-The resolution of these pulses has first-order effectsTo gain better resolution of these peaks we follow a wave field decomposition approach to To better focus this energy the North and East channels are rotated into radial and transverse components.
-Then the radial and vertical components are then decomposed into P, SV and SH components with a wave field decomposition transfer matrix [Bostock, 1998].
+Resolving these peaks is essential to the success of this method, therefore it is advantageous to rotate the seismogram components to seperate the P and S wave energy.
+This is accomplished by first rotating the N and E coordinates into radial and transverse dimensions and then performing a wave field decomposition on the radial and vertical channels [Bostock, 1998].
+The direct arrival of the signal on the resulting P wave componenent is used as an approximation to the source function as P waves are simpler and have a response closer to that of a delta function.
+This windowed source coda is deconvolved from the S wave component computed from the wave field decomposition.
 
- Two derivations of receiver function analysis have been used in this study, one well tested stacking approach and the other a recent method that has not yet been employed at scale.
-Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts.
-Deconvolution of the P-wave energy arrival, used as an estimate for the source function, from the S-wave signal produces a Green's function with energy peaks at the times of the main and reflected arrivals.
-For better separation of P and S wave energy all components are rotated into radial and transverse dimensions and transformed into P, SV and SH components with a wave field decomposition transfer matrix (Bostock, 1998).
+  An $L_2$ frequency domain deconvolution approach is used which has the advantages of computational efficiency as well as not requiring any assumptions about the noise in the data.
+This method performs a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $g(t)$.
 
- Deconvolution is performed by minimizing the general cross validation function $GCV(\delta)$ where $\delta$ is used the regularizer (Bostock, 1998; Golub et al., 1979).
-This method is an $L_2$ frequency domain deconvolution which performs quickly and makes no assumptions on the noise in the data.
-All deconvolved signals are filtered between 0.04hz and 3.0hz and are stacked with weights $w_1=0.5$, $w_1=0.3$, $w_1=0.2$ for the Ps, PpPs and PpSs phases respectively.
-Error calculations are performed with bootstrap resampling by rerunning the processing with randomly chosen receiver functions - allowing multiples - 1024 times.
-The error is obtained by taking the standard deviation of the results.
+\begin{equation}
+  g(t) = F^{-1} \left[ G(\omega) \right] = F{^-1}
+ \left[ \frac {\sum_n^N S_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta} \right ],
+\end{equation}
+
+where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ S wave component, $P_n$ is the windowed P wave component, $^*$ denotes the complex conjugatte and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
+The parameter $\delta$ is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ which is given by
+
+\begin{equation}
+  GCV(\delta) = \frac {\sum_n^N\sum_m^M \left( S_n(\omega_m) - P_n(\omega_m)G(\omega_m) \right)^2 }
+                      { \left( NM - \sum_m^M X(\omega_m) \right)^2 },
+\end{equation}
+
+where
+
+\begin{equation}
+  X(\omega) = \frac {\sum_n^N P_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta},
+\end{equation}
+
+and $\omega_m$ is the $m^{th}$ frequency bin in the discrete Fourier transform.
+All receiver functions are filtered between 0.04Hz and 3.0Hz and comprise the data for the following parameter inversions.
 

### 2013-06-25 paper/draftBP.tex
paper

@@
 In its generic form this transformation involves deconvolving an approximation of the earthquake source from channels representing ground motion.
-The resulting waveform contains discrete pulses which corrispond to arrivals of S-wave energy scattered from subsurface discontinuities.
-Resolving these peaks is essential to the success of this method, therefore it is advantageous to rotate the seismogram components to seperate the P and S wave energy.
+The resulting waveform contains discrete pulses corresponding to the arrivals of S-wave energy scattered from subsurface discontinuities.
+Resolving these peaks is essential to the success of the following inversion, therefore it is advantageous to rotate the seismogram components to separate the P and S wave energy.
 This is accomplished by first rotating the N and E coordinates into radial and transverse dimensions and then performing a wave field decomposition on the radial and vertical channels [Bostock, 1998].
-The direct arrival of the signal on the resulting P wave componenent is used as an approximation to the source function as P waves are simpler and have a response closer to that of a delta function.
+The direct arrival of the signal on the resulting P wave component is used as an approximation to the source function as P waves are simpler and have a response closer to that of a delta function.
 This windowed source coda is deconvolved from the S wave component computed from the wave field decomposition.
@@
 This method performs a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $g(t)$.
-
+% Deconvolution equations
 \begin{equation}
@@
 
-where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ S wave component, $P_n$ is the windowed P wave component, $^*$ denotes the complex conjugatte and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
+where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ S wave component, $P_n$ is the windowed P wave component, $^*$ denotes the complex conjugate and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
 The parameter $\delta$ is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ which is given by
@@
 and $\omega_m$ is the $m^{th}$ frequency bin in the discrete Fourier transform.
-All receiver functions are filtered between 0.04Hz and 3.0Hz and comprise the data for the following parameter inversions.
+
+  All resulting receiver functions, $g(t)$, are filtered between 0.04Hz and 3.0Hz.
 
 \subsection{Vp/Vs method}
-   A well tested approach for calculating $R=\frac{V_P}{V_S}$ and $H$ was employed by Zhu and Kanamori (2000).
-After deconvolution the signals are stacked and a gridsearch is performed using the travel time equations for different values of $R$ and $H$ and a spectrum of slowness.
-For each parameter combination the travel time functions output estimates for the differential arrival time of the S-component main and reflected phases.
-These times are used to pick the energy in each receiver function - with the large amplitudes at arrivals stacking to provide an estimate for the optimal parameter estimation.
 
-\subsection{Vp method}
-   The stacking approach outlined above provides estimates for the parameters $R=\frac{V_P}{V_S}$ as well as crustal thickness $H$ - which gives the depth of the Moho.
-A method which estimates $V_P$ and $V_S$ separately (Bostock, 2010) is also employed for select stations.
-This method makes use of the fact the dependence on $H$ in the travel time equations can be removed if we divide the reflected phases by the main arrival.
-$$ t_{Pps}(p_i) = \frac{ \sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
-$$ t_{Pss}(p_i) = \frac{2\sqrt{R^2-p_i^2V_P^2}} {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
-This has the advantage that no assumptions on $V_P$ are necessary to perform the gridsearch and stack.
-A similar stacking method as the Zhu and Kanamori approach may be employed for $t_{Pps}$ and $t_{Pss}$ as long as an estimate for $t_{Ps}$ exists.
-With estimates for $R$ and $V_P$ a simple line search along $H$ can be made using all three undivided travel time equations.
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{reflectedPhases}
+  \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
+  \label{fig:reflectedPhases}
+\end{figure}
 
-(MOST LIKELY UNNECESSARY INFO)
-Several methods for choosing $t_{Ps}$ are available.
-Direct arrival $t_{Ps}$ should contain the largest fraction of energy out of the other phases and the moveout, depending on Moho depth, should be in the order of 3 to 6 seconds.
-Given this it is trivial to define a window and select the time corresponding to maximum amplitude.
-Another approach is to use max amplitude estimates as data in a non-linear optimization to find the $R$, $V_P$, and $H$ which minimizes the residual between $t_{Ps}$ and the data.
-The travel time equations are twice-differentiable so the quadratically convergent Newton's method may be employed.
-This approach has the advantage that noise leading to poor maximum amplitude picks are effectively collapsed onto the curve corresponding to the travel time function.
-A third approach which offers picks along a travel time curve but better stability than the non-linear method is to perform a gridsearch with all three travel time functions, a full Kanamori stack.
-With the best estimates for $R$ and $H$ the $t_{Ps}$ function can be found and used as input into the Bostock method.
-The trade-off is that the requirement for an initial $V_P$ estimate for the initial stack introduces secondary $V_P$ dependence into the system.
+  A well tested and widely published method for extracting the ratio, $R=\frac{V_P}{V_S}$, where $V_P$ is P-wave velocity and $V_S$ is S-wave velocity and crustal thickness, or depth to Moho, $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
+This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$ $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$, Figure ~\ref{fig:reflectedPhases}.
+Note that $PsPs$ and $PpSs$ have an equal number of $P$ and $S$ phases, and the energy for these two phases arrive simultaneously and are indistinguishable.
+For a range of slowness values, $p$, the differential arrival times, $t(p)$, trace moveout curves for each phase arrival given by
 
-\subsection{Vp database}
+% Travel time equations
+\begin{equation}
+t_{Ps}(p_i)=H \left[ \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2} - \sqrt{1 - p_i^2V_P^2} \right]
+\end{equation}
+
+\begin{equation}
+t_{Pps}(p_i)=H \left[ \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2} + \sqrt{1 - p_i^2V_P^2} \right]
+\end{equation}
+
+\begin{equation}
+t_{Pss}(p_i)= 2H  \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2}
+\end{equation}
+
+where $p_i$ is the slowness for the $i^{th}$ receiver function.
+Since strong reflected phases occur at sharp velocity contrasts, the Moho, the boundary targeted by ZK, tends to be well represented on most RF's.
+
+  The travel time equations demand an apriori assumption on crustal P-wave velocity, $V_P$, and will trade-off to some degree with crustal thickness $H$.
+For this study, each station is assigned a $V_P$ values corrisponding to the Crust 2.0 value for the $2^o$ containing cell.
+In the specific cases where data is being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
+
+  For a range of candidate models of $R$ and $H$ the RF's are stacked along trial moveout curves.
+During the stacking procedure each phase is assigned a weight to account for a general trend in the quality of the phases, with direct arrival usually carrying the best signal followed by $PpPs$ and finally $PpSs$.
+The weights chosen during the stacking are $w1 = 0.5$, $w2 = 0.3$ , $w3 = 0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases respectively.
+In addition Semblance weighting [Eaton, 2006] is used to reduce the effect of spurious large amplitude noise in the data.
+The model which stacks the most coherent energy is used to provide the best estimates for the bulk crustal parameters $R$ and $H$ under a given seismic station.
+
+\subsection{Full Gridsearch Method}
+
+\subsection{Vp data}
+   Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources.
+The data was compiled by Mooney (personal communication, 2012).
+
+\subsection{Crust 2.0 data}
    Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources.

### 2013-06-26 paper/draftBP.tex
paper

@@
   An $L_2$ frequency domain deconvolution approach is used which has the advantages of computational efficiency as well as not requiring any assumptions about the noise in the data.
-This method performs a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $g(t)$.
+This method performs a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $r(t)$.
 % Deconvolution equations
 \begin{equation}
-  g(t) = F^{-1} \left[ G(\omega) \right] = F{^-1}
+  r(t) = F^{-1} \left[ G(\omega) \right] = F{^-1}
  \left[ \frac {\sum_n^N S_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta} \right ],
@@
 
-  All resulting receiver functions, $g(t)$, are filtered between 0.04Hz and 3.0Hz.
+  All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
 
@@
   The travel time equations demand an apriori assumption on crustal P-wave velocity, $V_P$, and will trade-off to some degree with crustal thickness $H$.
-For this study, each station is assigned a $V_P$ values corrisponding to the Crust 2.0 value for the $2^o$ containing cell.
+For this study, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^o$ containing cell.
 In the specific cases where data is being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
@@
   For a range of candidate models of $R$ and $H$ the RF's are stacked along trial moveout curves.
-During the stacking procedure each phase is assigned a weight to account for a general trend in the quality of the phases, with direct arrival usually carrying the best signal followed by $PpPs$ and finally $PpSs$.
-The weights chosen during the stacking are $w1 = 0.5$, $w2 = 0.3$ , $w3 = 0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases respectively.
-In addition Semblance weighting [Eaton, 2006] is used to reduce the effect of spurious large amplitude noise in the data.
+During the stacking procedure each phase is assigned a weight to account for a general trend in the quality of the phases, with direct arrival usually carrying the best signal followed by $PpPs$ and $PpSs$.
+The weights chosen during the stacking are $w1 = 0.5$, $w2 = 0.3$ , $w3 = -0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases respectively.
+A negative weight for the combined $PpSs$ and $PsPs$ phases is required as the polarity of the signal is reversed.
+Semblance weighting [Eaton, 2006] is employed to reduce the effect of spurious large amplitude noise in the data.
+The semblance function assigns a weight between zero (incoherent noise) and one (coherent signal).
+The stacking function is therefore defined as
+
+\begin{equation}
+s(H,R) = \sum_{j=1}^{3} S_j \sum_{i=1}^N w_jr_i(t_j)
+\end{equation}
+
+where
+
+\begin{equation}
+S_j(H,R) = \frac {\left[ \sum_{i=1}^N r_i(t_j) \right]^2}
+                 { \sum_{i=1}^N r_i^2(t_j) }
+\end{equation}
+
+and $S_j$ is the semblance weight for the $j^{th}$ phase, time $t_j$ is calculated from the corrisponding travel time function for a given $H$, $R$ as a function of slowness $p_i$ and $N$ is the number of receiver functions.
+Multiplying by the semblance weighting sharpens the stacked image, $s(H,R)$, and exhibits better resolution when selecting between stacked models.
 The model which stacks the most coherent energy is used to provide the best estimates for the bulk crustal parameters $R$ and $H$ under a given seismic station.
 
+  For a given synthetic receiver function
 \subsection{Full Gridsearch Method}
+
+  Owing to increasing computational availability it is now reasonably efficient to extend the ZK approach and perform the stack along all three seismic parameters, $V_P$, $R$ (therefore by definition $V_S$) and $H$.
 

### 2013-06-26 matlab/functions/gridsearch3D.m

@@ -1,0 +1,91 @@
+function [ results ] = gridsearch3D(rec, dt, pslow)
+%GSEARCHKANAMORI Grid Search as implemented in Kanamori, Zhu 2000
+%
+% Grid search over Tps TPpPs and TPpSs with parameters R (Vp/Vs) and H
+% Input Mooneys v
+
+
+%% Grip parameters
+% Vp/Vs Ratio
+nr = 100;
+r1 = 1.65; %1.65
+r2 = 1.95;
+dr = (r2 - r1)/(nr - 1);
+r = r1:dr:r2;
+
+% Thickness H
+nh = 100;
+h1 = 25;
+h2 = 50;
+dh = (h2 - h1)/(nh - 1);
+h = h1:dh:h2;
+
+
+% Vp
+nv = 100;
+v1 = 5;
+v2 = 8;
+dv = (v2 - v1)/(nv - 1);
+v = v1:dv:v2;
+
+
+% Misc
+%v = 6.38; % bulk vp pwave velocity.
+p2 = pslow.^2;
+np = length(pslow);
+nt = length(rec);
+w1 = 0.5;
+w2 = 0.3;
+w3 =-0.2;
+
+% Reshape for fast access
+gvr = rec'; % rotate
+gvr = gvr(:); % vectorize
+
+%% Grid serarch for r and H
+stackhr = zeros(nr,nh,nv);
+parfor ir = 1:nr
+    for ih = 1:nh
+        for iv = 1:nv
+            
+            f1 = sqrt((r(ir) / v(iv))^2 - p2);
+            f2 = sqrt((1 / v(iv))^2 - p2);
+            t1 = h(ih) * (f1 - f2);
+            t2 = h(ih) * (f1 + f2);
+            t3 = h(ih) * 2*(f1);
+            r1 = mean(gvr(round(t1/dt)+1+[0:np-1]*nt));
+            r2 = mean(gvr(round(t2/dt)+1+[0:np-1]*nt));
+            r3 = mean(gvr(round(t3/dt)+1+[0:np-1]*nt));
+        
+            stackhr(ir,ih, iv) = w1*r1 + w2*r2 + w3*r3;
+        end
+    end
+end
+
+% Select best/highest values and recalculate
+smax = max(max(max(stackhr)));
+[ir,ih,iv] = ind2sub(size(stackhr),find(stackhr == smax));
+
+hbest = h(ih);
+rbest = r(ir);
+vbest = v(iv);
+
+f1 = sqrt((rbest / vbest)^2 - p2);
[…]

### 2013-06-26 matlab/synthetic.m

@@ -1,0 +1,118 @@
+% Synthetic Recs
+close all; clear all
+loadtools;
+addpath functions
+databasedir = '/media/TerraS/database';
+station = 'ULM'; %
+sacfolder = '/media/TerraS/CN';
+
+dbfile = fullfile(databasedir, [station,'.mat'] );
+workingdir = fullfile(sacfolder,station);
+
+
+
+load(dbfile)
+
+s  = db.rec(7,:);
+p = db.pslow(7);
+dt = db.dt;
+
+t1 = db.hk.tps(7);
+t2 = db.hk.tpps(7);
+t3 = db.hk.tpss(7);
+p2 = p^2;
+
+%% R vs H
+v = db.hk.v;
+r = db.hk.rRange;
+
+f1 = sqrt((r / v).^2 - p2);
+f2 = sqrt((1 / v).^2 - p2);
+
+h1 = t1 ./ (f1 - f2);
+h2 = t2 ./ (f1 + f2);
+h3 = t3 ./ (2*f1);
+
+plot(h1, r, h2, r, h3, r)
+axis tight
+title('R Vs H')
+xlabel('H [km]')
+ylabel('R [Vp/Vs]')
+%% Vp vs R
+
+v = linspace(5,8,200);
+h = db.hk.hbest;
+
+
+f2 = sqrt((1 ./ v).^2 - p2);
+
+r1 = sqrt( v.^2 .* ((t1/h).^2 + 2 .* (t1/h) .* f2 + f2.^2 + p2) );
+r2 = sqrt( v.^2 .* ((t2/h).^2 - 2 .* (t2/h) .* f2 + f2.^2 + p2) );
+r3 = sqrt( v.^2 .* ( (t3 / (2*h)).^2 + p2) );
+
+figure()
+plot(r1, v, r2, v, r3, v)
+axis tight
+title('Vp Vs R')
+ylabel('Vp [km/s]')
+xlabel('R [Vp/Vs]')
+%% Vp vs H
+
+v = linspace(5,8,200);
+r = db.hk.rbest;
+
+f1 = sqrt((r ./ v).^2 - p2);
+f2 = sqrt((1 ./ v).^2 - p2);
+
+h1 = t1 ./ (f1 - f2);
+h2 = t2 ./ (f1 + f2);
+h3 = t3 ./ (2*f1);
+
+figure()
+plot(h1, v, h2, v, h3, v)
+axis tight
+title('Vp Vs H')
+xlabel('H [km]')
+ylabel('Vp [km/s]')
+
[…]

### 2013-06-27 paper/draftBP.tex
paper

@@
   The primary data utilized in this study are computed from teleseismic P-wave seismograms representing discrete seismic events.
-These events are comprised of seismic records representing more than 700 earthquake sources occuring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
+These events are comprised of seismic records representing more than 700 earthquake sources occurring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
 Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
@@
 
-\subsection{Vp/Vs method}
+\subsection{Vp/Vs method} \label{section:VpVsMethod}
 
@@
   A well tested and widely published method for extracting the ratio, $R=\frac{V_P}{V_S}$, where $V_P$ is P-wave velocity and $V_S$ is S-wave velocity and crustal thickness, or depth to Moho, $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
-This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$ $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$, Figure ~\ref{fig:reflectedPhases}.
+This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$ $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
 Note that $PsPs$ and $PpSs$ have an equal number of $P$ and $S$ phases, and the energy for these two phases arrive simultaneously and are indistinguishable.
@@
 % Travel time equations
-\begin{equation}
-t_{Ps}(p_i)=H \left[ \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2} - \sqrt{1 - p_i^2V_P^2} \right]
+\begin{equation} \label{eq:tps}
+t_{Ps}(p_i)=H \left[ \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2} - \sqrt{\frac{1}{V_P^2} - p_i^2} \right]
 \end{equation}
@@
 \begin{equation}
-t_{Pps}(p_i)=H \left[ \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2} + \sqrt{1 - p_i^2V_P^2} \right]
+t_{Pps}(p_i)=H \left[ \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2} + \sqrt{\frac{1}{V_P^2} - p_i^2} \right]
 \end{equation}
@@
 \begin{equation}
-t_{Pss}(p_i)= 2H  \sqrt{ \left(\frac{V_p}{V_s}\right)^2 - p_i^2V_p^2}
+t_{Pss}(p_i)= 2H  \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2}
 \end{equation}
@@
 
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{kinematicRvsH}
+  \caption{Kinematic curves traced out by the stacking function, $s(H,R)$, corrisponding to the mapping of fixed travel times $t_{Ps}$, $t_{PpPs}$ and $t_{PsPs}$ into $H$ and $R$ space.}
+  \label{fig:kinematicRvsH}
+\end{figure}
+
   For a range of candidate models of $R$ and $H$ the RF's are stacked along trial moveout curves.
@@
 
-\begin{equation}
+\begin{equation}  \label{eq:stack}
 s(H,R) = \sum_{j=1}^{3} S_j \sum_{i=1}^N w_jr_i(t_j)
@@
 Multiplying by the semblance weighting sharpens the stacked image, $s(H,R)$, and exhibits better resolution when selecting between stacked models.
-The model which stacks the most coherent energy is used to provide the best estimates for the bulk crustal parameters $R$ and $H$ under a given seismic station.
+The function $s(H,R)$ can be thought of as a transformation that maps the $Ps$, $PpPs$ and $PpSs+PsPs$ pulses of the receiver function, $r(t)$,  into $R$ and $H$ space as positive bands, via Eq.
+\ref{eq:stack}.
+Constructive interference where these bands intersect will produce a maximum in the stacking function (Fig \ref{fig:kinematicRvsH}).
+The model, $R$ and $H$, which produce this maximum provides the best estimate for the bulk crustal parameters for a given seismic station.
+The elongate shape about the point of intersection, with the long axis tilted mostly in the $R$ coordinate, conveys the inherent trade-off between $R$ and $H$ and the lower resolution exhibited by the seismic velocity ratio $R$.
 
-  For a given synthetic receiver function
 \subsection{Full Gridsearch Method}
 
-  Owing to increasing computational availability it is now reasonably efficient to extend the ZK approach and perform the stack along all three seismic parameters, $V_P$, $R$ (therefore by definition $V_S$) and $H$.
+  Owing to increasing computational availability it is now reasonably efficient to extend the ZK approach and perform the stack along a range of all three seismic parameters, $V_P$, $R$ (therefore by definition $V_S$) and $H$.
+This has the advantage of not requiring an assumption on $V_P$ and should produce more accurate estimates, especially above crustal units with anomalous $V_P$.
+The inherent trade-off between $V_P$ and $R$ is limited, offering reasonable resolution between the two parameters.
+However, a significant trade-off between $V_P$ and $H$, which is a source of significant correlated error, requires that recovery of $V_P$ should be attempted only in cases where high frequency and high signal-to-noise ratio signals are present in the receiver function data.
+[[ Could go into more depth here, show the kinematic curves and poor trade-off between $V_P$ and $R$, Figure showing correlation between $V_P$ and $H$]]
 
-\subsection{Vp data}
-   Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources.
-The data was compiled by Mooney (personal communication, 2012).
+  These trade-offs, along with other sources of error such as low signal-to-noise signals resulting from complex geology, are quantified by bootstrap resampling [Efron and Tibshirani, 1986].
+Estimates are reprocessed using a random selection of receiver functions, with replacement, 1024 times.
+An estimate for the error is computed by taking the standard deviation from these trial runs.
 
-\subsection{Crust 2.0 data}
-   Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources.
-The data was compiled by Mooney (personal communication, 2012).
+\subsection{Additional data}
+   Accompanying the processed estimates are data from controlled source experiments collected and compiled from GSC (Geological Survey of Canada) references by Walter Mooney (personal communication, 2012).
+This data provides $V_P$ and a few $V_S$ estimates for hundreds of experiment locations across Canada, many along traversals across geological provincial boundaries, faults and discontinuities.
+Some of these data are located within close enough proximity to seismic stations being utilized in this study that they may be used to compare with the $V_P$ and $V_S$ estimates resulting from the full parameter gridsearch.
+
+  Whereas the active source data provide high resolution values at specific sites, the Crust 2.0 model, a global dataset widely cited in academic literature, provides complete coverage across Canada with an attendant trade-off in resolution.
+The data includes estimates of $V_P$ and $V_S$ for $2^o$ longitude and latitude cells as well as information regarding sedimentary thickness and crustal depth.
+While Crust 2.0 source data is a superset of the aforementioned active source data, the crustal model is also composed of statistically aggregated data from geographically distant but geologically similar regions.
+It therefore forms a unique dataset well suited for comparison with the crustal estimates calculated in this study.
 
@@
 \subsection{Comparisons}
+
+  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod} provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.
+
 As regional studies exist which utilize similar methods and have corresponding parameter estimates for some of the stations used in this study it is possible to directly compare values to those previously published.
@@
 
-% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
+\begin{thebibliography}
+Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21,183-21,200.
 
-\begin{thebibliography}{}
+Bostock, M. G., M. R. Kumar (2010), Bias in seismic estimates of crustal properties, J. Geophys. Int., 182, 403-407.
+
+Christensen, N. I. (1996), Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156.
+
+Darbyshire, F.
+A., D.
+W.
+Eaton, A.
+W.
+Frederiksen, E.
+Leila (2006), New insights into the lithosphere beneath the Superior Province from Rayleigh wave dispersion and receiver function analysis, J.
+Geophys.
+Int., 169, 1043-1068.
+
+Durrheim, R. J., W. D. Mooney (1991), Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
+
+Eaton, D.
+W., S.
+Dineva, R.
+Mereu (2005), Crustal thickness and Vp/Vs variations in the Grenville orogen (Ontario, Canada) from analysis of teleseismic receiver functions, Tectonophysics, 420, 223-238.
+
+Efron, B., R. Tibshirani (1986), Bootstrap Methods for Standard Errors, Confidence Intervals, and Other Measures of Statistical Accuracy, Statistical Science, 1, 54-75.
+
+Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
+Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.
+
+Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002) LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+
+Thompson, D.
+A., I.
+D.
+Bastow, G.
+Helffrich, J.-M.
+Kendall, J.
+Wookey, D.
+B.
+Snyder, D.
+W.
+Eaton (2010), Precambrian crustal evolution: Seismic constraints from the Canadian Shield, Earth and Planetary Science Letters, 297, 655–666.
+
+Zhu, L., H. Kanamori (2000), Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys. Res., 105, 2969-2980.
+
+Zandt, G., C. J. Ammon (1995), Continental crust composition constrained by measurements of crustal Poisson's ratio, Nature, 374, 152-154.
 
 \end{thebibliography}
-
-%Reference citation examples:
-
-%...as shown by \textit{Kilby} [2008].
-%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
-%...has been shown [\textit{Kilby et al.}, 2008].
-%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
-%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
-
-%...as shown by \citet{jskilby}.
-%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
-%...has been shown \citep{jskilbye}.
-%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.
-%
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-
-%% ------------------------------------------------------------------------ %%
 
@@
 %% ------------------------------------------------------------------------ %%
-
-M. G. Bostock, M. R. Kumar, 2010. Bias in seismic estimates of crustal properties, Geophys. J. Int., 182, 403-407.
-N. I. Christensen, 1996. Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156,
-R. J. Durrheim, W. D. Mooney, 1991. Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
-W. D. Mooney. 2012. Personal communication. Compiled GSC active source data for the Canada.
-D.A.
-Thompson, I.D.
-Bastow, G.
-Helffrich, J-M.
-Kendall, J.
-Wookey, D.B.
-Snyder, D.W.
-Eaton, 2010.
-Precambrian crustal evolution: Seismic constraints from the Canadian Shield, Earth and Planetary Science Letters, Volume 297, Issues 3–4, 1 September 2010, Pages 655–666
-L. Zhu, H. Kanamori, 2000. Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys Res, 105, 2969-2980.
-G. Zandt, C. J. Ammon, 1995. Continental crust composition constrained by measurements of crustal Poisson's ratio, Nature, 374, 152-154.
-
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-ABSTRACT
-   It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
-Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
-Durrheim, W.
-Mooney, 1991).
-Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
-We apply a new stacking method (Bostock and Kumar, 2010) to retrieve estimates of bulk crustal velocities (Vp, Vs) and thickness H from these data under the assumption of 1-D structure.
-Bootstrap error analysis is performed for each station dataset and results are compared to the results produced from alternate methods (Zhu and Kanamori, 2000).
-Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
-These results will be used to evaluate competing models of early crustal formation.
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
-% BREAK THE EQUATIONS INTO TWO OR MORE LINES
-% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
-% using double backslashes (\\).
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  LANDSCAPE/SIDEWAYS FIGURE AND TABLE EXAMPLES
-%
-%% ------------------------------------------------------------------------ %%
-%
-% For figures, add \usepackage{lscape} to the file and the landscape.sty style file
-% to the paper folder.
-%
-% \begin{figure*}[p]
-% \begin{landscapefigure*}
-% Illustration here.
-% \caption{caption here}
-% \end{landscapefigure*}
-% \end{figure*}
-%
-% For tables, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
-%
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaystable}
-% \caption{}
-% \begin{tabular}
-% Table layout here.
-% \end{tabular}
-% \end{sidewaystable}
-%
-%

### 2013-06-28 paper/draftBP.tex
paper

@@
 \usepackage[final]{graphicx}
+\usepackage{multirow}
 \usepackage{color}
@@
   Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod} provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.
+Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06.
+For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).
 
-As regional studies exist which utilize similar methods and have corresponding parameter estimates for some of the stations used in this study it is possible to directly compare values to those previously published.
-A comprehensive study in the Hudson Bay region of the Canadian Shield (Thomson et.
-al., 2010) uses an approach similar to the Zhu and Kanamori method published estimates for $H$ and $R=\frac{V_P}{V_S}$ for 35 stations.
-5 stations with $R$ estimates above 1.8 are removed due to the uncertainty surrounding these high values.
-There is strong correlation of 0.95 between crustal thickness values for both datasets (Figure~\ref{fig:thompsonCompH}).
-The velocity ratio data has a lower correlation of 0.5 (Figure~\ref{fig:thompsonCompR}).
-[Compare this value to uncertainty / deviation in the data?
-].
+  A comprehensive study in the Hudson Bay region of the Canadian Shield [Thomson et.
+al., 2010] employs data from 35 stations, of which, 27 stations share an error acceptable for comparison in both datasets.
+There is strong correlation, 0.97, between crustal thickness values (Fig \ref{fig:thompsonCompH}).
+The velocity ratio data has a lower correlation of 0.49 (Fig \ref{fig:thompsonCompR}) and a mean difference of 0.02.
+As indicated in section \ref{section:VpVsMethod} there is less resolution in $R$ so the lower correlation for this parameter is not unexpected.
 
-Directly comparing $V_P$ estimates processed with the MB algorithm is not possible as this method has not been employed in publication.
-However, comparisons can be made to active source records for experiments within close proximity to a given seismic station (Figure~\ref{fig:activeComp}.
-The correlation between these datasets is 0.235.
-[This is low, explain here?].
-Another method of determining the reliability of estimates is to check the $\frac{V_P}{V_S}$ ratio taken from the Kanamori approach to the $\frac{V_P}{V_S}$ value computed from the MB approach.
-These values, working with the same data, preprocessed using the same methods and tools should be equal.
-Selecting only those stations with an error of less than +/-0.05 $R$, 132 stations, we get a correlation of 0.57.
-Reducing the error tolerance and selecting stations with less than +/-0.01 $R$, leaving 31 of the cleanest stations, we get a correlation coefficient of 0.95.
+ Another study conducted in the Grenville Orogen [Eaton et al., 2006] utilizes data from 29 stations, with two stations being excluded due to low quality and high error.
+Again, there is strong correlation in $H$, 0.89, and lower correlation in $R$, 0.56 (Figs \ref{fig:eatonCompH}, \ref{fig:eatonCompR}).
+The mean difference in $R$ between the two datasets is 0.04.
+
+\begin{table}
+  \begin{tabular}{ l l l l l l }
+    \cline{3-6}
+    & & \multicolumn{2}{ c }{Correlation} & \multicolumn{2}{ c }{Mean Difference} \\
+    \hline
+    Study Authors & Num. of Stns & $H$ & $R$ & $H$(km) & $R$ \\
+    \hline
+    Thompson et. al. (2010)   & 0.97 & 0.49 & 3.36 & 0.024 \\
+    Eaton et. al. (2006)      & 0.90 & 0.61 & 2.65 & 0.04  \\
+    Darbyshire et. al. (2007) & 0.95 & 0.43 & 4.16 & 0.39  \\
+    \hline
+  \end{tabular}
+  \caption{Comparison of $R$ and $H$ estimates with three published studies}
+\label{table:comparison}
+
+\end{table}
 
@@
     \includegraphics[width=\textwidth]{thompsonComparisonH}
-  \caption{Comparison of crustal thickness $H$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.95}
+  \caption{Comparison of crustal thickness $H$ data from this study with data from Thompson et. al. (2010). Data shows a Pearson correlation of 0.95}
   \label{fig:thompsonCompH}
@@
   \centering
+    \includegraphics[width=\textwidth]{eatonComparisonH}
+  \caption{Comparison of crustal thickness $H$ data from this study with from Eaton et. al. (2010). Data shows a Pearson correlation of 0.95}
+  \label{fig:eatonCompH}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{eatonComparisonR}
+  \caption{Comparison of $V_P / V_R$ data from this study with data from Eaton et. al. (2010). Data shows a Pearson correlation of 0.5}
+  \label{fig:eatonCompR}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{darbyshireComparisonH}
+  \caption{Comparison of crustal thickness $H$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.95}
+  \label{fig:darbyshireCompH}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{darbyshireComparisonR}
+  \caption{Comparison of $V_P / V_R$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.5}
+  \label{fig:darbyshireCompR}
+\end{figure}
+
+\begin{figure}
+  \centering
   \includegraphics[width=\textwidth]{activeSourceComparison}
@@
 
-\subsection{Crustal Thickness}
+\subsection{Regional Bulk Crustal Parameters}
 (MAY NOT INCLUDE)
@@
   \item Appears to be secular variation in the North Eastern Churchill. As we move North and East we have thickening of continental crust.
-\end{itemize}
-
-\subsection{Vp/Vs}
-\begin{itemize}
-  \item reference Figure~\ref{fig:VpVsMap}
-  \item For perspective, the value for Quartzite is 1.65 and Basalt is 1.85.
-  \item North East of the Slave province, has cluster of low Vp/Vs values in a semi circle of high values.
-  \item Appears to be an shift throughout the Superior province from high Vp/Vs in the South to low Vp/Vs in the North East.
-  \item Compare to Crust 2.0 overlay (Figure~\ref{fig:VpVsMapCrust}).  Average value for Churchill province Vp/Vs is 1.73 while average value computed from Crust 2.0 is 1.77.
 \end{itemize}

### 2013-06-28 python/compareResults.py

@@ -48,7 +48,19 @@ title = 18 / ratio
 leg = 16 / ratio
 
 
-for i in range(3):
+
+def meanDifference(x, y):
+    assert (len(x) == len(y))
+    n = len(x)
+    md = 0
+    for i in range(n):
+        for j in range(n):
+            md += abs(x[i] - y[j])
+
+    return md/(n**2)
+
+
+for i in range(1,2):
 
     pro = os.environ['HOME'] + data[2*i + 1]
     pub = os.environ['HOME'] + data[2*i]
@@ -56,33 +68,48 @@ for i in range(3):
     arg = Args().addQuery("stdR", "lt", 0.06)
 
     p0 = Params(pro, ["H", "R", "stdR", "stdH"], arg)
-    p2 = Params(pub, ["H", "R", "stdR", "stdH"])
-    arg = Args().addQuery("hk::stdR", "lt", 0.06)
+    p2 = Params(pub, ["H", "R", "stdR", "stdH"], arg)
+
     p1 = Params(stnfile, ["hk::H","hk::R", "hk::stdR", "hk::stdH"])
 
-    p1.sync(p2)
-    # p2.sync(p0)
-    # p0.sync(p1)
-    #p0.sync(p2)
+    p0.sync(p2)
 
-    Rcorr = pearsonr(p1.hk_R, p2.R)
-    print "Correlation between", leglabel[i], "Vp/Vs datasets is {0:.2f}".format(Rcorr[0]), "using", len(p2.stns), "stations"
-    Hcorr = pearsonr(p1.hk_H, p2.H)
+    stns = p0.stns
+
+    R1 = p0.R
+    R2 = p2.R
+
+    H1 = p0.H
+    H2 = p2.H
+
+    R1std = p0.stdR
+    R2std = p2.stdR
+
+    H1std = p0.stdH
+    H2std = p2.stdH
+
+
+    Rcorr = pearsonr(R1, R2)
+
+    Hcorr = pearsonr(H1, H2)
     print "Correlation between", leglabel[i], "H datasets is {0:.2f}".format(Hcorr[0]), "using", len(p2.stns), "stations"
+    print "Correlation between", leglabel[i], "Vp/Vs datasets is {0:.2f}".format(Rcorr[0]), "using", len(p2.stns), "stations"
 
+    print "Mean Difference between", leglabel[i], "H datasets is {0:.2f}".format(meanDifference(H1,H2)), "using", len(p2.stns), "stations"
[…]

### 2013-06-28 python/compareResults.py
paper

@@ -60,7 +60,7 @@ def meanDifference(x, y):
     return md/(n**2)
 
 
-for i in range(1,2):
+for i in range(0,3):
 
     pro = os.environ['HOME'] + data[2*i + 1]
     pub = os.environ['HOME'] + data[2*i]
@@ -144,7 +144,7 @@ for i in range(1,2):
     plt.axis("tight")
 
 
-plt.show()
+#plt.show()
 
 
 

### 2013-06-28 paper/draftBP.tex
paper

@@
 
-  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod} provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.
+  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod}, ZK + semblance weighting, provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.
 Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06.
+This error value best partitions the data between those receiver function stacks with visually recognisable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve.
 For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).
-
-  A comprehensive study in the Hudson Bay region of the Canadian Shield [Thomson et.
-al., 2010] employs data from 35 stations, of which, 27 stations share an error acceptable for comparison in both datasets.
-There is strong correlation, 0.97, between crustal thickness values (Fig \ref{fig:thompsonCompH}).
-The velocity ratio data has a lower correlation of 0.49 (Fig \ref{fig:thompsonCompR}) and a mean difference of 0.02.
-As indicated in section \ref{section:VpVsMethod} there is less resolution in $R$ so the lower correlation for this parameter is not unexpected.
-
- Another study conducted in the Grenville Orogen [Eaton et al., 2006] utilizes data from 29 stations, with two stations being excluded due to low quality and high error.
-Again, there is strong correlation in $H$, 0.89, and lower correlation in $R$, 0.56 (Figs \ref{fig:eatonCompH}, \ref{fig:eatonCompR}).
-The mean difference in $R$ between the two datasets is 0.04.
 
@@
     \hline
-    Thompson et. al. (2010)   & 0.97 & 0.49 & 3.36 & 0.024 \\
-    Eaton et. al. (2006)      & 0.90 & 0.61 & 2.65 & 0.04  \\
-    Darbyshire et. al. (2007) & 0.95 & 0.43 & 4.16 & 0.39  \\
+    Thompson et. al. (2010)   & 27 & 0.97 & 0.49 & 3.36 & 0.024 \\
+    Eaton et. al. (2006)      & 26 & 0.90 & 0.61 & 2.65 & 0.04  \\
+    Darbyshire et. al. (2007) & 10 & 0.95 & 0.43 & 4.16 & 0.39  \\
     \hline
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{thompsonComparisonH}
-  \caption{Comparison of crustal thickness $H$ data from this study with data from Thompson et. al. (2010). Data shows a Pearson correlation of 0.95}
-  \label{fig:thompsonCompH}
-\end{figure}
+  All three studies provide $H$ estimates that correlate strongly with the computed crustal thickness parameters.
+[[maximum mean difference, students ttest ?
+what would be a good stat for actual dataset difference?
+]].
+The correlation between the seismic velocity ratio is lower, though due to the lower resolution in $R$ this result is not unexpected.
+Overall there is good correlation between the estimates from the three studies and the computed estimates.
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{thompsonComparisonR}
-  \caption{Comparison of $V_P / V_R$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.5}
-  \label{fig:thompsonCompR}
-\end{figure}
+[[Comparison between $V_P$ estimates Mooney Active Source estimates]]
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{eatonComparisonH}
-  \caption{Comparison of crustal thickness $H$ data from this study with from Eaton et. al. (2010). Data shows a Pearson correlation of 0.95}
-  \label{fig:eatonCompH}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{eatonComparisonR}
-  \caption{Comparison of $V_P / V_R$ data from this study with data from Eaton et. al. (2010). Data shows a Pearson correlation of 0.5}
-  \label{fig:eatonCompR}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{darbyshireComparisonH}
-  \caption{Comparison of crustal thickness $H$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.95}
-  \label{fig:darbyshireCompH}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{darbyshireComparisonR}
-  \caption{Comparison of $V_P / V_R$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.5}
-  \label{fig:darbyshireCompR}
-\end{figure}
+\subsection{Regional Bulk Crustal Parameters}
+[[ Are we using ZK + semblance, or Full Grid Search + semblance ]]
+  Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error.
 
@@
 \end{figure}
-
-\subsection{Regional Bulk Crustal Parameters}
-(MAY NOT INCLUDE)
-\begin{itemize}
-  \item Jump in crustal thickness over the fault as you go from the Slave to the Rae or Churchill Province.
-  \item Trend in Southern Ontario as we move north and up the St Lawrence Seaway.
-  \item Appears to be secular variation in the North Eastern Churchill. As we move North and East we have thickening of continental crust.
-\end{itemize}
 
@@
 
-\begin{thebibliography}
+\begin{bibliography}
 Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21,183-21,200.
@@
 
-\end{thebibliography}
+\end{bibliography}
 

### 2013-06-28 paper/draftBPextra.tex
paper

  A comprehensive study in the Hudson Bay region of the Canadian Shield [Thomson et. al., 2010] employs data from 35 stations, of which, 27 stations share an error acceptable for comparison in both datasets. There is strong correlation, 0.97, between crustal thickness values (Fig \ref{fig:thompsonCompH}). The velocity ratio data has a lower correlation of 0.49 (Fig \ref{fig:thompsonCompR}) and a mean difference of 0.02. As indicated in section \ref{section:VpVsMethod} there is less resolution in $R$ so the lower correlation for this parameter is not unexpected.

 Another study conducted in the Grenville Orogen [Eaton et al., 2006] utilizes data from 29 stations, with two stations being excluded due to low quality and high error. Again, there is strong correlation in $H$, 0.89, and lower correlation in $R$, 0.56 (Figs \ref{fig:eatonCompH}, \ref{fig:eatonCompR}). The mean difference in $R$ between the two datasets is 0.04.

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{thompsonComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with data from Thompson et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:thompsonCompH}
\end{figure}

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{thompsonComparisonR}
  \caption{Comparison of $V_P / V_R$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.5}
  \label{fig:thompsonCompR}
\end{figure}

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{eatonComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with from Eaton et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:eatonCompH}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{eatonComparisonR}
  \caption{Comparison of $V_P / V_R$ data from this study with data from Eaton et. al. (2010). Data shows a Pearson correlation of 0.5}
  \label{fig:eatonCompR}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{darbyshireComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:darbyshireCompH}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{darbyshireComparisonR}
  \caption{Comparison of $V_P / V_R$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.5}
  \label{fig:darbyshireCompR}
\end{figure}

### 2013-07-04 paper/draftBP.tex
paper

@@
 \subsection{Regional Bulk Crustal Parameters}
-[[ Are we using ZK + semblance, or Full Grid Search + semblance ]]
+[[ Move from ZK + semblance to Full Grid Search + semblance ]]
   Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error.
+To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region.
+Wieghts are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection.
+A Voronoi diagram is calculated for all projected station locations.
+The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is computed and used as the weighting value.
+
+\begin{table}
+  \begin{tabular}{ l l l l l l }
+    \cline{2-6}
+    & \multicolumn{2}{ c }{ZK + semblance} & \multicolumn{2}{ c }{Crust 2.0} & \multicolumn{1}{ c }{Active Source} \\
+    \hline
+    Region & $R$ & $H [km]$ & $R$ & $H [km]$ & $V_P [km/s]$ \\
+    \hline
+    Canada     & 37.0 & 1.75 & 38.2 & 1.77 & 6.33\\
+    Shield     & 39.2 & 1.74 & 38.8 & 1.77 & 6.42\\
+    Slave      & 38.2 & 1.74 & 37.0 & 1.76 & 6.44\\
+    Churchill  & 38.9 & 1.73 & 37.8 & 1.77 & 6.39\\
+    Superior   & 39.7 & 1.74 & 39.1 & 1.76 & 6.44\\
+    Grenville  & 41.7 & 1.78 & 40.5 & 1.78 & 6.48\\
+    \hline
+  \end{tabular}
+  \caption{Comparison of $R$ and $H$ estimates with three published studies}
+\label{table:regionParameters}
+
+\end{table}
+
+  Moving South from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness.
+The trend in increasing values is also seen in the weighted averaged active source $V_P$ data.
+Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeble.
+The trend in the seismic velocity data is not as apparent in the Crust 2.0 data.
 
@@
 
-\subsection{Slave Province}
-<Discussion on Slave Province>
-
 %% -----------------------------------------------%%
 \section{Conclusions}
-<Conclusions here>
+Conclusions here
 

### 2013-07-04 matlab/ProcessTraces.m

@@ -23,7 +23,7 @@ printinfo = 1; % On and off flag to print out processing results
     ConvertFilterTraces(dlist, pfile, sfile,...
     picktol, printinfo, splitAzimuth, cluster);
 %fclose('all'); % Close all open files from reading
-clear picktol printinfo dlist splitAzimuth cluster
+%clear picktol printinfo dlist splitAzimuth cluster
 %% 3) Bin by p value (build pIndex)
 %
 numbin = round((1/npb) * size(ptrace, 1));
@@ -48,7 +48,7 @@ clear printwdiag wdiag
 %% Setup parallel toolbox
 if ~matlabpool('size')
     workers = 4;
-    matlabpool('local', workers)
+    matlabpool('test1');%, workers)
 end
 %% 5)  Window with Taper and fourier transform signal.
 adj = 0.1; % This adjusts the Tukey window used.
@@ -61,10 +61,13 @@ clear adj
 % p and put them into bins, all need to be length n
 % Now fft windowed traces
 Rec = zeros(nbins,size(wft,2));
+tic
 parfor ii = 1:nbins
     [r,~,betax(ii)] = simdecf(wft(pIndex(:,ii),:), vft(pIndex(:,ii),:), -1); %#ok<PFBNS>
     Rec(ii,:) = real(ifft(r));
 end
+toc
+return
 % if discardBad flag set simdecf will return Nan arrays where it did not
 % find a minimum, the following strips NaNs out and strips out appropriate
 % Pslow indices.

### 2013-07-04 matlab/dataDriver.m

@@ -13,14 +13,14 @@ addpath([userdir,'/programming/matlab/toolbox_general'])
 addpath([userdir,'/programming/matlab/toolbox_signal'])
 addpath([userdir,'/programming/matlab/jsonlab'])
 %% Variables
-sacfolder = '/media/TerraS/CN';
-databasedir = '/media/TerraS/database';
+sacfolder = '/media/bpostlet/TerraS/CN';
+databasedir = '/media/bpostlet/TerraS/database';
 pfile = 'stack_P.sac';
 sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'PECO'; %
+station = 'KGNO'; %
 % KGNO
 %{
 

### 2013-07-07 paper/draftBP.tex
paper

@@
 Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06.
-This error value best partitions the data between those receiver function stacks with visually recognisable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve.
+This error value best partitions the data between those receiver function stacks with visually recognizable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve.
 For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).
@@
 To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region.
-Wieghts are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection.
+Weights are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection.
 A Voronoi diagram is calculated for all projected station locations.
@@
 The trend in increasing values is also seen in the weighted averaged active source $V_P$ data.
-Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeble.
+Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeable.
 The trend in the seismic velocity data is not as apparent in the Crust 2.0 data.
+
+\subsection{Conrad Discontinuity}
+Results and figures showing statistics regarding depth of crustal reflectors.
 
@@
   \includegraphics[width=\textwidth]{activeSourceComparison}
-  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longtitude.}
+  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longitude.}
   \label{fig:activeComp}
@@
 
-  The preliminary interrogation of the data set yields the observation that the bulk Canadian Shield has lower Vp/Vs than anticipated.
-Previous experiments show crustal averages of 1.77 (Christensen and Mooney, 1995) and 1.78 (Zandt and Ammon, 1995).
-This compares to computed values of 1.73, 1.74, 1.795 for the Churchill, Superior and Grenville Provinces Vp/Vs ratios respectively.
+  - Comparison Discussion
+    - Crustal Thickness
+      - Very good correlation in thickness estimates with all three datasets.
+    - Vp/Vs
+      - Lower correlation with the Vp/Vs values.
+        - Expect lower correlation with Vp/Vs since the travel time equations are
+          less sensitive to seismic velocities than they are to crustal thickness.
+          This means that there will be more uncertainty during the parameter
+          search for Vp/Vs and hence higher degree of relative error.
+        - ??? Utilizing different sources, applying different deconvolution and
+          filtering techniques will therefore have a greater impact on the
+          uncertainty of picking Vp/Vs than that of crustal thickness H. ???
 
-(MIGHT NOT INCLUDE)  The data bares out earlier results showing Proterozoic crust to have a higher seismic velocity ratio than Archean crust.
-This follows previous studies on crustal formation that have noted this trend in other regions (Durrheim and Mooney, 1991) .
-The increased crustal thickness of Proterozoic crust is clearly seen in the active source data while it not visible in data from processed seismic stations.
-
-  Further work on the Bostock-Kumar stacking approach is warranted from a look at the cleanest stations.
-Before it can be employed in large scale analysis additional denoising methods or alternative deconvolution techniques will need to be investigated to reduce the noise in the data.
+   - Explanation of results
+    - Canada H and Vp/Vs
+      - Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
+        2.0 average is 38km as well, however this is expected since Crust2.0 data for
+        Canada is based on the same data as that of Lith5.0
+      - ??? Make some comparison to global average ???
 

### 2013-07-07 matlab/functions/gridsearchKanC.cpp

@@ -2,6 +2,7 @@
 #include <math.h>
 #include <matrix.h>
 
+
 void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
 {
 
@@ -37,7 +38,6 @@ void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
   R = mxGetPr(plhs[0]);
   H = mxGetPr(plhs[1]);
   HRx = mxGetPr(plhs[2]); // Output actual value of gridsearch cell for setting contour level
-
 
   // Grid parameters.
   double adjtps = 0.5; // Weights chosen from Precambrian crustal evolution - D.A. Thompson

### 2013-07-18 paper/draftBP.tex
giant

@@
 \section{Data and methods}
-  Analysis of bulk Canadian continental crust requires estimates of crustal properties such as seismic-wave velocity and crustal thickness.
+Our analysis of bulk Canadian continental crust is based on estimates of crustal properties that can be accessed via seismic techniques, namely seismic wave velocities (or velocity ratio) and crustal thickness.
 This study draws on three sources for these crustal properties.
-The first and primary source for these estimates are data computed by processing raw seismograms into receiver functions and then inverting these RF's for parameter estimates.
-The other two datasets, the statistically compiled Crust 2.0 dataset and a compilation of pre-processed active source data, are used to qualify and compare with the primary dataset.
+The first and primary source are teleseismic receiver functions (hereafter RFs).
+The remaining two datasets are Crust 2.0, a widely published statistical average of similar regions with a global two degree resolution and a compilation of pre-processed active source data (Mooney, 2012).
+Both Crust 2.0 and the active source compilation are used to qualify and compare with the primary dataset.
 
 \subsection{Teleseismic Data Set}
-
-  The primary data utilized in this study are computed from teleseismic P-wave seismograms representing discrete seismic events.
-These events are comprised of seismic records representing more than 700 earthquake sources occurring between the years 2000 and 2012 at 343 broadband seismic stations across Canada.
+The primary data utilized in this study are computed from teleseismic P-wave seismograms originating from more than 700 earthquake sources.
+These earthquakes occurred between the years 2000 and 2012 at and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
 Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
-Events are filtered by the epicentral distance from source to receiver with only the events within a 30 to 100 degree window being included.
-Seismic events are further filtered by hand picking those with reasonable signal to noise ratio and of sufficient impulsiveness that arrival times can be accurately measured.
+Seismograms are included for analysis if they within a 30 to 100$^\circ$ epicentral distance window.
+Quality control is performed by admitting only those seismograms with high enough signal to noise ratios to permit visual observation of the dominant P-wave energy and with sufficient impulsive first arrivals that allow arrival times to be accurately measured.
 After selection and filtering more than 80,000 events are available for further processing.
@@
 The first stage of processing requires the transformation of teleseismic data into receiver functions.
-In its generic form this transformation involves deconvolving an approximation of the earthquake source from channels representing ground motion.
-The resulting waveform contains discrete pulses corresponding to the arrivals of S-wave energy scattered from subsurface discontinuities.
-Resolving these peaks is essential to the success of the following inversion, therefore it is advantageous to rotate the seismogram components to separate the P and S wave energy.
-This is accomplished by first rotating the N and E coordinates into radial and transverse dimensions and then performing a wave field decomposition on the radial and vertical channels [Bostock, 1998].
-The direct arrival of the signal on the resulting P wave component is used as an approximation to the source function as P waves are simpler and have a response closer to that of a delta function.
-This windowed source coda is deconvolved from the S wave component computed from the wave field decomposition.
+In general terms, this transformation involves deconvolving an approximation of the earthquake source from horizontal recordings of ground motion [Langston, 1979].
+The resulting waveforms contain discrete pulses corresponding to the S-wave arrivals scattered from subsurface discontinuities including the base of the crust, or Moho.
+Confident identification of these peaks to the estimate of crustal parameters, and improved results can be obtained by transforming the seismogram channels into P and S contributions based on a 1-D model of propagation.
+This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels [Bostock, 1998].
+The direct arrival of the signal on the resulting P wave component is used as an approximation to the source function as P-wave impulse response for teleseismic P approximates a delta function.
+This windowed source estimate is deconvolved from the S wave component computed from the wave field decomposition.
 
-  An $L_2$ frequency domain deconvolution approach is used which has the advantages of computational efficiency as well as not requiring any assumptions about the noise in the data.
-This method performs a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $r(t)$.
+We employ $L_2$, multichannel, frequency domain approach to perform the deconvolution which has the advantage of computational efficiency and does not require a priori assumptions about the noise in the data.
+More specifically, we employ a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $r(t)$.
+
 % Deconvolution equations
@@
 
-  A well tested and widely published method for extracting the ratio, $R=\frac{V_P}{V_S}$, where $V_P$ is P-wave velocity and $V_S$ is S-wave velocity and crustal thickness, or depth to Moho, $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
-This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$ $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
-Note that $PsPs$ and $PpSs$ have an equal number of $P$ and $S$ phases, and the energy for these two phases arrive simultaneously and are indistinguishable.
+A well tested and widely published method for extracting the $\frac{V_P}{V_S}$ ratio $R$ (where $V_P$ is P wave velocity and $V_S$ is S-wave velocity) and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
+This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
+Note that $PsPs$ and $PpSs$ are kinematic analogues, such that the energy for these two phases arrive simultaneously for a 1-D model {\it COMMENT: note that the first phase is 2nd order whereas the 2nd is first
+order in scattering strength, which is hwy PsPs is not always considered}.
 For a range of slowness values, $p$, the differential arrival times, $t(p)$, trace moveout curves for each phase arrival given by
+ve
 
@@
 
-  The travel time equations demand an apriori assumption on crustal P-wave velocity, $V_P$, and will trade-off to some degree with crustal thickness $H$.
-For this study, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^o$ containing cell.
-In the specific cases where data is being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
+The travel time equations, as employed by ZK, assume a value for crustal P-wave velocity, $V_P$, that
+in practice will trade-off to some degree with crustal thickness $H$.
+In our implementation of the ZK approac, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^\circ$ containing cell.
+In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
 
@@
 
-  For a range of candidate models of $R$ and $H$ the RF's are stacked along trial moveout curves.
-During the stacking procedure each phase is assigned a weight to account for a general trend in the quality of the phases, with direct arrival usually carrying the best signal followed by $PpPs$ and $PpSs$.
-The weights chosen during the stacking are $w1 = 0.5$, $w2 = 0.3$ , $w3 = -0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases respectively.
+RFs are stacked along trial moveout curves for a range of candidate models of $R$ and $H$ to generate a fitness function $s(H,R)$.
+Within the stacking procedure, each phase is assigned a weight to account for a general trend in the quality of the phases, with direct conversion $Ps$ usually characterized by the highest quality signal followed by $PpPs$ and $PpSs$.
+We chose weights of $w1 = 0.5$, $w2 = 0.3$ , $w3 = -0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases, respectively.
 A negative weight for the combined $PpSs$ and $PsPs$ phases is required as the polarity of the signal is reversed.
@@
 
-and $S_j$ is the semblance weight for the $j^{th}$ phase, time $t_j$ is calculated from the corrisponding travel time function for a given $H$, $R$ as a function of slowness $p_i$ and $N$ is the number of receiver functions.
-Multiplying by the semblance weighting sharpens the stacked image, $s(H,R)$, and exhibits better resolution when selecting between stacked models.
+is the semblance weight for the $j^{th}$ phase, time $t_j$ is calculated from the corresponding travel time function for a given $H$, $R$ pair as a function of slowness $p_i$ and $N$ is the total number of receiver functions.
+Multiplying by the semblance weighting sharpens the stacked image, $s(H,R)$, and results in better resolution when discriminating between different models.
 The function $s(H,R)$ can be thought of as a transformation that maps the $Ps$, $PpPs$ and $PpSs+PsPs$ pulses of the receiver function, $r(t)$, into $R$ and $H$ space as positive bands, via Eq.
@@
 
-  Owing to increasing computational availability it is now reasonably efficient to extend the ZK approach and perform the stack along a range of all three seismic parameters, $V_P$, $R$ (therefore by definition $V_S$) and $H$.
-This has the advantage of not requiring an assumption on $V_P$ and should produce more accurate estimates, especially above crustal units with anomalous $V_P$.
-The inherent trade-off between $V_P$ and $R$ is limited, offering reasonable resolution between the two parameters.
-However, a significant trade-off between $V_P$ and $H$, which is a source of significant correlated error, requires that recovery of $V_P$ should be attempted only in cases where high frequency and high signal-to-noise ratio signals are present in the receiver function data.
-[[ Could go into more depth here, show the kinematic curves and poor trade-off between $V_P$ and $R$, Figure showing correlation between $V_P$ and $H$]]
-
-  These trade-offs, along with other sources of error such as low signal-to-noise signals resulting from complex geology, are quantified by bootstrap resampling [Efron and Tibshirani, 1986].
-Estimates are reprocessed using a random selection of receiver functions, with replacement, 1024 times.
-An estimate for the error is computed by taking the standard deviation from these trial runs.
+In addition to employing the ZK approach, we also investigate the feasibility of separately estimating $V_P$ and $H$ through a full 3 parameter grid search involving these parameters and $R$.
+The trade-off between $V_P$ and $H$ (and other parameter combinations) and the effects of various errors, are quantified for both the full grid search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
+Bootstrap standard deviations are produced using 1024 random selections of receiver functions, with replacement, per station.
 
 \subsection{Additional data}
-   Accompanying the processed estimates are data from controlled source experiments collected and compiled from GSC (Geological Survey of Canada) references by Walter Mooney (personal communication, 2012).
-This data provides $V_P$ and a few $V_S$ estimates for hundreds of experiment locations across Canada, many along traversals across geological provincial boundaries, faults and discontinuities.
-Some of these data are located within close enough proximity to seismic stations being utilized in this study that they may be used to compare with the $V_P$ and $V_S$ estimates resulting from the full parameter gridsearch.
 
-  Whereas the active source data provide high resolution values at specific sites, the Crust 2.0 model, a global dataset widely cited in academic literature, provides complete coverage across Canada with an attendant trade-off in resolution.
-The data includes estimates of $V_P$ and $V_S$ for $2^o$ longitude and latitude cells as well as information regarding sedimentary thickness and crustal depth.
-While Crust 2.0 source data is a superset of the aforementioned active source data, the crustal model is also composed of statistically aggregated data from geographically distant but geologically similar regions.
-It therefore forms a unique dataset well suited for comparison with the crustal estimates calculated in this study.
+We supplement receiver function estimates of crustal properties with data from controlled source experiments collected and compiled from GSC (Geological Survey of Canada) references by Walter Mooney (personal communication, 2012).
+The active source data provide $V_P$ and a few $V_S$ estimates for hundreds of locations across Canada, many along transects that cross geological provincial boundaries, faults and discontinuities.
+Some of these data are located within reasonable ($<$100km) proximity to seismic stations used in this study and can be used to compare with the $V_P$ and $V_S$ estimates resulting from the RF full gridsearch.
+
+The Crust 2.0 model [Bassin, 2000] employs many of these same active source data to provide regular, complete coverage across all of Canada.
+The model includes estimates of $V_P$ and $V_S$ information with depth at
+$2^\circ$ resolution in latitude and longitude.
+In addition to employing local active source data, Crust 2.0 is also constrained by statistically aggregated data from geographically distant but geologically similar regions.
+It therefore affords a somewhat independent dataset suited for comparison with the crustal estimates calculated in this study.
 
@@
 
-[[Comparison between $V_P$ estimates Mooney Active Source estimates]]
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{PCA}
+  \caption{Trade-off between $H$ and $V_P$ captures by principal component analysis (PCA) of the full parameter search $V_P$ and $H$ estimates.
+The principal vector captures 89\% of the variance in the dataset and the correlation between $V_P$ and $H$ is 0.79.
+Independent random parameters will have zero correlation and principal components of roughly equal variance.}
+  \label{fig:PCA}
+\end{figure}
+
+\subsection{Full Parameter Search Comparisons}
+The primary limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
+This can be resolved by computing the stacking function Eq.
+\ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
+Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time for this method can be measured in seconds.
+
+Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of only 0.51.
+Since $t_{Ps}$ represents the differential travel time between the S and P wave, the dependence of $H$ on $V_P$ is weak.
+Error in $V_P$ will trade-off with $H$ (Figure \ref{fig:PCA} and can be removed by dividing out $V_P$ from $H$ in both datasets.
+The $H/V_P$ ratio between the two datasets shows a high correlation of 0.96 as expected.
+
+If the error in $V_P$ is low the trade-off between $V_P$ and $H$ will be kept to a minimum.
+However, $V_P$ is difficult to resolve and even for clean stations the trade-off is quite large (Figure [slice through Vp space from 3D stack]).
+Taking the best three stations
+
+\begin{table}
+  \begin{tabular}{ l l }
+    \hline
+    Parameter & correlation coeff \\
+    \hline
+    $R$ (km) &  0.76 \\
+    $H$      &  0.51 \\
+    $H/V_P$  &  0.96 \\
+    \hline
+  \end{tabular}
+  \caption{Correlation between parameters computed from ZK + semblance and full parameter search methods}
+\label{table:ZKvsFG}
+
+\end{table}
 
 \subsection{Regional Bulk Crustal Parameters}
-[[ Move from ZK + semblance to Full Grid Search + semblance ]]
 Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error.
@@
 Results and figures showing statistics regarding depth of crustal reflectors.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{activeSourceComparison}
-  \caption{Comparison of $V_P$ estimates from the MB algorithm with active source $V_P$ recordings. Stations selected if active source experiment locations within 1 degree of latitude / longitude.}
-  \label{fig:activeComp}
-\end{figure}
 
@@
 \begin{bibliography}
+Bassin, C., G. Laske, G. Masters (2000), The Current Limits of Resolution for Surface Wave Tomography in North America, EOS Trans AGU, 81.
+
 Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21,183-21,200.
@@
 
+Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
+
 Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002) LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.

### 2013-07-18 matlab/functions/gridsearch3DC_stack.cpp

@@ -9,7 +9,7 @@ void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
   mxArray *recIN, *pslowIN;
   const mwSize *dims;
   double *rec, *pslow, *Tps, *R, *H, *V, *HRx;
-  double dt, *stack3d;
+  double dt, *h, *v, *r, *stack3d;
   int nrecs, i, ir , ih, iv;
   int N, lim;
 
@@ -30,7 +30,11 @@ void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
   plhs[1] = mxCreateDoubleMatrix(1, 1, mxREAL);
   plhs[2] = mxCreateDoubleMatrix(1, 1, mxREAL);
   plhs[3] = mxCreateDoubleMatrix(1, 1, mxREAL);
-
+  plhs[4] = mxCreateDoubleMatrix(lim, 1, mxREAL);
+  plhs[5] = mxCreateDoubleMatrix(lim, 1, mxREAL);
+  plhs[6] = mxCreateDoubleMatrix(lim, 1, mxREAL);
+  plhs[7] = mxCreateDoubleMatrix(lim*lim*lim, 1, mxREAL);
+  
   //associate pointers
   rec = mxGetPr(recIN);
   pslow = mxGetPr(pslowIN);
@@ -39,7 +43,11 @@ void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
   R = mxGetPr(plhs[1]);
   H = mxGetPr(plhs[2]);
   HRx = mxGetPr(plhs[3]); // Output actual value of gridsearch cell for setting contour level
-
+  v = mxGetPr(plhs[4]);
+  r = mxGetPr(plhs[5]);
+  h = mxGetPr(plhs[6]);
+  stack3d = mxGetPr(plhs[7]);
+          
 
   // Grid parameters.
   double adjtps = 0.5; // Weights chosen from Precambrian crustal evolution - D.A. Thompson
@@ -51,21 +59,21 @@ void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
   double r1 = 1.65;
   double r2 = 1.95;
   double dr = (r2 - r1) / ((double) nr - 1.);
-  double r[nr];
+  //double r[nr];
 
   // H limits
   int nh = lim;
   double h1 = 25.;
   double h2 = 50.;
[…]

### 2013-07-18 matlab/mergePlotFullGridSearch.m

@@ -3,8 +3,8 @@ close all
 
 addpath functions
 
-database = '/media/TerraS/database';
-dbfullgrid = '/media/TerraS/dbfullgrid';
+database = '/media/bpostlet/TerraS/database';
+dbfullgrid = '/media/bpostlet/TerraS/dbfullgrid';
 
 flist=dir([dbfullgrid,'/*.mat']); 
 

### 2013-07-18 matlab/process3D.m

@@ -31,10 +31,10 @@ for ii = ((NUM - 1) * ns + 1) : ns * NUM
     
     load(fullfile(databasedir, station));
         
-    [ v, r, h, ~] = gridsearch3DC(db.rec', db.dt, db.pslow, 150);
-    [Vp, R, H] = bootstrap3D(db.rec, db.dt, db.pslow, 150, 1);    
+    [ v, r, h, smaxss] = gridsearch3DC(db.rec', db.dt, db.pslow, 150);
+    [Vp, R, H, SMax] = bootstrap3D(db.rec, db.dt, db.pslow, 150, 1);    
  
-    save(['data/',station,'.mat'], 'v', 'r', 'h', 'Vp', 'R', 'H')
+    save(['data/',station,'.mat'], 'v', 'r', 'h', 'Vp', 'R', 'H', 'SMax')
 
 end
 

### 2013-07-18 python/vpPlots.py

@@ -15,27 +15,41 @@ from scipy.stats import pearsonr, spearmanr
 stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
 csfile = os.environ['HOME'] + '/thesis/data/csStations.json'
 
+
 def princomp(A,numpc=0):
      # computing eigenvalues and eigenvectors of covariance matrix
-    M = (A - np.mean(A.T,axis=1)).T # subtract the mean (along columns)
-    [latent,coeff] = np.linalg.eig( np.cov(M))
-    p = np.size(coeff,axis=1)
-    idx = np.argsort(latent) # sorting the eigenvalues
+    C = np.dot(A, A.T) / A.shape[1]
+    [eigValue,eigVect] = np.linalg.eig( C )
+    p = np.size(eigVect,axis=1)
+    idx = np.argsort(eigValue) # sorting the eigenvalues
     idx = idx[::-1]       # in ascending order
     # sorting eigenvectors according to the sorted eigenvalues
-    coeff = coeff[:,idx]
-    latent = latent[idx] # sorting eigenvalues
+    eigVect = eigVect[:,idx]
+    eigValue = eigValue[idx] # sorting eigenvalues
     if numpc < p or numpc >= 0:
-        coeff = coeff[:,range(numpc)] # cutting some PCs
-        score = np.dot(coeff.T, M) # projection of the data in the new space
-        return coeff, score, latent
+        eigVect = eigVect[:,range(numpc)] # cutting some PCs
+        score = np.dot(eigVect, A) # projection of the data in the new space
+        return eigVect, score, eigValue, C
+
+def getPCAvect(pc, xscale):
+    yb = pc[1] #+ am[1]
+    xb = pc[0] #+ am[0]
+    slope = -pc[1] / pc[0]
+    ypt = np.array([slope * xscale[0], slope * xscale[1]]) #+ 1.2
+    return xscale, ypt
+
+
+# def getPCAvect(pc, xscale):
+#     yb = pc[1] #+ am[1]
+#     xb = pc[0] #+ am[0]
+#     slope = -pc[1] / pc[0]
+#     b = yb - slope * xb
+#     ypt = np.array([slope * xscale[0] + b, slope * xscale[1] + b]) #+ 1.2
+#     return xscale, ypt
+
 
 if __name__  == "__main__":
 
-    maxerr = 0.06
-    arg = Args().addQuery("hk::stdR", "lt", str(maxerr))
[…]

### 2013-07-18 matlab/distroBoot.m
boot distro probability mixture

@@ -1,0 +1,52 @@
+% distibution analysis
+close all
+clear all
+addpath functions
+
+x = [trnd(20,1,50) trnd(4,1,100)+3 trnd(2,1,100)-2];
+%hist(x,-2.25:.5:7.25);
+%bar(bins,histc(x,bins)/(length(x)*.5),'histc');
+
+n = 3;
+breaks = [.25 .55 .85];
+%a = nModeDist(x, n, breaks);
+
+%%
+[pn, weightFnc] = mixtureWeights(n);
+
+%pdf_normmixture = @(x,p,mu1,mu2,sigma1,sigma2) ...
+%    weightFnc(p,1)*normpdf(x,mu1,sigma1) + ...
+%    weightFnc(p,2)*normpdf(x,mu2,sigma2);
+
+% pdf_normmixture = @pdf_Mixture;
+% 
+% pStart = 1/n;
+% muStart = quantile(x, breaks);
+% sigmaStart = sqrt(var(x) - .25*diff(muStart).^2);
+% start = [pStart muStart sigmaStart sigmaStart];
+% 
+% lb = [0 -Inf -Inf 0 0];
+% ub = [1 Inf Inf Inf Inf];
+% 
+% 
+% options = statset('MaxIter',500, 'MaxFunEvals',600);
+% paramEsts = mle(x, 'pdf',pdf_normmixture, 'start',start, ...
+%     'lower',lb, 'upper',ub, 'options',options);
+
+paramEsts = {nModeDist(x, n, breaks)};
+
+
+bins = -2.5:.5:7.5;
+h = bar(bins,histc(x,bins)/(length(x)*.5),'histc');
+set(h,'FaceColor',[.9 .9 .9]);
+xgrid = linspace(1.1*min(x),1.1*max(x),200);
+pdfgrid = pdf_Mixture(xgrid, paramEsts{:});
+hold on; plot(xgrid,pdfgrid,'-'); hold off
+xlabel('x'); ylabel('Probability Density');
+
+acov = mlecov(paramEsts, x, 'pdf', pdf_Mixture);
+se = sqrt(diag(acov));
+
+% n probs
+%%
+

### 2013-07-18 matlab/functions/mixtureWeights.m

@@ -1,0 +1,18 @@
+function [pn, func] = mixtureWeights(n)
+% Function to pick different combination of weights between n parameters
+
+% Change spacing from 11 to 21 for finer resolution
+a = linspace(0,1,11);
+args = {a};
+for ii = 2:n
+    args{end+1} = a;
+end
+
+c = combvec(args{:});
+w = sort(c(:, sum(c) == 1));
+pn = length(w);
+
+func = @(p, ind) w(ind,floor(p*pn));
+
+
+end

### 2013-07-18 matlab/functions/nModeDist.m

@@ -1,0 +1,22 @@
+function paramEsts = nModeDist(x, n, breaks)
+
+
+pStart = 1/n;
+muStart = quantile(x, breaks);
+d = diff(muStart);
+sigmaStart = sqrt(var(x) - 0.5*pStart*[d, min(d)].^2);
+start = [pStart muStart sigmaStart];
+
+assert(length(start) == 2*n + 1)
+
+lb = [0 -inf(1,n) zeros(1,n)];
+ub = [1 inf(1,2*n)];
+
+paramEsts = 0;
+options = statset('MaxIter',500, 'MaxFunEvals',600);
+
+
+
+paramEsts = mle(x, 'pdf',@pdf_Mixture, 'start',start, ...
+    'lower',lb, 'upper',ub, 'options',options);
+

### 2013-07-18 python/avgpipe.py

@@ -14,9 +14,16 @@ import re, sys
 
 if __name__== '__main__' :
 
+    numbers = re.findall(r'\d+', sys.stdin.read() )
+    numbers = map(int, numbers)
 
-    numbers = sys.stdin.read().split()
-    numbers = map(float, numbers)
-    print sum(numbers)/len(numbers)
-
-
+    sortn = [numbers.pop()]
+    while numbers:
+        hit = False
+        n = numbers.pop()
+        for i, ns in enumerate(sortn):
+            if n < ns:
+                sortn.insert(ind, n)
+                hit = True
+        if not hit:
+            sortn.append(n)

### 2013-07-18 python/conradPlots.py

@@ -1,0 +1,66 @@
+#!/usr/bin/python2
+#
+# python-mode indent C-c < or C-c >
+# python-mode comment/uncomment region M-;
+###########################################################################
+# IMPORTS
+###########################################################################
+import os, json
+import numpy as np
+import matplotlib.pyplot as plt
+from plotTools import Args, Params
+
+
+stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
+
+def extract(d, key):
+    a = []
+    for k in d.keys():
+        if key in d[k]:
+            a.extend(d[k][key])
+    return np.array(a)
+
+if __name__  == "__main__":
+
+
+    ## Figure Properties #######
+    figwidth = 12
+    figheight = figwidth / 1.618
+    ratio = 1.5
+    lw = 4 / ratio# line width
+    ms = 12 / ratio# marker size
+    caplen = 7 / ratio
+    capwid = 2 / ratio
+    elw = 2 / ratio
+    ticks = 16 / ratio
+    label = 16 / ratio
+    title = 18 / ratio
+    leg = 16 / ratio
+
+    #############################################################################
+    # FIG 1: Vp estimates against H
+    ##############################################################################
+
+    data = json.loads( open(stnfile).read() )
+    c = { stn:data[stn]["conrad"] for stn in data.keys() if "conrad" in data[stn] }
+    h = extract(c, "hdisc")
+    hmin = 10
+    hmax = 35
+    nbins = 30
+    h = h[(h >= hmin) & (h <= hmax)]
+
+    bins = np.linspace(hmin, hmax, nbins)
+    n, bins, patches = plt.hist(h, bins = bins, facecolor='green', alpha=0.75, label = "asdfads")
+    plt.axis('tight')
+
+    plt.show()
+    # fig = plt.figure( figsize = (figwidth, figheight) )
+    # plt.plot(M[0], M[1], 'ob', lw = lw, ms = ms, label = "normalized Vp vs H")
+    # plt.plot(x1, y1,'--r', lw = lw, label="Primary PCA")
+    # plt.plot(x2, y2,'--g', lw = lw, label="Secondary PCA")
+
+    # plt.legend(loc= 2, prop={'size': leg})
+    # plt.ylabel("normalized Vp", size = label)
+    # plt.xlabel("normalized H", size = label)
+    # plt.grid(True)
+    # plt.axis("equal")

### 2013-07-25 paper/draftBP.tex
boot distro probability mixture
conrad PCP, vpPlots

@@
 \subsection{Full Gridsearch Method}
+The primary limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
+This can be resolved by computing the stacking function Eq.
+\ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
+Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method scalable to hundreds of stations.
 
-In addition to employing the ZK approach, we also investigate the feasibility of separately estimating $V_P$ and $H$ through a full 3 parameter grid search involving these parameters and $R$.
-The trade-off between $V_P$ and $H$ (and other parameter combinations) and the effects of various errors, are quantified for both the full grid search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
-Bootstrap standard deviations are produced using 1024 random selections of receiver functions, with replacement, per station.
+[[ Maybe have a subsection for error? ]]
+ The trade-off between $V_P$ and $H$ (and other parameter combinations) and error resulting from data quality are quantified for both the full grid search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
+An estimate for the error is calculated in a bootstrap resampling approach by taking the standard deviation of 1024 parameter estimates each produced with randomly selected data (RFs), with replacement.
 
@@
 \subsection{Full Parameter Search Comparisons}
-The primary limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
-This can be resolved by computing the stacking function Eq.
-\ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
-Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time for this method can be measured in seconds.
-
 Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of only 0.51.
@@
 \subsection{Regional Bulk Crustal Parameters}
-Data from X seismic stations across Canada are analyzed, using only those stations with less than 0.06 standard error.
+Parameter estimates for stations assigned less than a 0.6 bootstrap error in the velocity ratio, $R$, are grouped regionally and analyzed.
 To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region.
 Weights are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection.
-A Voronoi diagram is calculated for all projected station locations.
+A Voronoi diagram is then calculated using all projected station locations.
 The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is computed and used as the weighting value.
@@
 
-Moving South from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness.
-The trend in increasing values is also seen in the weighted averaged active source $V_P$ data.
+Moving South and East from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness.
+The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
+Similarly the weighted average Moho depth increases from 38.9 km in the Churchill to 39.7 in the Superior and 41.7km in the Grenville.
+The Southeastern trend towards increasing values is also seen in the weighted averaged active source $V_P$ data.
 Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeable.
@@
 Results and figures showing statistics regarding depth of crustal reflectors.
+[[ Idea: Statistical test of distributions? How much confidence in a range of modes n = 1->5
+   Winning number of modes - partition the dataset for those stations with those modes
+   map spatially and see if it is a geographic density bias / trend ]]
 

### 2013-07-25 matlab/functions/plotSlice.m

@@ -1,0 +1,21 @@
+function plotSlice(station, x, y, xl, yl, s, xstd, ystd, sstd, X, Y, slice)
+
+
+figure()
+set(gca,'FontName','Helvetica','FontSize',16,...
+    'Clipping','off','layer','top');
+imagesc(X, Y, slice);
+axis xy
+axis square
+colorbar
+hold on
+plot(x, y, 'w+')
+plot(x, y, 'ko')
+contour(X, Y, slice, [s - sstd, s - sstd], 'k-')
+hold off
+xlab=xlabel(xl);
+ylab=ylabel(yl);
+set(xlab,'FontName','Helvetica','FontSize',16);
+set(ylab,'FontName','Helvetica','FontSize',16);
+title(sprintf('%s\n%s = %1.3f +/- %1.3f km/s\n%s = %1.3f +/- %1.3f',...
+    station, xl, x, xstd, yl, y, ystd));

### 2013-07-25 matlab/mergePlotFullGridSearch.m

@@ -4,12 +4,12 @@ close all
 addpath functions
 
 database = '/media/bpostlet/TerraS/database';
-dbfullgrid = '/media/bpostlet/TerraS/dbfullgrid';
+dbfullgrid = '/media/bpostlet/TerraS/data3D';
 
 flist=dir([dbfullgrid,'/*.mat']); 
 
-mergeFLAG = false;
-plotFLAG = true;
+mergeFLAG = true;
+plotFLAG = false;
 
 for ii=1:length(flist)
 
@@ -31,7 +31,7 @@ if mergeFLAG
     db.fg.stdVp = std(S.Vp);
     db.fg.stdR = std(S.R);
     db.fg.stdH = std(S.H);
-    
+    db.fg.stdS = std(S.SMax);
     save(dbfile,'db') % Save .mat file with bulk data
     fprintf('Saved data\n')
     

### 2013-07-25 python/vpPlots.py

@@ -70,7 +70,7 @@ if __name__  == "__main__":
     # FIG 1: Vp estimates against H
     ##############################################################################
 
-    if 0:
+    if 1:
         f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdVp", "hk::stdR"])
 
         corrfg = pearsonr(f.fg_Vp, f.fg_H)
@@ -96,8 +96,8 @@ if __name__  == "__main__":
 
         fig = plt.figure( figsize = (figwidth, figheight) )
         plt.plot(M[0], M[1], 'ob', lw = lw, ms = ms, label = "normalized Vp vs H")
-        plt.plot(x1, y1,'--r', lw = lw, label="Primary PCA")
-        plt.plot(x2, y2,'--g', lw = lw, label="Secondary PCA")
+        plt.plot(x2, y2,'--r', lw = lw, label="Primary PCA: variance = {:2.1f}%".format(var[0]*100))
+        plt.plot(x1, y1,'--g', lw = lw, label="Secondary PCA: variance = {:2.1f}%".format(var[1]*100))
 
         plt.legend(loc= 2, prop={'size': leg})
         plt.ylabel("normalized Vp", size = label)
@@ -109,7 +109,7 @@ if __name__  == "__main__":
     # FIG 2: Vp estimates and controlled source
     ##############################################################################
     if 1:
-        maxerr = 0.4#0.21
+        maxerr = 0.21#0.24#0.21
         arg = Args().addQuery("fg::stdVp", "lt", str(maxerr))
         f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdH", "fg::stdVp", "hk::stdR"], arg)
 
@@ -128,10 +128,10 @@ if __name__  == "__main__":
         ax1 = fig.add_subplot(111)
 
 
-        ax1.plot(t, f.fg_Vp, '-ob', lw = lw, ms = ms, label = "Full Gridsearch Vp estimate")
-        ax1.errorbar(t, f.fg_Vp, yerr=stdVp, xerr=None, fmt=None, ecolor = 'blue',
+        l1=ax1.plot(t, f.fg_Vp, '-ob', lw = lw, ms = ms, label = "Full Gridsearch Vp estimate")
+        l2=ax1.errorbar(t, f.fg_Vp, yerr=stdVp, xerr=None, fmt=None, ecolor = 'blue',
[…]

### 2013-07-26 paper/draftBP.tex
paper

@@
 The primary data utilized in this study are computed from teleseismic P-wave seismograms originating from more than 700 earthquake sources.
-These earthquakes occurred between the years 2000 and 2012 at and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
+These earthquakes occurred between the years 2000 and 2012 and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
 Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
@@
 \begin{equation}
-  r(t) = F^{-1} \left[ G(\omega) \right] = F{^-1}
+  r(t) = F^{-1} \left[ G(\omega) \right] = F^{-1}
  \left[ \frac {\sum_n^N S_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta} \right ],
@@
 
-The travel time equations, as employed by ZK, assume a value for crustal P-wave velocity, $V_P$, that
-in practice will trade-off to some degree with crustal thickness $H$.
-In our implementation of the ZK approac, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^\circ$ containing cell.
+The travel time equations, as employed by ZK, assume a value for crustal P-wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
+In our implementation of the ZK approach, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^\circ$ containing cell.
 In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth]{kinematicRvsH}
-  \caption{Kinematic curves traced out by the stacking function, $s(H,R)$, corrisponding to the mapping of fixed travel times $t_{Ps}$, $t_{PpPs}$ and $t_{PsPs}$ into $H$ and $R$ space.}
-  \label{fig:kinematicRvsH}
-\end{figure}
 
@@
 \ref{eq:stack}.
-Constructive interference where these bands intersect will produce a maximum in the stacking function (Fig \ref{fig:kinematicRvsH}).
+Constructive interference where these bands intersect will produce a maximum in the stacking function.
 The model, $R$ and $H$, which produce this maximum provides the best estimate for the bulk crustal parameters for a given seismic station.
-The elongate shape about the point of intersection, with the long axis tilted mostly in the $R$ coordinate, conveys the inherent trade-off between $R$ and $H$ and the lower resolution exhibited by the seismic velocity ratio $R$.
 
-\subsection{Full Gridsearch Method}
-The primary limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
+%% [[\subsection{Full Parameter Method}]]
+A limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
 This can be resolved by computing the stacking function Eq.
 \ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
-Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method scalable to hundreds of stations.
+Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method easily scalable to hundreds of stations.
+This approach, denoted FG, is analyzed to determine the viability of recovering $V_P$.
 
-[[ Maybe have a subsection for error? ]]
- The trade-off between $V_P$ and $H$ (and other parameter combinations) and error resulting from data quality are quantified for both the full grid search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
-An estimate for the error is calculated in a bootstrap resampling approach by taking the standard deviation of 1024 parameter estimates each produced with randomly selected data (RFs), with replacement.
+The trade-off between $V_P$ and $H$ (and other parameter combinations) and error resulting from data quality are quantified for both the full parameter search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
+An estimate for the error is calculated in a bootstrap resampling approach by taking the standard deviation of 1024 parameter estimates each produced with randomly selected RFs, with replacement.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{PCA}
-  \caption{Trade-off between $H$ and $V_P$ captures by principal component analysis (PCA) of the full parameter search $V_P$ and $H$ estimates.
-The principal vector captures 89\% of the variance in the dataset and the correlation between $V_P$ and $H$ is 0.79.
-Independent random parameters will have zero correlation and principal components of roughly equal variance.}
-  \label{fig:PCA}
-\end{figure}
-
-\subsection{Full Parameter Search Comparisons}
+\subsection{Full Parameter Search Analysis}
 Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of only 0.51.
 Since $t_{Ps}$ represents the differential travel time between the S and P wave, the dependence of $H$ on $V_P$ is weak.
-Error in $V_P$ will trade-off with $H$ (Figure \ref{fig:PCA} and can be removed by dividing out $V_P$ from $H$ in both datasets.
-The $H/V_P$ ratio between the two datasets shows a high correlation of 0.96 as expected.
+This means that for a given uncertainty in the three dimensional stack $s(H,R,V_P)$, the gradient in the stack-space will be lowest along the kinematic curve in the $H$ and $V_P$ axis (Figure \ref{fig:HVp_ULM}).
+This conveys the inherent trade-off in resolution between $H$ and $V_P$ and contributes to the low correlation between the two methods.
+Parity between the ZK and FG methods can be regained by dividing out $V_P$ from $H$.
+The trade-off between the other parameters is lower, owing to greater dependence between the parameters in the travel time equations (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
 
-If the error in $V_P$ is low the trade-off between $V_P$ and $H$ will be kept to a minimum.
-However, $V_P$ is difficult to resolve and even for clean stations the trade-off is quite large (Figure [slice through Vp space from 3D stack]).
-Taking the best three stations
+For all but the cleanest stations, the trade-off between $H$ and $V_P$ makes the recoverability of $V_P$ unfeasible using the FG approach.
+A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this challenge.
+The horizontally stacked RFs for both stations are ordered by increasing slowness and are shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
+Cross sections of the three dimensional stack are given for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
+The lower resolution conveyed in the $H$ and $V_P$ cross section for DORN translates into a bootstrap error of $V_P +/- 0.64 km/s$.
+The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corrisponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
 
@@
 \end{table}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{RF_ULM}
+  \caption{}
+  \label{fig:RF_ULM}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackHR_ULM}
+  \caption{}
+  \label{fig:HR_ULM}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackRVp_ULM}
+  \caption{}
+  \label{fig:RVp_ULM}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackHVp_ULM}
+  \caption{}
+  \label{fig:HVp_ULM}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{RF_DORN}
+  \caption{}
+  \label{fig:RF_DORN}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackHR_DORN}
+  \caption{}
+  \label{fig:HR_DORN}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackRVp_DORN}
+  \caption{}
+  \label{fig:RVp_DORN}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stackHVp_DORN}
+  \caption{}
+  \label{fig:HVp_DORN}
+\end{figure}
 

### 2013-07-26 matlab/runFullGridsearch.m

@@ -1,65 +1,0 @@
-% Run fullgrid search
-
-clear all;
-close all
-loadtools;
-
-
-addpath functions
-addpath ../sac
-addpath functions
-addpath sourceStack
-addpath([userdir,'/programming/matlab/toolbox_general'])
-addpath([userdir,'/programming/matlab/toolbox_signal'])
-addpath([userdir,'/programming/matlab/jsonlab'])
-%% Variables
-sacfolder = '/media/bpostlet/TerraS/CN';
-database = '/media/TerraS/database';
-databasedir = '/media/bpostlet/TerraS/database';
-
-station = 'ULM';
-
-dbfile = fullfile(databasedir, [station,'.mat'] );
-workingdir = fullfile(sacfolder,station);
-clear db dbold
-load(dbfile)
-
-%% Run FullGridsearch
-lim = 150;
-[V, R, H, HRx, v, r, h, stack3d] = gridsearch3DC_stack( db.rec', db.dt, db.pslow', lim );
-
-
-G = reshape(stack3d, lim, lim, lim); % G(h,r,v)
-
-%% Find optimum R slice
-
-tmp = abs(r - db.fg.rbest);
-[~, idx] = min(tmp);
-clear tmp
-
-Rslice = squeeze(G(:,idx,:));
-
-%% Plot section
-
-figure()
-set(gca,'FontName','Helvetica','FontSize',16,...
-    'Clipping','off','layer','top');
-imagesc(h, v, Rslice);
-axis xy
-axis square
-colorbar
-hold on
-plot(db.fg.hbest, db.fg.vbest, 'w+')
-plot(db.fg.hbest, db.fg.vbest, 'ko')
-contour(h, v, Rslice,...
-    [HRx - db.mb.stdsmax, HRx - db.mb.stdsmax], 'k-')
-hold off
-xlab=xlabel('R');
-ylab=ylabel('V_P [km/s]');
-set(xlab,'FontName','Helvetica','FontSize',16);
-set(ylab,'FontName','Helvetica','FontSize',16);
-title(sprintf('%s\nVp = %1.3f +/- %1.3f km/s\nR = %1.3f +/- %1.3f',...
-    db.station, db.mb.vbest, db.mb.stdVp, db.mb.rbest, db.mb.stdR));
-
-
-

### 2013-07-26 paper/draftBP.tex
conrad

@@
 \subsection{Full Parameter Search Analysis}
-Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of only 0.51.
-Since $t_{Ps}$ represents the differential travel time between the S and P wave, the dependence of $H$ on $V_P$ is weak.
+Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of 0.51.
+The low correlation between crustal thickness estimates can be explained by the weak dependence of $H$ on $V_P$ resulting from $t_{Ps}$.
 This means that for a given uncertainty in the three dimensional stack $s(H,R,V_P)$, the gradient in the stack-space will be lowest along the kinematic curve in the $H$ and $V_P$ axis (Figure \ref{fig:HVp_ULM}).
@@
 Parity between the ZK and FG methods can be regained by dividing out $V_P$ from $H$.
-The trade-off between the other parameters is lower, owing to greater dependence between the parameters in the travel time equations (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
+The trade-off between the other parameters is lower as there is greater dependence between the parameters in the travel time equations (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
 
@@
 
-\subsection{Conrad Discontinuity}
-Results and figures showing statistics regarding depth of crustal reflectors.
-[[ Idea: Statistical test of distributions? How much confidence in a range of modes n = 1->5
-   Winning number of modes - partition the dataset for those stations with those modes
-   map spatially and see if it is a geographic density bias / trend ]]
+\subsection{Crustal Discontinuities}
 

### 2013-07-26 matlab/conradPCP.m

@@ -24,7 +24,7 @@ for ii = 1 : length(s)
     
     dbfile = fullfile(databasedir, [station,'.mat'] );
     
-    if  numel(strfind(json.(station).status, 'processed'))
+    if  numel(strfind(json.(station).status, 'processed-ok'))
         if exist(dbfile, 'file')
             disp(station)
             load(dbfile)
@@ -110,7 +110,7 @@ end
 
 
 %%
-
+close all
 %X = zscore(X);
 figure()
 imagesc([1:length(s)], H, X')
@@ -120,18 +120,33 @@ E = diag(S*S');
  
 disp(E(1:10) ./ sum(E))
  
-%biplot()
- 
-[pc,score, eigs] = princomp(X);
-% biplot(pc(:,1:2),'Scores',score(:,1:2))
+ne = 6;
+
+scale  = 2;%1%5e2;
+Vs = [];
+for ii = 1:ne
+    Vs(:,ii) = S(ii,ii)^2 * V(:,ii);%/(max(abs(V(:,ii)))) ; %
+end
+
+% Flip first component cause its negative
+Vs(:,1) = Vs(:,1) * -1;
+
+shift = kron([1:ne]*scale, ones(length(X), 1));
+
 var = E./sum(E);
 
-figure()
-plot(var(1:10))
-title('Variance captured by first 10 Principal components')
+% figure()
+% plot(var(1:10))
+% title('Variance captured by first 10 Principal components')
+
+Xs = sum(S.^2'*V, 2);
+
 
 figure(34)
-plot(H, V(:,1:5))%, H, pc(:,2))
+plot(Xs, H,'k', 'LineWidth', 3)
+hold on
+%plot(H, sum(Vs'))
+plot(Vs + shift, H, 'LineWidth', 2)%, H, pc(:,2))
 legend(sprintf('PC 1 = %2.1f%%', var(1) *100),...
     sprintf('PC 2 = %2.1f%%', var(2) *100),...
     sprintf('PC 3 = %2.1f%%', var(3) *100),...
@@ -140,4 +155,6 @@ legend(sprintf('PC 1 = %2.1f%%', var(1) *100),...
     'Location', 'Best')
 xlabel('Principal Component Vector')
 ylabel('H [km]')
+set(gca,'YDir','reverse');
+hold off
  

### 2013-07-29 paper/draftBP.tex
conrad

@@
 The travel time equations, as employed by ZK, assume a value for crustal P-wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
-In our implementation of the ZK approach, each station is assigned a $V_P$ value corrisponding to the Crust 2.0 value for the $2^\circ$ containing cell.
+In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 2.0 value for the $2^\circ$ containing cell.
 In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
@@
 The lower resolution conveyed in the $H$ and $V_P$ cross section for DORN translates into a bootstrap error of $V_P +/- 0.64 km/s$.
-The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corrisponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
+The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{VpVsMap}
-  \caption{}
-  \label{fig:VpVsMap}
+  \includegraphics[width=\textwidth]{conradPCA}
+  \caption{The average of all station depth profiles (black) and the SVD singular vectors of the station depth profile matrix.
+Each vector represents a mode that captures some part of the total variance in the dataset.
+Vertical lines are the axis each profile and vector is normalized about.
+A grey horizontal window highlights a region of low energy.}
+  \label{fig:conradPCA}
 \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{VpVsMapCrust}
-  \caption{}
-  \label{fig:VpVsMapCrust}
-\end{figure}
+Another feature of interest is the average crustal structure across Canada.
+We perform a stack along $H$ using the best estimate for $R$ computed via the ZK + semblance method given above.
+This depth profile contains energy at depths corresponding to velocity discontinuities.
+Accumulating all station depth profiles into a matrix we then decompose the data via SVD into its major modes.
+Figure \ref{fig:conradPCA} contains the average depth profile for all stations and the first five major singular vectors or modes accounting for 78\% of the variance in the data.
+Additional modes have not been included as mode six contains energy at similar depths to modes two and three while modes seven and above appear noisy and represent variances in the dataset that do not account for significant variance or structure in the dataset.
+Each vector is normalized to its maximum for clarity and centred on an axis given as vertical grey lines.
+The first profile (black) is the average of all stations depth stacks.
+The second mode capturing 50\% of the variance in the dataset represent the mean and as expected appears similar to the average depth stack.
+The second and third modes are dominated by a shallow discontinuity and a deeper discontinuity associated with sedimentary layers and the Moho respectively.
+The fourth and fifth modes, accounting for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10km to 20km range.
+All modes display lower energy within a window roughly stretching from 20km depth to 30km.
+
+ %% \begin{figure}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{VpVsMap}
+%%   \caption{}
+%%   \label{fig:VpVsMap}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{VpVsMapCrust}
+%%   \caption{}
+%%   \label{fig:VpVsMapCrust}
+%% \end{figure}
 

### 2013-07-29 matlab/conradPCP.m

@@ -112,17 +112,17 @@ end
 %%
 close all
 %X = zscore(X);
-figure()
-imagesc([1:length(s)], H, X')
+%figure()
+%imagesc([1:length(s)], H, X')
 
 [U,S,V] = svd(X);
 E = diag(S*S');
  
 disp(E(1:10) ./ sum(E))
  
-ne = 6;
+ne = 5;
 
-scale  = 2;%1%5e2;
+scale  = 1.5;%1%5e2;
 Vs = [];
 for ii = 1:ne
     Vs(:,ii) = V(:,ii)/(max(abs(V(:,ii)))) ; %%S(ii,ii)^2 * V(:,ii);
@@ -145,16 +145,85 @@ figure(34)
 plot(Xs, H,'k', 'LineWidth', 3)
 hold on
 %plot(H, sum(Vs'))
-plot(Vs + shift, H, 'LineWidth', 2)%, H, pc(:,2))
-legend(sprintf('Average Stack', var(1) *100),...
-sprintf('PC 1 = %2.1f%%', var(1) *100),...
-    sprintf('PC 2 = %2.1f%%', var(2) *100),...
-    sprintf('PC 3 = %2.1f%%', var(3) *100),...
-    sprintf('PC 4 = %2.1f%%', var(4) *100),...
-    sprintf('PC 5 = %2.1f%%', var(5) *100),...
-    'Location', 'Best')
-xlabel('Principal Component Vector')
+plot(Vs + shift, H, 'LineWidth', 2)
+
+axis tight
 ylabel('H [km]')
 set(gca,'YDir','reverse');
+set(gca,'TickLength', [0 0]);
+
+hold on
+% Box 
+lim = xlim;
+H1 = 20;
+H2 = 30;
+p=patch([lim(1) lim(1) lim(2) lim(2)],[H1 H2 H2 H1],'k',...
+     'EdgeColor', 'none','FaceColor',[0.9, 0.9, 0.9],...
+    'FaceAlpha',0.5);
+
+% Vertical Base lines
+line([0,0],[H(1), H(end)], 'Color', [0.7,0.7,0.7])
+for ii = 1:ne
+  line([shift(1,ii),shift(1,ii)],[H(1), H(end)], 'Color', [0.7,0.7,0.7])
+end
+hold off
+
+
+pos = get(gca,'Position');
+set(gca,'Position',[pos(1), .25, pos(3) .65])
+
+
[…]

### 2013-08-02 paper/draftBP.tex
paper paper paper

--- 2013-07-26 paper/draftBP.tex
@@
 
-and $\omega_m$ is the $m^{th}$ frequency bin in the discrete Fourier transform.
-
-  All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
+and $\omega_m$ is the $m^{th}$ frequency bin in the discrete Fourier transform. All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
 
@@
 This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
-Note that $PsPs$ and $PpSs$ are kinematic analogues, such that the energy for these two phases arrive simultaneously for a 1-D model {\it COMMENT: note that the first phase is 2nd order whereas the 2nd is first
-order in scattering strength, which is hwy PsPs is not always considered}.
+Note that $PsPs$ and $PpSs$ are kinematic analogues, such that the energy for these two phases arrive simultaneously for a 1-D model.
 For a range of slowness values, $p$, the differential arrival times, $t(p)$, trace moveout curves for each phase arrival given by
@@
 
-Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod}, ZK + semblance weighting, provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.
-Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06.
-This error value best partitions the data between those receiver function stacks with visually recognizable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to resolve.
-For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).
+Three published studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
+The most recent, a study [Thompson et.
+al., 2010] focused on the Northern Canadian Shield, employs both the linear stacking approach used by ZK as well as a phase weighted method.
+A 2007 study of the Superior Province [Darbyshire et.
+al, 2007] and a study of the Grenville Orogen [Eaton et.
+al., 2006] utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
+Comparisons between crustal $R$ estimates for particular seismic stations are made for those values with a corresponding standard error of less than $\pm$0.06.
+This error threshold is chosen empirically as that which best partitions those receiver function stacks with visually recognizable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to distinguish.
+For each study under comparison, data are reprocessed using the $V_P$ chosen by the study authors.
+For the 2006 and 2007 studies the ZK + semblance method is used, for the Thompson et.
+al study the linear stacking ZK apprach is employed.
+Both correlation coefficient and mean difference are provided in (Table \ref{table:comparison}) as measures of similarity.
 
@@
     \hline
-    Study Authors & Num. of Stns & $H$ & $R$ & $H$(km) & $R$ \\
+    Study Authors &   Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
     \hline
-    Thompson et. al. (2010)   & 27 & 0.97 & 0.49 & 3.36 & 0.024 \\
-    Eaton et. al. (2006)      & 26 & 0.90 & 0.61 & 2.65 & 0.04  \\
-    Darbyshire et. al. (2007) & 10 & 0.95 & 0.43 & 4.16 & 0.39  \\
+    Thompson et. al. (2010)   & 29 & 0.97 & 0.70 & 0.49 & 0.012 \\
+    Darbyshire et. al. (2007) & 10 & 0.95 & 0.43 & 1.21 & 0.033 \\
+    Eaton et. al. (2006)      & 26 & 0.90 & 0.61 & 0.79 & 0.025 \\
     \hline
@@
 
-All three studies provide $H$ estimates that correlate strongly with the computed crustal thickness parameters.
-[[maximum mean difference, students ttest ?
-what would be a good stat for actual dataset difference?
-]].
-The correlation between the seismic velocity ratio is lower, though due to the lower resolution in $R$ this result is not unexpected.
-Overall there is good correlation between the estimates from the three studies and the computed estimates.
+All three studies provide $H$ estimates that correlate strongly with the crustal thickness parameters computed here.
+The data from the Northern Canadian Shield most closely correlates in both $H$ and $R$.
+The lower correlation in $R$ than $H$ may be attributed to a number of factors such as the choice of deconvolution algorithm and selection and scaling of data.
+The greater uncertainty in $R$ inherent in the ZK method will cause larger fluctuations in computed estimates for differences in procedure.
+The data from Thompson et.
+al.
+[2010] also align most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
+Results published by Eaton show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that are roughly equal to the averaged uncertainty in the data.
+Parameter estimates from the Superior Province [Darbyshire et.
+al., 2007] show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
+These differences are well over the averaged computed error, and can not be easily explained by deviation related to noise in the data.
 
 \subsection{Full Parameter Search Analysis}
-Comparing the ZK approach with the full parameter search (Table \ref{table:ZKvsFG}) we see that the velocity ratio $R$ has a correlation between the two datasets of 0.76 while $H$ has a correlation of 0.51.
-The low correlation between crustal thickness estimates can be explained by the weak dependence of $H$ on $V_P$ resulting from $t_{Ps}$.
-This means that for a given uncertainty in the three dimensional stack $s(H,R,V_P)$, the gradient in the stack-space will be lowest along the kinematic curve in the $H$ and $V_P$ axis (Figure \ref{fig:HVp_ULM}).
-This conveys the inherent trade-off in resolution between $H$ and $V_P$ and contributes to the low correlation between the two methods.
-Parity between the ZK and FG methods can be regained by dividing out $V_P$ from $H$.
+
+We now proceed to compare our results computed using the ZK approach with those from the full parameter search (Table \ref{table:ZKvsFG}).
+We note that the velocity ratio $R$ has a correlation between the two datasets of 0.76 whereas $H$ has a correlation of 0.51.
+The low correlation between crustal thickness estimates can be explained by the strong tradeoff between $H$ and $V_P$ through the quantity $(t_{Pps}-t_{Ps})/2=t_{Pss}/2-t_{Ps}=H\sqrt{1-p^2V^2_P)}/V_P$.
+As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:HVp_ULM}).
+This trade-off contributes to the low correlation in $H$ between the two methods.
+Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$.
 The trade-off between the other parameters is lower as there is greater dependence between the parameters in the travel time equations (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
 
-For all but the cleanest stations, the trade-off between $H$ and $V_P$ makes the recoverability of $V_P$ unfeasible using the FG approach.
-A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this challenge.
-The horizontally stacked RFs for both stations are ordered by increasing slowness and are shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
-Cross sections of the three dimensional stack are given for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
-The lower resolution conveyed in the $H$ and $V_P$ cross section for DORN translates into a bootstrap error of $V_P +/- 0.64 km/s$.
-The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corrisponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
+For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ infeasible using the FG approach.
+A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty.
+RFs are arranged vertically, ordered by increasing slowness, for both stations and are shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
+Cross sections of the three dimensional stacks $s(H,R,V_P)$  are displayed for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
+The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P +/- 0.64 km/s$.
+The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
+
+Active source experiments located near broadband seismic stations provide independant crustal thickness and P-wave velocity estimates to test against data derived from the full parameter search method.
+We select stations with low error in $V_P$ and active source experiment locations within a 100km radius.
+If several active source experiments are located within this radius estimates are averaged.
+Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and are presented in figure \ref{fig:activeVp}-\ref{fig:activeH}.
+Estimates computed from the FG approach are roughly 0.25 km/s higher than the active source data.
+This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting a bias.
+However, the limited number of data means this bias is poorly constrained.
+
+There is less correlation between the crustal thickness data, owing largely to a single station ARVN.
+The reason for this discrepency, roughly 10km, is unknown though both the ZK method and and a published study [Thompson et.
+al., 2010] provide estimates for ARVN of around 41km, much nearer the FG value of 42km.
+Despite the lower correlation the apparent bias towards higher values computed from teleseismic data is maintained.
+A positive bias in one parameter will tend to bias the other, owing to the $H/Vp$ relationship in $t_{ps}$.
+Taking the average active source estimates from the two best stations, ULM and WHY, a $V_P$ of 6.3 km/s, a $V_P/V_S$ of 1.75 and a 36km thick crust, a 0.25 km/s change in $V_P$ will result in a 1.3km change in $H$.
+The actual mean difference for the two stations, 1.9km, is close but slightly greater than this value.
+
+[[ SOURCES REGARDING BIAS ]] [[ WHAT DOES THIS MEAN, IS VP USEABLE FOR THE HIGHEST QUALITY STATIONS?]]
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{activeSourceCompVp}
+  \caption{P-wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
+  \label{fig:activeVp}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{activeSourceCompH}
+  \caption{Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
+  \label{fig:activeH}
+\end{figure}
+
+\subsection{Crustal Discontinuities}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{conradPCA}
+  \caption{The SVD singular vectors (principal components) of the station depth profile matrix.
+Each principal component represents a mode that captures some part of the total variance in the dataset.
+Vertical lines are the axis each profile and vector is normalized about.
+A grey horizontal window highlights a region of low energy.}
+  \label{fig:conradPCA}
+\end{figure}
+
+We first consider the average crustal structure across Canada.
+We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to XX km in increments of DX km)  using the estimates for $R$ computed via the ZK + semblance method.
+The resulting profiles contain energy at depths corresponding to velocity discontinuities.
+By assembling all depth profiles into a matrix, we may decompose the data via SVD into its principal components.
+Figure \ref{fig:conradPCA} contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles.
+Higher modes corresponding to diminishing contributions to total variance.
+Each vector is normalized to unit magnitude.
+The first mode (blue) is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset.
+It features a maximum near 38.7 km that may be taken as an average Moho depth.
+The second and third modes are dominated by a high amplitude shallow structures (e.g.
+sedimentary layers) within the top 10 km and Moho variability between 30 and 46 km .
+The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10km to 20km range.
+All 5 modes display markedly lower energy within a depth window roughly between 20km and 30km.
+
 \subsection{Regional Bulk Crustal Parameters}
-Parameter estimates for stations assigned less than a 0.6 bootstrap error in the velocity ratio, $R$, are grouped regionally and analyzed.
-To compensate for the uneven geographic distribution of seismic stations, weights are applied to all crustal parameter estimates before averaging over a given region.
-Weights are calculated by first projecting station locations onto the 2D plane using the Albers equal-area conic projection.
-A Voronoi diagram is then calculated using all projected station locations.
-The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is computed and used as the weighting value.
+
+We proceed to consider velocity ratio, $R$, estimates for stations with assigned bootstrap errors below $\pm 0.06$ in a geographic context.
+To compensate for the uneven distribution of seismic stations, weights are applied to both $R$ and $H$ estimates before averaging over a given region.
+Weights are calculated by projecting station locations onto a 2D plane using the Albers equal-area conic projection.
+A Voronoi diagram is then computed using all projected station locations.
+The ratio between the Voronoi cell surface encompassing each station and the total area of the convex hull bounding region is employed as that station's weight.
+
+Table \ref{table:regionParameters} displays the results of our regionalization in $R$ and $H$.
+The weighted average crustal thickness for all stations (Canada) is 37km, about two kilometres thinner than the average for the Canadian Shield.
+Crust 2.0 provides slightly thicker crustal average for Canada (38.2km) and a slightly thinner average for the Shield (38.8km).
+The Slave Province averages are similar to those for the Churchill province and has a deeper crust than Crust2.0 and a smaller seismic velocity ratio.
+Moving south and east from the Churchill Province through the Superior Province into the Grenville, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$.
+The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
+In a similar fashion, weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7km in the Grenville Province.
+The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data.
+The Crust 2.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable.
+The trend in the seismic velocity is not, however, as apparent.
 
@@
     \hline
-    Region & $R$ & $H [km]$ & $R$ & $H [km]$ & $V_P [km/s]$ \\
+    Region  & $H [km]$ &$R$ & $H [km]$&$R$ &$V_P [km/s]$ \\
     \hline
@@
 
-Moving South and East from the Churchill through the Superior and the Grenville there is an increase in the ZK + semblance computed seismic velocity ratio as well as crustal thickness.
-The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
-Similarly the weighted average Moho depth increases from 38.9 km in the Churchill to 39.7 in the Superior and 41.7km in the Grenville.
-The Southeastern trend towards increasing values is also seen in the weighted averaged active source $V_P$ data.
-Generally the Crust 2.0 values show a slightly thinner crust though the southward thickening trend is noticeable.
-The trend in the seismic velocity data is not as apparent in the Crust 2.0 data.
+ %% \begin{figure}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{VpVsMap}
+%%   \caption{}
+%%   \label{fig:VpVsMap}
+%% \end{figure}
 
-\subsection{Crustal Discontinuities}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{VpVsMap}
-  \caption{}
-  \label{fig:VpVsMap}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{VpVsMapCrust}
-  \caption{}
-  \label{fig:VpVsMapCrust}
-\end{figure}
+%% \begin{figure}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{VpVsMapCrust}
+%%   \caption{}
+%%   \label{fig:VpVsMapCrust}
+%% \end{figure}
 

### 2013-08-02 data/publishedData.py

@@ -1,23 +1,40 @@
 import json
 
-data = """ATKO 44 0.46 1.71 0.011
-EPLO 41 0.45 1.73 0.015
-HSMO 32 0.83 1.76 0.038
-KAPO 48 0.62 1.71 0.020
-KILO 35 0.45 1.87 0.012
-LDIO 41 0.49 1.82 0.021
-MALO 37 0.36 1.77 0.024
-MUMO 38 0.31 1.74 0.015
-NANO 42 0.67 1.75 0.022
-OTRO 41 0.81 1.79 0.029
-PKLO 37 0.35 1.81 0.012
-PNPO 44 0.54 1.69 0.017
-RLKO 42 0.32 1.77 0.011
-RSPO 34 0.40 1.80 0.015
-SILO 38 0.29 1.73 0.011
-SUNO 34 0.39 1.79 0.018
-VIMO 44 0.66 1.73 0.010
-WEMQ 39 0.38 1.72 0.014"""
+data = """AP3N 69.459 -84.462 POLARIS 60 40.1 0.8 1.711 0.017 40.6 0.5 1.714 0.013
+ARVN 61.098 -94.070 POLARIS 41 41.0 1.5 1.751 0.037 41.4 0.7 1.740 0.022
+B1NU 68.462 -71.588 CHASME 40 43.5 2.0 1.706 0.032 43.9 2.2 1.695 0.040
+B2NU 68.922 -73.197 CHASME 14 42.3 7.3 1.709 0.039 43.3 5.8 1.714 0.070
+BULN 66.397 -93.125 POLARIS 85 37.5 0.3 1.696 0.010 38.1 0.4 1.692 0.011
+CRLN 64.189 -83.348 NERC 31 35.4 0.6 1.740 0.018 36.0 0.4 1.732 0.018
+CTSN 62.852 -82.485 NERC 58 36.1 0.4 1.757 0.017 36.6 0.5 1.757 0.024
+DORN 64.230 -76.531 NERC 51 36.2 0.3 1.768 0.010 36.5 0.4 1.770 0.011
+FCC 58.762 -94.087 CNSN 160 40.4 0.8 1.764 0.016 40.6 1.8 1.761 0.034
+FRB 63.747 -68.547 CNSN 247 43.3 0.2 1.760 0.006 43.5 0.2 1.748 0.006
+GIFN 69.994 -81.638 POLARIS 138 41.2 0.1 1.726 0.005 41.5 0.2 1.723 0.007
[…]

### 2013-08-02 matlab/ZKderivs.m

@@ -1,0 +1,18 @@
+% ZK derivative calcs
+
+H = 36
+V = 6.26
+R = 1.75
+p = 0.06;
+t = H * ( sqrt( (R^2/V^2) - p^2) - sqrt( 1/V^2 - p^2 ) );
+
+dHdV = -t * ( sqrt( (R^2/V^2) - p^2) - sqrt( 1/V^2 - p^2 ) )^(-2) * ...
+    ( ( R^2/V^2 - p^2)^(-1/2) * (-R^2/V^3) - ...
+      (1/V^2 - p^2)^(-1/2) * (-1/V^3) )
+
+
+dv = 0.25;
+dh = dHdV * dv;
+
+
+fprintf('for a dv of %1.2f get a %2.2f change in dh\n', dv, dh);

### 2013-08-02 matlab/conradPCP.m

@@ -122,14 +122,12 @@ disp(E(1:10) ./ sum(E))
  
 ne = 5;
 
-scale  = 1.5;%1%5e2;
-Vs = [];
-for ii = 1:ne
-    Vs(:,ii) = V(:,ii)/(max(abs(V(:,ii)))) ; %%S(ii,ii)^2 * V(:,ii);
-end
 
 % Flip first component cause its negative
-Vs(:,1) = Vs(:,1) * -1;
+Vs = V(:, 1:ne);
+vmax = max(max(abs(Vs)));
+
+scale  = vmax + 0.2*vmax;
 
 shift = kron([1:ne]*scale, ones(length(X), 1));
 
@@ -142,8 +140,8 @@ Xs = sum(X);
 Xs = Xs/max(abs(Xs));
 
 figure(34)
-plot(Xs, H,'k', 'LineWidth', 3)
-hold on
+%plot(Xs, H,'k', 'LineWidth', 3)
+%hold on
 %plot(H, sum(Vs'))
 plot(Vs + shift, H, 'LineWidth', 2)
 
@@ -162,7 +160,7 @@ p=patch([lim(1) lim(1) lim(2) lim(2)],[H1 H2 H2 H1],'k',...
     'FaceAlpha',0.5);
 
 % Vertical Base lines
-line([0,0],[H(1), H(end)], 'Color', [0.7,0.7,0.7])
+%line([0,0],[H(1), H(end)], 'Color', [0.7,0.7,0.7])
 for ii = 1:ne
   line([shift(1,ii),shift(1,ii)],[H(1), H(end)], 'Color', [0.7,0.7,0.7])
 end
@@ -170,16 +168,15 @@ hold off
 
 
 pos = get(gca,'Position');
-set(gca,'Position',[pos(1), .25, pos(3) .65])
+set(gca,'Position',[pos(1), .15, pos(3) .75])
 
-
-tickLabels ={ 'Average Stack'};
+clear tickLabels
 for ii = 1:ne
-    tickLabels{ii+1} = sprintf('PC %1.0d = %2.1f%%', ii, var(ii) *100);
+    tickLabels{ii} = sprintf('PC %1.0d \n %2.1f%%', ii, var(ii) *100);
 end
 
 % Set xtick points
-Xt = [0, shift(1,:)];
+Xt = shift(1,:) + 0.2*vmax;
 set(gca,'XTick',Xt);
 
 ax = axis; % Current axis limits
[…]

### 2013-08-02 matlab/dataDriver.m

@@ -19,9 +19,9 @@ pfile = 'stack_P.sac';
 sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
-method = 'kanamori';
-station = 'KGNO'; %
-% KGNO
+method = 'fullgrid';
+station = 'ARVN'; %
+
 %{
 
 %}
@@ -84,13 +84,25 @@ if strcmp(method, 'kanamori')
     end
 end
 
+if strcmp(method, 'fullgrid')
+    fprintf('Vp is %f +/- %1.3f \n',db.fg.vbest, db.fg.stdVp )
+    fprintf('R is %f +/- %1.3f \n',db.fg.rbest, db.fg.stdR )
+    fprintf('H is %f +/- %1.3f \n',db.fg.hbest, db.fg.stdH )
+    if exist('dbold','var')
+        if isfield(dbold,'fg')
+            fprintf('Old FG Vp is %f +/- %1.3f \n',dbold.fg.vbest, dbold.fg.stdVp )
+            fprintf('Old FG R is %f +/- %1.3f \n',dbold.fg.rbest, dbold.fg.stdR )
+            fprintf('Old FG H is %f +/- %1.3f \n',dbold.fg.hbest, dbold.fg.stdH )
+        end
+        if isfield(dbold,'hk')
+            fprintf('Old hk R is %f +/- %1.3f \n',dbold.hk.rbest, dbold.hk.stdR )
+            fprintf('Old hk H is %f +/- %1.3f \n',dbold.hk.hbest, dbold.hk.stdH )
+        end
+    end
+end
+
 %% Enter Finishing commands:
-%notes = input('Enter Processing Notes: ', 's');
-%db.processnotes = notes;
-% Enter use / ignore flag
-%db.usable = input('Enter 1 to set as usable or 0 to set as unusable: ');
-%db.usable = 1;
-% Save entrydbold.mb.
+
 saveit = input('Save data to .mat file? (y|n): ','s');
 switch lower(saveit)
     case 'y'

### 2013-08-02 matlab/functions/gridsearchKanC_linear.cpp

@@ -1,0 +1,136 @@
+#include <mex.h>
+#include <math.h>
+#include <matrix.h>
+
+
+void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[] )
+{
+
+  //declare variables
+  mxArray *recIN, *pslowIN;
+  const mwSize *dims;
+  double *rec, *pslow, *Tps, *R, *H, *HRx;
+  double dt, v;
+  int nrecs;
+  int N;
+
+  //associate inputs
+  recIN = mxDuplicateArray(prhs[0]);
+  dt = mxGetScalar(prhs[1]);
+  pslowIN = mxDuplicateArray(prhs[2]);
+  v = mxGetScalar(prhs[3]);
+
+  //figure out dimensions
+  dims = mxGetDimensions(prhs[0]);
+  //numdims = mxGetNumberOfDimensions(prhs[0]);
+  N = (int)dims[0];
+  nrecs = (int)dims[1];
+
+  // associate outputs
+  plhs[0] = mxCreateDoubleMatrix(1, 1, mxREAL);
+  plhs[1] = mxCreateDoubleMatrix(1, 1, mxREAL);
+  plhs[2] = mxCreateDoubleMatrix(1, 1, mxREAL);
+
+  //associate pointers
+  rec = mxGetPr(recIN);
+
+  pslow = mxGetPr(pslowIN);
+  R = mxGetPr(plhs[0]);
+  H = mxGetPr(plhs[1]);
+  HRx = mxGetPr(plhs[2]); // Output actual value of gridsearch cell for setting contour level
+
+  // Grid parameters.
+  double adjtps = 0.5; // Weights chosen from Precambrian crustal evolution - D.A. Thompson
+  double adjtpps = 0.3; // I.D. Bastow, G. Helffrich et al.
+  double adjtpss = -0.2;
+  int i, ir , ih;
+
+  // R limits
+  int nr = 200;
+  double r1 = 1.65;
+  double r2 = 1.95;
+  double dr = (r2-r1)/(nr-1);
+  double r[nr];
+
+  // H limits
+  int nh = 200;
+  double h1 = 25.;
+  double h2 = 50.;
+  double dh = (h2-h1)/(nh-1);
+  double h[nh];
+
+  // Build 1-D arrays
+  r[0] = r1;
+  h[0] = h1;
+  // This loop only works if nv = nr = nh.
[…]

### 2013-08-02 python/compareResults.py

@@ -56,10 +56,9 @@ def meanDifference(x, y):
     n = len(x)
     md = 0
     for i in range(n):
-        for j in range(n):
-            md += abs(x[i] - y[j])
+        md += abs(x[i] - y[i])
 
-    return md/(n**2)
+    return md/(n)
 
 
 for i in range(0,3):
@@ -94,11 +93,6 @@ for i in range(0,3):
     Rcorr = pearsonr(R1, R2)
 
     Hcorr = pearsonr(H1, H2)
-
-    ttR  = ttest(R1, R2)
-    ttH  = ttest(H1, H2)
-    print ttR
-    print ttH
 
     print "Correlation between", leglabel[i], "H datasets is {0:.2f}".format(Hcorr[0]), "using", len(p2.stns), "stations"
     print "Correlation between", leglabel[i], "Vp/Vs datasets is {0:.2f}".format(Rcorr[0]), "using", len(p2.stns), "stations"

### 2013-08-02 python/vpPlots.py

@@ -16,6 +16,16 @@ stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
 csfile = os.environ['HOME'] + '/thesis/data/csStations.json'
 
 
+def meanDifference(x, y):
+    assert (len(x) == len(y))
+    n = len(x)
+    md = 0
+    for i in range(n):
+        md += abs(x[i] - y[i])
+
+    return md/(n)
+
+
 def princomp(A,numpc=0):
      # computing eigenvalues and eigenvectors of covariance matrix
     C = np.dot(A, A.T) / A.shape[1]
@@ -70,7 +80,7 @@ if __name__  == "__main__":
     # FIG 1: Vp estimates against H
     ##############################################################################
 
-    if 1:
+    if 0:
         f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdVp", "hk::stdR"])
 
         corrfg = pearsonr(f.fg_Vp, f.fg_H)
@@ -108,8 +118,8 @@ if __name__  == "__main__":
     #############################################################################
     # FIG 2: Vp estimates and controlled source
     ##############################################################################
-    if 1:
-        maxerr = 0.21#0.24#0.21
+    if 0:
+        maxerr = 0.25#0.24#0.21
         arg = Args().addQuery("fg::stdVp", "lt", str(maxerr))
         f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdH", "fg::stdVp", "hk::stdR"], arg)
 
@@ -169,35 +179,82 @@ if __name__  == "__main__":
 
 
 ########################
-        # fig = plt.figure( figsize = (figwidth, figheight) )
-        # ax = plt.subplot(111)
+    if 1:
 
-        # plt.plot(t, f.fg_Vp, '-ob', lw = lw, ms = ms, label = "Full Gridsearch Vp estimate")
-        # plt.errorbar(t, f.fg_Vp, yerr=stdVp, xerr=None, fmt=None, ecolor = 'blue',
-        #              elinewidth = elw, capsize = caplen, mew = capwid, label = "2 std dev Bootstrap")
-        # plt.plot(t, c.Vp, '-og', lw = lw, ms = ms, label = "Proximal active source estimate")
+        maxerr = 0.21#0.24#0.21
+        arg = Args().addQuery("fg::stdVp", "lt", str(maxerr))
+        f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdH", "fg::stdVp", "hk::stdR"], arg)
[…]

### 2013-08-07 paper/draftBP.tex
paper aperp perpa erpap rpape paper

@@
 
-Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
+%% Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
 The primary reason for this is the poor resolution afforded by the seismic networks currently and historically deployed across Canada.
@@
 
-This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
+%% This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
 It begins with a discussion of the raw data itself followed by a review of the receiver function method and a more detailed explanation of the inversion algorithms used to produce the dataset.
@@
 \subsection{Geological and tectonic summary}
+\begin{itemize}
+
+\item Large scale overview/summary of regions under study
+\item Justification for selecting these regions?
+
+\end{itemize}
 
 \subsection{Previous geophysical studies}
+\begin{itemize}
 
+\item Lithoprobe and attendent papers
+\item Teleseismic studies
+
+\end{itemize}
 %% -----------------------------------------------%%
@@
 This study draws on three sources for these crustal properties.
-The first and primary source are teleseismic receiver functions (hereafter RFs).
-The remaining two datasets are Crust 2.0, a widely published statistical average of similar regions with a global two degree resolution and a compilation of pre-processed active source data (Mooney, 2012).
+The first and primary source is teleseismic receiver function (hereafter RF) analysis.
+The remaining two datasets are Crust 2.0, a widely published statistical average of similar regions with a global two degree resolution, and a compilation of pre-processed active source data (Mooney, 2012).
 Both Crust 2.0 and the active source compilation are used to qualify and compare with the primary dataset.
@@
 \subsection{Teleseismic Data Set}
-The primary data utilized in this study are computed from teleseismic P-wave seismograms originating from more than 700 earthquake sources.
-These earthquakes occurred between the years 2000 and 2012 and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
-Seismic stations are selected from all available regional and national networks including CNSN, Polaris, FedNor and Chasme.
-Seismograms are included for analysis if they within a 30 to 100$^\circ$ epicentral distance window.
-Quality control is performed by admitting only those seismograms with high enough signal to noise ratios to permit visual observation of the dominant P-wave energy and with sufficient impulsive first arrivals that allow arrival times to be accurately measured.
+The primary data utilized in this study are computed from teleseismic {\it P} wave seismograms originating from more than 700 earthquake sources.
+These earthquakes occurred between 2000 and 2012 and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
+Seismic stations are selected from all available regional and national networks including CNSN, POLARIS, FedNor and Chasme.
+Seismograms are included for analysis if they fall within a 30 to 100$^\circ$ epicentral distance window.
+Quality control is performed by admitting only those seismograms with sufficiently high signal to noise ratios to permit visual observation of the dominant {\it P} wave energy and with sufficiently impulsive first arrivals to allow arrival onsets to be accurately measured.
 After selection and filtering more than 80,000 events are available for further processing.
@@
 In general terms, this transformation involves deconvolving an approximation of the earthquake source from horizontal recordings of ground motion [Langston, 1979].
-The resulting waveforms contain discrete pulses corresponding to the S-wave arrivals scattered from subsurface discontinuities including the base of the crust, or Moho.
-Confident identification of these peaks to the estimate of crustal parameters, and improved results can be obtained by transforming the seismogram channels into P and S contributions based on a 1-D model of propagation.
-This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels [Bostock, 1998].
-The direct arrival of the signal on the resulting P wave component is used as an approximation to the source function as P-wave impulse response for teleseismic P approximates a delta function.
-This windowed source estimate is deconvolved from the S wave component computed from the wave field decomposition.
+The resulting waveforms contain discrete pulses corresponding to {\it S} waves scattered from subsurface discontinuities including the base of the crust, or Moho.
+Confident identification of these peaks is required to estimate crustal parameters, and improved results can be obtained by transforming the displacement seismogram into their {\it P} and {\it S} contributions based on a 1-D model of propagation.
+This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels [Kennet, 1991].
+The direct {\it P} arrival of the signal on the resulting {\it P} wave component is used as an approximation to the source function as {\it P} component impulse response for teleseismic P approximates a delta function.
+This windowed source estimate is deconvolved from the {\it S} wave component computed from the wave field decomposition.
 
-We employ $L_2$, multichannel, frequency domain approach to perform the deconvolution which has the advantage of computational efficiency and does not require a priori assumptions about the noise in the data.
+We employ an $L_2$, multichannel, frequency domain approach to perform the deconvolution that has the advantage of computational efficiency and does not require {\it a priori} assumptions concerning the noise levels in the data.
 More specifically, we employ a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $r(t)$.
@@
 \begin{equation}
-  r(t) = F^{-1} \left[ G(\omega) \right] = F^{-1}
+  r(t) = {\cal F}^{-1} \left[ G(\omega) \right] = {\cal F}^{-1}
  \left[ \frac {\sum_n^N S_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta} \right ],
@@
 
-where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ S wave component, $P_n$ is the windowed P wave component, $^*$ denotes the complex conjugate and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
-The parameter $\delta$ is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ which is given by
+\noindent where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ {\it S} wave seismogram, $P_n$ is the corresponding windowed {\it P} wave seismogram, $^*$ denotes complex conjugation and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
+The parameter $\delta$ which is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ is given by
 
@@
 
-where
+\noindent where
 
 \begin{equation}
-  X(\omega) = \frac {\sum_n^N P_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta},
+  X(\omega_m) = \frac {\sum_n^N P_n(\omega_m)P_n^*(\omega_m)} {\sum_n^N P_n(\omega_m)P_n^*(\omega_m) + \delta},
 \end{equation}
 
-and $\omega_m$ is the $m^{th}$ frequency bin in the discrete Fourier transform. All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
+\noindent and $\omega$ is the frequency bin in the discrete Fourier transform. All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
 
@@
 
-A well tested and widely published method for extracting the $\frac{V_P}{V_S}$ ratio $R$ (where $V_P$ is P wave velocity and $V_S$ is S-wave velocity) and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
-This method takes advantage of the differential arrival times between the S-wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct P-wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
+A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, herafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
+This method exploits the dependence of these parameters to the differential arrival times between the {\it S} wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct {\it P} wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
 Note that $PsPs$ and $PpSs$ are kinematic analogues, such that the energy for these two phases arrive simultaneously for a 1-D model.
 For a range of slowness values, $p$, the differential arrival times, $t(p)$, trace moveout curves for each phase arrival given by
-ve
 
@@
 \begin{equation} \label{eq:tps}
-t_{Ps}(p_i)=H \left[ \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2} - \sqrt{\frac{1}{V_P^2} - p_i^2} \right]
+t_{Ps}(p_n) = \frac{H}{V_P} \left[ \sqrt{ R^2 - V_P^2p_n^2} - \sqrt{1 - V_P^2p_n^2} \right]
 \end{equation}
@@
 \begin{equation}
-t_{Pps}(p_i)=H \left[ \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2} + \sqrt{\frac{1}{V_P^2} - p_i^2} \right]
+t_{Pps}(p_n) = \frac{H}{V_P} \left[ \sqrt{ R^2 - V_P^2p_n^2} + \sqrt{1 - V_P^2p_n^2} \right]
 \end{equation}
@@
 \begin{equation}
-t_{Pss}(p_i)= 2H  \sqrt{ \left(\frac{R}{V_P}\right)^2 - p_i^2}
+t_{Pss}(p_n)= \frac{2H}{V_P} \sqrt{ R^2 - V_P^2p_n^2}
 \end{equation}
 
-where $p_i$ is the slowness for the $i^{th}$ receiver function.
+\noindent where $p_n$ is the slowness for the $n^{th}$ receiver function.
 Since strong reflected phases occur at sharp velocity contrasts, the Moho, the boundary targeted by ZK, tends to be well represented on most RF's.
 
-The travel time equations, as employed by ZK, assume a value for crustal P-wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
-In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 2.0 value for the $2^\circ$ containing cell.
+The travel time equations, as employed by ZK, assume a value for crustal {\it P} wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
+In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 2.0 value for the corrisponding $2^\circ$ containing cell.
 In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
@@
 RFs are stacked along trial moveout curves for a range of candidate models of $R$ and $H$ to generate a fitness function $s(H,R)$.
-Within the stacking procedure, each phase is assigned a weight to account for a general trend in the quality of the phases, with direct conversion $Ps$ usually characterized by the highest quality signal followed by $PpPs$ and $PpSs$.
+Within the stacking procedure, each phase is assigned a weight to account for the expected relative amplitudes of the phases for typical velocity contrasts, with the direct conversion $Ps$ usually characterized by the highest quality signal followed by $PpPs$ and $PpSs$.
 We chose weights of $w1 = 0.5$, $w2 = 0.3$ , $w3 = -0.2$ for the $Ps$, $PpPs$ and $PpSs$ phases, respectively.
@@
 \begin{equation}  \label{eq:stack}
-s(H,R) = \sum_{j=1}^{3} S_j \sum_{i=1}^N w_jr_i(t_j)
+s(H,R) = \sum_{m=1}^{3} S_m \sum_{n=1}^N w_mr_n(t_m)
 \end{equation}
 
-where
+\noindent where
 
 \begin{equation}
-S_j(H,R) = \frac {\left[ \sum_{i=1}^N r_i(t_j) \right]^2}
-                 { \sum_{i=1}^N r_i^2(t_j) }
+S_m(H,R) = \frac {\left[ \sum_{n=1}^N r_n(t_m) \right]^2}
+                 { \sum_{n=1}^N r_n^2(t_m) }
 \end{equation}
 
-is the semblance weight for the $j^{th}$ phase, time $t_j$ is calculated from the corresponding travel time function for a given $H$, $R$ pair as a function of slowness $p_i$ and $N$ is the total number of receiver functions.
+\noindent is the semblance weight for the $m^{th}$ phase, time $t_m$ is calculated from the corresponding travel time function for a given $H$, $R$ pair as a function of slowness $p_n$ and $N$ is the total number of receiver functions.
 Multiplying by the semblance weighting sharpens the stacked image, $s(H,R)$, and results in better resolution when discriminating between different models.
-The function $s(H,R)$ can be thought of as a transformation that maps the $Ps$, $PpPs$ and $PpSs+PsPs$ pulses of the receiver function, $r(t)$, into $R$ and $H$ space as positive bands, via Eq.
+The function $s(H,R)$ can be thought of as a transformation that maps times of the $Ps$, $PpPs$ and $PpSs+PsPs$ phases into $R$ and $H$ space as positive bands, via Eq.
 \ref{eq:stack}.
-Constructive interference where these bands intersect will produce a maximum in the stacking function.
-The model, $R$ and $H$, which produce this maximum provides the best estimate for the bulk crustal parameters for a given seismic station.
+Constructive interference where these bands intersect produce a maximum in the stacking function.
+The model, $R$ and $H$, that produces this maximum provides the best estimate for the bulk crustal parameters for a given seismic station.
 
 %% [[\subsection{Full Parameter Method}]]
-A limitation of the ZK approach is the requirement for an initial $V_P$ estimate.
-This can be resolved by computing the stacking function Eq.
+A limitation of the ZK approach is the requirement of an initial $V_P$ estimate.
+In principal we may compute the stacking function Eq.
 \ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
 Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method easily scalable to hundreds of stations.
-This approach, denoted FG, is analyzed to determine the viability of recovering $V_P$.
+This ``fullgrid'' approach is analyzed to determine the viability of recovering $V_P$.
 
@@
 We supplement receiver function estimates of crustal properties with data from controlled source experiments collected and compiled from GSC (Geological Survey of Canada) references by Walter Mooney (personal communication, 2012).
-The active source data provide $V_P$ and a few $V_S$ estimates for hundreds of locations across Canada, many along transects that cross geological provincial boundaries, faults and discontinuities.
-Some of these data are located within reasonable ($<$100km) proximity to seismic stations used in this study and can be used to compare with the $V_P$ and $V_S$ estimates resulting from the RF full gridsearch.
+The active source data provide $V_P$ and a few $V_S$ estimates for hundreds of locations across Canada, many along transects that cross major geological boundaries, faults and discontinuities.
+Some of these data are located within reasonable ($<$100 km) proximity to seismic stations used in this study and can be used to compare with the $V_P$ and $V_S$ estimates resulting from the RF full gridsearch.
 
@@
 Three published studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
-The most recent, a study [Thompson et.
-al., 2010] focused on the Northern Canadian Shield, employs both the linear stacking approach used by ZK as well as a phase weighted method.
+The most recent study [Thompson et.
+al., 2010] focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
 A 2007 study of the Superior Province [Darbyshire et.
@@
 al., 2006] utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
-Comparisons between crustal $R$ estimates for particular seismic stations are made for those values with a corresponding standard error of less than $\pm$0.06.
+Comparisons between crustal estimates of $R$ for particular seismic stations are made for those values with corresponding standard errors of less than $\pm$0.06.
 This error threshold is chosen empirically as that which best partitions those receiver function stacks with visually recognizable energy along phase moveout curves from those where lack of data or poor quality make these moveout curves difficult to distinguish.
 For each study under comparison, data are reprocessed using the $V_P$ chosen by the study authors.
-For the 2006 and 2007 studies the ZK + semblance method is used, for the Thompson et.
-al study the linear stacking ZK apprach is employed.
+For the 2006 and 2007 studies the ZK + semblance method is used whereas for the Thompson et.
+al study the linear stacking ZK approach is employed.
 Both correlation coefficient and mean difference are provided in (Table \ref{table:comparison}) as measures of similarity.
@@
 
-All three studies provide $H$ estimates that correlate strongly with the crustal thickness parameters computed here.
-The data from the Northern Canadian Shield most closely correlates in both $H$ and $R$.
-The lower correlation in $R$ than $H$ may be attributed to a number of factors such as the choice of deconvolution algorithm and selection and scaling of data.
-The greater uncertainty in $R$ inherent in the ZK method will cause larger fluctuations in computed estimates for differences in procedure.
-The data from Thompson et.
+All three studies reported $H$ estimates that correlate strongly with the crustal thickness parameters computed here.
+The data from the northern Canadian Shield shows strong correlation in both $H$ and $R$.
+The lower correlation observed in estimates of $R$ than $H$ is not unexpected as the range of $R$ is relatively narrow and small changes ($\approx 2\%$) are physically meaningful.
+A corollary of this is greater sensitivity of $R$ to differences in procedure such as the choice of deconvolution algorithm and selection and scaling of data.
+The results from Thompson et.
 al.
-[2010] also align most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
-Results published by Eaton show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that are roughly equal to the averaged uncertainty in the data.
+[2010] also corrispond most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
+Results published by {\it Eaton et.
+al.} [2006] show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that is roughly equal to the averaged uncertainty in the data.
 Parameter estimates from the Superior Province [Darbyshire et.
@@
 
-\subsection{Full Parameter Search Analysis}
+\subsection{Full Grid Search Analysis}
 
-We now proceed to compare our results computed using the ZK approach with those from the full parameter search (Table \ref{table:ZKvsFG}).
+We now proceed to compare our results computed using the ZK approach with those from the full grid search (Table \ref{table:ZKvsFG}).
 We note that the velocity ratio $R$ has a correlation between the two datasets of 0.76 whereas $H$ has a correlation of 0.51.
-The low correlation between crustal thickness estimates can be explained by the strong tradeoff between $H$ and $V_P$ through the quantity $(t_{Pps}-t_{Ps})/2=t_{Pss}/2-t_{Ps}=H\sqrt{1-p^2V^2_P)}/V_P$.
-As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:HVp_ULM}).
+The low correlation between crustal thickness estimates can be explained by the strong tradeoff between $H$ and $V_P$ through the quantity
+
+$$\frac{t_{Pps}-t_{Ps}}{2}=\frac{t_{Pss}}{2} - t_{Ps}= \frac{H}{V_P}\sqrt{1-p^2V^2_P)}.$$
+
+\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:HVp_ULM}).
 This trade-off contributes to the low correlation in $H$ between the two methods.
 Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$.
-The trade-off between the other parameters is lower as there is greater dependence between the parameters in the travel time equations (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
+The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
 
-For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ infeasible using the FG approach.
+For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ unfeasible using the full grid approach.
 A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty.
-RFs are arranged vertically, ordered by increasing slowness, for both stations and are shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
-Cross sections of the three dimensional stacks $s(H,R,V_P)$  are displayed for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
-The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P +/- 0.64 km/s$.
-The station ULM, with cleaner data has a lower absolute trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P +/- 0.16 km/s$.
+RFs are arranged vertically, ordered by increasing slowness, and shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
+Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
+The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P \pm 0.64 {\rm km/s}$.
+Station ULM, with cleaner data, has less trade-off between $H$ and $V_P$ as indicated by the corresponding cross section and the lower bootstrap error of $V_P \pm 0.16 {\rm km/s}$.
 
-Active source experiments located near broadband seismic stations provide independant crustal thickness and P-wave velocity estimates to test against data derived from the full parameter search method.
-We select stations with low error in $V_P$ and active source experiment locations within a 100km radius.
-If several active source experiments are located within this radius estimates are averaged.
-Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and are presented in figure \ref{fig:activeVp}-\ref{fig:activeH}.
-Estimates computed from the FG approach are roughly 0.25 km/s higher than the active source data.
-This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting a bias.
+Active source experiments located near broadband seismic stations provide independent crustal thickness and {\it P} wave velocity estimates to test against data derived from the full parameter search method.
+We select stations with low error in $V_P$ and active source experiment locations within a 100 km radius.
+If several active source experiments are located within this radius, estimates are averaged.
+Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and the parameter estimates are presented in figure \ref{fig:activeVp}-\ref{fig:activeH}.
+Estimates computed from the full grid approach are roughly 0.25 km/s higher than the active source data.
+This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting possible bias.
 However, the limited number of data means this bias is poorly constrained.
 
-There is less correlation between the crustal thickness data, owing largely to a single station ARVN.
-The reason for this discrepency, roughly 10km, is unknown though both the ZK method and and a published study [Thompson et.
-al., 2010] provide estimates for ARVN of around 41km, much nearer the FG value of 42km.
-Despite the lower correlation the apparent bias towards higher values computed from teleseismic data is maintained.
-A positive bias in one parameter will tend to bias the other, owing to the $H/Vp$ relationship in $t_{ps}$.
-Taking the average active source estimates from the two best stations, ULM and WHY, a $V_P$ of 6.3 km/s, a $V_P/V_S$ of 1.75 and a 36km thick crust, a 0.25 km/s change in $V_P$ will result in a 1.3km change in $H$.
-The actual mean difference for the two stations, 1.9km, is close but slightly greater than this value.
+There is less correlation between the crustal thickness data, though this is largely attributed to an approximate 10 km difference between the ARVN estimates.
+The reason for this discrepancy is unknown though both the ZK method and and a published study [Thompson et.
+al., 2010] provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
+Despite lower correlation the apparent bias towards higher values computed from teleseismic data is also maintained for $H$.
+A positive bias in one parameter will tend to bias the other, owing to the relationship between thickness, velocity and times of a scattered phase.
+For a constant $t_{ps}$ one may compute how much $H$ must change for a 0.25 km/s bias in $V_P$ [Zhu and Kanamori, 2000].
+Using an average of the parameter estimates from the two cleanest stations, ULM and WHY, $\delta H$ will be 1.3 km, this fits well with the actual bias in the values.
 
-[[ SOURCES REGARDING BIAS ]] [[ WHAT DOES THIS MEAN, IS VP USEABLE FOR THE HIGHEST QUALITY STATIONS?]]
+[[ SOURCES REGARDING BIAS ]] [[ WHAT DOES THIS MEAN, IS VP USABLE FOR THE HIGHEST QUALITY STATIONS?]]
 
@@
   \includegraphics[width=\textwidth]{activeSourceCompVp}
-  \caption{P-wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
+  \caption{\it{P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
   \label{fig:activeVp}
@@
 
-We first consider the average crustal structure across Canada.
-We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to XX km in increments of DX km)  using the estimates for $R$ computed via the ZK + semblance method.
-The resulting profiles contain energy at depths corresponding to velocity discontinuities.
-By assembling all depth profiles into a matrix, we may decompose the data via SVD into its principal components.
+We now consider the average crustal structure across Canada.
+We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method.
+The resulting profiles contain energy at depths corresponding to major velocity discontinuities.
+By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components.
 Figure \ref{fig:conradPCA} contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles.
-Higher modes corresponding to diminishing contributions to total variance.
+Higher modes correspond to diminishing contributions to total variance.
 Each vector is normalized to unit magnitude.
@@
 sedimentary layers) within the top 10 km and Moho variability between 30 and 46 km .
-The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10km to 20km range.
-All 5 modes display markedly lower energy within a depth window roughly between 20km and 30km.
+The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10 km to 20 km range.
+All 5 modes display markedly lower energy within a depth window roughly between 20 km and 30 km.
 
@@
 
-We proceed to consider velocity ratio, $R$, estimates for stations with assigned bootstrap errors below $\pm 0.06$ in a geographic context.
+We proceed to consider crustal thickness $H$ and velocity ratio, $R$, estimates for stations with assigned $R$ bootstrap errors below $\pm 0.06$ in a geographic context (figure \ref{map:stationMap}).
 To compensate for the uneven distribution of seismic stations, weights are applied to both $R$ and $H$ estimates before averaging over a given region.
@@
 Table \ref{table:regionParameters} displays the results of our regionalization in $R$ and $H$.
-The weighted average crustal thickness for all stations (Canada) is 37km, about two kilometres thinner than the average for the Canadian Shield.
-Crust 2.0 provides slightly thicker crustal average for Canada (38.2km) and a slightly thinner average for the Shield (38.8km).
-The Slave Province averages are similar to those for the Churchill province and has a deeper crust than Crust2.0 and a smaller seismic velocity ratio.
-Moving south and east from the Churchill Province through the Superior Province into the Grenville, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$.
+The weighted average crustal thickness for all stations (Canada) is 37 km, about two kilometres thinner than the average for the Canadian Shield.
+Crust 2.0 provides a slightly thicker crustal average for Canada (38.2 km) and a slightly thinner average for the Shield (38.8 km).
+The Slave Province averages are similar to those for the Churchill province and a thicker crust than Crust2.0 and a smaller seismic velocity ratio.
+Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$, figures \ref{map:Hmap}-\ref{map:Rmap}.
 The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
-In a similar fashion, weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7km in the Grenville Province.
+In a similar fashion, the weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7 km in the Grenville Province.
 The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data.
@@
 
- %% \begin{figure}
-%%   \centering
-%%   \includegraphics[width=\textwidth]{VpVsMap}
-%%   \caption{}
-%%   \label{fig:VpVsMap}
-%% \end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stationMap}
+  \caption{}
+  \label{map:stationMap}
+\end{figure}
 
-%% \begin{figure}
-%%   \centering
-%%   \includegraphics[width=\textwidth]{VpVsMapCrust}
-%%   \caption{}
-%%   \label{fig:VpVsMapCrust}
-%% \end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{Hmap}
+  \caption{}
+  \label{map:Hmap}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{Rmap}
+  \caption{}
+  \label{map:Rmap}
+\end{figure}
 
@@
 
-  - Comparison Discussion
-    - Crustal Thickness
-      - Very good correlation in thickness estimates with all three datasets.
-    - Vp/Vs
-      - Lower correlation with the Vp/Vs values.
-        - Expect lower correlation with Vp/Vs since the travel time equations are
-          less sensitive to seismic velocities than they are to crustal thickness.
-          This means that there will be more uncertainty during the parameter
-          search for Vp/Vs and hence higher degree of relative error.
-        - ??? Utilizing different sources, applying different deconvolution and
-          filtering techniques will therefore have a greater impact on the
-          uncertainty of picking Vp/Vs than that of crustal thickness H. ???
+\begin{itemize}
 
-   - Explanation of results
-    - Canada H and Vp/Vs
-      - Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
-        2.0 average is 38km as well, however this is expected since Crust2.0 data for
-        Canada is based on the same data as that of Lith5.0
-      - ??? Make some comparison to global average ???
+\item Quiet zone in discontinuity data
+
+\item Compare data to Crust 2.0. What do the differences mean.
+  \begin{itemize}
+  \item Lith 5.0 data?
+  \end{itemize}
+
+\item Discuss spatial trends -> high R -> Low R moving North West through Ontario
+  \begin{itemize}
+  \item this is part of larger trend of changing R values as we move from Churchill to Grenville
+  \item in the literature? Yes. See notes on lithoprobe studies
+  \end{itemize}
+
+\item Secular variation in NE Churchill from thinner crusts to thicker as we move East
+  \begin{itemize}
+  \item in the literature? See Lith 5.0: A revised crustal model for Canada based on lithoprobe results.
+  \end{itemize}
+
+ \item Explanation of results
+
+ \item Canada H and Vp/Vs
+
+ \item Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
+   2.0 average is 38 km as well, however this is expected since Crust2.0 data for
+   Canada is based on the same data as that of Lith5.0 .... is it definately?
+
+ \item Make some comparison to global average?
+
+\end{itemize}
 
@@
 \section{Conclusions}
-Conclusions here
 
@@
 
-\begin{bibliography}
+\section{bibliography}
+
 Bassin, C., G. Laske, G. Masters (2000), The Current Limits of Resolution for Surface Wave Tomography in North America, EOS Trans AGU, 81.
 
-Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21,183-21,200.
+Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.
 
@@
 
+Kennet, B. L. N., (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
+
 Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
@@
 
-\end{bibliography}
-
 \end{document}

### 2013-08-07 python/codecounter.py

@@ -11,7 +11,7 @@ def countLinesOfCode(directory):
     for fname in listing:
         subtotal = 0
         Fname = os.path.join(directory,fname)
-        if fname[-2:] == "py" or fname[-1:] == "m" or fname[-2:] == "sh":
+        if fname[-2:] == "py" or fname[-1:] == "m" or fname[-2:] == "sh" or fname[-3:] == 'cpp' or fname[-2:] == 'js':
             with open(Fname) as f:
                 for i, l in enumerate(f):
                     pass
@@ -27,5 +27,5 @@ if __name__ == "__main__":
     if (len(sys.argv) < 2):
         print "Please supply a directory"
         exit()
-    
+
     countLinesOfCode(sys.argv[1])

### 2013-08-08 paper/draftBP.tex
crust1

@@
 
-\item Quiet zone in discontinuity data
+\item Crustal discontinuities
+  \begin{itemize}
+    \item No observed reflector patterns outside of Sedimentary and Moho Layers.
+    \item Upper crust appears to have more consistent structure, though only marginally 4th and 5th principal components
+    \item Lower crust appears more heterogenous.
+    \item Brittle Ductile transition? See Roy Hynmann
+    \item Conrad may be last reflector in an more reflective upper crust, similar to the reflection moho. This interpretation may be more consistent with the dataset
+    \item
+  \end{itemize}
 
+\item Crustal Evolution
+  \begin{itemize}
+    \item Archean vs Proterozoic
+    \item Canada Averages
+    \item Regional Averages
     \item Compare data to Crust 2.0. What do the differences mean.
@@
       \item Lith 5.0 data?
+      \item Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
+        2.0 average is 38 km as well, however this is expected since Crust2.0 data for
+        Canada is based on the same data as that of Lith5.0 .... is it definately?
+
       \end{itemize}
@@
       \end{itemize}
-
- \item Explanation of results
-
- \item Canada H and Vp/Vs
-
- \item Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
-   2.0 average is 38 km as well, however this is expected since Crust2.0 data for
-   Canada is based on the same data as that of Lith5.0 .... is it definately?
-
- \item Make some comparison to global average?
+  \end{itemize}
 

### 2013-08-13 paper/draftBP.tex
narg

@@
 
+\item Analysis of the discontuity data for the Canadian crust indicates a heterogenous crust with a
 \item Crustal discontinuities
   \begin{itemize}
-    \item No observed reflector patterns outside of Sedimentary and Moho Layers.
-    \item Upper crust appears to have more consistent structure, though only marginally 4th and 5th principal components
-    \item Lower crust appears more heterogenous.
+
+    \item Upper crust appears to have more consistent structure, though only marginally... 4th and 5th principal components
+    \item Lower crust appears more heterogenous (Fewer consistent structure throughout Canada)
     \item Brittle Ductile transition? See Roy Hynmann
     \item Conrad may be last reflector in an more reflective upper crust, similar to the reflection moho. This interpretation may be more consistent with the dataset
-    \item
+
   \end{itemize}
 
-\item Crustal Evolution
+% \item Crustal Evolution
   \begin{itemize}

### 2013-08-13 matlab/conradPCA.m

@@ -1,0 +1,237 @@
+% conrad pricipal componenet analysis
+
+%clear all; close all
+loadtools;
+addpath functions
+addpath ../sac
+addpath([userdir,'/programming/matlab/jsonlab'])
+%% Variables
+clearvars -except json
+sacfolder = '/media/bpostlet/TerraS/CN';
+databasedir = '/media/bpostlet/TerraS/database';
+if ~exist('json', 'var')
+    json = loadjson('../data/stations.json');
+end
+%%  Select Station to Process and load station data
+s = fieldnames(json);
+
+PLOT = false;
+idx = 1;
+
+for ii = 1 : length(s)
+    
+    station = s{ii};
+    
+    dbfile = fullfile(databasedir, [station,'.mat'] );
+    
+    if  numel(strfind(json.(station).status, 'processed-ok'))
+        if exist(dbfile, 'file')
+            disp(station)
+            load(dbfile)
+            
+            if db.hk.stdR > 0.06
+                fprintf('skipping %s\n', station)
+                continue
+            end
+        else
+            continue
+        end
+        
+    else
+        fprintf('skipping %s\n', station)
+        continue
+    end
+    %% Application logic
+    % Set up constants
+    p2 = db.pslow.^2;
+    f1 = sqrt( (db.hk.rbest/db.hk.v)^2 - p2);
+    f2 = sqrt( (1/db.hk.v)^2 - p2);
+    np = length(p2);
+    nt = length(db.rec);
+    
+    %% Line search for H.
+    nh = 500;
+    h1 = .1; %db.dt / (f1(1) - f2(1));
+    h2 = 50; %;db.hk.hbest + 2;
+    dh = (h2-h1)/(nh-1);
+    H = h1:dh:h2;
+%     
+    gvr = db.rec'; %rotate
+    gvr = gvr(:); %vectorize
+    
+    % Stack
+    for ih= 1:length(H)
+        tps = H(ih)*(f1-f2);
+        %ind = round(tps/db.dt)+1+[0:np-1]*nt;
+        %disp(ind)
+        stackh(ih) = mean(gvr(round(tps/db.dt)+1+[0:np-1]*nt));
+    end
+    
+   
+    
[…]

### 2013-08-13 python/compareResults.py

@@ -66,7 +66,7 @@ for i in range(0,3):
     pro = os.environ['HOME'] + data[2*i + 1]
     pub = os.environ['HOME'] + data[2*i]
 
-    arg = Args().addQuery("stdR", "lt", 0.06)
+    arg = Args().addQuery("stdR", "lt", 0.08)
 
     p0 = Params(pro, ["H", "R", "stdR", "stdH"], arg)
     p2 = Params(pub, ["H", "R", "stdR", "stdH"], arg)
@@ -146,7 +146,7 @@ for i in range(0,3):
     plt.axis("tight")
 
 
-#plt.show()
+plt.show()
 
 
 

### 2013-08-13 python/dbutils.py

@@ -280,7 +280,7 @@ def setStatus(s, stdict):
             status = "picked"
         # CHANGE THIS BELOW TO A KANAMORI STATISTIC
         if 'hk' in s[k]: # Set processing status by Kan R result
-            if (s[k]['hk']['stdR'] < 0.06) and s[k]['usable']:
+            if (s[k]['hk']['stdR'] < 0.08):
                 status = "processed-ok"
             else:
                 status = "processed-notok"

### 2013-08-26 matlab/conradPCP.m
thwat

@@ -1,191 +1,0 @@
-% conrad pricipal componenet analysis
-
-%clear all; close all
-loadtools;
-addpath functions
-addpath ../sac
-addpath([userdir,'/programming/matlab/jsonlab'])
-%% Variables
-clear X
-sacfolder = '/media/bpostlet/TerraS/CN';
-databasedir = '/media/bpostlet/TerraS/database';
-if ~exist('json', 'var')
-    json = loadjson('../data/stations.json');
-end
-%%  Select Station to Process and load station data
-s = fieldnames(json);
-
-PLOT = false;
-idx = 1;
-
-for ii = 1 : length(s)
-    
-    station = s{ii};
-    
-    dbfile = fullfile(databasedir, [station,'.mat'] );
-    
-    if  numel(strfind(json.(station).status, 'processed-ok'))
-        if exist(dbfile, 'file')
-            disp(station)
-            load(dbfile)
-        else
-            continue
-        end
-        
-    else
-        fprintf('skipping %s\n', station)
-        continue
-    end
-    %% Application logic
-    % Set up constants
-    p2 = db.pslow.^2;
-    f1 = sqrt( (db.hk.rbest/db.hk.v)^2 - p2);
-    f2 = sqrt( (1/db.hk.v)^2 - p2);
-    np = length(p2);
-    nt = length(db.rec);
-    
-    %% Line search for H.
-    nh = 500;
-    h1 = .1; %db.dt / (f1(1) - f2(1));
-    h2 = 50; %;db.hk.hbest + 2;
-    dh = (h2-h1)/(nh-1);
-    H = h1:dh:h2;
-%     
-    gvr = db.rec'; %rotate
-    gvr = gvr(:); %vectorize
-    
-    % Stack
-    for ih= 1:length(H)
-        tps = H(ih)*(f1-f2);
-        %ind = round(tps/db.dt)+1+[0:np-1]*nt;
-        %disp(ind)
-        stackh(ih) = mean(gvr(round(tps/db.dt)+1+[0:np-1]*nt));
-    end
-    
-   
-    
-    % Note, that we are using all processed stations not processed-ok
-    X(idx, :) = stackh;
-    idx = idx + 1;
-    
-    if (PLOT)
-
-        t1 = H * (f1(1) - f2(1));
[…]

### 2013-08-26 python/lithologyPlots.py

@@ -1,0 +1,152 @@
+#!/usr/bin/python2
+#
+# python-mode indent C-c < or C-c >
+# python-mode comment/uncomment region M-;
+###########################################################################
+# IMPORTS
+###########################################################################
+import os, json
+import numpy as np
+import matplotlib
+#matplotlib.use('agg')
+from matplotlib.patches import Polygon
+from matplotlib.collections import PatchCollection
+import matplotlib.pyplot as plt
+from plotTools import Args, Params
+from histplots import poisson
+
+
+stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
+regionf = os.environ['HOME'] + '/thesis/data/voronoi.json'
+
+def lithVp(vs, poisson):
+    return vs * np.sqrt( (2 - 2*poisson) / (1 - 2 * poisson)  )
+
+
+left, width = .25, .5
+bottom, height = .25, .5
+right = left + width
+top = bottom + height
+
+def addtext(ax, x, y, text, rotation):
+    ax.text(x-0.02, y, text,
+            color = 'black',
+            horizontalalignment = 'left',
+            verticalalignment = 'bottom',
+            rotation = rotation,
+            zorder = 10)
+
+
+liths = {
+    "granite gneiss": {
+        "Vp": 6.208,
+        "Vs": 3.583,
+        "R": 1.732,
+        "poisson": 0.250
+        },
+    "tonalite gneiss": {
+        "Vp": 6.302,
+        "Vs": 3.606,
+        "R": 1.747,
+        "poisson": 0.257
+        },
+    "mafic granulite": {
+        "Vp": 6.942,
+        "Vs": 3.820,
+        "R": 1.817,
+        "poisson": 0.283
+        },
+    "mafic garnet granulite": {
+        "Vp": 7.249,
+        "Vs": 4.026,
+        "R": 1.801,
+        "poisson": 0.277
+        },
+    "amphibolite": {
+        "Vp": 6.983,
+        "Vs": 3.959,
+        "R": 1.764,
+        "poisson": 0.263
+        },
+    "hornblendite": {
+        "Vp": 7.261,
+        "Vs": 4.144,
+        "R": 1.752,
+        "poisson": 0.258
+        }
+    }
+
+
+if __name__  == "__main__":
+
+
+    ## Figure Properties #######
[…]

### 2013-08-26 python/vpPlots.py

@@ -16,14 +16,14 @@ stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
 csfile = os.environ['HOME'] + '/thesis/data/csStations.json'
 
 
-def meanDifference(x, y):
+def RMSDifference(x, y):
     assert (len(x) == len(y))
     n = len(x)
     md = 0
     for i in range(n):
-        md += abs(x[i] - y[i])
+        md += (x[i] - y[i])**2
 
-    return md/(n)
+    return np.sqrt(md/(n))
 
 
 def princomp(A,numpc=0):
@@ -199,8 +199,8 @@ if __name__  == "__main__":
         corr = pearsonr(f.fg_Vp, c.Vp)
         print "Active Source vs FG Vp with {} stations: correlation = {}".format(len(f.stns), corr[0])
 
-        md = meanDifference(f.fg_Vp, c.Vp)
-        print "Active Source vs FG Vp with {} stations: Mean Difference = {}".format(len(f.stns), md)
+        md = RMSDifference(f.fg_Vp, c.Vp)
+        print "Active Source vs FG Vp with {} stations: RMS Difference = {}".format(len(f.stns), md)
 
         fig = plt.figure( figsize = (figwidth, figheight) )
         ax = plt.subplot(111)

### 2013-09-08 paper/draftBP.tex
thwat
disucssion

--- 2013-08-07 paper/draftBP.tex
@@
 \begin{table}
-  \begin{tabular}{ l l l l l l }
-    \cline{3-6}
-    & & \multicolumn{2}{ c }{Correlation} & \multicolumn{2}{ c }{Mean Difference} \\
+  \begin{tabular}{ l l l l l l l}
+    \cline{4-7}
+    & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
     \hline
-    Study Authors &   Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
+    Study Authors             & Region         & Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
     \hline
-    Thompson et. al. (2010)   & 29 & 0.97 & 0.70 & 0.49 & 0.012 \\
-    Darbyshire et. al. (2007) & 10 & 0.95 & 0.43 & 1.21 & 0.033 \\
-    Eaton et. al. (2006)      & 26 & 0.90 & 0.61 & 0.79 & 0.025 \\
+    Thompson et. al. (2010)   & Northern Canadian Sheild & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
+    Darbyshire et. al. (2007) & Superior Province        & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
+    Eaton et. al. (2006)      & Grenville Orogen         & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
     \hline
@@
 \section{Discussion}
-\subsection{Canada}
 
-\begin{itemize}
+Mean crustal composition is an important parameter in the literature studying the continental crust.
+Mean composition should not be confused as an estimate of real physical crustal lithological and mineral assemblages.
+Rather, mean crustal composition is a representative quantity that is useful for constraining crustal genesis and evolution in space and time.
+Early attempts at estimating mean crustal composition were largely accomplished through extrapolations of upper crustal rocks [Rudnick and Gao, 2003].
+After the plate-tectonic revolution, Taylor [1967, 1977], developed the 'island arc' or 'andesite' model which posited that bulk crustal composition should be equivalent to the average convergent-margin andesite.
+Several problems with this theory, such as the fact that island arcs are more basaltic than andesitic and therefore produce basaltic crustal accretions and cannot account for the intermediate composition of post-Archean crust, led to the introduction of seismic methods to estimate mid and lower crustal composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
+This study differs from these later efforts in that it follows more closely to the complete crustal seismic analysis in the work of Smithson et.
+al.
+[1981] and Zandt and Ammon [1995].
 
-\item Quiet zone in discontinuity data
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{lithology}
+  \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
+  \label{fig:lith}
+\end{figure}
 
-\item Compare data to Crust 2.0. What do the differences mean.
-  \begin{itemize}
-  \item Lith 5.0 data?
-  \end{itemize}
+The mean seismic velocity ratio computed in this study is combined with values for $V_P$ from the Crust 1.0 model to enable estimates for $V_P$ and $V_S$ to be used to estimate mean crustal composition.
+Figure \ref{fig:lith} shows mineral assemblages and weighted Canadian Shield estimates plotted against $V_P$ and $V_S$.
+Mineral assemblages are chosen as useful representations of common crustal rocks and, specifically, for crust in Shield-like regions.
+Examples of common shield lithologies can be seen in the 30km of paleo-depth represented in the Kapuskasing uplift in the Superior province [Percival and West, 1994].
+This uplift is largely comprised of Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed to greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites.
+Tonalites and granodiorites are also part of the TGG suite (for tonalite-trondhjemite-granodiorite) which dominate Archaean cratons [Rudnick, 1995].
+Seismic velocity values for these lithologies are taken from the work of Christensen [1996] and are selected at average crustal pressures of 600 MPa.
+Lines of constant Poisson's Ratio are plotted for reference.
 
-\item Discuss spatial trends -> high R -> Low R moving North West through Ontario
-  \begin{itemize}
-  \item this is part of larger trend of changing R values as we move from Churchill to Grenville
-  \item in the literature? Yes. See notes on lithoprobe studies
-  \end{itemize}
-
-\item Secular variation in NE Churchill from thinner crusts to thicker as we move East
-  \begin{itemize}
-  \item in the literature? See Lith 5.0: A revised crustal model for Canada based on lithoprobe results.
-  \end{itemize}
-
- \item Explanation of results
-
- \item Canada H and Vp/Vs
-
- \item Lith5.0 (Perry, 2002) estimates for Canadian crustal averages is 38 km. The Crust
-   2.0 average is 38 km as well, however this is expected since Crust2.0 data for
-   Canada is based on the same data as that of Lith5.0 .... is it definately?
-
- \item Make some comparison to global average?
-
-\end{itemize}
+The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in composition.
+These three provinces cluster around the average for the Canadian Shield.
+The Grenville Province on the other hand differs significantly in mean composition.
+One interpretation is that the crust in the Grenville province has a mean crustal composition corrisponding to a more mafic lower crustal composition than other Shield provinces.
+This interpretation consistent with a current model for the tectonic evolution of the central region of the Grenville Province that
 
@@
 Christensen, N. I. (1996), Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156.
+
+Christensen, N. I., W. D. Mooney (1995) Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
 
@@
 
-Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002) LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+Percival, J. A., Gordon F. W. (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Canadian. J. of Earth Sciences, 31, 1256-1286.
+
+Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+
+Rudnick, R. L., D. M. Fountain (1995), Nature and composition of the continental crust: A lower crustal perspective, Rev. Geophys., 33(3), 267–309, doi:10.1029/95RG01302.
+
+Rudnick, R. L., S. Gao (2003), Composition of the Continental Crust, Treatise on Geochem., 3, 1-64.
+
+Smithson, S. B., R. A. Johnson, Y. K. Wong (1981), Mean crustal velocity: a critical parameter for interpreting crustal structure and crustal growth, Earth and Planetary Science Letters, 53, 323-332.
+
+Taylor, S. R. (1967), The origin and growth of continents, Tectonophysics, 4, 17–34.
+
+Taylor, S.
+R.
+(1977), Island arc models and the composition of the continental crust, in Island Arcs, Deep Sea Trenches and Back-Arc Basins, Maurice Ewing Ser., vol.
+1, edited by M.
+Talwani and W.
+C.
+Pitman III, pp.
+325–335, AGU, Washington, D.
+C., doi:10.1029/ME001p0325.
 

### 2013-09-09 paper/draftBP.tex
discussion A

@@
 The first and primary source is teleseismic receiver function (hereafter RF) analysis.
-The remaining two datasets are Crust 2.0, a widely published statistical average of similar regions with a global two degree resolution, and a compilation of pre-processed active source data (Mooney, 2012).
-Both Crust 2.0 and the active source compilation are used to qualify and compare with the primary dataset.
+The remaining two datasets are Crust 1.0, a widely published statistical average of similar regions with a global two degree resolution, and a compilation of pre-processed active source data (Mooney, 2012).
+Both Crust 1.0 and the active source compilation are used to qualify and compare with the primary dataset.
 
@@
 The travel time equations, as employed by ZK, assume a value for crustal {\it P} wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
-In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 2.0 value for the corrisponding $2^\circ$ containing cell.
+In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 1.0 value for the corrisponding $2^\circ$ containing cell.
 In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
@@
 
-The Crust 2.0 model [Bassin, 2000] employs many of these same active source data to provide regular, complete coverage across all of Canada.
-The model includes estimates of $V_P$ and $V_S$ information with depth at
-$2^\circ$ resolution in latitude and longitude.
-In addition to employing local active source data, Crust 2.0 is also constrained by statistically aggregated data from geographically distant but geologically similar regions.
+The Crust 1.0 model [Laske et.
+al., 2013] employs many of these same active source data to provide regular, complete coverage across all of Canada.
+The model includes estimates of $V_P$ and $V_S$ information with depth at $1^\circ$ resolution in latitude and longitude.
+In addition to employing local active source data, Crust 1.0 is also constrained by statistically aggregated data from geographically distant but geologically similar regions.
 It therefore affords a somewhat independent dataset suited for comparison with the crustal estimates calculated in this study.
@@
   \includegraphics[width=\textwidth]{activeSourceCompVp}
-  \caption{\it{P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
+  \caption{\it{P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
   \label{fig:activeVp}
@@
   \includegraphics[width=\textwidth]{activeSourceCompH}
-  \caption{Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 2.0}
+  \caption{Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
   \label{fig:activeH}
@@
 The weighted average crustal thickness for all stations (Canada) is 37 km, about two kilometres thinner than the average for the Canadian Shield.
-Crust 2.0 provides a slightly thicker crustal average for Canada (38.2 km) and a slightly thinner average for the Shield (38.8 km).
-The Slave Province averages are similar to those for the Churchill province and a thicker crust than Crust2.0 and a smaller seismic velocity ratio.
+Crust 1.0 provides a slightly thicker crustal average for Canada (38.1 km) and a slightly thinner average for the Shield (38.3 km).
+The Slave Province averages are similar to those for the Churchill province with a crustal thickness and seismic velocity ratio comparable to Crust 1.0.
 Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$, figures \ref{map:Hmap}-\ref{map:Rmap}.
@@
 In a similar fashion, the weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7 km in the Grenville Province.
-The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data.
-The Crust 2.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable.
-The trend in the seismic velocity is not, however, as apparent.
+The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data as well as as the Crust 1.0 $V_P$ data.
+The Crust 1.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable.
+The trend in the seismic velocity ratio is not, however, as apparent.
 
 \begin{table}
-  \begin{tabular}{ l l l l l l }
-    \cline{2-6}
-    & \multicolumn{2}{ c }{ZK + semblance} & \multicolumn{2}{ c }{Crust 2.0} & \multicolumn{1}{ c }{Active Source} \\
+  \begin{tabular}{ l l l l l l l }
+    \cline{2-7}
+    & \multicolumn{2}{ c }{ZK + semblance} & \multicolumn{3}{ c }{Crust 1.0} & \multicolumn{1}{ c }{Active Source} \\
     \hline
-    Region  & $H [km]$ &$R$ & $H [km]$&$R$ &$V_P [km/s]$ \\
+    Region  & $H [km]$& $R$ &$H [km]$& $R$  &$V_P [km/s]$&$V_P [km/s]$ \\
     \hline
-    Canada     & 37.0 & 1.75 & 38.2 & 1.77 & 6.33\\
-    Shield     & 39.2 & 1.74 & 38.8 & 1.77 & 6.42\\
-    Slave      & 38.2 & 1.74 & 37.0 & 1.76 & 6.44\\
-    Churchill  & 38.9 & 1.73 & 37.8 & 1.77 & 6.39\\
-    Superior   & 39.7 & 1.74 & 39.1 & 1.76 & 6.44\\
-    Grenville  & 41.7 & 1.78 & 40.5 & 1.78 & 6.48\\
+    Canada    & 37.0 & 1.75 & 38.1   & 1.74 &   6.41     & 6.33\\
+    Shield    & 39.2 & 1.74 & 38.3   & 1.74 &   6.47     & 6.42\\
+    Slave     & 38.2 & 1.74 & 38.0   & 1.73 &   6.47     & 6.44\\
+    Churchill & 38.9 & 1.73 & 37.9   & 1.74 &   6.46     & 6.39\\
+    Superior  & 39.7 & 1.74 & 38.9   & 1.74 &   6.49     & 6.44\\
+    Grenville & 41.7 & 1.78 & 39.7   & 1.74 &   6.50     & 6.48\\
     \hline
@@
 al.
-[1981] and Zandt and Ammon [1995].
+[1981].
 
@@
 
-The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in composition.
+The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in average composition.
 These three provinces cluster around the average for the Canadian Shield.
 The Grenville Province on the other hand differs significantly in mean composition.
-One interpretation is that the crust in the Grenville province has a mean crustal composition corrisponding to a more mafic lower crustal composition than other Shield provinces.
-This interpretation consistent with a current model for the tectonic evolution of the central region of the Grenville Province that
+One interpretation is that the crust in the Grenville province has a mean crustal composition corresponding to a more mafic lower crustal composition than other Shield provinces.
+This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally [Eaton et.
+al., 2005; Rivers, 2008].
+However, many stations in the South-East of the province also have seismic ratios greater than 1.78, similar to the high values common in the adjacent St.
+Lawrence Platform (where the geometric mean seismic ratio is 1.79).
+
+ [Any other reasons why Grenville and St. Lawrence Platform have higher R values than rest of Shield???]
+
+[ Continue with crustal composition discussion ]
+[ Note: I should include Platforms and Orogens in this paper: will work on this ]
+
+Other researchers have focused on Poisson's Ratio [Zandt and Ammon, 1995] as an indicator of bulk crustal composition as well as crustal thickness [Durrheim and Mooney, 1991] to constrain models of crustal formation.
+Durrheim and Mooney [1991] compile crustal thickness data and P-wave velocity data for crustal units divided into geochronological ages.
+They show that Proterozoic crust is thicker than Archean crust and has a higher P-wave velocity.
+The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as drastic as the greater than 5km difference reported by Durrheim and Mooney.
+Though $Vp$ data was not explicity estimated in this study, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for a thicker basal layer in Proterozoic crustal units.
+
+In contrast, Zandt and Ammon [1995] report high Poisson's Ratio values for Archean Shields and conclude that no systematic difference between Archaean and Proterozoic crust exists.
+Their average, transformed into $Vp/Vs$, is 1.84, much higher than the wieghted average for the Canadian Shield of 1.74 computed herein.
+The Authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust, 1.73 compared to the value calculated here of 1.77.
+Zandt and Ammon interpret these results showing high values for older crust and lower values for younger crust as indicating two end member scenarios.
+In the uniformitarian model, a delamination type process operates to remove mafic lower crust during continental collisions and is subsequently followed by a stabilizing process involving basaltic underplating.
+This implies that younger crust has not yet completed the crustal genesis cycle.
+In their second scenario, Precambrian cratons evolve through unique processes and younger orogens involve recycling and reworking existing material with little crustal growth.
+
+Whereas Zandt and Ammon show increasing $V_P/V_S$ values with crustal age the estimates computed in this study indicate an opposite trend, that is lower seismic velocity ratios with crustal age.
+This aligns more closely to the results of Durrheim and Mooney [1991] and seems to provide evidence for an evolution of crustal processes throughout geological time.
+Though Durrheim and Mooney accept the hypothesis that Proterozoic crust is more prone to basaltic underplating, thereby explaining higher velocities and thicker crust, Roberta Rudnick [1995; Rucknick and Gao, 2003] has argued persuasively that a range of possible processes may explain this data.
+
+[ Use crustal composition plus increasing Vp/Vs with younger crust result to make ``best'' pick among the 4 main processes Rudnick proposes to complete section.]
 
@@
 \section{bibliography}
-
-Bassin, C., G. Laske, G. Masters (2000), The Current Limits of Resolution for Surface Wave Tomography in North America, EOS Trans AGU, 81.
 
@@
 
+Laske, G., G. Masters, Z. Ma, M. Pasyanos (2013), Update on CRUST1.0 - A 1-degree Global Model of Earth's Crust, Geophys. Res. Abstracts, 15, Abstract EGU2013-2658.
+
 Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
@@
 Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+
+Rudnick, R. L. (1995), Making Continental Crust, Nature, 378, 571-578.
 

### 2013-09-09 python/compareResults.py

@@ -60,7 +60,7 @@ def rmsDifference(x, y):
     return np.sqrt(md/(n))
 
 
-for i in range(0,3):
+for i in range(1,2):
 
     pro = os.environ['HOME'] + data[2*i + 1]
     pub = os.environ['HOME'] + data[2*i]
@@ -98,6 +98,8 @@ for i in range(0,3):
 
     print "RMS Difference between", leglabel[i], "H datasets is {0:.2f}".format(rmsDifference(H1,H2)), "using", len(p2.stns), "stations"
     print "RMS Difference between", leglabel[i], "Vp/Vs datasets is {0:.3f}".format(rmsDifference(R1,R2)), "using", len(p2.stns), "stations"
+
+    print np.mean(R1), np.mean(R2)
 
     t = np.arange(len(R1))
 

### 2013-09-09 python/crust1json.py

@@ -22,6 +22,8 @@ def shape2json(shapeRecs):
         m[ix]["R"] = shapeRec.record[1]
         m[ix]["H"] = shapeRec.record[2]
         m[ix]["geoprov"] = shapeRec.record[3]
+        m[ix]["era"] = shapeRec.record[4].lower()
+        m[ix]["period"] = shapeRec.record[5].lower()
         (lon, lat) = lonlatbbox(shapeRec.shape.bbox)
         m[ix]["lat"] = lat
         m[ix]["lon"] = lon
@@ -34,7 +36,7 @@ def lonlatbbox(bbox):
 if __name__== '__main__' :
 
     c2jfile = os.environ['HOME'] + '/thesis/data/crust1.json'
-    sf = shapefile.Reader(os.environ['HOME'] + '/thesis/mapping/crust1/crust1geology')
+    sf = shapefile.Reader(os.environ['HOME'] + '/thesis/mapping/crust1GeoChron')
     shapeRecs = sf.shapeRecords()
 
     jdict = shape2json(shapeRecs)

### 2013-09-09 python/dbfExtract.py

@@ -45,41 +45,41 @@ import dbf
 
 
 ##Extracting Station Chrons
-# datafile = os.environ['HOME'] + "/thesis/data/stnChrons.json"
-epochdata = open(os.environ['HOME'] + '/thesis/data/epoch.json')
-epochdict = json.loads( epochdata.read() )
+datafile = os.environ['HOME'] + "/thesis/data/stnChrons.json"
+epochdata = open(datafile)
+stndict = json.loads( open(os.environ['HOME'] + "/thesis/data/stations.json").read() )
 
 ##     #Load geological time data
-# db = dbf.Table(os.environ['HOME'] + "/thesis/mapping/stationGeology.dbf")
+db = dbf.Table(os.environ['HOME'] + "/thesis/mapping/stationGeoChron.dbf")
 
- stnChrons = {}
-# for rec in db:
-#     stnChrons[rec["station"]] = {}
-#     stnChrons[rec["station"]]["lower"] = epochdict[rec["epoch"]][0]
-#     stnChrons[rec["station"]]["upper"] = epochdict[rec["epoch"]][1]
-#     stnChrons[rec["station"]]["era"] = rec["era"].lower()
+stnChrons = {}
+for rec in db:
+    stnChrons[rec[0]] = {}
+    stnChrons[rec[0]]["era"] = rec[11].lower()
+    stnChrons[rec[0]]["period"] = rec[12].lower()
+    stnChrons[rec[0]]["epoch"] = rec[13].lower()
 
 # open(datafile,'w').write( json.dumps(stnChrons, sort_keys = True, indent = 2 ))
-#print json.dumps(stnChrons, sort_keys = True, indent = 2 )
+print json.dumps(stnChrons, sort_keys = True, indent = 2 )
 
 
 # # Extracting VpmoonGeology data -> VpmoonShots.json
-moonvpGeology = os.environ['HOME'] + "/thesis/data/moonvpGeology.json"
-moonjson = os.environ['HOME'] + "/thesis/data/moonShots.json"
-d = json.loads( open(moonjson).read() )
+# moonvpGeology = os.environ['HOME'] + "/thesis/data/moonvpGeology.json"
+# moonjson = os.environ['HOME'] + "/thesis/data/moonShots.json"
+# d = json.loads( open(moonjson).read() )
 
-#Load geological time data
[…]

### 2013-09-09 python/histplots.py

@@ -167,7 +167,7 @@ def getdata(dtype, d, m, vdict, vdictregion):
             bins = np.arange(vpmin, vpmax, dvp)
             if vdictregion in vdict:
                 pt.avgmethod = "weighted"
-                avgd = vdict[vdictregion]['mooney']['Vp']
+                avgd = vdict[vdictregion]['crust1']['Vp']
             else:
                 avgd = np.mean(m.Vp)
     if pt.H:
@@ -184,7 +184,7 @@ def getdata(dtype, d, m, vdict, vdictregion):
             bins = np.arange(hmin, hmax, dh)
             if vdictregion in vdict:
                 pt.avgmethod = "weighted"
-                avgd = vdict[vdictregion]['mooney']['H']
+                avgd = vdict[vdictregion]['crust1']['H']
             else:
                 avgd = np.mean(m.H)
 
@@ -552,10 +552,10 @@ if __name__  == "__main__":
     if plotnum[5]:
         plt.figure( figsize = (width, height) )
         arg = Args()
-        arg.addQuery("wm::type", "in", "Archean")
+        arg.addQuery("geochron::period", "in", "archean")
         d2 = copy.deepcopy(d)
         arg2 = Args()
-        arg2.addQuery("wm::type", "in", "Proter")
+        arg2.addQuery("geochron::period", "in", "proterozoic")
 
         d.filter(arg)
         d2.filter(arg2)
@@ -583,20 +583,25 @@ if __name__  == "__main__":
 
         #plt.figure( figsize = (width, height) )
         ax = plt.subplot(122)
+
         arg = Args()
-        arg.addQuery("wm::type", "in", "Archean")
+        arg.addQuery("period", "in", "archean")
+
+        arg2 = Args()
+
+        crustfile = os.environ['HOME'] + '/thesis/data/crust1.json'
+        m = Params(crustfile, ["H","Vp"])
 
         m2 = copy.deepcopy(m)
 
-        arg2 = Args()
-        arg2.addQuery("wm::type", "in", "Proter")
-
         m.filter(arg)
-        m2.filter(arg2)
 
-        dm, avgdm, bins, ptype= getdata("moon", d, m, vdict, "Archean")
+        m2.filter(arg2.addQuery("period", "not in", "archean"))
+        m2.filter(arg2.addQuery("period", "in", "proterozoic"))
+
[…]

### 2013-09-09 python/lithologyPlots.py

@@ -17,7 +17,7 @@ from histplots import poisson
 
 
 stnfile = os.environ['HOME'] + '/thesis/data/stations.json'
-regionf = os.environ['HOME'] + '/thesis/data/voronoi.data'
+regionf = os.environ['HOME'] + '/thesis/data/voronoi.json'
 
 def lithVp(vs, poisson):
     return vs * np.sqrt( (2 - 2*poisson) / (1 - 2 * poisson)  )

### 2013-09-09 python/mergeJson.py

@@ -15,7 +15,7 @@ import json, os, argparse, sys
 
 
 # CONFIGS
-dbfile = os.environ['HOME'] + '/thesis/stations.json'
+dbfile = os.environ['HOME'] + '/thesis/data/stations.json'
 
 
 def mergeJson(target, source, pfx):
@@ -36,7 +36,7 @@ if __name__== '__main__' :
     group = parser.add_mutually_exclusive_group()
 
     # Create query parser
-    parser.add_argument('prefix', nargs = 1,
+    parser.add_argument('-p','--prefix', nargs = 1,
                         help = 'The prefix to use to to merge dictionary')
 
     parser.add_argument('mergeFile', nargs = '?',

### 2013-09-11 paper/draftBP.tex
Composition plots

@@
 
-Mean crustal composition is an important parameter in the literature studying the continental crust.
-Mean composition should not be confused as an estimate of real physical crustal lithological and mineral assemblages.
-Rather, mean crustal composition is a representative quantity that is useful for constraining crustal genesis and evolution in space and time.
-Early attempts at estimating mean crustal composition were largely accomplished through extrapolations of upper crustal rocks [Rudnick and Gao, 2003].
-After the plate-tectonic revolution, Taylor [1967, 1977], developed the 'island arc' or 'andesite' model which posited that bulk crustal composition should be equivalent to the average convergent-margin andesite.
-Several problems with this theory, such as the fact that island arcs are more basaltic than andesitic and therefore produce basaltic crustal accretions and cannot account for the intermediate composition of post-Archean crust, led to the introduction of seismic methods to estimate mid and lower crustal composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
-This study differs from these later efforts in that it follows more closely to the complete crustal seismic analysis in the work of Smithson et.
+Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state
+of the crust. Rather, mean crustal composition is a representative quantity that serves as a constraint on crustal structure and evolution in space and time. Early
+attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003].
+After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal composition is equivalent to the average convergent-margin andesite.
+Several problems with this theory subsequently emerged; for example, island arcs are more basaltic than andesitic and their accretions will lead to basaltic crustal growth and therefore cannot account for the intermediate composition of post-Archean crust.
+The recognition of these shortcomings led to the development of seismic methods to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
+Unlike these latter studies, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
 al.
-[1981].
+[1981] that combines $V_P/V_S$ from teleseismic measurements with contraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
 
@@
 
-The mean seismic velocity ratio computed in this study is combined with values for $V_P$ from the Crust 1.0 model to enable estimates for $V_P$ and $V_S$ to be used to estimate mean crustal composition.
-Figure \ref{fig:lith} shows mineral assemblages and weighted Canadian Shield estimates plotted against $V_P$ and $V_S$.
-Mineral assemblages are chosen as useful representations of common crustal rocks and, specifically, for crust in Shield-like regions.
-Examples of common shield lithologies can be seen in the 30km of paleo-depth represented in the Kapuskasing uplift in the Superior province [Percival and West, 1994].
-This uplift is largely comprised of Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed to greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites.
-Tonalites and granodiorites are also part of the TGG suite (for tonalite-trondhjemite-granodiorite) which dominate Archaean cratons [Rudnick, 1995].
-Seismic velocity values for these lithologies are taken from the work of Christensen [1996] and are selected at average crustal pressures of 600 MPa.
-Lines of constant Poisson's Ratio are plotted for reference.
+In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$.
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996].
+Dashed lines delimit isocontours in $V_P/V_S$.
+The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province [Percival and West, 1994].
+The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites.
+Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].
+Seismic velocity values for these lithologies are taken from the work of Christensen [1996] and are quoted at average crustal pressures of 600 MPa.
 
@@
 Lawrence Platform (where the geometric mean seismic ratio is 1.79).
+The average $V_P/V_S$ value for the Grenville Province is 1.77 marginally higher than the average value computed by the study of Eaton et.
+al.
+[2005].
 
- [Any other reasons why Grenville and St. Lawrence Platform have higher R values than rest of Shield???]
+We now take the
 
 [ Continue with crustal composition discussion ]
-[ Note: I should include Platforms and Orogens in this paper: will work on this ]
 
-Other researchers have focused on Poisson's Ratio [Zandt and Ammon, 1995] as an indicator of bulk crustal composition as well as crustal thickness [Durrheim and Mooney, 1991] to constrain models of crustal formation.
-Durrheim and Mooney [1991] compile crustal thickness data and P-wave velocity data for crustal units divided into geochronological ages.
-They show that Proterozoic crust is thicker than Archean crust and has a higher P-wave velocity.
-The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as drastic as the greater than 5km difference reported by Durrheim and Mooney.
-Though $Vp$ data was not explicity estimated in this study, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for a thicker basal layer in Proterozoic crustal units.
-
-In contrast, Zandt and Ammon [1995] report high Poisson's Ratio values for Archean Shields and conclude that no systematic difference between Archaean and Proterozoic crust exists.
-Their average, transformed into $Vp/Vs$, is 1.84, much higher than the wieghted average for the Canadian Shield of 1.74 computed herein.
+Other researchers have focused on Poisson's Ratio [Zandt and Ammon, 1995] as an indicator of bulk crustal composition as well as crustal thickness and P-wave velocity [Durrheim and Mooney, 1991] to constrain models of crustal formation.
+Zandt and Ammon [1995] report high $V_P/V_S$ values for Archean Shields and conclude that no systematic difference between Archaean and Proterozoic crust exists.
+Their average, 1.84, much higher than the wieghted average for the Canadian Shield of 1.74 computed herein.
 The Authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust, 1.73 compared to the value calculated here of 1.77.
@@
 Whereas Zandt and Ammon show increasing $V_P/V_S$ values with crustal age the estimates computed in this study indicate an opposite trend, that is lower seismic velocity ratios with crustal age.
-This aligns more closely to the results of Durrheim and Mooney [1991] and seems to provide evidence for an evolution of crustal processes throughout geological time.
+This aligns more closely to the results of Durrheim and Mooney [1991] who provide results showing a thick higher velocity Proterozoic crust compared to a thinner lower velocity Archean crust.
+The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as drastic as the greater than 5km difference reported by Durrheim and Mooney.
+Though $Vp$ data was not explicity estimated in this study, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
 Though Durrheim and Mooney accept the hypothesis that Proterozoic crust is more prone to basaltic underplating, thereby explaining higher velocities and thicker crust, Roberta Rudnick [1995; Rucknick and Gao, 2003] has argued persuasively that a range of possible processes may explain this data.
@@
 
+Smithson, S. B. (1978) Modeling continental crust-structural and chemical constraints, Geophys. Res. Lett. 5(9), 749-775
+
 Smithson, S. B., R. A. Johnson, Y. K. Wong (1981), Mean crustal velocity: a critical parameter for interpreting crustal structure and crustal growth, Earth and Planetary Science Letters, 53, 323-332.

### 2013-09-11 matlab/dataDriver.m

@@ -20,7 +20,7 @@ sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'STCO'; %
+station = 'MRHQ'; %
 
 %{
 

### 2013-09-11 python/lithologyPlots.py

@@ -23,10 +23,21 @@ def lithVp(vs, poisson):
     return vs * np.sqrt( (2 - 2*poisson) / (1 - 2 * poisson)  )
 
 
-left, width = .25, .5
-bottom, height = .25, .5
-right = left + width
-top = bottom + height
+def project(ax, ay, bx, by, cx, cy):
+    m = (by - ay) / (bx - ax)
+    y = m * (vs - ax) + ay
+    y2 = -m * (vs - cx) + cy
+
+    ix = 0.5 * ( (cy - ay)/m + ax + cx)
+    iy = m * (ix - ax) + ay
+
+    lengthA = np.sqrt((bx - ax)**2 + (by - ay)**2 )
+    lengthD = np.sqrt((bx - cx)**2 + (by - cy)**2 )
+
+    distp = 1 - lengthD / lengthA
+
+    return (ix, iy, distp)
+
 
 def addtext(ax, x, y, text, rotation):
     ax.text(x-0.02, y, text,
@@ -75,17 +86,11 @@ liths = {
         "R": 1.752,
         "poisson": 0.258
         },
-    "Diorite": {
+    "diorite": {
         "Vp": 6.611,
         "Vs": 3.733,
         "R": 1.771,
         "poisson": 0.266
-        },
-    "Granodiorite": {
-        "Vp": 6.327,
-        "Vs": 3.706,
-        "R": 1.707,
-        "poisson": 0.239
         }
 
     }
@@ -112,63 +117,156 @@ if __name__  == "__main__":
 
 
 
-    #############################################################################
-    # FIG 1: Vp estimates against H
-    ##############################################################################
 
-    if 1:
-        #f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdVp", "hk::stdR"])
-        fig = plt.figure( figsize = (figwidth, figheight) )
-        ax = plt.subplot(111)
+    fig = plt.figure( figsize = (figwidth, figheight) )
+    ax = plt.subplot(111)
 
-        vs = np.linspace(2.5, 5., 100)
-        sigmas = [0.3, 0.25, 0.2]
-        px = [3.7, 3.9, 4.1]
+    vs = np.linspace(2.5, 5., 100)
+    vpvs = [1.88, 1.73, 1.63]
+    px = [3.7, 3.9, 4.1]
 
-        rd = {}
+    rd = {}
 
 
-        with open(regionf, 'r') as f:
-            j = json.load(f)
-            for k in j.keys():
[…]

### 2013-09-12 paper/draftBP.tex
near end of discussion A

@@
 
+The estimates for Canada and the Canadian Shield are, to a first-order approximation, located between two candidate end-members, granite geniss and diorite.
+This intermediate or andesitic composition is in general agreement with other recent studies on crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
 The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in average composition.
@@
 
-We now take the
+\begin{table}
+  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} }
+    \hline
+    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work \\
+    \hline
+    $SiO_2$   & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 \\
+    $TiO_2$   & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 \\
+    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 \\
+    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 \\
+    $FeO$     & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 \\
+    $MnO$     & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
+    $MgO$     & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 \\
+    $CaO$     & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 \\
+    $Na_2O$   & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 \\
+    $K_2O$    & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 \\
+    $H_2O$    & --- & --- & 0.8 & --- & 1.1 \\
+    $P_2O_5$  & --- & 0.2 & 0.2 & 0.1 & 0.2 \\
+    $CO_2$    & --- & --- & --- & --- & 0.1 \\
+    \hline
+  \end{tabular}
+  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Fountain [1995] and Rudnick and Gao [2003] use a combined $Fe$ quantity and is listed under $FeO$}
+\label{table:composition}
 
-[ Continue with crustal composition discussion ]
+\end{table}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{composition}
+  \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}
+  \label{fig:composition}
+\end{figure}
+
+The $V_P$ and $V_S$ estimate for Canada lies on a line connecting diorite with granite gneiss (\ref{fig:lith}) and this can be used to estimate the bulk crustal composition.
+Using this analysis the bulk crustal lithology of the Canadian crust is 51\% diorite and 49\% granitic.
+The elemental breakdown is provided in table \ref{table:composition} along with previously published global crustal studies.
+The elemental crustal estimates produced in this study show a crust with marginally higher $Sio_2$ composition and lower $Fe$ and $Mg$ content.
+It is noteworthy that both studies involving greater reliance on seismic analysis, this study and the P-wave velocity study by Smithson [1978] closely follow this trend.
+Figure \ref{fig:composition} provides the major crustal elements from the data in table \ref{table:composition} but normalizes the estimates to the Candian crustal composition model produced here.
+Those estimates that are within within 30\% of model are highlighted by the gray shaded region.
+Also included in this figure is the crustal composition estimate for the Canadian Shield.
+The estimate for this Archean shield is much closer to the global crustal estimates of Christensen {\it et al.} and Smithson and indeed is bracketed by estimates from these two studies.
+These data suggest that the extensive Platform and Orogenic regions of Canada skew the bulk composition away from global averages and are comparatively better represented by granite gneiss lithologies than diorites.
+In contrast, the Canadian Shield crust shares a composition in greater alignment to the global averages which is best represented as 68\% dioritic and 32\% granitic in the formulation given above.
+Interestingly, the root-mean-squared error between the model proposed here and the other models presented is minimized by choosing a diorite component of 69\%, which results in an rms of 1.19\%.
+
+[[[ Can I actually say any of that?
+Isn't the data on elemental composition likely one of the main differentiators?
+The foundations seem a little weak, I don't want to start building castles, thoughts?
+]]]
 
@@
 This aligns more closely to the results of Durrheim and Mooney [1991] who provide results showing a thick higher velocity Proterozoic crust compared to a thinner lower velocity Archean crust.
-The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as drastic as the greater than 5km difference reported by Durrheim and Mooney.
+The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as evident as the greater than 5km difference reported by Durrheim and Mooney.
 Though $Vp$ data was not explicity estimated in this study, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
-Though Durrheim and Mooney accept the hypothesis that Proterozoic crust is more prone to basaltic underplating, thereby explaining higher velocities and thicker crust, Roberta Rudnick [1995; Rucknick and Gao, 2003] has argued persuasively that a range of possible processes may explain this data.
+The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney [1991], has been persuasively augmented by recent work [1995; Rucknick and Gao, 2003] to now include a range of possible processes that may explain the data.
 
-[ Use crustal composition plus increasing Vp/Vs with younger crust result to make ``best'' pick among the 4 main processes Rudnick proposes to complete section.]
+One possible process is the preferential recycling of soluble ions such as $Mg^{2+}$ and $Ca^{2+}$.
+These ions are weathered and carried into the oceans leaving behind silicon and aluminum [Rudnick and Gao, 2003] producing a net change to crustal composition over time.
+In another model, deep crustal roots expose lower crust to processes leading to eclogitization and subsequent delamination due to a density inversion with the underlying mantle peridoties.
+This model does not directly address the question as to why there is an apparent variance in average velocities between younger and older crusts, unless the delamination process has evolved over time.
+However, knoweldge of the spatial or temporal extent of delamination is difficult to document [Rudnick, 1995].
+A third model suggests that a mixture of silicic melts from subducting oceanic crust and basaltic melts from mantle peridotites were important in early crustal formation.
+This process is thought to be more common in the hotter Archean and is related to the TGG suites of Archean granitoids [Rudnick and Gao, 2003] and therefore provides a possible explanation for the data.
+It is important to state that all of these processes, as well as others, may have been operating throughout geologic time.
+
+[It is very possible that all of these as well as other processes have been operating thoughout geologic history, [Rudnick, 1995]]
+
+This study provides results that are in general agreement with other recent crustal studies and similarily argues for an intermediate or andesitic crust.
+There are several processes that have been suggested to explain this intermediate composition ...
+
+... Taking a break ...
 

### 2013-09-20 paper/draftBP.tex
interpolated stack plot

--- 2013-09-11 paper/draftBP.tex
@@
 
-Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state
-of the crust. Rather, mean crustal composition is a representative quantity that serves as a constraint on crustal structure and evolution in space and time. Early
-attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003].
+Seismic reflectivity of the continental crust is a valuable constraint on the processes governing crustal evolution.
+Research into the Mohoveric discontinuity and crustal thickness is ongoing and providing new insights into ...
+
+1. introductory paragraph along the lines of the one you did for crustal composition, identifying the
+importance of 1-D crustal structure for models of evolution. Introduce Conrad discontinuity and
+general seismic reflection subdivision of crust into more transparent upper half and reflective
+lower half, ie, summarize basic ideas from Brown EOS article on Conrad and Brocher and Mooney
+GJRAS article
+
+2. cite station sorted figure + perhaps normalized stack figure (as per above) and discuss where
+energy lies, most pronounced near surface and Moho, lower energy levels in bottom 1/3
+
+3. draw association between bottom 1/3 of low teleseismic reflectivity and high lower crustal  (seismic
+reflection) reflectivity, explain effect of frequency in different response
+
+4. discuss evidence for Conrad as distinct discontinuity (doesn't appear to be) versus the more
+general subdivision of upper and lower crusts with distinct properties.
+
+Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state of the crust.
+Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time.
+Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003].
 After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal composition is equivalent to the average convergent-margin andesite.
-Several problems with this theory subsequently emerged; for example, island arcs are more basaltic than andesitic and their accretions will lead to basaltic crustal growth and therefore cannot account for the intermediate composition of post-Archean crust.
-The recognition of these shortcomings led to the development of seismic methods to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
+Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
+Hence this model cannot account for the intermediate composition of post-Archean crust.
+Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
 Unlike these latter studies, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
@@
 In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$.
-Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996].
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996], all quoted at average crustal pressures of 600 MPa.
 Dashed lines delimit isocontours in $V_P/V_S$.
@@
 Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].
-Seismic velocity values for these lithologies are taken from the work of Christensen [1996] and are quoted at average crustal pressures of 600 MPa.
 
-The Slave, Superior and Churchill provinces share similar mean crustal compositions, with the Slave province being slightly more tonalitic in average composition.
-These three provinces cluster around the average for the Canadian Shield.
-The Grenville Province on the other hand differs significantly in mean composition.
-One interpretation is that the crust in the Grenville province has a mean crustal composition corresponding to a more mafic lower crustal composition than other Shield provinces.
+Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite.
+This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
+The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more tonalitic in average composition.
+All three provinces cluster about the average value for the Canadian Shield.
+The Grenville Province, in contrast, displays markedly lower $V_S$ and higher $V_P/V_S$ implying a significant difference in mean composition.
+The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study [Eaton et.
+al., 2005] that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces.
 This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally [Eaton et.
 al., 2005; Rivers, 2008].
-However, many stations in the South-East of the province also have seismic ratios greater than 1.78, similar to the high values common in the adjacent St.
+However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St.
 Lawrence Platform (where the geometric mean seismic ratio is 1.79).
-The average $V_P/V_S$ value for the Grenville Province is 1.77 marginally higher than the average value computed by the study of Eaton et.
-al.
-[2005].
+Moreover, and more importantly, the average $V_P$ values for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition.
+Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces.
+(NIK??
+any others?)
 
-We now take the
+\begin{table}
+  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} }
+    \hline
+    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work \\
+    \hline
+    $SiO_2$   & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 \\
+    $TiO_2$   & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 \\
+    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 \\
+    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 \\
+    $FeO$     & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 \\
+    $MnO$     & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
+    $MgO$     & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 \\
+    $CaO$     & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 \\
+    $Na_2O$   & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 \\
+    $K_2O$    & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 \\
+    $H_2O$    & --- & --- & 0.8 & --- & 1.1 \\
+    $P_2O_5$  & --- & 0.2 & 0.2 & 0.1 & 0.2 \\
+    $CO_2$    & --- & --- & --- & --- & 0.1 \\
+    \hline
+  \end{tabular}
+  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Fountain [1995] and Rudnick and Gao [2003] use a combined $Fe$ quantity and is listed under $FeO$}
+\label{table:composition}
 
-[ Continue with crustal composition discussion ]
+\end{table}
 
-Other researchers have focused on Poisson's Ratio [Zandt and Ammon, 1995] as an indicator of bulk crustal composition as well as crustal thickness and P-wave velocity [Durrheim and Mooney, 1991] to constrain models of crustal formation.
-Zandt and Ammon [1995] report high $V_P/V_S$ values for Archean Shields and conclude that no systematic difference between Archaean and Proterozoic crust exists.
-Their average, 1.84, much higher than the wieghted average for the Canadian Shield of 1.74 computed herein.
-The Authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust, 1.73 compared to the value calculated here of 1.77.
-Zandt and Ammon interpret these results showing high values for older crust and lower values for younger crust as indicating two end member scenarios.
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{composition}
+  \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}
+  \label{fig:composition}
+\end{figure}
+
+Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite.
+The corresponding chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
+The chemical estimates produced in our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies.
+It is noteworthy that that the two studies involving greatest reliance on seismic attributes, ours and the {\it P}-wave velocity study by Smithson [1978], bear the closest similarity.
+Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition.
+The gray-shaded region indicates the $\pm 30\%$ range from our reference average Canadian crustal model.
+Also included in this figure is the our average crustal composition estimate for the Canadian Shield.
+Its values closely track the global crustal estimates of Christensen \& Mooney [1995] and Smithson [1978].
+These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
+In contrast, the Canadian Shield crust aligns more closely with these previous global estimates and is represented as 68\% dioritic and 32\% granitic using the two end-member model.
+
+Other researchers have employed Poisson's ratio [Zandt and Ammon, 1995] alone as an indicator of bulk crustal composition, and crustal thickness and {\it P}-wave velocity [Durrheim and Mooney, 1991] to constrain models of crustal formation.
+Zandt and Ammon [1995] report high $V_P/V_S$ values for Archean Shields and concluded that no systematic difference between Archean and Proterozoic crust exists.
+Their average, $V_P/V_S=1.84$, is much higher than our weighted average for the Canadian Shield of 1.74.
+These authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust, 1.73 compared to the corresponding value for our data of 1.77.
+They interpret these high values for older crust and low values for younger crust as favouring either of two limiting models.
 In the uniformitarian model, a delamination type process operates to remove mafic lower crust during continental collisions and is subsequently followed by a stabilizing process involving basaltic underplating.
-This implies that younger crust has not yet completed the crustal genesis cycle.
-In their second scenario, Precambrian cratons evolve through unique processes and younger orogens involve recycling and reworking existing material with little crustal growth.
+This model implies that younger crust has yet to complete the full crustal genesis cycle.
+In their second model, Precambrian cratons evolve through unique processes and younger orogens involve the recycling and reworking of existing material with little net crustal growth.
 
-Whereas Zandt and Ammon show increasing $V_P/V_S$ values with crustal age the estimates computed in this study indicate an opposite trend, that is lower seismic velocity ratios with crustal age.
-This aligns more closely to the results of Durrheim and Mooney [1991] who provide results showing a thick higher velocity Proterozoic crust compared to a thinner lower velocity Archean crust.
-The crustal thickness data attained in this study also show thicker crusts for Proterozoic units (38.9 km) than Archean units (38.0 km) though this difference is not as drastic as the greater than 5km difference reported by Durrheim and Mooney.
-Though $Vp$ data was not explicity estimated in this study, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
-Though Durrheim and Mooney accept the hypothesis that Proterozoic crust is more prone to basaltic underplating, thereby explaining higher velocities and thicker crust, Roberta Rudnick [1995; Rucknick and Gao, 2003] has argued persuasively that a range of possible processes may explain this data.
+Whereas Zandt and Ammon present evidence for increasing $V_P/V_S$ with crustal age, our estimates signify the opposite trend.
+Our results are more consistent with those of Durrheim and Mooney [1991] who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
+The crustal thickness data presented in this study is only marginally thicker for Proterozoic crust (38.9 km) than Archean crust (38.0 km) and this difference is significantly smaller than the 5 km difference reported by Durrheim and Mooney.
+Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
+The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney [1991], has been persuasively augmented by recent work [Rudnick 1995; Rucknick and Gao, 2003] to now include a range of possible processes that may explain the data.
 
-[ Use crustal composition plus increasing Vp/Vs with younger crust result to make ``best'' pick among the 4 main processes Rudnick proposes to complete section.]
+One possible process is the preferential recycling of soluble ions such as $Mg^{2+}$ and $Ca^{2+}$.
+These ions are weathered and carried into the oceans leaving behind silicon and aluminum [Rudnick and Gao, 2003] producing a net change to crustal composition over time.
+In another model, deep crustal roots expose lower crust to processes leading to eclogitization and subsequent delamination due to a density inversion with the underlying mantle peridoties.
+This model does not directly address the question as to why there is an apparent variance in average velocities between younger and older crusts, unless the delamination process has evolved over time.
+However, knowledge of the spatial or temporal extent of delamination is difficult to document [Rudnick, 1995].
+A third model suggests that a mixture of silicic melts from subducting oceanic crust and basaltic melts from mantle peridotites were important in early crustal formation.
+This process is thought to be more common in the hotter Archean and is related to the TGG suites of Archean granitoids [Rudnick and Gao, 2003] and therefore provides a possible explanation for the data.
+It is important to state that all of these processes, as well as others, may have been operating throughout geologic time.
+
+This study provides results that are in general agreement with other recent crustal studies and similarly argues for an intermediate or andesitic crust.
+There are several processes that have been suggested to explain this intermediate composition ...
 

### 2013-09-20 matlab/conradPCA.m

@@ -24,7 +24,7 @@ for ii = 1 : length(s)
     
     dbfile = fullfile(databasedir, [station,'.mat'] );
     
-    if  numel(strfind(json.(station).status, 'processed-ok'))
+    if  numel(strfind(json.(station).status, 'processed'))
         if exist(dbfile, 'file')
             disp(station)
             load(dbfile)
@@ -153,7 +153,7 @@ hwin2 = H(end);
 
 win1 = find(H >= hwin1, 1, 'first');
 win2 = find(H <= hwin2, 1, 'last');
-
+%{
 for ii = 1:length(stations)
     mask(ii, ii) = S(ii, ii);
     
@@ -172,6 +172,92 @@ for ii = 1:length(stations)
         title(sprintf('adding %i modes capturing %2.1f%% variance', ii, var))
     pause()
 end
+%}
+%% Interpolate and Stretch
+
+midx = zeros(1, length(moho));
+
+ninterp = 500;
+
+xi = 1:500; % Set for xor'ing later
+
+XI = zeros(length(stations), ninterp);
+
+maxq = 1;
+
+for qq = 1:maxq
+    
+    for ii = 1:length(moho)
+        midx(ii) = find(H >= moho(ii), 1, 'first');
+    end
+
+    for ii = 1:length(midx)
+
+        ntoshift = ninterp - midx(ii);
+        x = 1:midx(ii); % Vector representing indicies of profile
+        vx = X(ii, 1:midx(ii));
+
+        for jj = 1:ntoshift
+            ridx = randi([2, midx(ii) - 1], 1);
+            x(ridx:end) = x(ridx:end) + 1; % Shift indices one for each insert
+        end
+
+        qx = setxor(x, xi);
+
+        vq = interp1(x, vx, qx);
+
+        XI(ii, x) = XI(ii, x) + vx;
+        XI(ii, qx) = XI(ii, qx) + vq;
+
+    end
+end
+
+
+profile =  sum(sqrt(XI.^2),1)';
+profile =  sum(abs(XI),1)';
+
+profile = profile;
+
+figure(23)
+h(1) = subplot(1,2,1);
+    imagesc(XI')
+
+    set(gca, 'XTick', []);
+    set(gca, 'YTick', []);
+    set(gca,'YTickLabel','')
+    set(gca,'XTickLabel','')
[…]

### 2013-09-20 matlab/dataDriver.m

@@ -20,7 +20,7 @@ sfile = 'stack_S.sac';
 load stnsjson.mat
 %%  Select Station to Process and load station data
 method = 'kanamori';
-station = 'MRHQ'; %
+station = 'ULM'; %
 
 %{
 

### 2013-09-21 paper/draftBP.tex
discontinuity discussion

@@
 %% -----------------------------------------------%%
+
 \section{Data and methods}
@@
 %% -----------------------------------------------%%
+
 \section{Results}
@@
 
-\subsection{Crustal Discontinuities}
+\subsection{Crustal Discontinuities} \label{section:discontinuities}
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{discontinuities}
+  \caption{}
+  \label{fig:disc}
+\end{figure}
+
 We now consider the average crustal structure across Canada.
 We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method.
-The resulting profiles contain energy at depths corresponding to major velocity discontinuities.
+The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:disc} arranged by increasing crustal thickness.
 By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components.
@@
 Seismic reflectivity of the continental crust is a valuable constraint on the processes governing crustal evolution.
-Research into the Mohoveric discontinuity and crustal thickness is ongoing and providing new insights into ...
+Early efforts to characterize continental crust revolved around a slab model with a lower velocity upper crust and a higher velocity lower crust [Conrad, 1925; Jeffreys, 1926].
+Evidence for this model was originally put forth by Conrad [1925] who suggested observable P arrivals were propogating along a mid-crustal interface similar to the deeper crust-mantle interface imaged by Andrija Mohorovičić [Litak and Brown, 1989].
+Both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence [Litak and Brown, 1989].
+The development of reflection seismology allowed workers to further argue for a complex heterogenous crust and propose a model whereby smooth random lateral inhomogenities superimposed on a linearly increasing velocity gradient could produce seismic responses similar to that of a simple two layer model [Litak and Brown, 1989; Mereu and Ojo, 1981].
 
-1. introductory paragraph along the lines of the one you did for crustal composition, identifying the
-importance of 1-D crustal structure for models of evolution. Introduce Conrad discontinuity and
-general seismic reflection subdivision of crust into more transparent upper half and reflective
-lower half, ie, summarize basic ideas from Brown EOS article on Conrad and Brocher and Mooney
-GJRAS article
+Despite the fading support for the Conrad Discontinuity as a discrete geological phenomena the generalization of the crust into seismically distinguishable upper and lower components continued.
+Mooney and Brocher [1987] argued for a model of the continental crust seperated into a seismically reflective lower crust overlain by transparent upper crust.
+Their physical interpretion is of a poorly laminated upper crust dominated by features of short length scale contrasted with a lower crust composed of thin (100 metre) lamellae of significant length scales.
+Supporting evidence for a reflective lower crust [Ito, 1998; Mayer, 1996; Meissner et.
+al., 2006] indicate that this is likely regional and representational of extented as well as thicker warmer crusts.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{discontinuityInterp}
+  \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
+  \label{fig:discInterp}
+\end{figure}
+
+The lower energy noted in Section \ref{section:discontinuities} in the 20 to 30 km depth window for the station profiles given in figure \ref{fig:disc} provides some evidence for seismically distinguishable crustal layers.
+To investigate this further we stretch and interpolate the profiles so that the Moho of each profile lines up on a common axis (Figure \ref{fig:discInterp}, left) and then take the magnitude of each profile and stack (Figure \ref{fig:discInterp}, right).
+The energy of the stack is most pronounced near the surface, associated with sedimentary layers, and at the Moho.
+There is a marked decrease in energy in the bottom third of the profile.
+
+(Jackson 2002 - strength of lithosphere - mirrors decrease in energy)
 
@@
 
+[[ADD GRANITE DIORITE REFERENCE]]
+
 Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.

### 2013-09-22 paper/draftBP.tex
discussion draft

@@
   \centering
-  \includegraphics[width=\textwidth]{discontinuities}
+  \includegraphics[width=\textwidth]{discontinuity}
   \caption{}
@@
 Early efforts to characterize continental crust revolved around a slab model with a lower velocity upper crust and a higher velocity lower crust [Conrad, 1925; Jeffreys, 1926].
-Evidence for this model was originally put forth by Conrad [1925] who suggested observable P arrivals were propogating along a mid-crustal interface similar to the deeper crust-mantle interface imaged by Andrija Mohorovičić [Litak and Brown, 1989].
+Evidence for this model was originally put forth by Conrad [1925] who suggested observable P arrivals were propagating along a mid-crustal interface similar to the deeper crust-mantle interface imaged by Andrija Mohorovičić [Litak and Brown, 1989].
 Both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence [Litak and Brown, 1989].
-The development of reflection seismology allowed workers to further argue for a complex heterogenous crust and propose a model whereby smooth random lateral inhomogenities superimposed on a linearly increasing velocity gradient could produce seismic responses similar to that of a simple two layer model [Litak and Brown, 1989; Mereu and Ojo, 1981].
+The development of reflection seismology allowed workers to further argue for a complex heterogeneous crust and propose a model whereby smooth random lateral inhomogeneities superimposed on a linearly increasing velocity gradient could produce seismic responses similar to that of a simple two layer model [Litak and Brown, 1989; Mereu and Ojo, 1981].
 
 Despite the fading support for the Conrad Discontinuity as a discrete geological phenomena the generalization of the crust into seismically distinguishable upper and lower components continued.
-Mooney and Brocher [1987] argued for a model of the continental crust seperated into a seismically reflective lower crust overlain by transparent upper crust.
-Their physical interpretion is of a poorly laminated upper crust dominated by features of short length scale contrasted with a lower crust composed of thin (100 metre) lamellae of significant length scales.
-Supporting evidence for a reflective lower crust [Ito, 1998; Mayer, 1996; Meissner et.
-al., 2006] indicate that this is likely regional and representational of extented as well as thicker warmer crusts.
+Mooney and Brocher [1987] argued for a model of the continental crust separated into a seismically reflective lower crust overlain by transparent upper crust.
+Their physical interpretation is of a poorly laminated upper crust dominated by features of short length scale contrasted with a lower crust composed of thin (100 metre) lamellae of significant length scales.
+Supporting evidence for a reflective lower crust confirms widespread laminated structure in the lower crust with thicknesses of 100m to 300m though these structures have been found to be localized and representational of extended as well as thicker warmer crusts [Ito, 1998; Mayer, 1996; Meissner et.
+al., 2006].
 
@@
 
-The lower energy noted in Section \ref{section:discontinuities} in the 20 to 30 km depth window for the station profiles given in figure \ref{fig:disc} provides some evidence for seismically distinguishable crustal layers.
+The lower energy noted in Section \ref{section:discontinuities} in the 20km to 30km depth window for the station profiles given in figure \ref{fig:disc} provides some evidence for seismically distinguishable crustal layers.
 To investigate this further we stretch and interpolate the profiles so that the Moho of each profile lines up on a common axis (Figure \ref{fig:discInterp}, left) and then take the magnitude of each profile and stack (Figure \ref{fig:discInterp}, right).
@@
 There is a marked decrease in energy in the bottom third of the profile.
+This study employs S-wave energy using both direct and reflected arrivals and it has been shown that these waves interact with velocity gradients on scales lengths in the order of $\lambda / 2$ for direct phases and $\lambda / 4$ for reflected phases where $\lambda$ is wavelength [Bostock, 1999].
+For 1Hz teleseismic S-waves crustal structure in the scale lengths of 750m to 1500m is resolvable.
+Higher frequency active source data (10Hz - 40Hz) employing P-wave sources would be sensitive to structure roughly in the ranges of 75m to 300m.
+Lithoprobe transects [Hammer et.
+al., 2010] show significant heterogeneity throughout the crust including several transects, such as the Western Superior transect, that appear to contain higher reflectivity in the lower crust.
+One interpretation for these data, taken together, is that in general the Canadian lower crustal structure contains fewer thick (~1km) velocity contrasts as compared to the upper crust.
+This interpretation allows for high reflectivity in the lower crust arising from thinner, less than 750m, contrasts.
 
-(Jackson 2002 - strength of lithosphere - mirrors decrease in energy)
+Though the Conrad Discontinuity continues to be invoked, for example in a study of the Rhine Graben [Mayer, 1996], as terminology to describe the interface between a transparent upper crust and reflective lower crust, it does not appear to be a prevalent feature of the Canadian crust.
+The results from this study do suggest however, that in general, and with the seismic methods employed here, a distinction between the seismic properties of the upper crust and the bottom third of the crust can be resolved.
+Studies on crustal strength [Jackson, 2002] and on brittle and ductile models [Burov and Watts, 2006; Meisner and Rabbel, 1999] as well as the widely noted reduced seismicity of the lower crust indicate changes in crustal processes at similar depths, though see [Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999] for competing evidence.
 
-2. cite station sorted figure + perhaps normalized stack figure (as per above) and discuss where
-energy lies, most pronounced near surface and Moho, lower energy levels in bottom 1/3
-
-3. draw association between bottom 1/3 of low teleseismic reflectivity and high lower crustal  (seismic
-reflection) reflectivity, explain effect of frequency in different response
-
-4. discuss evidence for Conrad as distinct discontinuity (doesn't appear to be) versus the more
-general subdivision of upper and lower crusts with distinct properties.
-
-Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state of the crust.
+Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
 Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time.
@@
 al.
-[1981] that combines $V_P/V_S$ from teleseismic measurements with contraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
+[1981] that combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
 

### 2013-10-10 paper/draftBP.tex
add Nik and MB edits

--- 2013-09-20 paper/draftBP.tex
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{discontinuity}
+  \caption{}
+  \label{fig:disc}
+\end{figure}
+
 We now consider the average crustal structure across Canada.
 We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method.
-The resulting profiles contain energy at depths corresponding to major velocity discontinuities.
+The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:disc} arranged by increasing crustal thickness.
 By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components.
@@
 
-Seismic reflectivity of the continental crust is a valuable constraint on the processes governing crustal evolution.
-Research into the Mohoveric discontinuity and crustal thickness is ongoing and providing new insights into ...
+\section{Discussion}
 
-1. introductory paragraph along the lines of the one you did for crustal composition, identifying the
-importance of 1-D crustal structure for models of evolution. Introduce Conrad discontinuity and
-general seismic reflection subdivision of crust into more transparent upper half and reflective
-lower half, ie, summarize basic ideas from Brown EOS article on Conrad and Brocher and Mooney
-GJRAS article
+Seismic refraction and reflection studies of continental crust have afforded valuable constraints on the processes governing crustal evolution.
+Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust [Conrad, 1925; Jeffreys, 1926].
+Evidence supporting this model was originally presented by Conrad [1925] who cited observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovičić in 1909 [Litak and Brown, 1989].
+Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence [Litak and Brown, 1989].
+Technological improvements in reflection seismology allowed researchers to further argue for a complex heterogeneous crust and propose models wherein smooth random lateral inhomogeneities are superimposed on simple velocity gradient models to reproduce the range of seismic observations [Litak and Brown, 1989; Mereu and Ojo, 1981].
 
-2. cite station sorted figure + perhaps normalized stack figure (as per above) and discuss where
-energy lies, most pronounced near surface and Moho, lower energy levels in bottom 1/3
+Despite diminishing support in recent years for the Conrad Discontinuity as a universal crustal feature, the division of the crust into seismically distinguishable upper and lower components remains a common assertion.
+In their review, Mooney and Brocher [1987] argued for a model of the continental crust separated intoa seismically reflective lower crust overlain by transparent upper crust.
+Their physical interpretation is that of a poorly laminated upper crust dominated by features of short length scale that contrasts with a lower crust composed of thin ($\sim$ 100 m) lamellae of significantly longer horizontal length scale.
+The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust [e.g.
+Ito, 1998; Mayer, 1996; Meissner et.
+al., 2006].
 
-3. draw association between bottom 1/3 of low teleseismic reflectivity and high lower crustal  (seismic
-reflection) reflectivity, explain effect of frequency in different response
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{discontinuityInterp}
+ \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
+ \label{fig:discInterp}
+\end{figure}
 
-4. discuss evidence for Conrad as distinct discontinuity (doesn't appear to be) versus the more
-general subdivision of upper and lower crusts with distinct properties.
+The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in figure \ref{fig:disc} provides some supporting evidence for seismically distinct upper and lower crustal layers.
+To better compare results across stations, we normalize each depth profile such that the Moho falls on a common ordinate (Figure \ref{fig:discInterp}, left) and stack the absolute values (Figure \ref{fig:discInterp}, right).
+The stacked signal has highest amplitude near the surface, where energy is associated with sedimentary layers, and at the Moho.
+There is a steady and marked decrease in energy over the bottom third of the profile that cannot be ascribed to sidelobe structure associated with the Moho pulse.
+This observation is ostensibly at odds with the Mooney and Brocher [1987] high reflectivity lower crustal model.
+Moreover, reflection profiles from {\sc Lithoprobe} transects [Hammer et.
+al., 2010] frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced pronounced reflectivity within the lower crust.
+We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive.
+Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength [Bostock, 1999].
+For 1 Hz teleseismic waves, crustal layering with scale lengths of 1000-2000 m should be resolvable.
+Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of 75-300 m.
+We infer then that the vertical scales of lower crustal reflectivity are limited to scale lengths under 100 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.
+Some authors, for example Mayer et .al.
+[1996] in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust.
+It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such
+signals, but the lower energy levels ($<5\%$) in this mode, would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
+The increased transparency of the lower crust would appear to be better defined and would appear to relate to changes in rheology.
+Studies on crustal strength [Jackson, 2002] and on brittle and ductile models [Burov and Watts, 2006; Meisner and Rabbel, 2006] as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention, though see [Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999] for competing evidence.
 
-Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state of the crust.
+Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
 Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time.
 Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003].
-After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal composition is equivalent to the average convergent-margin andesite.
+After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
 Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
@@
 Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
-Unlike these latter studies, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
+Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
 al.
-[1981] that combines $V_P/V_S$ from teleseismic measurements with contraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
+[1981] though whereas Smithson utilizes mean crustal $V_P$ this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
 
@@
 The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province [Percival and West, 1994].
-The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites.
+The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures comprised largely of tonalites and their metamorphosed equivalents.
 Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].
@@
 This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
-The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more tonalitic in average composition.
+The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
 All three provinces cluster about the average value for the Canadian Shield.
@@
 Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces.
-(NIK??
-any others?)
+[[Double checkout less quartzite argument, this implies higher Vp ...
+but the data does not show it ]]
 
 \begin{table}
-  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} }
+  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
     \hline
-    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work \\
+    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work: Canada & This Work: Shield\\
     \hline
-    $SiO_2$   & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 \\
-    $TiO_2$   & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 \\
-    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 \\
-    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 \\
-    $FeO$     & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 \\
-    $MnO$     & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
-    $MgO$     & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 \\
-    $CaO$     & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 \\
-    $Na_2O$   & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 \\
-    $K_2O$    & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 \\
-    $H_2O$    & --- & --- & 0.8 & --- & 1.1 \\
-    $P_2O_5$  & --- & 0.2 & 0.2 & 0.1 & 0.2 \\
-    $CO_2$    & --- & --- & --- & --- & 0.1 \\
+    $SiO_2$ & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 & 61.9 \\
+    $TiO_2$ & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 & 0.7 \\
+    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 & 15.9 \\
+    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 & 2.1 \\
+    $FeO$ & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 & 3.9 \\
+    $MnO$ & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
+    $MgO$ & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 & 2.8 \\
+    $CaO$ & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 & 5.1 \\
+    $Na_2O$ & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 & 3.6 \\
+    $K_2O$ & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 & 2.5 \\
+    $H_2O$ & --- & --- & 0.8 & --- & 1.1 & 1.2 \\
+    $P_2O_5$ & --- & 0.2 & 0.2 & 0.1 & 0.2 & 0.2 \\
+    $CO_2$  & --- & --- & --- & --- & 0.1 & 0.1 \\
     \hline
@@
 Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite.
-The corresponding chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
-The chemical estimates produced in our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies.
+The lithological estimate for the Canadian Shield is 68\% diorite and 32\% granite.
+This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre [1976] to produce an estimate of bulk crustal composition.
+The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
+Our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies.
 It is noteworthy that that the two studies involving greatest reliance on seismic attributes, ours and the {\it P}-wave velocity study by Smithson [1978], bear the closest similarity.
@@
 These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
-In contrast, the Canadian Shield crust aligns more closely with these previous global estimates and is represented as 68\% dioritic and 32\% granitic using the two end-member model.
+In contrast, the Canadian Shield crust aligns more closely with these previous global estimates.
 
@@
 
+[[ADD GRANITE DIORITE REFERENCE]]
+
 Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.
 
+Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophysical Journal International, 138, 747-756.
+
 Bostock, M. G., M. R. Kumar (2010), Bias in seismic estimates of crustal properties, J. Geophys. Int., 182, 403-407.
+
+Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere:" jelly sandwich" or" crème brûlée"?, GSA today, 16, 4.
 
@@
 Christensen, N. I., W. D. Mooney (1995) Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
+
+Conrad, V. (1925), Laufzeitkurven des Tauernbebens vom 28. November 1923, Mitt. Erdb. Komm. Wiener Akad. Wiss., 59, 1–23.
 
@@
 
+Deichmann, N. (1992), Structural and rheological implications of lower-crustal earthquakes below northern Switzerland, Physics of the earth and planetary interiors, 69, 270-280.
+
 Durrheim, R. J., W. D. Mooney (1991), Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
@@
 
+Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
+
+Hammer, P.
+T., R.
+M.
+Clowes, F.
+A.
+Cook, A.
+J.
+van der Velden, K.
+Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Canadian Journal of Earth Sciences, 47, 821-857.
+
+Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.
+
+Ito, K. (1999), Seismogenic layer, reflective lower crust, surface heat flow and large inland earthquakes, Tectonophysics, 306, 423-433.
+
+Jackson, J. (2002), Strength of the continental lithosphere: Time to abandon the jelly sandwich?, GSA today, 12, 4-9.
+
+Jeffreys, H. (1926), On compressional waves in two superposed layers, Proc. Cambridge Philos. Soc., 23, 472–481.
+
+Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
+
+Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
+
 Laske, G., G. Masters, Z. Ma, M. Pasyanos (2013), Update on CRUST1.0 - A 1-degree Global Model of Earth's Crust, Geophys. Res. Abstracts, 15, Abstract EGU2013-2658.
 
-Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
+Litak, R. K., L. D. Brown (1989), A modern perspective on the Conrad discontinuity, Eos, 70, 713-725.
+
+Mayer, G., P.
+M.
+Mai, T.
+Plenefisch, H.
+Echtler, E.
+Lüschen, V.
+Wehrle, B.
+Müller, K-P.
+Bonjer, C.
+Prodehl, K.
+Fuchs (1997), The deep crust of the Southern Rhine Graben: reflectivity and seismicity as images of dynamic processes, Tectonophysics, 275, 15-40.
+
+Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.
+
+Mereu, R.
+F., S.
+B.
+Ojo (1981), The scattering of seismic waves through a crust and upper mantle with random lateral and vertical inhomogeneities, Physics of the Earth and Planetary Interiors, 26, 233-240.
+
 Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.
 
-Kennet, B. L. N., (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
-
-Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
+Mooney, W. D., T. M. Brocher (1987) Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.
 
@@
 Rudnick, R. L., S. Gao (2003), Composition of the Continental Crust, Treatise on Geochem., 3, 1-64.
+
+Simpson, F. (1999), Stress and seismicity in the lower continental crust: A challenge to simple ductility and implications for electrical conductivity mechanisms. Surveys in Geophysics, 20, 201-227.
 

### 2013-11-11 paper/draftBP.tex
nik edits

@@
 Moreover, reflection profiles from {\sc Lithoprobe} transects [Hammer et.
-al., 2010] frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced pronounced reflectivity within the lower crust.
+al., 2010] frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
 We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive.
@@
 [1996] in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust.
-It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such
-signals, but the lower energy levels ($<5\%$) in this mode, would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
+It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
+This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by Rolf Meisnner [Meisnner, 1986].
+It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data [Hammer et.
+al., 2010].
 The increased transparency of the lower crust would appear to be better defined and would appear to relate to changes in rheology.
@@
 Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite.
-This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
+This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition e.g., [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
 The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
@@
     \hline
-    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work: Canada & This Work: Shield\\
+    Oxide, wt\% & Pakiser and Robinson [1966] & Smithson [1978] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work: Canada & This Work: Shield\\
     \hline
-    $SiO_2$ & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 & 61.9 \\
-    $TiO_2$ & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 & 0.7 \\
-    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 & 15.9 \\
-    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 & 2.1 \\
-    $FeO$ & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 & 3.9 \\
-    $MnO$ & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
-    $MgO$ & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 & 2.8 \\
-    $CaO$ & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 & 5.1 \\
-    $Na_2O$ & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 & 3.6 \\
-    $K_2O$ & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 & 2.5 \\
-    $H_2O$ & --- & --- & 0.8 & --- & 1.1 & 1.2 \\
-    $P_2O_5$ & --- & 0.2 & 0.2 & 0.1 & 0.2 & 0.2 \\
-    $CO_2$  & --- & --- & --- & --- & 0.1 & 0.1 \\
+    $SiO_2$ & 57.9 & 63.0 & 61.7 & 60.6 & 64.3 & 61.9 \\
+    $TiO_2$ & 1.2 & 0.7 & 0.9 & 0.7 & 0.6 & 0.7 \\
+    $Al_2O_3$ & 15.2 & 15.8 & 14.7 & 15.9 & 15.5 & 15.9 \\
+    $Fe_2O_3$ & 2.3 & 2.0 & 1.9 & --- & 1.9 & 2.1 \\
+    $FeO$ & 5.5 & 3.4 & 5.1 & 6.7 & 3.3 & 3.9 \\
+    $MnO$ & 0.2 & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
+    $MgO$ & 5.3 & 2.8 & 3.1 & 4.7 & 2.2 & 2.8 \\
+    $CaO$ & 7.1 & 4.6 & 5.7 & 6.4 & 4.3 & 5.1 \\
+    $Na_2O$ & 3.0 & 4.0 & 3.6 & 3.1 & 3.6 & 3.6 \\
+    $K_2O$ & 2.1 & 2.7 & 2.1 & 1.8 & 2.9 & 2.5 \\
+    $P_2O_5$ & 0.3 & --- & 0.2 & 0.1 & 0.2 & 0.2 \\
     \hline
   \end{tabular}
-  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Fountain [1995] and Rudnick and Gao [2003] use a combined $Fe$ quantity and is listed under $FeO$}
+  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Gao [2003] use total $Fe$ expressed as $FeO$.}
 \label{table:composition}
@@
 Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite.
-The lithological estimate for the Canadian Shield is 68\% diorite and 32\% granite.
+A similar estimate for the Canadian Shield is 68\% diorite and 32\% granite.
 This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre [1976] to produce an estimate of bulk crustal composition.
 The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
-Our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies.
-It is noteworthy that that the two studies involving greatest reliance on seismic attributes, ours and the {\it P}-wave velocity study by Smithson [1978], bear the closest similarity.
+Our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $MgO$ content than these previous studies.
 Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition.
@@
 
+Meissner, R., (Ed.) (1986) \it{The Continental Crust: A Geophysical Approach}, 426, Academic, San Diego, Calif.
+
 Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.

### 2013-11-11 python/lithologyPlots.py

@@ -53,43 +53,57 @@ liths = {
         "Vp": 6.208,
         "Vs": 3.583,
         "R": 1.732,
-        "poisson": 0.250
+        "poisson": 0.250,
+        "x": -0.08,
+        "y": 0.0
         },
     "tonalite gneiss": {
         "Vp": 6.302,
         "Vs": 3.606,
         "R": 1.747,
-        "poisson": 0.257
+        "poisson": 0.257,
+        "x": 0.11,
+        "y": 0.
         },
     "mafic granulite": {
         "Vp": 6.942,
         "Vs": 3.820,
         "R": 1.817,
-        "poisson": 0.283
+        "poisson": 0.283,
+        "x": -0.08,
+        "y": 0.05
         },
     "mafic garnet granulite": {
         "Vp": 7.249,
         "Vs": 4.026,
         "R": 1.801,
-        "poisson": 0.277
+        "poisson": 0.277,
+        "x": -0.15,
+        "y": 0.0
         },
     "amphibolite": {
         "Vp": 6.983,
         "Vs": 3.959,
         "R": 1.764,
-        "poisson": 0.263
+        "poisson": 0.263,
+        "x": 0.,
+        "y": 0.05
         },
     "hornblendite": {
         "Vp": 7.261,
         "Vs": 4.144,
         "R": 1.752,
-        "poisson": 0.258
+        "poisson": 0.258,
+        "x": 0.,
+        "y": 0.05
         },
     "diorite": {
         "Vp": 6.611,
         "Vs": 3.733,
         "R": 1.771,
-        "poisson": 0.266
+        "poisson": 0.266,
+        "x": 0.,
+        "y": 0.05
         }
 
     }
@@ -105,28 +119,23 @@ def RMSDifference(x, y):
     return np.sqrt(md/(n))
 
 
-def minimizeRatio(lith1, lith2):
-
-    RF = np.array([52.6, 0.8, 16.6, np.nan, 6.6, 0.11,
-                   4.4, 6.4, 3.2, 1.88, np.nan, 0.2, np.nan])
-    CM = np.array([61.7, 0.9, 14.7, 1.9, 5.1, 0.1, 3.1,
-                   5.7, 3.6, 2.1, 0.8, 0.2, np.nan])
-    RG = np.array([60.6, 0.7, 15.9, np.nan, 6.7, 0.1,
[…]

### 2013-11-18 paper/draftBP.tex
half intro complete

@@
 
+The Lithoprobe project has provided researches with a rich dataset from which detailed cross-sections have been derived.
+These cross sections form a near continuous transect from the Pacific to the Atlantic oceans have and enabled many insights into crustal structure and genesis.
+Recent examples include Cook et al [2009] analysis of the Moho where they present the Moho as having varied and complex origins and as evolving through thermal and pressure changes, deformation and fluid migration.
+In another recent Lithoprobe study, Hammer and Clowes [2011] discuss broad structural characteristics of the Canadian crust.
+They note a remarkable consistency in orogenic style such as the similarity between geolgical structures contained within Archean cratons and their boundaries as well as those observed in Proterozoic and Phanerozoic orogens.
+Also noted is the striking consistency in moho depth, which is imaged as lying between 33km and 43km depth along the full 20,000km of Lithoprobe transects, despite the diversity of overlying crust.
+Both Cook et al.
+and Hammer and Clowes raise the possibility that in some situations the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary.
+A final example of a Canada wide crustal study utilizing Lithoprobe data is the Lith5.0 revised crustal model presented by Perry, Eaton and Forte [2002].
+This work updates the Crust5.1 model [Mooney et al, 1998] to include the new data from Lithoprobe and introduces broad crustal features of the Canadian crust such as the relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
+These studies and other Canada wide studies employ active source Lithoprobe data which records higher frequency (10-40hz) {\it P}-wave velocity whereas this study uses lower frequency (0.4-3hz) passive source scattering techniques which provides estimates of the $V_P/V_S$ ratio and images relatively larger scale structure.
+Studies of the Canadian crust employing a similar teleseismic receiver function method as this work include a study of the Great Slave Lake shear zone [Eaton and Hope, 2002] which utilize receiver functions, shear-wave splitting results and gravity data to characterize the shear zone.
+Finally, three teleseismic studies discussed in greater detail in later sections are presented in chronological order: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen [Eaton et al., 2005], a structural and compositional characterization of the Superior Province [Darbyshire et al.
+2006] and a crustal evolution analysis of the Hudson Bay region [Thompson et al., 2010].
+This work differs from these studies in that it focuses on interregional comparisons and Canada wide structural characteristics, as does some of the Lithoprobe literature, but also on Canadian bulk crustal composition.
+
+Darbyshire, F.
+A., D.
+W.
+Eaton, A.
+W.
+Frederiksen, E.
+Leila (2006), New insights into the lithosphere beneath the Superior Province from Rayleigh wave dispersion and receiver function analysis, J.
+Geophys.
+Int., 169, 1043-1068.
+
+Eaton, D.
+W., S.
+Dineva, R.
+Mereu (2005), Crustal thickness and Vp/Vs variations in the Grenville orogen (Ontario, Canada) from analysis of teleseismic receiver functions, Tectonophysics, 420, 223-238.
+
+Hammer, P.
+T., R.
+M.
+Clowes, F.
+A.
+Cook, A.
+J.
+van der Velden, K.
+Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Canadian Journal of Earth Sciences, 47, 821-857.
+
+Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+
+Thompson, D.
+A., I.
+D.
+Bastow, G.
+Helffrich, J.-M.
+Kendall, J.
+Wookey, D.
+B.
+Snyder, D.
+W.
+Eaton (2010), Precambrian crustal evolution: Seismic constraints from the Canadian Shield, Earth and Planetary Science Letters, 297, 655–666.
+
+While there is a great diversity and body of seismic research on the Canadian crust those employing similar teleseismic methods and those encompassing the whole Canadian landmass require mention here.
+Recent teleseismic receiver function studies in the Candian Shield include
+
+In the following sections we describe the data and methods that we employ to interrogate the Canadian crust.
+We then proceed to present the major results beginning with a comparison between our seismic parameter estimates and the estimates from three published for matching seismic stations.
+We then present the results of a modified approach to our parameter estimation method and follow this with a presentation of the main dataset in two sections, one focusing on crustal discontinuities and structure and the other on regional variation and characterization.
+Finally we conclude with a discussion of implications of the results for our understanding of the Conrad discontinuity and bulk crustal composition of the Canadian Crust.
 %% Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
@@
 
-Meissner, R., (Ed.) (1986) \it{The Continental Crust: A Geophysical Approach}, 426, Academic, San Diego, Calif.
+Meissner, R., (Ed.) (1986) {\it The Continental Crust: A Geophysical Approach}, 426, Academic, San Diego, Calif.
 

### 2013-11-24 paper/draftBP.tex
intro almost complete

@@
 
+The Canadian continental lithosphere is composed of more than fifteen large geological provinces representing crustal formation and evolution from the early Archean to the present accretionary growth and orogenic deformation in the Western Cordillera.
+The Precambrian Canadian Shield and the extensive platforms of North America's interior form part of the assemblage that comprises the cratonic core, Laurentia.
+The Canadian Shield is itself composed of several distinct geological provinces brought together during a series of Paleoproterozoic orogens [Hoffman, 1998].
+The largest of these provinces (the world's largest Archean craton) is the Superior Province.
+Bounded by Proterozoic orogenic belts the western and sourthern Superior is composed of roughly eastwest trending geological subprovinces that include the granite-greestones of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga [Darbyshire et al., 2006].
+In the northwest of the Superior Province on the Eastern shores of the Hudson Bay is the Nuvvuagittuq greenstone belt which recent geochronology work has suggested formation times as early as ~4.28Ga making it the only known remnant of Hadean crust preserved on Earth [O'Neil et al., 2011].
+To the east is the relatively younger, 1.2-0.98Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision [Eaton et al., 2006].
+To the north of the Superior Province and separated by the Paleoproterozoic Trans-Hudson Orogen are the Rae and Hearne domains which together form the Western extent of the Churchill Province.
+Work in the the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga [Thompson et al., 2010].
+To the northwest of the Superior is the Slave Province.
+A relatively small cratonic fragment, the Slave is an Archean granite-greenstone block with geochronology dating between 4.0 Ga in the western Slave Province, the remnant of a larger Archean continent, to 2.6Ga in the eastern Slave, a younger paired accretionary prism and island arc [Bostock, 1998].
+To the west of the Superior are the extensive Interior Platforms comprising the southern extents of the Trans-Hudson Orogen and Hearne Domain as well as the Alberta Orogen.
+The Interior platform is largely Archean reworked during the Paleoproterzoic with bore-hole results showing amphibolite to granulite-grade Proterozoic metamorphism [Hammer et al., 2010].
+Extending west from the Interior platform the Western Cordillera is composed of amalgamated island arcs accreted after the onset of subduction during the Mesozoic.
+Although there is considerable variation in structure and tectonic history roughly half of the Cordillera is underlain by Proterozoic-margin crust which forms a crustal-scale décollement beneath interleaved and duplexed crust.
+Although there is significant orogeny in the southeast studies have shown that the crust is thin and relatively uniform in thickness with no crustal root being observed [Hammer et al., 2010].
+The rich tectonic history of the Canadian continental crust affords an excellent opportunity to study crustal properties and composition with geological age.
+[[ THAT LAST SENTENCE WAS KILLING ME, I GAVE UP...
+IT ALL SOUNDED TRITE TO ME ]]
+
 The Lithoprobe project has provided researches with a rich dataset from which detailed cross-sections have been derived.
-These cross sections form a near continuous transect from the Pacific to the Atlantic oceans have and enabled many insights into crustal structure and genesis.
-Recent examples include Cook et al [2009] analysis of the Moho where they present the Moho as having varied and complex origins and as evolving through thermal and pressure changes, deformation and fluid migration.
+These cross sections, which form a near continuous transect from the Pacific to the Atlantic Oceans, have enabled many insights into crustal structure and genesis.
+Recent research utilizing lithoprobe data include the analysis of Cook et al [2009] where the Moho is argued as having varied and complex origins and as evolving through thermal and pressure changes, deformation and fluid migration.
 In another recent Lithoprobe study, Hammer and Clowes [2011] discuss broad structural characteristics of the Canadian crust.
@@
 and Hammer and Clowes raise the possibility that in some situations the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary.
-A final example of a Canada wide crustal study utilizing Lithoprobe data is the Lith5.0 revised crustal model presented by Perry, Eaton and Forte [2002].
+A final example of lithoprobe work is the Lith5.0 revised crustal model presented by Perry, Eaton and Forte [2002].
 This work updates the Crust5.1 model [Mooney et al, 1998] to include the new data from Lithoprobe and introduces broad crustal features of the Canadian crust such as the relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
-These studies and other Canada wide studies employ active source Lithoprobe data which records higher frequency (10-40hz) {\it P}-wave velocity whereas this study uses lower frequency (0.4-3hz) passive source scattering techniques which provides estimates of the $V_P/V_S$ ratio and images relatively larger scale structure.
-Studies of the Canadian crust employing a similar teleseismic receiver function method as this work include a study of the Great Slave Lake shear zone [Eaton and Hope, 2002] which utilize receiver functions, shear-wave splitting results and gravity data to characterize the shear zone.
-Finally, three teleseismic studies discussed in greater detail in later sections are presented in chronological order: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen [Eaton et al., 2005], a structural and compositional characterization of the Superior Province [Darbyshire et al.
+This work shares some features with the Lithoprobe research in that it is both regionally comparitive and concerned with Canadian bulk crustal characteristics.
+However, the focus on crustal reflectivity patterns and crustal composition is unique to these other works.
+Additionally, whereas the studies and other Lithoprobe research rely on the relatively high frequency (10-40hz) {\it P}-waves as the primary imaging source this study uses lower frequency (0.4-3hz) passive source scattering techniques.
+Therefore Lithoprobe data represents velocity contrasts in tens to hundreds of meters and resolves {\it P}-wave velocity while the data employed in this study is sensitive to larger scale lengths (1000-2000m) and resolves the {\it P} and {\it S} wave ratio.
+Studies of the Canadian crust employing similar teleseismic scattering techniques as this work include a study of the Great Slave Lake shear zone [Eaton and Hope, 2002] which utilize receiver functions, shear-wave splitting results and gravity data to characterize the shear zone.
+Finally, three teleseismic studies discussed in greater detail in later sections are here presented in chronological order: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen [Eaton et al., 2005], a structural and compositional characterization of the Superior Province [Darbyshire et al.
 2006] and a crustal evolution analysis of the Hudson Bay region [Thompson et al., 2010].
-This work differs from these studies in that it focuses on interregional comparisons and Canada wide structural characteristics, as does some of the Lithoprobe literature, but also on Canadian bulk crustal composition.
+This work differs from these teleseismic studies in both scale and primary focus.
+Whereas the teleseismic studies tend to interrogate a particular geological region or set of related regions this study looks at both differences across regions and the aggregate bulk Canadian crust.
+Finally, this study does not seek to define particular geological sequences but rather ...
+bulk crustal reflectivity, discontinuity and composition.
 
-Darbyshire, F.
-A., D.
-W.
-Eaton, A.
-W.
-Frederiksen, E.
-Leila (2006), New insights into the lithosphere beneath the Superior Province from Rayleigh wave dispersion and receiver function analysis, J.
-Geophys.
-Int., 169, 1043-1068.
-
-Eaton, D.
-W., S.
-Dineva, R.
-Mereu (2005), Crustal thickness and Vp/Vs variations in the Grenville orogen (Ontario, Canada) from analysis of teleseismic receiver functions, Tectonophysics, 420, 223-238.
-
-Hammer, P.
-T., R.
-M.
-Clowes, F.
-A.
-Cook, A.
-J.
-van der Velden, K.
-Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Canadian Journal of Earth Sciences, 47, 821-857.
-
-Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
-
-Thompson, D.
-A., I.
-D.
-Bastow, G.
-Helffrich, J.-M.
-Kendall, J.
-Wookey, D.
-B.
-Snyder, D.
-W.
-Eaton (2010), Precambrian crustal evolution: Seismic constraints from the Canadian Shield, Earth and Planetary Science Letters, 297, 655–666.
-
-While there is a great diversity and body of seismic research on the Canadian crust those employing similar teleseismic methods and those encompassing the whole Canadian landmass require mention here.
-Recent teleseismic receiver function studies in the Candian Shield include
+Studies on bulk crustal composition are better represented in the global continental crust literature.
+Two widely cited studies [Christensen and Mooney, 1995, Ruknick and Fountain, 1995] use refraction data to develop a velocity model which is used to constrain ratios of rock units assigned to different crustal depths.
+The chemical compositions of these ratios are aggregated to develop compositional estimates for the lower, mid and upper crust as well as for the entire crust.
+This work is different in that it employs updated rock velocity measurements [Christensen, 1996] and utilizes the reciever function method, which constrains the compositional analysis to the aggregated bulk crust.
+Two studies of crustal evolution use seismic parameter estimation and  by Zandt and Ammon [1995] employs teleseismic reflections to estimate Poisson's Ratio at 114 global broadband stations which are then sorted by geological age.
 
@@
 Finally we conclude with a discussion of implications of the results for our understanding of the Conrad discontinuity and bulk crustal composition of the Canadian Crust.
-%% Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
-The primary reason for this is the poor resolution afforded by the seismic networks currently and historically deployed across Canada.
-The Canadian continental landmass is composed of at least fifteen large geological provinces as recognized by the Geological Survey of Canada (GSC).
-Each of these regions is itself complex and heterogeneous and often larger than most European nations.
-On top of this there is poor seismic coverage, roughly one seismic station per 25,000 $km^2$.
-Many of these stations are clustered near areas of geologic interest such as the Cascadia Subduction zone or population and research centres such as Southern Ontario or areas of significant resource interest like the diamondiferous Great Slave Lake area.
-This leaves vast areas of the Canadian landmass completely unsampled.
-Despite these limitations, a low resolution comprehensive and comparative study is still feasible for some of the geological regions comprising the Canadian continental crust.
-Direct comparison between the bulk average seismic properties of geological regions and between aggregated regions and global averages provide some new insight into the crustal composition of the Canadian landmass.
-The accumulated dataset also affords investigations into variations of bulk crustal parameters such as Poisson's Ratio and crustal thickness with age and tectonic environment as well as some basic statistical data on interesting crustal features.
 
-%% This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
+This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
 It begins with a discussion of the raw data itself followed by a review of the receiver function method and a more detailed explanation of the inversion algorithms used to produce the dataset.

### 2013-12-04 paper/draftBP.tex
intro revisions, still need to add eaton multigenic moho

--- 2013-11-11 paper/draftBP.tex
@@
 \section{Introduction}
+\section{Introduction}
 
-%% Seismic studies of the Canadian continental crust predominately analyze specific geological features or regions rather than taking a comprehensive and comparative inter-regional analysis.
-The primary reason for this is the poor resolution afforded by the seismic networks currently and historically deployed across Canada.
-The Canadian continental landmass is composed of at least fifteen large geological provinces as recognized by the Geological Survey of Canada (GSC).
-Each of these regions is itself complex and heterogeneous and often larger than most European nations.
-On top of this there is poor seismic coverage, roughly one seismic station per 25,000 $km^2$.
-Many of these stations are clustered near areas of geologic interest such as the Cascadia Subduction zone or population and research centres such as Southern Ontario or areas of significant resource interest like the diamondiferous Great Slave Lake area.
-This leaves vast areas of the Canadian landmass completely unsampled.
-Despite these limitations, a low resolution comprehensive and comparative study is still feasible for some of the geological regions comprising the Canadian continental crust.
-Direct comparison between the bulk average seismic properties of geological regions and between aggregated regions and global averages provide some new insight into the crustal composition of the Canadian landmass.
-The accumulated dataset also affords investigations into variations of bulk crustal parameters such as Poisson's Ratio and crustal thickness with age and tectonic environment as well as some basic statistical data on interesting crustal features.
+The Canadian continental lithosphere is composed of at least fifteen large geological provinces representing crustal evolution from the early Archean to the present accretionary growth and orogenic deformation in the Western Cordillera.
+The Precambrian Canadian Shield and the bounding platforms of North America's interior form part of the assemblage that comprises the cratonic core of Laurentia [Hammer et al., 2010].
+Several distinct geological provinces brought together during a series of Paleoproterozoic orogens make up the Canadian Shield [Hoffman, 1988, Figure 1].
+The largest of these provinces, and the world's largest Archean craton, is the Superior Province.
+Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly eastwest trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga [Thurston et al., 1991].
+In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth [O'Neil et al., 2011].
+To the east lies the younger, 1.2-0.98Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision [Eaton et al., 2006].
+North of the Superior Province and separated by the Paleoproterozoic Trans-Hudson Orogen are the Rae and Hearne domains which together form the western extent of the Churchill Province.
+Studies of the the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga [Thompson et al., 2010].
+The Slave Province is a relatively small Archean granite-greenstone block lying to the northwest of the Churchill province.
+Its western portion is the 4.0 remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc [Kusky, 1989].
+To the west of the Superior are the extensive Interior Platforms comprising the southern extents of the Trans-Hudson Orogen and Hearne Domain as well as the Alberta Orogen.
+The Interior platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism [Hammer et al., 2010].
+Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean to Paleoproterozoic crystalline basement.
+Between the middle of the Jurasic and Paleocene intense metamorphic, plutonic and uplifting events occured as the result of Paleozoic and younger intraoceanic-arc accretionary events.
+The Coastal mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock [Clowes et al, 1995].
+Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale décollement beneath interleaved and duplexed crust [Hammer et al., 2010].
+The rich tectonic tapestry of the Canadian continental crust affords an excellent opportunity to study variations in bulk crustal properties with geological age.
 
-%% This paper presents a comparative tour through the dataset accumulated from processing more than a decade worth of data from all available Canadian seismic stations.
-It begins with a discussion of the raw data itself followed by a review of the receiver function method and a more detailed explanation of the inversion algorithms used to produce the dataset.
-Three previous publications utilizing similar processing schemes provide unique subsets of data to compare with the values computed in this survey for quality assurance.
-...
+Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling [Clowes et al., 1999].
+The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and have enabled many insights into crustal structure and genesis.
+Recent summary studies utilizing Lithoprobe data include the analysis of Cook et al [2009] who argue that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
+In another recent Lithoprobe study, Hammer and Clowes [2011] discuss broad structural characteristics of the Canadian crust such as their observation that Achean cratons and their boundaries contain structures very similar to those found in Proterozoic and Phanerozoic orogens.
+Another key observation is the lack of evidence for significant crustal disruption from magmatic instrusion related to subduction, post-orogenic extension or orogenic collapse.
+The authors also remark upon the striking uniformity in Moho depth that is imaged between 33km and 43km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
+Both Cook et al.
+and Hammer and Clowes raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistguishable from the upper mantle.
+If this is the case, the seismic Moho would represent the top of a metamorphic front and would be shallower than the crust-mantle boundary.
+The Lith5.0 revised crustal model presented by Perry, Eaton and Forte [2002] was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
+This work updates the global Crust5.1 model [Mooney et al, 1998] and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
+
+In this study, we examine the additional constraints that analyses of teleseismic receiver functions place upon Canadian bulk crustal properties when combined with available active source results.
+It is widely appreciated that receiver functions are sensitive primarily to crustal thickness and Poisson's ratio (or equivalently $V_P/V_S$) (Zhu and Kanomori, 20xx), and that the latter quantity provides some ability to discriminate among possible bulk compositions [Zandt and Ammon 1995].
+The constraint on Poisson's ratio and the greater averaging performed by low frequency (0.04-3 Hz) scattered teleseismic waves versus high frequency (10-40 Hz) {\it P}-waves employed in active source studies render the two techniques somewhat complementary.
+Our data set comprises the majority of teleseismic data available for the Canadian landmass including both permanent and temporary deployments.
+Many of the temporary deployments are described in [Bostock, Eaton, Snyder].
+Three previous studies employing similar methodologies but with more regional emphases deserve special mention: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen [Eaton et al., 2005], a structural and compositional characterization of the Superior Province [Darbyshire et al.
+2006] and a crustal evolution analysis of the Hudson Bay region [Thompson et al., 2010].
+Results from these studies will serve as benchmarks for our expanded effort.
+In the following sections we describe the data and methods employed to interrogate the Canadian crust.
+We proceed to present the major results beginning with a comparison between our seismic parameter estimates and the estimates from three studies mentioned above for identical seismic stations.
+We then present the results of a modified approach to our parameter estimation method and follow this with a presentation of the resulting crustal property estimates.
+First presented are the average crustal discontinuity and structural patterns across Canada.
+Second, we take a tour through those geological provinces for which sufficient data exists, focusing on regional variation and characterization.
+We conclude with a discussion on the implications for our understanding of the composition and evolution of the Canadian Crust.
 
@@
 
+Clowes, R.
+M., F.
+Cook, Z.
+Hajnal, J.
+Hall, J.
+Lewry, S.
+Lucas, R.
+Wardle (1999), Canada's Lithoprobe Project (collaborative, multidisciplinary geoscience research leads to new understanding of continental evolution), Episodes, 22, 3-20.
+
+Clowes, R. M., C. A. Zelt, J. R. Amor, R. M. Ellis (1995), Lithospheric structure in the southern Canadian Cordillera from a network of seismic refraction lines, Can. J. Earth Sci., 32, 1485-1513.
+
 Conrad, V. (1925), Laufzeitkurven des Tauernbebens vom 28. November 1923, Mitt. Erdb. Komm. Wiener Akad. Wiss., 59, 1–23.
@@
 
-Efron, B., R. Tibshirani (1986), Bootstrap Methods for Standard Errors, Confidence Intervals, and Other Measures of Statistical Accuracy, Statistical Science, 1, 54-75.
+Efron, B., R. Tibshirani (1986), Bootstrap methods for standard errors, confidence intervals, and other measures of statistical accuracy, Statistical Science, 1, 54-75.
 
@@
 
+Hammer, P. T., R. M. Clowes (2011), The big picture: a lithospheric cross section of the North American continent, GSA Today, 21, 4-9.
+
 Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.
+
+Hoffman, P. F. (1998), United plates of America, the birth of a craton: Early Proterozoic Assembly and Growth of Laurentia, Ann. Rev. Earth Planet, 16, 543-603.
 
@@
 Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
+
+Kusky, T. M. (1989), Accretion of the Archean Slave province Geology, 17, 63-67.
 
@@
 
+Thurston, P. C., H. R. Williams, R. H. Sutcliffe, G. M. Scott (Eds.) (1991), Geology of Ontario, Ontario Geological Survey Special, 4.
+
 Zhu, L., H. Kanamori (2000), Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys. Res., 105, 2969-2980.

### 2013-12-08 paper/bptectono.tex
figures much better

--- 2013-12-04 paper/draftBP.tex
@@
-\documentclass[draft, 12pt]{article}
-\usepackage[english]{babel}
-\usepackage[final]{graphicx}
-\usepackage{multirow}
-\usepackage{color}
+\documentclass{elsart}
+\usepackage{graphicx,multirow,color,natbib,amssymb}
+\journal{Tectonophysics}
+
 \graphicspath{ {figures/} }
@@
 
-\title{Seismic parameter estimation and the Canadian crust}
-\author{Ben Postlethwaite}
+\begin{document}
+\begin{frontmatter}
+
+\title{Stroboscopic Doppler tomography of FO Aqr}
+%% \author[South]{T.R. Marsh\corauthref{cor}}
+%% \corauth[cor]{Corresponding author.}
+%% \ead{trm@astro.soton.ac.uk}
+%% \author[Oxford]{S.R. Duck\thanksref{now}}
+%% \thanks[now]{Present address: Systems Engineering and Assessment Ltd., Beckington Castle, PO Box 800, Bath BA3 6TB, UK.}
+%% \ead{srd@sea.co.uk}
+%% \address[South]{University of Southampton, Department of Physics, Highfield, Southampton SO17 1BJ, UK}
+%% \address[Oxford]{University of Oxford, Department of Physics, Nuclear Physics Laboratory, Keble Road, Oxford, OX1 3RH, UK}
 
 %% ------------------------------------------------------------------------ %%
-\begin{document}
-
 \begin{abstract}
@@
 
+\end{frontmatter}
 %% -----------------------------------------------%%
-\section{Introduction}
 \section{Introduction}
@@
   \centering
-    \includegraphics[width=\textwidth]{reflectedPhases}
+    \includegraphics[width=\textwidth]{reflectedPhases.pdf}
   \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
@@
     \cline{4-7}
-    & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
+    & & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
     \hline
@@
 
-\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:HVp_ULM}).
+\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:kcurves} a).
 This trade-off contributes to the low correlation in $H$ between the two methods.
 Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$.
-The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figures \ref{fig:HR_ULM}-\ref{fig:RVp_ULM}.
+The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).
 
@@
 A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty.
-RFs are arranged vertically, ordered by increasing slowness, and shown in figures \ref{fig:RF_ULM} and \ref{fig:RF_DORN}.
-Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figures \ref{fig:HR_ULM}-\ref{fig:HVp_ULM}) and figures \ref{fig:HR_DORN}-\ref{fig:HVp_DORN}.
+RFs are arranged vertically, ordered by increasing slowness, and shown in figure \ref{fig:rfs}.
+Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figure \ref{fig:kcurves}.
 The lower resolution apparent in the $H$ - $V_P$ cross section for DORN translates into a bootstrap error of $V_P \pm 0.64 {\rm km/s}$.
@@
 If several active source experiments are located within this radius, estimates are averaged.
-Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and the parameter estimates are presented in figure \ref{fig:activeVp}-\ref{fig:activeH}.
+Four stations (ARVN, TYNO, ULM and WHY) meet these criteria and the parameter estimates are presented in figure \ref{fig:vpactive}.
 Estimates computed from the full grid approach are roughly 0.25 km/s higher than the active source data.
@@
 Using an average of the parameter estimates from the two cleanest stations, ULM and WHY, $\delta H$ will be 1.3 km, this fits well with the actual bias in the values.
-
-[[ SOURCES REGARDING BIAS ]] [[ WHAT DOES THIS MEAN, IS VP USABLE FOR THE HIGHEST QUALITY STATIONS?]]
 
@@
   \centering
-  \includegraphics[width=\textwidth]{RF_ULM}
+  \includegraphics[width=\textwidth]{kinematicCurves.eps}
   \caption{}
-  \label{fig:RF_ULM}
+  \label{fig:kcurves}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{stackHR_ULM}
+  \includegraphics[width=\textwidth]{rfDORNULM.eps}
   \caption{}
-  \label{fig:HR_ULM}
+  \label{fig:rfs}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{stackRVp_ULM}
-  \caption{}
-  \label{fig:RVp_ULM}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{stackHVp_ULM}
-  \caption{}
-  \label{fig:HVp_ULM}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{RF_DORN}
-  \caption{}
-  \label{fig:RF_DORN}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{stackHR_DORN}
-  \caption{}
-  \label{fig:HR_DORN}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{stackRVp_DORN}
-  \caption{}
-  \label{fig:RVp_DORN}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{stackHVp_DORN}
-  \caption{}
-  \label{fig:HVp_DORN}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{activeSourceCompVp}
-  \caption{\it{P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
-  \label{fig:activeVp}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{activeSourceCompH}
-  \caption{Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
-  \label{fig:activeH}
+  \includegraphics[width=\textwidth]{vpactive.eps}
+  \caption{Top: {\it P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0.
+Bottom: Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
+  \label{fig:vpactive}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{conradPCA}
+  \includegraphics[width=\textwidth]{conradPCA.eps}
   \caption{The SVD singular vectors (principal components) of the station depth profile matrix.
@@
   \centering
-  \includegraphics[width=\textwidth]{discontinuity}
+  \includegraphics[width=\textwidth]{discontinuity.eps}
   \caption{}
@@
 The Slave Province averages are similar to those for the Churchill province with a crustal thickness and seismic velocity ratio comparable to Crust 1.0.
-Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$, figures \ref{map:Hmap}-\ref{map:Rmap}.
+Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$.
 The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
@@
   \centering
-  \includegraphics[width=\textwidth]{stationMap}
+  \includegraphics[width=\textwidth]{stationMap.pdf}
   \caption{}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{Hmap}
-  \caption{}
-  \label{map:Hmap}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{Rmap}
-  \caption{}
-  \label{map:Rmap}
-\end{figure}
-
 %% -----------------------------------------------%%
-\section{Discussion}
-
 \section{Discussion}
@@
  \centering
- \includegraphics[width=\textwidth]{discontinuityInterp}
+ \includegraphics[width=\textwidth]{discontinuityInterp.eps}
  \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
@@
   \centering
-  \includegraphics[width=\textwidth]{lithology}
+  \includegraphics[width=\textwidth]{lithology.eps}
   \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
@@
   \centering
-  \includegraphics[width=\textwidth]{composition}
+  \includegraphics[width=\textwidth]{composition.eps}
   \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}

### 2013-12-08 matlab/conradPCA.m

@@ -226,8 +226,13 @@ h(1) = subplot(1,2,1);
     set(gca, 'YTick', []);
     set(gca,'YTickLabel','')
     set(gca,'XTickLabel','')
+    %set(gca,'position',[0.1 0.1 0.6 0.6]);
     %set(gca, 'TickDir', 'out')
-
+    xlabel('interpolated station depth profiles', 'FontSize', 14)
+    ylabel('Depth [km]', 'FontSize', 14)
+    
+    
+    
 h(2) = subplot(1,2,2);
     plot(profile, xi, 'LineWidth', 2)
     hold on
@@ -241,11 +246,14 @@ h(2) = subplot(1,2,2);
     %set(gca, 'TickDir', 'out')
     set(gca, 'XTick', []);
     set(gca, 'YTick', []);
+    xlabel('profile stack', 'FontSize', 14)
+    
     
 pos=get(h,'position');
+pos{1}(3) = 1.5*pos{1}(3);
 leftedge = pos{1}(1) + pos{1}(3);
 pos{2}(1) = leftedge;
-pos{2}(3) = 0.7 * pos{2}(3);
+pos{2}(3) = 0.8 * pos{2}(3);
 set(h(1),'position',pos{1});
 set(h(2),'position',pos{2});
 
@@ -257,15 +265,15 @@ set(h(2),'position',pos{2});
 %     plot( sum(sqrt(XI.^2),1)', 'LineWidth', 2)
 figure(19)
     imagesc(1:size(X,1), H, X')
-    xlabel('station depth profiles')
-    ylabel('Depth [km]')
+    xlabel('station depth profiles', 'FontSize', 14)
+    ylabel('Depth [km]', 'FontSize', 14)
     set(gca,'XTickLabel','')
     set(gca, 'XTick', []);
     set(gca,'TickDir','Out')
     
 %% Plot Modes
 
-ne = 10;
+ne = 5;
 
 Vs = V(:, 1:ne);
 vmax = max(max(abs(Vs)));
@@ -281,10 +289,10 @@ Xs = Xs/max(abs(Xs));
 
 figure(34)
 
-plot(Vs + shift, H, 'LineWidth', 2)
+plot(Vs + shift, H, 'Color', [0.25,0.25,0.25], 'LineWidth', 2)
 
 axis tight
-ylabel('H [km]')
+ylabel('H [km]', 'FontSize',14)
[…]

### 2013-12-08 matlab/functions/plotSlice.m

@@ -1,21 +1,24 @@
 function plotSlice(station, x, y, xl, yl, s, xstd, ystd, sstd, X, Y, slice)
 
-
-figure()
-set(gca,'FontName','Helvetica','FontSize',16,...
-    'Clipping','off','layer','top');
-imagesc(X, Y, slice);
+% set(gca,'FontName','Helvetica','FontSize',16,...
+%     'Clipping','off','layer','top');
+pcolor(X, Y, slice);
 axis xy
 axis square
-colorbar
+shading interp
 hold on
 plot(x, y, 'w+')
 plot(x, y, 'ko')
-contour(X, Y, slice, [s - sstd, s - sstd], 'k-')
+[~, H] = contour(X, Y, slice, [s - sstd, s - sstd], 'w-');
+set (H, 'LineWidth', 1.2);
 hold off
-xlab=xlabel(xl);
-ylab=ylabel(yl);
-set(xlab,'FontName','Helvetica','FontSize',16);
-set(ylab,'FontName','Helvetica','FontSize',16);
-title(sprintf('%s\n%s = %1.3f +/- %1.3f km/s\n%s = %1.3f +/- %1.3f',...
-    station, xl, x, xstd, yl, y, ystd));
+if ~isempty(xl)
+    xlab=xlabel(xl);
+end
+if ~isempty(yl)
+    ylab=ylabel(yl);
+end
+% set(xlab,'FontName','Helvetica','FontSize',16);
+% set(ylab,'FontName','Helvetica','FontSize',16);
+% title(sprintf('%s\n%s = %1.3f +/- %1.3f km/s\n%s = %1.3f +/- %1.3f',...
+%     station, xl, x, xstd, yl, y, ystd));

### 2013-12-08 python/vpPlots.py

@@ -64,7 +64,7 @@ if __name__  == "__main__":
     ## Figure Properties #######
     figwidth = 12
     figheight = figwidth / 1.618
-    ratio = 1.5
+    ratio = 1.3
     lw = 4 / ratio# line width
     ms = 12 / ratio# marker size
     caplen = 7 / ratio
@@ -73,7 +73,7 @@ if __name__  == "__main__":
     ticks = 16 / ratio
     label = 16 / ratio
     title = 18 / ratio
-    leg = 16 / ratio
+    leg = 14 / ratio
 
 
     #############################################################################
@@ -185,15 +185,14 @@ if __name__  == "__main__":
         arg = Args().addQuery("fg::stdVp", "lt", str(maxerr))
         f = Params(stnfile, ["fg::H","fg::Vp", "fg::stdH", "fg::stdVp", "hk::stdR"], arg)
         w = Params(stnfile, ["wm::Vp","wm::H"])
-
-
         c = Params(csfile, ["H","Vp"])
-        c.sync(f)
 
+        f.sync(w)
+        c.sync(f)
         w.sync(c)
 
-        stdVp = 2 * f.fg_stdVp # 2 stdError
-        stdH = 2 * f.fg_stdH
+        stdVp = 2*f.fg_stdVp # 2 stdError
+        stdH = 2*f.fg_stdH
         t = np.arange(len(f.fg_Vp))
 
         corr = pearsonr(f.fg_Vp, c.Vp)
@@ -202,57 +201,65 @@ if __name__  == "__main__":
         md = RMSDifference(f.fg_Vp, c.Vp)
         print "Active Source vs FG Vp with {} stations: RMS Difference = {}".format(len(f.stns), md)
 
-        fig = plt.figure( figsize = (figwidth, figheight) )
-        ax = plt.subplot(111)
-
-        plt.plot(t, f.fg_Vp, '-ob', lw = lw, ms = ms, label = "Full Gridsearch Vp estimate")
-        plt.errorbar(t, f.fg_Vp, yerr=stdVp, xerr=None, fmt=None, ecolor = 'blue',
+        #fig = plt.figure( figsize = (figwidth, figheight) )
+        fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True, figsize = (figwidth, figheight))
+        color = "0.25"
[…]

### 2013-12-09 paper/bptectono.tex
tyhor

@@
-\documentclass{elsart}
+\documentclass[review]{elsarticle}
 \usepackage{graphicx,multirow,color,natbib,amssymb}
@@
 \graphicspath{ {figures/} }
-\DeclareGraphicsExtensions{.png}
 
@@
 
-\title{Stroboscopic Doppler tomography of FO Aqr}
-%% \author[South]{T.R. Marsh\corauthref{cor}}
-%% \corauth[cor]{Corresponding author.}
-%% \ead{trm@astro.soton.ac.uk}
-%% \author[Oxford]{S.R. Duck\thanksref{now}}
-%% \thanks[now]{Present address: Systems Engineering and Assessment Ltd., Beckington Castle, PO Box 800, Bath BA3 6TB, UK.}
-%% \ead{srd@sea.co.uk}
-%% \address[South]{University of Southampton, Department of Physics, Highfield, Southampton SO17 1BJ, UK}
-%% \address[Oxford]{University of Oxford, Department of Physics, Nuclear Physics Laboratory, Keble Road, Oxford, OX1 3RH, UK}
+\title{Seismic parameter estimation and the Canadian crust}
+
+\author[ubc]{B. Postlethwaite\corref{cor1}}
+\ead{bpostlet@eos.ubc.ca}
+\author[ubc]{M. G. Bostock}
+\ead{bostock@eos.ubc.ca}
+\author[ubc]{N. I. Christensen}
+\author[gsc]{D. B. Snyder}
+
+\cortext[cor2]{Corresponding author}
+
+\address[ubc]{Earth and Ocean Sciences, UBC, Vancouver BC, Canada}
+\address[gsc]{Geological Survey Canada, Natural Resources Canada, Ottawa ON, Canada}
 
@@
 \begin{abstract}
-
+It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
+Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
+Durrheim, W.
+Mooney, 1991).
+Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
+We apply a stacking method to retrieve estimates of the bulk crustal velocity ratio $V_P/V_S$ and thickness $H$ from these data under the assumption of 1-D structure.
+Additionally, we investigate the feasibility of employing a variant on this method which allows resolution of $V_S$ and $V_P$ uniquely.
+Bootstrap error analysis is performed for each station dataset and subsets of these results are compared to the results for matching stations from previously published studies.
+Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
+These results will be used to make structural characterizations of the crust as well as evaluate competing models of crustal formation.
 \end{abstract}
@@
 The Canadian continental lithosphere is composed of at least fifteen large geological provinces representing crustal evolution from the early Archean to the present accretionary growth and orogenic deformation in the Western Cordillera.
-The Precambrian Canadian Shield and the bounding platforms of North America's interior form part of the assemblage that comprises the cratonic core of Laurentia [Hammer et al., 2010].
-Several distinct geological provinces brought together during a series of Paleoproterozoic orogens make up the Canadian Shield [Hoffman, 1988, Figure 1].
+The Precambrian Canadian Shield and the bounding platforms of North America's interior form part of the assemblage that comprises the cratonic core of Laurentia (Hammer et al., 2010).
+Several distinct geological provinces brought together during a series of Paleoproterozoic orogens make up the Canadian Shield (Hoffman, 1988, Figure 1).
 The largest of these provinces, and the world's largest Archean craton, is the Superior Province.
-Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly eastwest trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga [Thurston et al., 1991].
-In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth [O'Neil et al., 2011].
-To the east lies the younger, 1.2-0.98Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision [Eaton et al., 2006].
+Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly eastwest trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga (Thurston et al., 1991).
+In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth (O'Neil et al., 2011).
+To the east lies the younger, 1.2-0.98Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision (Eaton et al., 2006).
 North of the Superior Province and separated by the Paleoproterozoic Trans-Hudson Orogen are the Rae and Hearne domains which together form the western extent of the Churchill Province.
-Studies of the the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga [Thompson et al., 2010].
+Studies of the the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga (Thompson et al., 2010).
 The Slave Province is a relatively small Archean granite-greenstone block lying to the northwest of the Churchill province.
-Its western portion is the 4.0 remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc [Kusky, 1989].
+Its western portion is the 4.0 remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc (Kusky, 1989).
 To the west of the Superior are the extensive Interior Platforms comprising the southern extents of the Trans-Hudson Orogen and Hearne Domain as well as the Alberta Orogen.
-The Interior platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism [Hammer et al., 2010].
+The Interior platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism (Hammer et al., 2010).
 Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean to Paleoproterozoic crystalline basement.
 Between the middle of the Jurasic and Paleocene intense metamorphic, plutonic and uplifting events occured as the result of Paleozoic and younger intraoceanic-arc accretionary events.
-The Coastal mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock [Clowes et al, 1995].
-Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale décollement beneath interleaved and duplexed crust [Hammer et al., 2010].
+The Coastal mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock (Clowes et al, 1995).
+Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale décollement beneath interleaved and duplexed crust (Hammer et al., 2010).
 The rich tectonic tapestry of the Canadian continental crust affords an excellent opportunity to study variations in bulk crustal properties with geological age.
 
-Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling [Clowes et al., 1999].
+Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
 The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and have enabled many insights into crustal structure and genesis.
-Recent summary studies utilizing Lithoprobe data include the analysis of Cook et al [2009] who argue that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
-In another recent Lithoprobe study, Hammer and Clowes [2011] discuss broad structural characteristics of the Canadian crust such as their observation that Achean cratons and their boundaries contain structures very similar to those found in Proterozoic and Phanerozoic orogens.
+Recent summary studies utilizing Lithoprobe data include the analysis of Eaton (2005) who examines seismic evidence for several hypotheses of the origin of continental Moho: relic Moho, underplating, metamorphic front and regional décollement.
+Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis while the presence of high-velocity lower-crustal xenoliths in southern Alberta suggests thick mafic underplating and evidence for the second.
+Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large scale eclogitization of the lower crust.
+Finally, seismic imaging of subhorizontal reflections seperated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional dècollement.
+Continuing from this effort, a study by Cook et al.
+(2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
+In another recent Lithoprobe study, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust such as their observation that Achean cratons and their boundaries contain structures very similar to those found in Proterozoic and Phanerozoic orogens.
 Another key observation is the lack of evidence for significant crustal disruption from magmatic instrusion related to subduction, post-orogenic extension or orogenic collapse.
@@
 If this is the case, the seismic Moho would represent the top of a metamorphic front and would be shallower than the crust-mantle boundary.
-The Lith5.0 revised crustal model presented by Perry, Eaton and Forte [2002] was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
-This work updates the global Crust5.1 model [Mooney et al, 1998] and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
+The Lith5.0 revised crustal model presented by Perry, Eaton and Forte (2002) was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
+This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
 
 In this study, we examine the additional constraints that analyses of teleseismic receiver functions place upon Canadian bulk crustal properties when combined with available active source results.
-It is widely appreciated that receiver functions are sensitive primarily to crustal thickness and Poisson's ratio (or equivalently $V_P/V_S$) (Zhu and Kanomori, 20xx), and that the latter quantity provides some ability to discriminate among possible bulk compositions [Zandt and Ammon 1995].
+It is widely appreciated that receiver functions are sensitive primarily to crustal thickness and Poisson's ratio (or equivalently $V_P/V_S$) (Zhu and Kanomori, 2000), and that the latter quantity provides some ability to discriminate among possible bulk compositions (Zandt and Ammon, 1995).
 The constraint on Poisson's ratio and the greater averaging performed by low frequency (0.04-3 Hz) scattered teleseismic waves versus high frequency (10-40 Hz) {\it P}-waves employed in active source studies render the two techniques somewhat complementary.
 Our data set comprises the majority of teleseismic data available for the Canadian landmass including both permanent and temporary deployments.
-Many of the temporary deployments are described in [Bostock, Eaton, Snyder].
-Three previous studies employing similar methodologies but with more regional emphases deserve special mention: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen [Eaton et al., 2005], a structural and compositional characterization of the Superior Province [Darbyshire et al.
-2006] and a crustal evolution analysis of the Hudson Bay region [Thompson et al., 2010].
+Many of the temporary deployments are described in (Bostock et al, 2010).
+Three previous studies employing similar methodologies but with more regional emphases deserve special mention: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen (Eaton et al., 2005), a structural and compositional characterization of the Superior Province (Darbyshire et al.
+2006) and a crustal evolution analysis of the Hudson Bay region (Thompson et al., 2010).
 Results from these studies will serve as benchmarks for our expanded effort.
@@
 
-\subsection{Geological and tectonic summary}
-\begin{itemize}
-
-\item Large scale overview/summary of regions under study
-\item Justification for selecting these regions?
-
-\end{itemize}
-
-\subsection{Previous geophysical studies}
- \begin{itemize}
-
-\item Lithoprobe and attendent papers
-\item Teleseismic studies
-
-\end{itemize}
 %% -----------------------------------------------%%
@@
 \section{Data and methods}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{stationMap.pdf}
+  \caption{Map showing major Canadian geological provinces and seismic stations contributing to the data discussed in this study.}
+  \label{map:stationMap}
+\end{figure}
+
 Our analysis of bulk Canadian continental crust is based on estimates of crustal properties that can be accessed via seismic techniques, namely seismic wave velocities (or velocity ratio) and crustal thickness.
@@
 After selection and filtering more than 80,000 events are available for further processing.
+Seismic stations contributing to the data utilized in this study are shown in figure \ref{map:stationMap}.
 
 The first stage of processing requires the transformation of teleseismic data into receiver functions.
-In general terms, this transformation involves deconvolving an approximation of the earthquake source from horizontal recordings of ground motion [Langston, 1979].
+In general terms, this transformation involves deconvolving an approximation of the earthquake source from horizontal recordings of ground motion (Langston, 1979).
 The resulting waveforms contain discrete pulses corresponding to {\it S} waves scattered from subsurface discontinuities including the base of the crust, or Moho.
 Confident identification of these peaks is required to estimate crustal parameters, and improved results can be obtained by transforming the displacement seismogram into their {\it P} and {\it S} contributions based on a 1-D model of propagation.
-This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels [Kennet, 1991].
+This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels (Kennet, 1991).
 The direct {\it P} arrival of the signal on the resulting {\it P} wave component is used as an approximation to the source function as {\it P} component impulse response for teleseismic P approximates a delta function.
@@
 \begin{equation}
-  r(t) = {\cal F}^{-1} \left[ G(\omega) \right] = {\cal F}^{-1}
+  r(t) = {\cal F}^{-1} \left( G(\omega) \right) = {\cal F}^{-1}
  \left[ \frac {\sum_n^N S_n(\omega)P_n^*(\omega)} {\sum_n^N P_n(\omega)P_n^*(\omega) + \delta} \right ],
@@
 
-A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, herafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori [2000], hereafter ZK.
+A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, herafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori (2000), hereafter ZK.
 This method exploits the dependence of these parameters to the differential arrival times between the {\it S} wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct {\it P} wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
@@
 \begin{equation} \label{eq:tps}
-t_{Ps}(p_n) = \frac{H}{V_P} \left[ \sqrt{ R^2 - V_P^2p_n^2} - \sqrt{1 - V_P^2p_n^2} \right]
+t_{Ps}(p_n) = \frac{H}{V_P} \left( \sqrt{ R^2 - V_P^2p_n^2} - \sqrt{1 - V_P^2p_n^2} \right)
 \end{equation}
@@
 \begin{equation}
-t_{Pps}(p_n) = \frac{H}{V_P} \left[ \sqrt{ R^2 - V_P^2p_n^2} + \sqrt{1 - V_P^2p_n^2} \right]
+t_{Pps}(p_n) = \frac{H}{V_P} \left( \sqrt{ R^2 - V_P^2p_n^2} + \sqrt{1 - V_P^2p_n^2} \right)
 \end{equation}
@@
 A negative weight for the combined $PpSs$ and $PsPs$ phases is required as the polarity of the signal is reversed.
-Semblance weighting [Eaton, 2006] is employed to reduce the effect of spurious large amplitude noise in the data.
+Semblance weighting (Eaton, 2006) is employed to reduce the effect of spurious large amplitude noise in the data.
 The semblance function assigns a weight between zero (incoherent noise) and one (coherent signal).
@@
 \begin{equation}
-S_m(H,R) = \frac {\left[ \sum_{n=1}^N r_n(t_m) \right]^2}
+S_m(H,R) = \frac {\left( \sum_{n=1}^N r_n(t_m) \right)^2}
                  { \sum_{n=1}^N r_n^2(t_m) }
@@
 
-%% [[\subsection{Full Parameter Method}]]
 A limitation of the ZK approach is the requirement of an initial $V_P$ estimate.
@@
 
-The trade-off between $V_P$ and $H$ (and other parameter combinations) and error resulting from data quality are quantified for both the full parameter search and ZK methods by bootstrap resampling [Efron and Tibshirani, 1986].
+The trade-off between $V_P$ and $H$ (and other parameter combinations) and error resulting from data quality are quantified for both the full parameter search and ZK methods by bootstrap resampling (Efron and Tibshirani, 1986).
 An estimate for the error is calculated in a bootstrap resampling approach by taking the standard deviation of 1024 parameter estimates each produced with randomly selected RFs, with replacement.
@@
 
-The Crust 1.0 model [Laske et.
-al., 2013] employs many of these same active source data to provide regular, complete coverage across all of Canada.
+The Crust 1.0 model (Laske et.
+al., 2013) employs many of these same active source data to provide regular, complete coverage across all of Canada.
 The model includes estimates of $V_P$ and $V_S$ information with depth at $1^\circ$ resolution in latitude and longitude.
@@
 Three published studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
-The most recent study [Thompson et.
-al., 2010] focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
-A 2007 study of the Superior Province [Darbyshire et.
-al, 2007] and a study of the Grenville Orogen [Eaton et.
-al., 2006] utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
+The most recent study (Thompson et.
+al., 2010) focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
+A 2007 study of the Superior Province (Darbyshire et.
+al, 2007) and a study of the Grenville Orogen (Eaton et.
+al., 2006) utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
 Comparisons between crustal estimates of $R$ for particular seismic stations are made for those values with corresponding standard errors of less than $\pm$0.06.
@@
 \begin{table}
-  \begin{tabular}{ l l l l l l l}
-    \cline{4-7}
-    & & & \multicolumn{2}{ c }{Correlation Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
+  \begin{tabular}{ p{3cm} p{2cm} p{1cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm}}
+    & & & \multicolumn{2}{ c }{Corr. Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
     \hline
-    Study Authors             & Region         & Num. of Stns &  $H$ & $R$  &$H$(km)& $R$ \\
+    Study Authors & Region & Num. Stns &  $H$ & $R$  &$H$ (km)& $R$ \\
     \hline
-    Thompson et. al. (2010)   & Northern Canadian Sheild & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
+    Thompson et. al. (2010) & N. Canadian Shield & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
     Darbyshire et. al. (2007) & Superior Province & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
@@
 al.
-[2010] also corrispond most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
+(2010) also corrispond most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
 Results published by {\it Eaton et.
-al.} [2006] show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that is roughly equal to the averaged uncertainty in the data.
-Parameter estimates from the Superior Province [Darbyshire et.
-al., 2007] show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
+al.} (2006) show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that is roughly equal to the averaged uncertainty in the data.
+Parameter estimates from the Superior Province (Darbyshire et.
+al., 2007) show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
 These differences are well over the averaged computed error, and can not be easily explained by deviation related to noise in the data.
@@
 
-We now proceed to compare our results computed using the ZK approach with those from the full grid search (Table \ref{table:ZKvsFG}).
+We now proceed to compare our results computed using the ZK approach with those from the full grid search.
 We note that the velocity ratio $R$ has a correlation between the two datasets of 0.76 whereas $H$ has a correlation of 0.51.
@@
 This trade-off contributes to the low correlation in $H$ between the two methods.
-Higher correlation in this quantity can be restored by dividing out $V_P$ from $H$.
+Higher correlation in this quantity (0.96) can be restored by dividing out $V_P$ from $H$.
 The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).
@@
 There is less correlation between the crustal thickness data, though this is largely attributed to an approximate 10 km difference between the ARVN estimates.
-The reason for this discrepancy is unknown though both the ZK method and and a published study [Thompson et.
-al., 2010] provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
+The reason for this discrepancy is unknown though both the ZK method and and a published study (Thompson et.
+al., 2010) provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
 Despite lower correlation the apparent bias towards higher values computed from teleseismic data is also maintained for $H$.
 A positive bias in one parameter will tend to bias the other, owing to the relationship between thickness, velocity and times of a scattered phase.
-For a constant $t_{ps}$ one may compute how much $H$ must change for a 0.25 km/s bias in $V_P$ [Zhu and Kanamori, 2000].
+For a constant $t_{ps}$ one may compute how much $H$ must change for a 0.25 km/s bias in $V_P$ (Zhu and Kanamori, 2000).
 Using an average of the parameter estimates from the two cleanest stations, ULM and WHY, $\delta H$ will be 1.3 km, this fits well with the actual bias in the values.
-
-\begin{table}
-  \begin{tabular}{ l l }
-    \hline
-    Parameter & correlation coeff \\
-    \hline
-    $R$ (km) &  0.76 \\
-    $H$      &  0.51 \\
-    $H/V_P$  &  0.96 \\
-    \hline
-  \end{tabular}
-  \caption{Correlation between parameters computed from ZK + semblance and full parameter search methods}
-\label{table:ZKvsFG}
-
-\end{table}
 
@@
   \includegraphics[width=\textwidth]{vpactive.eps}
-  \caption{Top: {\it P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0.
-Bottom: Crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
+  \caption{Top, {\it P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0.
+Bottom, crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
   \label{fig:vpactive}
@@
   \centering
-  \includegraphics[width=\textwidth]{conradPCA.eps}
-  \caption{The SVD singular vectors (principal components) of the station depth profile matrix.
+  \includegraphics[width=\textwidth]{discPCA.eps}
+  \caption{Left, vertical station depth profiles.
+Right, the SVD singular vectors (principal components) of the station depth profile matrix.
 Each principal component represents a mode that captures some part of the total variance in the dataset.
@@
 A grey horizontal window highlights a region of low energy.}
-  \label{fig:conradPCA}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{discontinuity.eps}
-  \caption{}
-  \label{fig:disc}
+  \label{fig:discPCA}
 \end{figure}
@@
 We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method.
-The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:disc} arranged by increasing crustal thickness.
+The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:discPCA} (left) arranged by increasing crustal thickness.
 By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components.
-Figure \ref{fig:conradPCA} contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles.
+Figure \ref{fig:discPCA} (right) contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles.
 Higher modes correspond to diminishing contributions to total variance.
 Each vector is normalized to unit magnitude.
-The first mode (blue) is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset.
+The first mode is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset.
 It features a maximum near 38.7 km that may be taken as an average Moho depth.
@@
     \hline
-    Region  & $H [km]$& $R$ &$H [km]$& $R$  &$V_P [km/s]$&$V_P [km/s]$ \\
+    Region  & $H (km]$& $R$ &$H [km]$& $R$  &$V_P [km/s]$&$V_P [km/s)$ \\
     \hline
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{stationMap.pdf}
-  \caption{}
-  \label{map:stationMap}
-\end{figure}
-
 %% -----------------------------------------------%%
-\section{Discussion}
+\section{Discussion and Conclusions}
 
 Seismic refraction and reflection studies of continental crust have afforded valuable constraints on the processes governing crustal evolution.
-Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust [Conrad, 1925; Jeffreys, 1926].
-Evidence supporting this model was originally presented by Conrad [1925] who cited observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovičić in 1909 [Litak and Brown, 1989].
-Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence [Litak and Brown, 1989].
-Technological improvements in reflection seismology allowed researchers to further argue for a complex heterogeneous crust and propose models wherein smooth random lateral inhomogeneities are superimposed on simple velocity gradient models to reproduce the range of seismic observations [Litak and Brown, 1989; Mereu and Ojo, 1981].
+Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust (Conrad, 1925; Jeffreys, 1926).
+Evidence supporting this model was originally presented by Conrad (1925) who cited observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovičić in 1909 (Litak and Brown, 1989).
+Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence (Litak and Brown, 1989).
+Technological improvements in reflection seismology allowed researchers to further argue for a complex heterogeneous crust and propose models wherein smooth random lateral inhomogeneities are superimposed on simple velocity gradient models to reproduce the range of seismic observations (Litak and Brown, 1989; Mereu and Ojo, 1981).
 
 Despite diminishing support in recent years for the Conrad Discontinuity as a universal crustal feature, the division of the crust into seismically distinguishable upper and lower components remains a common assertion.
-In their review, Mooney and Brocher [1987] argued for a model of the continental crust separated intoa seismically reflective lower crust overlain by transparent upper crust.
+In their review, Mooney and Brocher (1987) argued for a model of the continental crust separated intoa seismically reflective lower crust overlain by transparent upper crust.
 Their physical interpretation is that of a poorly laminated upper crust dominated by features of short length scale that contrasts with a lower crust composed of thin ($\sim$ 100 m) lamellae of significantly longer horizontal length scale.
-The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust [e.g.
+The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust (e.g.
 Ito, 1998; Mayer, 1996; Meissner et.
-al., 2006].
+al., 2006).
 
@@
  \centering
- \includegraphics[width=\textwidth]{discontinuityInterp.eps}
- \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
+ \includegraphics[width=\textwidth]{discInterp.eps}
+ \caption{Left, vertical station depth profiles interpolated to a common Moho. Right, normalized stack of depth profiles, grey line shows mean}
  \label{fig:discInterp}
@@
 
-The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in figure \ref{fig:disc} provides some supporting evidence for seismically distinct upper and lower crustal layers.
+The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in figure \ref{fig:discPCA} (left) provides some supporting evidence for seismically distinct upper and lower crustal layers.
 To better compare results across stations, we normalize each depth profile such that the Moho falls on a common ordinate (Figure \ref{fig:discInterp}, left) and stack the absolute values (Figure \ref{fig:discInterp}, right).
@@
 There is a steady and marked decrease in energy over the bottom third of the profile that cannot be ascribed to sidelobe structure associated with the Moho pulse.
-This observation is ostensibly at odds with the Mooney and Brocher [1987] high reflectivity lower crustal model.
-Moreover, reflection profiles from {\sc Lithoprobe} transects [Hammer et.
-al., 2010] frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
+This observation is ostensibly at odds with the Mooney and Brocher (1987) high reflectivity lower crustal model.
+Moreover, reflection profiles from {\sc Lithoprobe} transects (Hammer et.
+al., 2010) frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
 We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive.
-Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength [Bostock, 1999].
+Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength (Bostock, 1999).
 For 1 Hz teleseismic waves, crustal layering with scale lengths of 1000-2000 m should be resolvable.
@@
 Some authors, for example Mayer et .al.
-[1996] in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust.
+(1996) in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust.
 It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
-This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by Rolf Meisnner [Meisnner, 1986].
-It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data [Hammer et.
-al., 2010].
+This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by Rolf Meisnner (Meisnner, 1986).
+It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data (Hammer et.
+al., 2010).
 The increased transparency of the lower crust would appear to be better defined and would appear to relate to changes in rheology.
-Studies on crustal strength [Jackson, 2002] and on brittle and ductile models [Burov and Watts, 2006; Meisner and Rabbel, 2006] as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention, though see [Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999] for competing evidence.
+Studies on crustal strength (Jackson, 2002) and on brittle and ductile models (Burov and Watts, 2006; Meisner and Rabbel, 2006) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention, though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
 
@@
 Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time.
-Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003].
-After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
+Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks (Rudnick and Gao, 2003).
+After the plate-tectonic revolution, Taylor (1967, 1977), proposed the 'island arc' or 'andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
 Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
 Hence this model cannot account for the intermediate composition of post-Archean crust.
-Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995].
+Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition (Christensen and Mooney, 1995; Rudnick and Fountain, 1995).
 Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
 al.
-[1981] though whereas Smithson utilizes mean crustal $V_P$ this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
+(1981) though whereas Smithson utilizes mean crustal $V_P$ this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
 
@@
 In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$.
-Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996], all quoted at average crustal pressures of 600 MPa.
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Le Maitre (1976] and Christensen [1996), all quoted at average crustal pressures of 600 MPa.
 Dashed lines delimit isocontours in $V_P/V_S$.
-The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province [Percival and West, 1994].
+The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province (Percival and West, 1994).
 The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures comprised largely of tonalites and their metamorphosed equivalents.
-Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].
+Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons (Rudnick, 1995).
 
 Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite.
-This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition e.g., [Rudnick and Fountain, 1995; Christensen and Mooney [1995].
+This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition e.g., (Rudnick and Fountain, 1995; Christensen and Mooney, 1995).
 The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
@@
 The Grenville Province, in contrast, displays markedly lower $V_S$ and higher $V_P/V_S$ implying a significant difference in mean composition.
-The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study [Eaton et.
-al., 2005] that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces.
-This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally [Eaton et.
-al., 2005; Rivers, 2008].
+The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study (Eaton et.
+al., 2005) that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces.
+This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally (Eaton et.
+al., 2005; Rivers, 2008).
 However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St.
@@
 Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces.
-[[Double checkout less quartzite argument, this implies higher Vp ...
-but the data does not show it ]]
 
 \begin{table}
-  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
+  \begin{tabular}{ p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm}}
     \hline
-    Oxide, wt\% & Pakiser and Robinson [1966] & Smithson [1978] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work: Canada & This Work: Shield\\
+    Oxide, wt\% & Pakiser et al. (1966) & Smithson (1978) & Christensen et al. (1995) & Rudnick et al. (2003) & This Work: Canada & This Work: Shield\\
     \hline
@@
   \end{tabular}
-  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Gao [2003] use total $Fe$ expressed as $FeO$.}
+  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Gao (2003) use total $Fe$ expressed as $FeO$.}
 \label{table:composition}
@@
 A similar estimate for the Canadian Shield is 68\% diorite and 32\% granite.
-This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre [1976] to produce an estimate of bulk crustal composition.
+This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre (1976) to produce an estimate of bulk crustal composition.
 The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
@@
 Also included in this figure is the our average crustal composition estimate for the Canadian Shield.
-Its values closely track the global crustal estimates of Christensen \& Mooney [1995] and Smithson [1978].
+Its values closely track the global crustal estimates of Christensen \& Mooney (1995] and Smithson [1978).
 These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
@@
 
-Other researchers have employed Poisson's ratio [Zandt and Ammon, 1995] alone as an indicator of bulk crustal composition, and crustal thickness and {\it P}-wave velocity [Durrheim and Mooney, 1991] to constrain models of crustal formation.
-Zandt and Ammon [1995] report high $V_P/V_S$ values for Archean Shields and concluded that no systematic difference between Archean and Proterozoic crust exists.
+Other researchers have employed Poisson's ratio (Zandt and Ammon, 1995) alone as an indicator of bulk crustal composition, and crustal thickness and {\it P}-wave velocity (Durrheim and Mooney, 1991) to constrain models of crustal formation.
+Zandt and Ammon (1995) report high $V_P/V_S$ values for Archean Shields and concluded that no systematic difference between Archean and Proterozoic crust exists.
 Their average, $V_P/V_S=1.84$, is much higher than our weighted average for the Canadian Shield of 1.74.
@@
 Whereas Zandt and Ammon present evidence for increasing $V_P/V_S$ with crustal age, our estimates signify the opposite trend.
-Our results are more consistent with those of Durrheim and Mooney [1991] who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
+Our results are more consistent with those of Durrheim and Mooney (1991) who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
 The crustal thickness data presented in this study is only marginally thicker for Proterozoic crust (38.9 km) than Archean crust (38.0 km) and this difference is significantly smaller than the 5 km difference reported by Durrheim and Mooney.
 Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
-The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney [1991], has been persuasively augmented by recent work [Rudnick 1995; Rucknick and Gao, 2003] to now include a range of possible processes that may explain the data.
+The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney (1991), has been persuasively augmented by recent work (Rudnick 1995; Rucknick and Gao, 2003) to now include a range of possible processes that may explain the data.
 
 One possible process is the preferential recycling of soluble ions such as $Mg^{2+}$ and $Ca^{2+}$.
-These ions are weathered and carried into the oceans leaving behind silicon and aluminum [Rudnick and Gao, 2003] producing a net change to crustal composition over time.
+These ions are weathered and carried into the oceans leaving behind silicon and aluminum (Rudnick and Gao, 2003) producing a net change to crustal composition over time.
 In another model, deep crustal roots expose lower crust to processes leading to eclogitization and subsequent delamination due to a density inversion with the underlying mantle peridoties.
 This model does not directly address the question as to why there is an apparent variance in average velocities between younger and older crusts, unless the delamination process has evolved over time.
-However, knowledge of the spatial or temporal extent of delamination is difficult to document [Rudnick, 1995].
+However, knowledge of the spatial or temporal extent of delamination is difficult to document (Rudnick, 1995).
 A third model suggests that a mixture of silicic melts from subducting oceanic crust and basaltic melts from mantle peridotites were important in early crustal formation.
-This process is thought to be more common in the hotter Archean and is related to the TGG suites of Archean granitoids [Rudnick and Gao, 2003] and therefore provides a possible explanation for the data.
+This process is thought to be more common in the hotter Archean and is related to the TGG suites of Archean granitoids (Rudnick and Gao, 2003) and therefore provides a possible explanation for the data.
 It is important to state that all of these processes, as well as others, may have been operating throughout geologic time.
@@
 
-%% -----------------------------------------------%%
-\section{Conclusions}
-
 %% ------------------------------------------------------------------------ %%
 
-\section{bibliography}
-
-[[ADD GRANITE DIORITE REFERENCE]]
+\section{References}
 
@@
 
-Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophysical Journal International, 138, 747-756.
+Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophys. J. Int., 138, 747-756.
+
+Bostock, M. G., D. W. Eaton, D. B. Snyder (2010), Teleseismic studies of the Canadian landmass: Lithoprobe and its legacy, Can. J. Earth Sci., 47, 445-461.
 
@@
 
-Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere:" jelly sandwich" or" crème brûlée"?, GSA today, 16, 4.
+Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere: jelly sandwich or crème brûlée?, GSA today, 16, 4.
 
@@
 
-Christensen, N. I., W. D. Mooney (1995) Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
+Christensen, N. I., W. D. Mooney (1995), Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
 
@@
 van der Velden, K.
-Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Canadian Journal of Earth Sciences, 47, 821-857.
+Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Can.
+J.
+of Earth Sciences, 47, 821-857.
 
@@
 Laske, G., G. Masters, Z. Ma, M. Pasyanos (2013), Update on CRUST1.0 - A 1-degree Global Model of Earth's Crust, Geophys. Res. Abstracts, 15, Abstract EGU2013-2658.
+
+Le Maitre, R. W. (1976), The chemical variability of some common igneous rocks, Journal of Petrology, 17, 589-637.
 
@@
 
-Percival, J. A., Gordon F. W. (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Canadian. J. of Earth Sciences, 31, 1256-1286.
+Percival, J. A., Gordon F. W. (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Can. J. of Earth Sciences, 31, 1256-1286.
 

### 2013-12-09 matlab/conradPCA.m

@@ -220,21 +220,17 @@ profile = profile;
 
 figure(23)
 h(1) = subplot(1,2,1);
-    imagesc(XI')
-
-    set(gca, 'XTick', []);
-    set(gca, 'YTick', []);
-    set(gca,'YTickLabel','')
+    imagesc(1:size(X,1), H, XI')
     set(gca,'XTickLabel','')
-    %set(gca,'position',[0.1 0.1 0.6 0.6]);
-    %set(gca, 'TickDir', 'out')
+    set(gca, 'XTick', []);
+    set(gca,'TickDir','Out')
     xlabel('interpolated station depth profiles', 'FontSize', 14)
     ylabel('Depth [km]', 'FontSize', 14)
     
     
     
 h(2) = subplot(1,2,2);
-    plot(profile, xi, 'LineWidth', 2)
+    plot(profile, xi, 'LineWidth', 2, 'Color',[0.25,0.25,0.25])
     hold on
     line([mean(profile)-0.5, mean(profile)-0.5],[xi(1), xi(end)], 'Color', [0.7,0.7,0.7])
     hold off
@@ -257,19 +253,12 @@ pos{2}(3) = 0.8 * pos{2}(3);
 set(h(1),'position',pos{1});
 set(h(2),'position',pos{2});
 
-
+colormap(flipud(colormap('gray')))
 % figure()
 % subplot(1,2,1)
 %     imagesc(XI')
 % subplot(1,2,2)
 %     plot( sum(sqrt(XI.^2),1)', 'LineWidth', 2)
-figure(19)
-    imagesc(1:size(X,1), H, X')
-    xlabel('station depth profiles', 'FontSize', 14)
-    ylabel('Depth [km]', 'FontSize', 14)
-    set(gca,'XTickLabel','')
-    set(gca, 'XTick', []);
-    set(gca,'TickDir','Out')
     
 %% Plot Modes
 
@@ -287,50 +276,75 @@ var = E./sum(E);
 Xs = sum(X);
 Xs = Xs/max(abs(Xs));
 
-figure(34)
 
-plot(Vs + shift, H, 'Color', [0.25,0.25,0.25], 'LineWidth', 2)
+figure()
+h(1) = subplot(1,2,1);
+    imagesc(1:size(X,1), H, X')
+    xlabel('station depth profiles', 'FontSize', 14)
+    ylabel('Depth [km]', 'FontSize', 14)
+    set(gca,'XTickLabel','')
[…]

### 2014-01-10 paper/bptectono.tex
recolor figures - changes in paper

@@
 \address[gsc]{Geological Survey Canada, Natural Resources Canada, Ottawa ON, Canada}
-
 %% ------------------------------------------------------------------------ %%
@@
 It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
-Existing models based on fractionating mantle composition or evolving mantle convection require accurate estimates of the geological and geophysical properties of crustal provinces to better understand the details of early continental formation (R.
-Durrheim, W.
-Mooney, 1991).
+The calibration of models based on evolving mantle fractionation or mantle convection style require accurate estimates of the geological and geophysical properties of crustal provinces to better constrain the details of crustal formation.
 Fifteen years of publicly accessible teleseismic data from all available Canadian seismic stations are binned in horizontal slowness and deconvolved into receiver functions.
-We apply a stacking method to retrieve estimates of the bulk crustal velocity ratio $V_P/V_S$ and thickness $H$ from these data under the assumption of 1-D structure.
-Additionally, we investigate the feasibility of employing a variant on this method which allows resolution of $V_S$ and $V_P$ uniquely.
-Bootstrap error analysis is performed for each station dataset and subsets of these results are compared to the results for matching stations from previously published studies.
-Cross-analyzing these results with the mineral and rock seismic property database of Christensen (1996) will afford improved constraints on bulk geological composition of the Canadian landmass.
-These results will be used to make structural characterizations of the crust as well as evaluate competing models of crustal formation.
+We apply a stacking method to retrieve estimates of the bulk crustal velocity ratio $V_P/V_S$ and thickness $H$ from these data under the assumption of locally 1-D structure.
+We also investigate modifications to this approach that can allow discrimination of $V_S$ and $V_P$ under certain conditions.
+Bootstrap error analysis is performed for each station dataset and subsets of these measurments are compared with results for matching stations from previous studies.
+Analysis of results in conjunction with additonal velocity estimates from active source seismic studies and a seismic property database affords improved constraints on bulk geological composition of the Canadian landmass that are used to evaluate competing models of crustal formation.
 \end{abstract}
@@
 The Canadian continental lithosphere is composed of at least fifteen large geological provinces representing crustal evolution from the early Archean to the present accretionary growth and orogenic deformation in the Western Cordillera.
-The Precambrian Canadian Shield and the bounding platforms of North America's interior form part of the assemblage that comprises the cratonic core of Laurentia (Hammer et al., 2010).
+The Precambrian Canadian Shield and the bounding platforms of North America's interior comprise the cratonic core of Laurentia (Hammer et al., 2010).
 Several distinct geological provinces brought together during a series of Paleoproterozoic orogens make up the Canadian Shield (Hoffman, 1988, Figure 1).
 The largest of these provinces, and the world's largest Archean craton, is the Superior Province.
-Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly eastwest trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga (Thurston et al., 1991).
+Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly east-west trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga (Thurston et al., 1991).
 In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth (O'Neil et al., 2011).
@@
 The Slave Province is a relatively small Archean granite-greenstone block lying to the northwest of the Churchill province.
-Its western portion is the 4.0 remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc (Kusky, 1989).
-To the west of the Superior are the extensive Interior Platforms comprising the southern extents of the Trans-Hudson Orogen and Hearne Domain as well as the Alberta Orogen.
-The Interior platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism (Hammer et al., 2010).
+Its western portion is the 4.0 Ga remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc (Kusky, 1989).
+To the west of the Superior is the extensive Interior Platform overlying the southern extensions of the Trans-Hudson Orogen, Hearne Domain and Alberta Orogen.
+The Interior Platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism (Hammer et al., 2010).
 Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean to Paleoproterozoic crystalline basement.
-Between the middle of the Jurasic and Paleocene intense metamorphic, plutonic and uplifting events occured as the result of Paleozoic and younger intraoceanic-arc accretionary events.
-The Coastal mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock (Clowes et al, 1995).
+Between the middle of the Jurassic and Paleocene intense metamorphic, plutonic and uplifting events occured as the result of Paleozoic and younger intraoceanic-arc accretionary events.
+The Coast Mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock (Clowes et al, 1995).
 Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale décollement beneath interleaved and duplexed crust (Hammer et al., 2010).
@@
 Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
-The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and have enabled many insights into crustal structure and genesis.
-Recent summary studies utilizing Lithoprobe data include the analysis of Eaton (2005) who examines seismic evidence for several hypotheses of the origin of continental Moho: relic Moho, underplating, metamorphic front and regional décollement.
-Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis while the presence of high-velocity lower-crustal xenoliths in southern Alberta suggests thick mafic underplating and evidence for the second.
-Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large scale eclogitization of the lower crust.
-Finally, seismic imaging of subhorizontal reflections seperated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional dècollement.
-Continuing from this effort, a study by Cook et al.
+The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and has enabled many insights into crustal structure and genesis.
+Recent summary studies utilizing Lithoprobe data include the analysis of Eaton (2005) who considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'ecollement.
+Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis whereas the presence of high-velocity, lower-crustal xenoliths in southern Alberta favours thick mafic underplating.
+Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large-scale eclogitization of the lower crust.
+Finally, seismic imaging of subhorizontal reflections separated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional d\'ecollement.
+A subsequent study by Cook et al.
 (2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
-In another recent Lithoprobe study, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust such as their observation that Achean cratons and their boundaries contain structures very similar to those found in Proterozoic and Phanerozoic orogens.
-Another key observation is the lack of evidence for significant crustal disruption from magmatic instrusion related to subduction, post-orogenic extension or orogenic collapse.
-The authors also remark upon the striking uniformity in Moho depth that is imaged between 33km and 43km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
+
+In another recent summary study of Lithoprobe data, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust and note that Archean cratons and their boundaries contain structures similar to those found in Proterozoic and Phanerozoic orogens.
+They also discuss the lack of evidence for significant crustal disruption from magmatic instrusion related to subduction, post-orogenic extension or orogenic collapse.
+The authors remark upon the striking uniformity in Moho depth that is imaged between 33km and 43km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
 Both Cook et al.
-and Hammer and Clowes raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistguishable from the upper mantle.
+(2009) and Hammer and Clowes (2011) raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistguishable from the upper mantle.
 If this is the case, the seismic Moho would represent the top of a metamorphic front and would be shallower than the crust-mantle boundary.
+
 The Lith5.0 revised crustal model presented by Perry, Eaton and Forte (2002) was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
-This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick (>42.5km) Phanerozoic crust in the south and west of the continental craton.
+This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick ($>$42.5km) Phanerozoic crust in the south and west of the continental craton.
 
@@
 Despite diminishing support in recent years for the Conrad Discontinuity as a universal crustal feature, the division of the crust into seismically distinguishable upper and lower components remains a common assertion.
-In their review, Mooney and Brocher (1987) argued for a model of the continental crust separated intoa seismically reflective lower crust overlain by transparent upper crust.
+In their review, Mooney and Brocher (1987) argued for a model of the continental crust separated into a seismically reflective lower crust overlain by transparent upper crust.
 Their physical interpretation is that of a poorly laminated upper crust dominated by features of short length scale that contrasts with a lower crust composed of thin ($\sim$ 100 m) lamellae of significantly longer horizontal length scale.
@@
 Some authors, for example Mayer et .al.
-(1996) in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust.
+(1996) in their study of the Rhine Graben, have employed the term Conrad discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
 It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
@@
 al., 2010).
-The increased transparency of the lower crust would appear to be better defined and would appear to relate to changes in rheology.
-Studies on crustal strength (Jackson, 2002) and on brittle and ductile models (Burov and Watts, 2006; Meisner and Rabbel, 2006) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention, though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
+The increased transparency of the lower crust would appear to be better defined and relate to changes in rheology.
+Studies of crustal strength (Jackson, 2002) and deformation (Burov and Watts, 2006; Meisner and Rabbel, 2006) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention; though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
 
 Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
-Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time.
+It serves as a representative, average quantity that affords constraint on crustal structure and evolution in space and time.
 Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks (Rudnick and Gao, 2003).
-After the plate-tectonic revolution, Taylor (1967, 1977), proposed the 'island arc' or 'andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
+After the plate-tectonic revolution, Taylor (1967, 1977), proposed the `island arc' or `andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
 Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
@@
 al.
-(1981) though whereas Smithson utilizes mean crustal $V_P$ this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.
+(1981) although, whereas Smithson utilizes mean crustal $V_P$, this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the Crust 1.0 model.
 
@@
 
-In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$.
-Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Le Maitre (1976] and Christensen [1996), all quoted at average crustal pressures of 600 MPa.
+In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a graph of $V_P$ versus $V_S$.
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Le Maitre (1976) and Christensen (1996), all quoted at average crustal pressures of 600 MPa.
 Dashed lines delimit isocontours in $V_P/V_S$.
@@
 Lawrence Platform (where the geometric mean seismic ratio is 1.79).
-Moreover, and more importantly, the average $V_P$ values for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition.
+Moreover, and more importantly, the average $V_P$ value for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition.
 Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces.
@@
 The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
-Our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $MgO$ content than these previous studies.
+Our analysis reveals a crust with marginally higher average $SiO_2$ composition and lower $MgO$ content than these previous studies.
 Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition.
 The gray-shaded region indicates the $\pm 30\%$ range from our reference average Canadian crustal model.
-Also included in this figure is the our average crustal composition estimate for the Canadian Shield.
-Its values closely track the global crustal estimates of Christensen \& Mooney (1995] and Smithson [1978).
+Also included in this figure is our average crustal composition estimate for the Canadian Shield.
+Its values closely track the global crustal estimates of Christensen \& Mooney (1995) and Smithson (1978).
 These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
@@
 Their average, $V_P/V_S=1.84$, is much higher than our weighted average for the Canadian Shield of 1.74.
-These authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust, 1.73 compared to the corresponding value for our data of 1.77.
+These authors also report a low $V_P/V_S$ value for Cenozoic-Mesozoic crust (1.73) compared to the corresponding value for our data of 1.77.
 They interpret these high values for older crust and low values for younger crust as favouring either of two limiting models.
@@
 Our results are more consistent with those of Durrheim and Mooney (1991) who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
-The crustal thickness data presented in this study is only marginally thicker for Proterozoic crust (38.9 km) than Archean crust (38.0 km) and this difference is significantly smaller than the 5 km difference reported by Durrheim and Mooney.
-Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) than in Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
+The crustal thickness data presented in this study reveals no significant difference between Proterozoic crust (38.9 km) and Archean crust (38.0 km) and this difference is significantly smaller than the 5 km difference reported by Durrheim and Mooney.
+Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) versus the Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
 The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney (1991), has been persuasively augmented by recent work (Rudnick 1995; Rucknick and Gao, 2003) to now include a range of possible processes that may explain the data.
-
-One possible process is the preferential recycling of soluble ions such as $Mg^{2+}$ and $Ca^{2+}$.
-These ions are weathered and carried into the oceans leaving behind silicon and aluminum (Rudnick and Gao, 2003) producing a net change to crustal composition over time.
-In another model, deep crustal roots expose lower crust to processes leading to eclogitization and subsequent delamination due to a density inversion with the underlying mantle peridoties.
-This model does not directly address the question as to why there is an apparent variance in average velocities between younger and older crusts, unless the delamination process has evolved over time.
-However, knowledge of the spatial or temporal extent of delamination is difficult to document (Rudnick, 1995).
-A third model suggests that a mixture of silicic melts from subducting oceanic crust and basaltic melts from mantle peridotites were important in early crustal formation.
-This process is thought to be more common in the hotter Archean and is related to the TGG suites of Archean granitoids (Rudnick and Gao, 2003) and therefore provides a possible explanation for the data.
-It is important to state that all of these processes, as well as others, may have been operating throughout geologic time.
 

### 2014-01-10 matlab/functions/plotSlice.m

@@ -9,8 +9,8 @@ shading interp
 hold on
 plot(x, y, 'w+')
 plot(x, y, 'ko')
-[~, H] = contour(X, Y, slice, [s - sstd, s - sstd], 'w-');
-set (H, 'LineWidth', 1.2);
+[~, H] = contour(X, Y, slice, [s - sstd, s - sstd], 'k-');
+set(H, 'LineWidth', 1.6);
 hold off
 if ~isempty(xl)
     xlab=xlabel(xl);

### 2014-01-28 paper/bpostlethwaite_Thesis.tex
thesis submission

--- 2014-01-10 paper/bptectono.tex
@@
 \usepackage{graphicx,multirow,color,natbib,amssymb}
+\usepackage[margin=1in]{geometry}
+
+\makeatletter
+\renewcommand\@biblabel[1]{}
+\renewenvironment{thebibliography}[1]
+     {\section*{\refname}%
+      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
+      \list{}%
+           {\leftmargin0pt
+            \@openbib@code
+            \usecounter{enumiv}}%
+      \sloppy
+      \clubpenalty4000
+      \@clubpenalty \clubpenalty
+      \widowpenalty4000%
+      \sfcode`\.\@m}
+     {\def\@noitemerr
+       {\@latex@warning{Empty `thebibliography' environment}}%
+      \endlist}
+\makeatother
+
 \journal{Tectonophysics}
@@
 
-\title{Seismic parameter estimation and the Canadian crust}
+\title{Seismic velocity estimation and the Canadian crust}
 
@@
 %% ------------------------------------------------------------------------ %%
+
 \begin{abstract}
@@
 We also investigate modifications to this approach that can allow discrimination of $V_S$ and $V_P$ under certain conditions.
-Bootstrap error analysis is performed for each station dataset and subsets of these measurments are compared with results for matching stations from previous studies.
-Analysis of results in conjunction with additonal velocity estimates from active source seismic studies and a seismic property database affords improved constraints on bulk geological composition of the Canadian landmass that are used to evaluate competing models of crustal formation.
+Bootstrap error analysis is performed for each station dataset and subsets of these measurements are compared with results for matching stations from previous studies.
+Analysis of results in conjunction with additional velocity estimates from active source seismic studies and a seismic property database affords improved constraints on bulk geological composition of the Canadian landmass that are used to evaluate competing models of crustal formation.
 \end{abstract}
@@
 The largest of these provinces, and the world's largest Archean craton, is the Superior Province.
-Bounded by Proterozoic orogenic belts, the western and sourthern Superior is composed of roughly east-west trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6Ga (Thurston et al., 1991).
+Bounded by Proterozoic orogenic belts, the western and southern Superior is composed of roughly east-west trending geological subprovinces that include the granite-greenstone terranes of the Abitibi Greenstone Belt, metasedimentary gneisses, high-grade gneisses and plutonic complexes assembled during the period between 3-2.6 Ga (Thurston et al., 1991).
 In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth (O'Neil et al., 2011).
-To the east lies the younger, 1.2-0.98Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision (Eaton et al., 2006).
+To the east lies the younger, 1.2-0.98 Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision (Eaton et al., 2006).
 North of the Superior Province and separated by the Paleoproterozoic Trans-Hudson Orogen are the Rae and Hearne domains which together form the western extent of the Churchill Province.
-Studies of the the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga (Thompson et al., 2010).
+Studies of the Rae suggest a Paleoarchean and Mesoarchean basement whereas the Hearne comprises granite-and-greenstone terranes of ages ~2.7Ga (Thompson et al., 2010).
 The Slave Province is a relatively small Archean granite-greenstone block lying to the northwest of the Churchill province.
@@
 Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean to Paleoproterozoic crystalline basement.
-Between the middle of the Jurassic and Paleocene intense metamorphic, plutonic and uplifting events occured as the result of Paleozoic and younger intraoceanic-arc accretionary events.
+Between the middle of the Jurassic and Paleocene intense metamorphic, plutonic and uplifting events occurred as the result of Paleozoic and younger intraoceanic-arc accretionary events.
 The Coast Mountains evolved during the late Cretaceous as a wide magmatic arc composed largely of granitic rock (Clowes et al, 1995).
-Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale décollement beneath interleaved and duplexed crust (Hammer et al., 2010).
+Although there is considerable variation in structure and tectonic history, roughly half of the Cordillera is underlain by Proterozoic-margin crust that forms a crustal-scale d\'{e}collement beneath interleaved and duplexed crust (Hammer et al., 2010).
 The rich tectonic tapestry of the Canadian continental crust affords an excellent opportunity to study variations in bulk crustal properties with geological age.
@@
 The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and has enabled many insights into crustal structure and genesis.
-Recent summary studies utilizing Lithoprobe data include the analysis of Eaton (2005) who considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'ecollement.
+Recent summary studies utilizing Lithoprobe data include the analysis of {\it Eaton} (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement.
 Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis whereas the presence of high-velocity, lower-crustal xenoliths in southern Alberta favours thick mafic underplating.
@@
 Finally, seismic imaging of subhorizontal reflections separated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional d\'ecollement.
-A subsequent study by Cook et al.
-(2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
+A subsequent study by {\it Cook et al.} (2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
 
 In another recent summary study of Lithoprobe data, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust and note that Archean cratons and their boundaries contain structures similar to those found in Proterozoic and Phanerozoic orogens.
-They also discuss the lack of evidence for significant crustal disruption from magmatic instrusion related to subduction, post-orogenic extension or orogenic collapse.
-The authors remark upon the striking uniformity in Moho depth that is imaged between 33km and 43km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
-Both Cook et al.
-(2009) and Hammer and Clowes (2011) raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistguishable from the upper mantle.
+They also discuss the lack of evidence for significant crustal disruption from magmatic intrusion related to subduction, post-orogenic extension or orogenic collapse.
+The authors remark upon the striking uniformity in Moho depth that is imaged between 33 km and 43 km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
+Both {\it Cook et al.} (2009) and {\it Hammer and Clowes} (2011) raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistinguishable from the upper mantle.
 If this is the case, the seismic Moho would represent the top of a metamorphic front and would be shallower than the crust-mantle boundary.
 
-The Lith5.0 revised crustal model presented by Perry, Eaton and Forte (2002) was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
-This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick ($>$42.5km) Phanerozoic crust in the south and west of the continental craton.
+The Lith5.0 revised crustal model presented by {\it Perry et al.} (2002) was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
+This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick ($>$42.5 km) Phanerozoic crust in the south and west of the continental craton.
 
@@
 Our data set comprises the majority of teleseismic data available for the Canadian landmass including both permanent and temporary deployments.
-Many of the temporary deployments are described in (Bostock et al, 2010).
+Many of the temporary deployments are described by {\it Bostock et al.} (2010).
 Three previous studies employing similar methodologies but with more regional emphases deserve special mention: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen (Eaton et al., 2005), a structural and compositional characterization of the Superior Province (Darbyshire et al.
@@
 Second, we take a tour through those geological provinces for which sufficient data exists, focusing on regional variation and characterization.
-We conclude with a discussion on the implications for our understanding of the composition and evolution of the Canadian Crust.
+We conclude with a discussion on the implications for our understanding of the composition and evolution of the Canadian crust.
 
@@
 The first and primary source is teleseismic receiver function (hereafter RF) analysis.
-The remaining two datasets are Crust 1.0, a widely published statistical average of similar regions with a global two degree resolution, and a compilation of pre-processed active source data (Mooney, 2012).
+The remaining two datasets are Crust 1.0, a widely published statistical average of similar regions with a global one degree resolution, and a compilation of preprocessed active source data (Mooney, 2012).
 Both Crust 1.0 and the active source compilation are used to qualify and compare with the primary dataset.
@@
 Seismic stations are selected from all available regional and national networks including CNSN, POLARIS, FedNor and Chasme.
-Seismograms are included for analysis if they fall within a 30 to 100$^\circ$ epicentral distance window.
+Seismograms are included for analysis if they fall within a 30$^\circ$-100$^\circ$ epicentral distance window.
 Quality control is performed by admitting only those seismograms with sufficiently high signal to noise ratios to permit visual observation of the dominant {\it P} wave energy and with sufficiently impulsive first arrivals to allow arrival onsets to be accurately measured.
@@
   \centering
-    \includegraphics[width=\textwidth]{reflectedPhases.pdf}
+    \includegraphics[width=0.7\textwidth]{reflectedPhases.pdf}
   \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
@@
 
-A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, herafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by Zhu and Kanamori (2000), hereafter ZK.
+A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, hereafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by {\it Zhu and Kanamori} (2000), hereafter ZK.
 This method exploits the dependence of these parameters to the differential arrival times between the {\it S} wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct {\it P} wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
@@
 The travel time equations, as employed by ZK, assume a value for crustal {\it P} wave velocity, $V_P$, that in practice will trade-off to some degree with crustal thickness $H$.
-In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 1.0 value for the corrisponding $2^\circ$ containing cell.
+In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 1.0 value for the corresponding $1^\circ$ containing cell.
 In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
@@
 \ref{eq:stack}, as a function of all three seismic parameters, $s(H,R,V_P)$.
-Taking advantage of modern multi-core systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method easily scalable to hundreds of stations.
-This ``fullgrid'' approach is analyzed to determine the viability of recovering $V_P$.
+Taking advantage of modern multicore systems with reasonable quantities of data (fifty to a few hundred RFs) and a moderate search space ($n^3$ parameter candidates where $n \approx 150$) the computational time makes this method easily scalable to hundreds of stations.
+This ``full grid'' approach is analyzed to determine the viability of recovering $V_P$.
 
@@
 
-The Crust 1.0 model (Laske et.
-al., 2013) employs many of these same active source data to provide regular, complete coverage across all of Canada.
+The Crust 1.0 model (Laske et al., 2013) employs many of these same active source data to provide regular, complete coverage across all of Canada.
 The model includes estimates of $V_P$ and $V_S$ information with depth at $1^\circ$ resolution in latitude and longitude.
@@
 Three published studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
-The most recent study (Thompson et.
-al., 2010) focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
-A 2007 study of the Superior Province (Darbyshire et.
-al, 2007) and a study of the Grenville Orogen (Eaton et.
-al., 2006) utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
+The most recent study (Thompson et al., 2010) focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
+A 2007 study of the Superior Province (Darbyshire et al, 2007) and a study of the Grenville Orogen (Eaton et al., 2006) utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
 Comparisons between crustal estimates of $R$ for particular seismic stations are made for those values with corresponding standard errors of less than $\pm$0.06.
@@
 For each study under comparison, data are reprocessed using the $V_P$ chosen by the study authors.
-For the 2006 and 2007 studies the ZK + semblance method is used whereas for the Thompson et.
-al study the linear stacking ZK approach is employed.
+For the 2006 and 2007 studies the ZK + semblance method is used whereas for the Thompson et al.
+study the linear stacking ZK approach is employed.
 Both correlation coefficient and mean difference are provided in (Table \ref{table:comparison}) as measures of similarity.
@@
 \begin{table}
-  \begin{tabular}{ p{3cm} p{2cm} p{1cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm}}
+  \begin{tabular}{ p{3.8cm} p{3cm} p{2cm} p{1.2cm} p{1.2cm} p{1.2cm} p{1.2cm}}
     & & & \multicolumn{2}{ c }{Corr. Coeff.} & \multicolumn{2}{ c }{RMS difference} \\
@@
     \hline
-    Thompson et. al. (2010) & N. Canadian Shield & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
-    Darbyshire et. al. (2007) & Superior Province & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
-    Eaton et. al. (2006) & Grenville Orogen & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
+    Thompson et al. (2010) & N. Canadian Shield & 29 & 0.97 & 0.70 & 0.78 & 0.017 \\
+    Darbyshire et al. (2007) & Superior Province & 10 & 0.95 & 0.43 & 1.71 & 0.037 \\
+    Eaton et al. (2006) & Grenville Orogen & 26 & 0.90 & 0.61 & 1.03 & 0.033 \\
     \hline
@@
 A corollary of this is greater sensitivity of $R$ to differences in procedure such as the choice of deconvolution algorithm and selection and scaling of data.
-The results from Thompson et.
-al.
-(2010) also corrispond most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
-Results published by {\it Eaton et.
-al.} (2006) show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that is roughly equal to the averaged uncertainty in the data.
-Parameter estimates from the Superior Province (Darbyshire et.
-al., 2007) show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
+The results from {\it Thompson et al.} (2010) also correspond most closely to the parameter estimates computed here, with the mean difference for both parameters being less than half of the averaged bootstrap error computed in this study.
+Results published by {\it Eaton et al.} (2006) show reasonable correlations in both $R$ and $H$ and a mean difference for both parameters that is roughly equal to the averaged uncertainty in the data.
+Parameter estimates from the Superior Province (Darbyshire et al., 2007) show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
 These differences are well over the averaged computed error, and can not be easily explained by deviation related to noise in the data.
@@
 We note that the velocity ratio $R$ has a correlation between the two datasets of 0.76 whereas $H$ has a correlation of 0.51.
-The low correlation between crustal thickness estimates can be explained by the strong tradeoff between $H$ and $V_P$ through the quantity
+The low correlation between crustal thickness estimates can be explained by the strong trade-off between $H$ and $V_P$ through the quantity
 
@@
 Higher correlation in this quantity (0.96) can be restored by dividing out $V_P$ from $H$.
-The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in an other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).
+The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in the other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).
 
@@
 There is less correlation between the crustal thickness data, though this is largely attributed to an approximate 10 km difference between the ARVN estimates.
-The reason for this discrepancy is unknown though both the ZK method and and a published study (Thompson et.
-al., 2010) provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
+The reason for this discrepancy may be an anomalously low quality active source reading as both the ZK method and a published study (Thompson et al., 2010) provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
 Despite lower correlation the apparent bias towards higher values computed from teleseismic data is also maintained for $H$.
@@
 In a similar fashion, the weighted average of $H$ increases from 38.9 km in the Churchill Province  to 39.7 in the Superior Province and 41.7 km in the Grenville Province.
-The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data as well as as the Crust 1.0 $V_P$ data.
+The southeastern trend towards increasing values is also expressed in weighted averaged active source $V_P$ data as well as the Crust 1.0 $V_P$ data.
 The Crust 1.0 model is characterized by slightly thinner crust though the southward thickening trend is noticeable.
@@
   \end{tabular}
-  \caption{Comparison of $R$ and $H$ estimates with three published studies}
+  \caption{Comparison of seismic velocity and crustal thickness estimates for Canada and selected geological provinces.}
 \label{table:regionParameters}
@@
 Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust (Conrad, 1925; Jeffreys, 1926).
-Evidence supporting this model was originally presented by Conrad (1925) who cited observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovičić in 1909 (Litak and Brown, 1989).
+Evidence supporting this model was originally presented by {\it Conrad} (1925) which cites observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovi\v{c}i\'{c} in 1909 (Litak and Brown, 1989).
 Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence (Litak and Brown, 1989).
@@
 Despite diminishing support in recent years for the Conrad Discontinuity as a universal crustal feature, the division of the crust into seismically distinguishable upper and lower components remains a common assertion.
-In their review, Mooney and Brocher (1987) argued for a model of the continental crust separated into a seismically reflective lower crust overlain by transparent upper crust.
+In their review, Mooney and Brocher (1987) argue for a model of the continental crust separated into a seismically reflective lower crust overlain by transparent upper crust.
 Their physical interpretation is that of a poorly laminated upper crust dominated by features of short length scale that contrasts with a lower crust composed of thin ($\sim$ 100 m) lamellae of significantly longer horizontal length scale.
 The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust (e.g.
-Ito, 1998; Mayer, 1996; Meissner et.
-al., 2006).
+Ito, 1998; Mayer, 1996; Meissner et al., 2006).
 
@@
 
-The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in figure \ref{fig:discPCA} (left) provides some supporting evidence for seismically distinct upper and lower crustal layers.
+The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in Figure \ref{fig:discPCA} (left) provides some supporting evidence for seismically distinct upper and lower crustal layers.
 To better compare results across stations, we normalize each depth profile such that the Moho falls on a common ordinate (Figure \ref{fig:discInterp}, left) and stack the absolute values (Figure \ref{fig:discInterp}, right).
@@
 There is a steady and marked decrease in energy over the bottom third of the profile that cannot be ascribed to sidelobe structure associated with the Moho pulse.
-This observation is ostensibly at odds with the Mooney and Brocher (1987) high reflectivity lower crustal model.
-Moreover, reflection profiles from {\sc Lithoprobe} transects (Hammer et.
-al., 2010) frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
+This observation is ostensibly at odds with the {\it Mooney and Brocher} (1987) high reflectivity lower crustal model.
+Moreover, reflection profiles from {\sc Lithoprobe} transects (Hammer et al., 2010) frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
 We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive.
 Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength (Bostock, 1999).
-For 1 Hz teleseismic waves, crustal layering with scale lengths of 1000-2000 m should be resolvable.
-Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of 75-300 m.
-We infer then that the vertical scales of lower crustal reflectivity are limited to scale lengths under 100 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.
-Some authors, for example Mayer et .al.
-(1996) in their study of the Rhine Graben, have employed the term Conrad discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
-It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust.
-This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by Rolf Meisnner (Meisnner, 1986).
-It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data (Hammer et.
-al., 2010).
+For 1 Hz teleseismic waves, crustal layering with scale lengths of $\approx$ 800-2000 m should be resolvable.
+Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of $\approx$ 75-300 m.
+We infer then that the vertical scales of lower crustal reflectivity are better represented by scale lengths under $\approx$ 800 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.
+Some authors, for example Mayer et al.
+(1996) in their study of the Rhine Graben, have employed the term Conrad Discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
+It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly represented within the Canadian crust.
+This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by {\it Meissner} (1986).
+It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data (Hammer et al., 2010).
 The increased transparency of the lower crust would appear to be better defined and relate to changes in rheology.
-Studies of crustal strength (Jackson, 2002) and deformation (Burov and Watts, 2006; Meisner and Rabbel, 2006) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention; though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
+Studies of crustal strength (Jackson, 2002), deformation (Burov and Watts, 2006; Meissner et al., 2006) and the observed $\sim$ 100 m lower crustal lamallae reported by Mooney and Brocher (1987) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention; though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
 
@@
 Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition (Christensen and Mooney, 1995; Rudnick and Fountain, 1995).
-Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et.
-al.
-(1981) although, whereas Smithson utilizes mean crustal $V_P$, this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the Crust 1.0 model.
+Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of {\it Smithson et al.} (1981) although, whereas {\it Smithson et al.} utilizes mean crustal $V_P$, this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the Crust 1.0 model.
 
@@
 
-In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a graph of $V_P$ versus $V_S$.
-Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Le Maitre (1976) and Christensen (1996), all quoted at average crustal pressures of 600 MPa.
+In Figure \ref{fig:lith} we plot our area-weighted seismic velocity estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a graph of $V_P$ versus $V_S$.
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilations of laboratory velocity measurements by Christensen (1996) at 600 MPa and chemical analyses by Le Maitre (1976), all quoted at average crustal pressures of 600 MPa.
 Dashed lines delimit isocontours in $V_P/V_S$.
-The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province (Percival and West, 1994).
+The choice of common shield lithologies is motivated by observations along 30 km of paleo-depth represented in the Kapuskasing uplift of the Superior province (Percival and West, 1994).
 The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures comprised largely of tonalites and their metamorphosed equivalents.
@@
 
-Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite.
+Seismic velocity estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two lithologies: granite geniss and diorite.
 This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition e.g., (Rudnick and Fountain, 1995; Christensen and Mooney, 1995).
-The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
+The Slave, Superior and Churchill provinces possess comparable seismic velocities implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
 All three provinces cluster about the average value for the Canadian Shield.
 The Grenville Province, in contrast, displays markedly lower $V_S$ and higher $V_P/V_S$ implying a significant difference in mean composition.
-The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study (Eaton et.
-al., 2005) that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces.
-This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally (Eaton et.
-al., 2005; Rivers, 2008).
+The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study (Eaton et al., 2005) that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition that is more mafic than other Shield provinces.
+This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally (Eaton et al., 2005; Rivers, 2008).
 However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St.
@@
 \begin{table}
-  \begin{tabular}{ p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm} p{1.4cm}}
+  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
     \hline
-    Oxide, wt\% & Pakiser et al. (1966) & Smithson (1978) & Christensen et al. (1995) & Rudnick et al. (2003) & This Work: Canada & This Work: Shield\\
+    Oxide, wt\% & Pakiser and Robinson (1966) & Smithson (1978) & Christensen and Mooney (1995) & Rudnick and Gao (2003) & This Work: Canada & This Work: Canadian Shield\\
     \hline
@@
   \includegraphics[width=\textwidth]{composition.eps}
-  \caption{Comparison of major crustal elements from different studies. All estimates have been normalized to the results computed in this study.}
+  \caption{Comparisons of crustal composition from selected studies. All estimates have been normalized to the results computed in this study.}
   \label{fig:composition}
@@
 
-Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite.
+Since our average seismic velocity estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (Figure \ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite.
 A similar estimate for the Canadian Shield is 68\% diorite and 32\% granite.
-This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre (1976) to produce an estimate of bulk crustal composition.
-The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies.
+This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by {\it Le Maitre} (1976) to produce an estimate of bulk crustal composition.
+The resulting chemical breakdown is provided in table \ref{table:composition} along with selected previous estimates from global crustal studies.
 Our analysis reveals a crust with marginally higher average $SiO_2$ composition and lower $MgO$ content than these previous studies.
-Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition.
+Figure \ref{fig:composition} displays the major crustal oxide contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition.
 The gray-shaded region indicates the $\pm 30\%$ range from our reference average Canadian crustal model.
 Also included in this figure is our average crustal composition estimate for the Canadian Shield.
-Its values closely track the global crustal estimates of Christensen \& Mooney (1995) and Smithson (1978).
+Its values closely track the global crustal estimates of {\it Christensen and Mooney} (1995) and {\it Smithson} (1978).
 These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
@@
 Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) versus the Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
-The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by Durrheim and Mooney (1991), has been persuasively augmented by recent work (Rudnick 1995; Rucknick and Gao, 2003) to now include a range of possible processes that may explain the data.
+The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by {\it Durrheim and Mooney} (1991), has been persuasively augmented by recent work (Rudnick 1995; Rudnick and Gao, 2003) to now include a range of possible processes that may explain the data.
 
-This study provides results that are in general agreement with other recent crustal studies and similarly argues for an intermediate or andesitic crust.
-There are several processes that have been suggested to explain this intermediate composition ...
+In summary, the teleseismic dataset employed in this study has provided an opportunity for comparison of crustal structure at regional scales and for characterization of the Canadian crust as a whole.
+A trend of increasing $V_P/V_S$ and crustal thickness proceeding southward from the Churchill Province to Grenville Province is identified, however more comprehensive sampling is required to constrain the full extent and significance of this trend.
+Analysis of our dataset supplies further support to the thesis that the Conrad Discontinuity, defined as a pervasive reflective discontinuity between the upper and lower crust, is not a ubiquitous feature of the continental crust.
+Despite the absence of evidence for such feature, the data do suggest that the upper and lower crust can be distinguished on the basis of contrasting scattering signatures at reflection-seismic and teleseismic scales.
+Constraints from $V_P/V_S$ estimates suggest that the composition of the Canadian Shield is comparable to the global continental crustal average whereas the Canadian crustal average is characterized by slightly lower $V_P$ and $V_S$ and higher $V_P/V_S$.
+Finally, the dataset supports the hypothesis that processes driving crustal formation in the Archean and Proterozoic were dissimilar.
 
@@
 
-\section{References}
+\begin{thebibliography}{1}
 
-Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.
+\bibitem{} Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.
 
-Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophys. J. Int., 138, 747-756.
+\bibitem{} Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophys. J. Int., 138, 747-756.
 
-Bostock, M. G., D. W. Eaton, D. B. Snyder (2010), Teleseismic studies of the Canadian landmass: Lithoprobe and its legacy, Can. J. Earth Sci., 47, 445-461.
+\bibitem{} Bostock, M. G., D. W. Eaton, D. B. Snyder (2010), Teleseismic studies of the Canadian landmass: Lithoprobe and its legacy, Can. J. Earth Sci., 47, 445-461.
 
-Bostock, M. G., M. R. Kumar (2010), Bias in seismic estimates of crustal properties, J. Geophys. Int., 182, 403-407.
+\bibitem{} Bostock, M. G., M. R. Kumar (2010), Bias in seismic estimates of crustal properties, J. Geophys. Int., 182, 403-407.
 
-Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere: jelly sandwich or crème brûlée?, GSA today, 16, 4.
+\bibitem{} Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere: jelly sandwich or crème brûlée?, GSA today, 16, 4.
 
-Christensen, N. I. (1996), Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156.
+\bibitem{} Christensen, N. I. (1996), Poisson's ratio and crustal seismology, J. Geophys. Res., 101, 3139–3156.
 
-Christensen, N. I., W. D. Mooney (1995), Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
+\bibitem{} Christensen, N. I., W. D. Mooney (1995), Seismic velocity structure and composition of the continental crust: A global view, J. Geophys. Res., 100, 9761-9788.
 
-Clowes, R.
+\bibitem{} Clowes, R.
 M., F.
@@
 
-Clowes, R. M., C. A. Zelt, J. R. Amor, R. M. Ellis (1995), Lithospheric structure in the southern Canadian Cordillera from a network of seismic refraction lines, Can. J. Earth Sci., 32, 1485-1513.
+\bibitem{} Clowes, R.
+M., C.
+A.
+Zelt, J.
+R.
+Amor, R.
+M.
+Ellis (1995), Lithospheric structure in the southern Canadian Cordillera from a network of seismic refraction lines, Can.
+J.
+Earth Sci., 32, 1485-1513.
 
-Conrad, V. (1925), Laufzeitkurven des Tauernbebens vom 28. November 1923, Mitt. Erdb. Komm. Wiener Akad. Wiss., 59, 1–23.
+\bibitem{} Conrad, V. (1925), Laufzeitkurven des Tauernbebens vom 28. November 1923, Mitt. Erdb. Komm. Wiener Akad. Wiss., 59, 1–23.
 
-Darbyshire, F.
+\bibitem{} Darbyshire, F.
 A., D.
@@
 
-Deichmann, N. (1992), Structural and rheological implications of lower-crustal earthquakes below northern Switzerland, Physics of the earth and planetary interiors, 69, 270-280.
+\bibitem{} Deichmann, N. (1992), Structural and rheological implications of lower-crustal earthquakes below northern Switzerland, Physics of the earth and planetary interiors, 69, 270-280.
 
-Durrheim, R. J., W. D. Mooney (1991), Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
+\bibitem{} Durrheim, R. J., W. D. Mooney (1991), Archean and Proterozoic crustal evolution, Geology, 19, 606-609.
 
-Eaton, D.
+\bibitem{} Eaton, D.
 W., S.
@@
 
-Efron, B., R. Tibshirani (1986), Bootstrap methods for standard errors, confidence intervals, and other measures of statistical accuracy, Statistical Science, 1, 54-75.
+\bibitem{} Efron, B., R. Tibshirani (1986), Bootstrap methods for standard errors, confidence intervals, and other measures of statistical accuracy, Statistical Science, 1, 54-75.
 
-Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
+\bibitem{} Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
 
-Hammer, P.
+\bibitem{} Hammer, P.
 T., R.
@@
 
-Hammer, P. T., R. M. Clowes (2011), The big picture: a lithospheric cross section of the North American continent, GSA Today, 21, 4-9.
+\bibitem{} Hammer, P. T., R. M. Clowes (2011), The big picture: a lithospheric cross section of the North American continent, GSA Today, 21, 4-9.
 
-Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.
+\bibitem{} Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.
 
-Hoffman, P. F. (1998), United plates of America, the birth of a craton: Early Proterozoic Assembly and Growth of Laurentia, Ann. Rev. Earth Planet, 16, 543-603.
+\bibitem{} Hoffman, P. F. (1998), United plates of America, the birth of a craton: Early Proterozoic Assembly and Growth of Laurentia, Ann. Rev. Earth Planet, 16, 543-603.
 
-Ito, K. (1999), Seismogenic layer, reflective lower crust, surface heat flow and large inland earthquakes, Tectonophysics, 306, 423-433.
+\bibitem{} Ito, K. (1999), Seismogenic layer, reflective lower crust, surface heat flow and large inland earthquakes, Tectonophysics, 306, 423-433.
 
-Jackson, J. (2002), Strength of the continental lithosphere: Time to abandon the jelly sandwich?, GSA today, 12, 4-9.
+\bibitem{} Jackson, J. (2002), Strength of the continental lithosphere: Time to abandon the jelly sandwich?, GSA today, 12, 4-9.
 
-Jeffreys, H. (1926), On compressional waves in two superposed layers, Proc. Cambridge Philos. Soc., 23, 472–481.
+\bibitem{} Jeffreys, H. (1926), On compressional waves in two superposed layers, Proc. Cambridge Philos. Soc., 23, 472–481.
 
-Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
+\bibitem{} Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
 
-Kusky, T. M. (1989), Accretion of the Archean Slave province Geology, 17, 63-67.
+\bibitem{} Kusky, T. M. (1989), Accretion of the Archean Slave province Geology, 17, 63-67.
 
-Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
+\bibitem{} Langston, C. A. (1979), Structure under Mount Rainier, Washington, inferred from teleseismic body waves, J. Geophys. Res., 84(B9), 4749–4762, doi:10.1029/JB084iB09p04749.
 
-Laske, G., G. Masters, Z. Ma, M. Pasyanos (2013), Update on CRUST1.0 - A 1-degree Global Model of Earth's Crust, Geophys. Res. Abstracts, 15, Abstract EGU2013-2658.
+\bibitem{} Laske, G., G. Masters, Z. Ma, M. Pasyanos (2013), Update on CRUST1.0 - A 1-degree Global Model of Earth's Crust, Geophys. Res. Abstracts, 15, Abstract EGU2013-2658.
 
-Le Maitre, R. W. (1976), The chemical variability of some common igneous rocks, Journal of Petrology, 17, 589-637.
+\bibitem{} Le Maitre, R. W. (1976), The chemical variability of some common igneous rocks, Journal of Petrology, 17, 589-637.
 
-Litak, R. K., L. D. Brown (1989), A modern perspective on the Conrad discontinuity, Eos, 70, 713-725.
+\bibitem{} Litak, R. K., L. D. Brown (1989), A modern perspective on the Conrad discontinuity, Eos, 70, 713-725.
 
-Mayer, G., P.
+\bibitem{} Mayer, G., P.
 M.
@@
 
-Meissner, R., (Ed.) (1986) {\it The Continental Crust: A Geophysical Approach}, 426, Academic, San Diego, Calif.
+\bibitem{} Meissner, R., (Ed.) (1986) {\it The Continental Crust: A Geophysical Approach}, 426, Academic, San Diego, Calif.
 
-Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.
+\bibitem{} Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.
 
-Mereu, R.
+\bibitem{} Mereu, R.
 F., S.
@@
 
-Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.
+\bibitem{} Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.
 
-Mooney, W. D., T. M. Brocher (1987) Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.
+\bibitem{} Mooney, W. D., T. M. Brocher (1987) Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.
 
-Percival, J. A., Gordon F. W. (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Can. J. of Earth Sciences, 31, 1256-1286.
+\bibitem{} Percival, J. A., G. F. West (1994), The Kapuskasing uplift: a geological and geophysical synthesis, Can. J. of Earth Sciences, 31, 1256-1286.
 
-Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
+\bibitem{} Perry, H. K. C., D. W. S. Eaton, A. M. Forte (2002), LITH5.0: a revised crustal model for Canada based on Lithoprobe results,  J. Geophys. Int., 150, 285-294.
 
-Rudnick, R. L. (1995), Making Continental Crust, Nature, 378, 571-578.
+\bibitem{} Rivers, T.
+(2008) Assembly and preservation of lower, mid, and upper orogenic crust in the Grenville Province: Implications for the evolution of large hot long-duration orogens, Precambrian Research, 167, 237–259.
 
-Rudnick, R. L., D. M. Fountain (1995), Nature and composition of the continental crust: A lower crustal perspective, Rev. Geophys., 33(3), 267–309, doi:10.1029/95RG01302.
+\bibitem{} Rudnick, R. L. (1995), Making Continental Crust, Nature, 378, 571-578.
 
-Rudnick, R. L., S. Gao (2003), Composition of the Continental Crust, Treatise on Geochem., 3, 1-64.
+\bibitem{} Rudnick, R. L., D. M. Fountain (1995), Nature and composition of the continental crust: A lower crustal perspective, Rev. Geophys., 33(3), 267–309, doi:10.1029/95RG01302.
 
-Simpson, F. (1999), Stress and seismicity in the lower continental crust: A challenge to simple ductility and implications for electrical conductivity mechanisms. Surveys in Geophysics, 20, 201-227.
+\bibitem{} Rudnick, R. L., S. Gao (2003), Composition of the Continental Crust, Treatise on Geochem., 3, 1-64.
 
-Smithson, S. B. (1978) Modeling continental crust-structural and chemical constraints, Geophys. Res. Lett. 5(9), 749-775
+\bibitem{} Simpson, F.
+(1999), Stress and seismicity in the lower continental crust: A challenge to simple ductility and implications for electrical conductivity mechanisms.
+Surveys in Geophysics, 20, 201-227.
 
-Smithson, S. B., R. A. Johnson, Y. K. Wong (1981), Mean crustal velocity: a critical parameter for interpreting crustal structure and crustal growth, Earth and Planetary Science Letters, 53, 323-332.
+\bibitem{} Smithson, S. B. (1978) Modeling continental crust-structural and chemical constraints, Geophys. Res. Lett. 5(9), 749-775.
 
-Taylor, S. R. (1967), The origin and growth of continents, Tectonophysics, 4, 17–34.
+\bibitem{} Smithson, S.
+B., R.
+A.
+Johnson, Y.
+K.
+Wong (1981), Mean crustal velocity: a critical parameter for interpreting crustal structure and crustal growth, Earth and Planetary Science Letters, 53, 323-332.
 
-Taylor, S.
+\bibitem{} Taylor, S. R. (1967), The origin and growth of continents, Tectonophysics, 4, 17–34.
+
+\bibitem{} Taylor, S.
 R.
@@
 
-Thompson, D.
+\bibitem{} Thompson, D.
 A., I.
@@
 
-Thurston, P. C., H. R. Williams, R. H. Sutcliffe, G. M. Scott (Eds.) (1991), Geology of Ontario, Ontario Geological Survey Special, 4.
+\bibitem{} Thurston, P. C., H. R. Williams, R. H. Sutcliffe, G. M. Scott (Eds.) (1991), Geology of Ontario, Ontario Geological Survey Special, 4.
 
-Zhu, L., H. Kanamori (2000), Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys. Res., 105, 2969-2980.
+\bibitem{} Zhu, L., H. Kanamori (2000), Moho depth variation in Southern California from teleseismic receiver functions, J. Geophys. Res., 105, 2969-2980.
 
-Zandt, G., C. J. Ammon (1995), Continental crust composition constrained by measurements of crustal Poisson's ratio, Nature, 374, 152-154.
+\bibitem{} Zandt, G., C. J. Ammon (1995), Continental crust composition constrained by measurements of crustal Poisson's ratio, Nature, 374, 152-154.
+
+\end{thebibliography}
 

### 2014-01-28 python/lithologyPlots.py

@@ -336,11 +336,11 @@ if __name__  == "__main__":
     plt.figure()
 
     # plt.plot(x, rf/pb, "-o", ms = 12, c = "gray", mfc = "black", label = "Rudnick and Fountain")
-    plt.plot(x, pr/pb, "-o", ms = 12, c = "gray", mfc = "black", label = "Pakiser and Robinson")
-    plt.plot(x, sm/pb, "-^", ms = 12, c = "gray", mfc = "white", label = "Smithson")
-    plt.plot(x, cm/pb, "-o", ms = 12, c = "gray", mfc = "white", label = "Christensen and Mooney")
-    plt.plot(x, rg/pb, "-D", ms = 12, c = "gray", mfc = "gray", label = "Rudnick and Gao")
-    plt.plot(x, cs/pb, "-s", ms = 12, c = "gray", mfc = "gray", label = "This study, Can. Shield")
+    plt.plot(x, pr/pb, "-o", ms = 12, c = "gray", mfc = "black", label = "Pakiser and Robinson (1966)")
+    plt.plot(x, sm/pb, "-^", ms = 12, c = "gray", mfc = "white", label = "Smithson (1978)")
+    plt.plot(x, cm/pb, "-o", ms = 12, c = "gray", mfc = "white", label = "Christensen and Mooney (1995)")
+    plt.plot(x, rg/pb, "-D", ms = 12, c = "gray", mfc = "gray", label = "Rudnick and Gao (1995)")
+    plt.plot(x, cs/pb, "-s", ms = 12, c = "gray", mfc = "gray", label = "This study, Canadian Shield")
 
     plt.xticks(x, labels, size = 14)
     plt.tick_params(

### 2014-04-14 lyx/tex2lyx2texSample.tex
thunk

%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,msc]{ubcthesis}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{longtable}
\usepackage{url}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[numbers]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%%
%% This is file `ubcsample.tex',
%% generated with the docstrip utility.
%
% The original source files were:
%
% ubcthesis.dtx  (with options: `ubcsampletex')
%% 
%% This file was generated from the ubcthesis package.
%% --------------------------------------------------------------
%% 
%% Copyright (C) 2001
%% Michael McNeil Forbes
%% mforbes@alum.mit.edu
%% 
%% This file may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License, either version 1.2
%% of this license or (at your option) any later version.
%% The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% LaTeX Project Public License for more details.
%% 
%% This program consists of the files ubcthesis.dtx, ubcthesis.ins, and
%% the sample figures fig.eps and fig.fig.
%% 
%% This file may be modified and used as a base for your thesis without
%% including the licence agreement as long as the content (i.e. textual
%% body) of the file is completely rewritten. You must, however, change
%% the name of the file.
%% 
%% This file may only be distributed together with a copy of this
%% program. You may, however, distribute this program without generated
%% files such as this one.
%%

% This Sample thesis requires \LaTeX2e
\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\ProvidesFile{ubcsample.tex}[2010/06/13 v1.66 ^^J
 University of British Columbia Sample Thesis]
% This is the \documentclass[]{} command.  The manditory argument
% specifies the "flavour" of thesis (ubcthesis for UBC).  The
% optional arguments (in []) specify options that affect how the
% thesis is displayed.  Please see the ubcthesis documentation for
% details about the options.

%
% To compile this sample thesis, issue the following commands:
% latex ubcsample
% bibtex ubcsample
% latex ubcsample
% latex ubcsample
% latex ubcsample
%
% To view use xdvi (on unix systems):
% xdvi ubcsample.dvi
%
% To make a postscript file, use dvips:
% dvips -o ubcsample.ps ubcsample.dvi
%
% To view the postscript file, use ghostview or gv (on unix systems):
% gv ubcsample.ps
%
%************************************************
% Optional packages.
%
% The use of these packages is optional, but they provide various
% tools for more flexible formating.  The sample thesis uses these,
% but if you remove the example code, you should be able to exclude
% these packages.  Only standard packages have been described here;
% they should be installed with any complete LaTeX instalation, but
% if not, you can find them at the Comprehensive TeX Archive Network
% (CTAN): http://www.ctan.org/
%

%******** afterpage ***************************
% This package allows you to issue commands at the end of the current
% page.  A good use for this is to use the command
% \afterpage{\clearpage} right after a figure.  This will cause the
% figure to be inserted on the page following the current one (or on
% the current page if it will fit) but will not break the page in the
% middle.
\usepackage{afterpage}

%******** float *********************************
% This package allows you to customize the style of
% "floats"---floating objects such as figures and tables.  In
% addition, it allows you to define additional floating objects which
% may be included in a list similar to that produces by \listoftables
% and \listoffigures.  Common uses include introducing floats for
% programs and other code bits in Compute Science and Chemical Schema.
\usepackage{float}

%******** tocloft *******************************
% This package allows you to customize and define custom lists such
% as a list of programs or Chemical Scheme.  Note: if you use the
% subfigure package, you must specify that you do as an option here.
% The title option uses the default formatting.  We do not use this
% here as the default formatting is acceptable.  Use the float
% package instead unless you need the extra formatting control
% provided by tocloft.
%\usepackage[subfigure, titles]{tocloft}

%******** alltt *********************************
% The alltt package allows you to include files and have them
% formatted in a verbatim fashion.  This is useful for including
% source code from an additional file.
%\usepackage{alltt}

%******** listings ******************************
% The listings package may be used to include chunks of source code
% and has facilities for pretty-printing many languages.
%\usepackage{listings}

%******** longtable *****************************
% The longtable package allows you to define tables that span
% multiple pages.
\usepackage{longtable}

%******** graphics and graphicx *****************
% This allows you to include encapsulated postscript files.  If you
% don't have this, comment the \includegraphics{} line following the
% comment "%includegraphics" later in this file.

%******** subfigure *****************************
% The subfigure package allows you to include multiple figures and
% captions within a single figure environment.
%\usepackage{subfigure}

%******** here **********************************
% The here package gives you more control over the placement of
% figures and tables.  In particular, you can specify the placement
% "H" which means "Put the figure here" rather than [h] which means
% "I would suggest that you put the figure here if you think it looks
% good."
%\usepackage{here}

%******** pdflscape ********************************
% This allows you to include landscape layout pages by using the
% |landscape| environment.  The use of |pdflscape| is preferred over
% the standard |lscape| package because it automatically rotates the
% page in the pdf file for easier reading.  (Thanks to Joseph Shea
% for pointing this out.)
\usepackage{pdflscape}

%******** natbib ********************************
% This is a very nice package for bibliographies.  It includes options
% for sorting and compressing bibliographic entries.

%******** psfrag ******************************
% This allows you to replace text in postscript pictures with formated
% latex text.  This allows you to use math in graph labels
% etc. Uncomment the psfrag lines following the "%psfrag" comment
% later in this file if you don't have this package.  The replacements
% will only be visible in the final postscript file: they will be
% listed in the .dvi file but not performed.
\usepackage{psfrag}

%******** hyperref *****************************
% Please read the manual:
% http://www.tug.org/applications/hyperref/manual.html
%
% This adds hyperlinks to your document: with the right viewers (later
% versions of xdvi, acrobat with pdftex, latex2html etc.) this will
% make your equation, figure, citation references etc. hyperlinks so
% that you can click on them.  Also, your table of contents will be
% able to take you to the appropriate sections.  In the viewers that
% support this, the links often appear with an underscore.  This
% underscore will not appear in printed versions.
%
% Note: if you do not use the hypertex option, then the dvips driver
% may be loaded by default.  This will cause the entries in the list
% of figures and list of tables to be on a single line because dvips
% does not deal with hyperlinks on broken lines properly.
%
% NOTE: HYPERREF is sensitive to the ORDER in which it is LOADED.
% For example, it must be loaded AFTER natbib but BEFORE newly
% defined float environments.  See the README file with the hyperref
% for some help with this.  If you have some very obscure errors, try
% first disabling hyperref.  If that fixes the problem, try various
% orderings.
%
% Note also that there is a bug with versions before 2003/11/30
% v6.74m that cause the float package to not function correctly.
% Please ensure you have a current version of this package.  A
% warning will be issued if you leave the date below but do not have
% a current version installed.
%
% Some notes on options: depending on how you build your files, you
% may need to choose the appropriate option (such as [pdftex]) for the
% backend driver (see the hyperref manual for a complete list).  Also,
% the default here is to make links from the page numbers in the table
% of contents and lists of figures etc.  There are other options:
% excluding the [linktocpage] option will make the entire text a
% hyperref, but for some backends will prevent the text from wrapping
% which can look terrible.  There is a [breaklinks=true] option that
% will be set if the backend supports (dvipdfm for example supports
% it but does not work with psfrag.)
%
% Finally, there are many options for choosing the colours of the
% links.  These will be included by default in future versions but
% you should probably consider changing some now for the electronic
% version of your thesis.
\usepackage[unicode=true,
  linktocpage,
  linkbordercolor={0.5 0.5 1},
  citebordercolor={0.5 1 0.5},
  linkcolor=blue]{hyperref}

% If you would like to compile this sample thesis without the
% hyperref package, then you will need to comment out the previous
% \usepackage command and uncomment the following command which will
% put the URL's in a typewriter font but not link them.
%\newcommand\url[1]{\texttt{#1}}

%******** setspace *******************************
% The setspace package allows you to manually set the spacing of the
% file.  UBC may require 1.5 spacing for microfilming of theses.  In
% this case you may obtain this by including this package and issuing
% one of the following commands:
%\usepackage{setspace}
%\singlespacing
%\onehalfspacing
%\doublespacing

% These commands are optional.  The defaults are shown.  You only
% need to include them if you need a different value
\institution{The University Of British Columbia}

% If you are at the Okanagan campus, then you should specify these
% instead.
%\faculty{The College of Graduate Studies}
%\institutionaddress{Okanagan}
\faculty{The Faculty of Graduate Studies}
\institutionaddress{Vancouver}

% You can issue as many of these as you have...
\previousdegree{B.Sc., The University of British Columbia, 1999}
\previousdegree{M.Sc., The University of British Columbia, 2001}
\previousdegree{Ph.D., Massachusetts Institute of Technology, 2005}

% You can override the option setting here.
% \degreetitle{Jack of All Trades}

% These commands are required.
\title{A Sample UBC Thesis}
\subtitle{With a Subtitle}
\author{Michael M$^{\rm c}$Neil Forbes}
\copyrightyear{2000}
\submitdate{\today} % The "\ " is required after
                                      % \monthname to prevent the
                                      % command from eating the space.
\program{Physics}

% These commands are presently not required for UBC theses as the
% advisor's name and title are not presently required anywhere.
%\advisor{Ariel R.~Zhitnitsky}
%\advisortitle{Professor of Physics}

% One might want to override the format of the section and chapter
% numbers.  This shows you how to do it.  Note that the current
% format is acceptable for submission to the FoGS: If you wish to modify
% these, you should check with the FoGS explicity. prior to making
% the modifications.
\renewcommand{\thepart}{\Roman{part}}
\renewcommand{\thechapter}{\arabic{chapter}}
\renewcommand{\thesection}{\thechapter.\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
\renewcommand{\theparagraph}{\thesubsubsection.\arabic{paragraph}}
\renewcommand{\thesubparagraph}{\theparagraph.\arabic{subparagraph}}

% Here is an example of a "Program" environment defined with the
% "float" package.  The list of programs will be stored in the file
% ubcsample.lop and the numbering will start with the chapter
% number.  The style will be "ruled".
\floatstyle{ruled}
\newfloat{Program}{htbp}{lop}[chapter]

% Here is the start of the document.

\makeatother

\usepackage{babel}

\begin{document}
% This starts numbering in Roman numerals as required for the thesis
% style.
\frontmatter % Mandatory

% The order of the following components should be preserved.  The order
% listed here is the order currently required by FoGS.
% Mandatory
% Mandatory -  maximum 350 words

\begin{abstract}
The \texttt{genthesis.cls} \LaTeX{} class file and accompanying documents,
such as this sample thesis, are distributed in the hope that it will
be useful but without any warranty (without even the implied warranty
of fitness for a particular purpose). For a description of this file's
purpose, and instructions on its use, see below.

These files are distributed under the GPL which should be included
here in the future. Please let the author know of any changes or improvements
that should be made.

Michael Forbes. mforbes@physics.ubc.ca 
\end{abstract}
\tableofcontents{}% Mandatory
\listoftables
% Mandatory if thesis has tables
\listoffigures
% Mandatory if thesis has figures
\listof{Program}{List of Programs} % Optional
% Any other lists should come here, i.e.
% Abbreviation schemes, definitions, lists of formulae, list of
% schemes, glossary, list of symbols etc.

\chapter{Acknowledgements}

% Optional
This is the place to thank professional colleagues and people who
have given you the most help during the course of your graduate work.

\chapter{Dedication}

% Optional
The dedication is usually quite short, and is a personal rather than
an academic recognition. The \emph{Dedication} does not have to be
titled, but it must appear in the table of contents. If you want to
skip the chapter title but still enter it into the Table of Contents,
use this command \verb|\chapter[Dedication]{}|.

\chapter{Statement of Co-Authorship}

% Manditory if thesis co-written
If any part of your thesis was co-written, you must include a Co-Authorship
statement which gives details of your contribution to the following
areas:
\begin{itemize}
\item identification and design of research program 
\item performing the research 
\item data analyses 
\item manuscript preparation 
\end{itemize}
Note that this section is the last of the preliminary pages (with
lowercase Roman numeral page numbers). It must be placed \emph{before}
the \verb|\mainmatter| command. After that, Arabic numbered pages
will begin.

% Any other unusual prefactory material should come here before the
% main body.

% Now regular page numbering begins.
\mainmatter

% Parts are the largest structural units, but are optional.
%\part{Thesis}

% Chapters are the next main unit.

\chapter{This is a Chapter}

% Sections are a sub-unit

\section{A Section}

Here is a section with some text. Equations look like this $y=x$.%
\footnote{Here is a footnote.%
}

This is an example of a second paragraph in a section so you can see
how much it is indented by.

% Subsections follow

\subsection{This is a Subsection}

Here is an example of a citation: \citep{Forbes:2006ba}. The actual
form of the citation is governed by the bibliographystyle. These citations
are maintained in a BIB\TeX{} file \texttt{sample.bib}. You could
type these directly into the file. For an example of the format to
use look at the file \texttt{ubcsample.bbl} after you compile this
file.%
\footnote{Here is another footnote.%
}

This is an example of a second paragraph in a subsection so you can
see how much it is indented by.

\subsubsection{This is a Subsubsection}

Here are some more citations \citep{LL3:1977,Peccei:1989,Turner:1999}.
If you use the \texttt{natbib} package with the \verb+sort&compress+
option, then the following citation will look the same as the first
citation in this section: \citep{Turner:1999,Peccei:1989,LL3:1977}.

This is an example of a second paragraph in a subsubsection so you
can see how much it is indented by.

\paragraph{This is a Paragraph}

Paragraphs and subparagraphs are the smallest units of text. There
is no subsubsubsection etc.

\subparagraph{This is a Subparagraph}

This is the last level of organisation. If you need more than this,
you should consider reorganizing your work\dots

\begin{equation}
\mathrm{f}(x)=\int_{-\infty}^{\int_{-\infty}^{x}e^{-\frac{y^{2}}{2}}\mathrm{d}{y}}e^{-z^{2}}\mathrm{d}z\end{equation}

In order to show you what a separate page would look like (i.e. without
a chapter heading) I must type some more text. Thus I will babble
a bit and keep babbling for at least one more page\ldots{} What you
should notice is that the chapter titles appear substantially lower
than the continuing text. Babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble.

Babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble babble babble babble babble babble babble
babble babble babble babble.

%
\begin{table}[t]
 %optional [t, b or h];
 \begin{tabular}{|r||r@{.}l}
\hline 
Phoenix  & \$960 & 35\tabularnewline
\hline 
Calgary  & \$250 & 00\tabularnewline
\hline
\end{tabular}\caption[Here is the caption for this wonderful table\ldots{}]{ \label{tab:Table1} Here is the caption for this wonderful table.
It has not been centered and the positioning has been specified to
be at the top of the page. Thus it appears above the babble rather
than below where it is defined in the source file.}

\end{table}

% Force a new page: without this, the quote would appear on the
% previous page.
\newpage{}

\section{Quote}

% It is centered

\begin{quote}
\begin{center}
This is a small poem,\\
 a little poem, a Haiku,\\
 to show you how to.\\
 ---Michael M$^{{\rm c}}$Neil Forbes. 
\par\end{center}
\end{quote}
This small poem shows several features: 
\begin{itemize}
\item The use of the \verb|quote| and \verb|center| environments. 
\item The \verb|\newpage| command has been used to force a page break.
(Sections do not usually start on a new page.) 
\item The pagestyle has been set to suppress the headers using the command
\verb|\thispagestyle{plain}|. Note that using \verb|\pagestyle{plain}|
would have affected all of the subsequent pages. 
\end{itemize}

\section{Programs}

Here we give an example of a new float as defined using the \texttt{float}
package. In the preamble we have use the commands \begin{verbatim}
\floatstyle{ruled} \newfloat{Program}{htbp}{lop}{[}chapter{]}
\end{verbatim} This creates a {}``Program'' environment that may
be use for program fragments. A sample \texttt{python} program is
show as Program~\ref{prog:fib}. (Note that Python places a fairly
restrictive limit on recursion so trying to call this with a large
$n$ before building up the cache is likely to fail unless you increase
the recursion depth.) \begin{Program} \caption{\label{prog:fib} Python program that computes the $n^{{\rm th}}$
Fibonacci number using memoization.}

\begin{verbatim} def fib(n,cache={}): if n < 2: return 1 if n in
cache: return cache{[}n{]} else: result = fib(n-1)+fib(n-2) cache{[}n{]}
= result return result \end{verbatim} \end{Program} Instead of using
a \texttt{verbatim} environment for your program chunks, you might
like to \texttt{include} them within an \texttt{alltt} envrironment
by including the \verb|\usepackage{alltt}| package (see page 187
of the \LaTeX{} book). Another useful package is the \verb|\usepackage{listings}|
which can pretty-print many different types of source code.

% Force a new page
\newpage{}

% Here we provide a short optional argument to \chapter[]{}.  This
% optional argument will appear in the table of contents.  For long
% titles, one should use this to give a single-line entry to the
% table of contents.

\chapter[Another Chapter\ldots{}]{Another Chapter with a Very Long Chapter-name that will Probably
Cause Problems}

This chapter name is very long and does not display properly in the
running headers or in the table of contents. To deal with this, we
provide a shorter version of the title as the optional argument to
the \verb|\chapter[]{}| command.

For example, this chapter's title and associated table of contents
heading and running header was created with\\
 \verb|\chapter[Another Chapter\ldots]{Another Chapter with a Very Long|\\
 \verb|Chapter-name that will Probably Cause Problems}|.

Note that, according to the thesis regulations, the heading included
in the table of contents must be a truncation of the actual heading.

\section{Another Section}

Another bunch of text to demonstrate what this file does. You might
want a list for example:%
\footnote{Here is a footnote in a different chapter. Footnotes should come after
punctuation.%
} 
\begin{itemize}
\item An item in a list. 
\item Another item in a list. 
\end{itemize}

\section*{An Unnumbered Section That is Not Included in the Table of Contents}

Here is an example of a figure environment. %
\begin{figure}[ht]
 %psfrag: comment the following line if not using the psfrag package

\centering{}\psfrag{pie makes me happy!}{$\pi$ makes me happy!}
%includegraphics: comment the following if not using the graphicx package
 \includegraphics[width=0.4\textwidth]{fig} \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of a happy face with a \texttt{psfrag}
replacement. The original figure (drawn in xfig and exported to a
.eps file) has the text {}``pie makes me happy!''. The \texttt{psfrag}
package replaces this with {}``$\pi$ makes me happy!''. Note that
we have used the optional argument for the caption command so that
only a short version of this caption occurs in the list of figures.}

\end{figure}

\afterpage{\clearpage{}} Perhaps I should say that the example
of a figure can be seen in Figure~\ref{fig:happy}. Figure placement
can be tricky with \LaTeX{}\ because figures and tables are treated
as {}``floats'': text can flow around them, but if there is not
enough space, they will appear later. To prevent figures from going
too far, the \verb|\afterpage{\clearpage}| command can be used. This
makes sure that the figure appears on the following page.

The \verb|\clearpage| forces a page break so that the figure can
be placed, but without the the \verb|\afterpage{}| command, the page
would be broken too early (at the \verb|\clearpage| statement). The
\verb|\afterpage{}| command tells \LaTeX{} to issue the command after
the present page has been rendered.

Figures can make a document more enjoyable as demonstrated by Figure~\ref{fig:happy}.

\section{Tables}

We have already included one table:~\ref{tab:Table1}. Another table
is plopped right here. %
\begin{table}[ht]

\centering{}\begin{tabular}{|l||l|l||l|l|}
\hline 
 & \multicolumn{2}{l|}{Singular} & \multicolumn{2}{l|}{Plural}\tabularnewline
\hline 
 & English & \textbf{Gaeilge} & English & \textbf{Gaeilge}\tabularnewline
\hline
\hline 
1st Person & at me & \textbf{agam} & at us & \textbf{againn}\tabularnewline
2nd Person & at you & \textbf{agat} & at you & \textbf{agaibh}\tabularnewline
3rd Person & at him & \textbf{aige} & at them & \textbf{acu}\tabularnewline
 & at her & \textbf{aici} &  & \tabularnewline
\hline
\end{tabular}\caption{ \label{tab:Table2} Another table.}

\end{table}

Well, actually, as with Figure~\ref{fig:happy}, the table does not
necessarily appear right {}``here'' because tables are also {}``floats''.
\LaTeX{} puts them where it can. Becuase of this, one should refer
to floats by their labels rather than by their location. This example
is demonstrated by Table~\ref{tab:Table2}. This one is pretty close,
however.

Another useful package is \verb|\usepackage{longtable}| which provides
the \texttt{longtable} environment. This is nice because it allows
tables to span multiple pages. Table~\ref{tab:longtable} has been
formatted this way.

\begin{center}
\begin{longtable}{|l|l|l|}
\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
\tabularnewline
\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
\endfirsthead
\hline 
\multicolumn{3}{c}{}\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
\hline
\endhead
\hline 
\multicolumn{3}{|r|}{{Continued on next page}}\tabularnewline
\hline
\endfoot
\hline 
0  & (1, 11, 13725)  & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \tabularnewline
274  & (1, 12, 10980)  & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \tabularnewline
5490  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
8235  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
10980  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
13725  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
16470  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
19215  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
21960  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
24705  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
27450  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
30195  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
32940  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
35685  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
38430  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
41175  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
43920  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
46665  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
49410  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
52155  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
54900  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
57645  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
60390  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
63135  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
65880  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
68625  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
71370  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
74115  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
76860  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
79605  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
82350  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
85095  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
87840  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
90585  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
93330  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
96075  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
98820  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
101565  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
104310  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
107055  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
109800  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
112545  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
115290  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
118035  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
120780  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
123525  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
126270  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
129015  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
131760  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
134505  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
137250  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
139995  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
142740  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
145485  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
148230  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
150975  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
153720  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
156465  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
159210  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
161955  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
164700  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
\end{longtable}
\par\end{center}

\subsection*{An Unnumbered Subsection}

Note that if you use subsections or further divisions under an unnumbered
section, then you should make them unnumbered as well otherwise you
will end up with zeros in the section numbering.

\chapter{Landscape Mode}

The landscape mode allows you to rotate a page through 90 degrees.
It is generally not a good idea to make the chapter heading landscape,
but it can be useful for long tables etc.

\begin{landscape} This text should appear rotated, allowing for formatting
of very wide tables etc. Note that this might only work after you
convert the \texttt{dvi} file to a postscript (\texttt{ps}) or \texttt{pdf}
file using \texttt{dvips} or \texttt{dvipdf} etc. This features is
provided by the \verb|lscape| and the \verb|pdflscape| packages.
The latter is preferred if it works as it also rotates the pages in
the pdf file for easier viewing. \end{landscape}

% This file is setup to use a bibtex file sample.bib and uses the
% plain style.  Other styles may be used depending on the conventions
% of your field of study.
% Note: the bibliography must come before the appendices.
\bibliographystyle{plain} \bibliographystyle{plain}
\bibliography{sample}

% Use this to reset the appendix counter.  Note that the FoGS
% requires that the word ``Appendices'' appear in the table of
% contents either before each appendix lable or as a division
% denoting the start of the appendices.  We take the latter option
% here.  This is ensured by making the \texttt{appendicestoc} option
% a default option to the UBC thesis class.

% If you only have one appendix, please uncomment the following line.

\appendix
%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty

\chapter{First Appendix}

Here you can have your appendices. Note that if you only have a single
appendix, you should issue \verb|\renewcommand{\appendicesname}{Appendix}|
before calling \verb|\appendix| to display the singular {}``Appendix''
rather than the default plural {}``Appendices'' some

\chapter{Second Appendix}

Here is the second appendix.

% Indices come here.

\chapter*{Additional Information}

This chapter shows you how to include additional information in your
thesis, the removal of which will not affect the submission. Such
material should be removed before the thesis is actually submitted.

First, the chapter is unnumbered and not included in the Table of
Contents. Second, it is the last section of the thesis, so its removal
will not alter any of the page numbering etc. for the previous sections.
Do not include any floats, however, as these will appear in the initial
lists.

The \texttt{ubcthesis} \LaTeX{} class has been designed to aid you
in producing a thesis that conforms to the requirements of The University
of British Columbia Faculty of Graduate Studies (FoGS).

Proper use of this class and sample is highly recommended---and should
produce a well formatted document that meets the FoGS requirement.
Notwithstanding, complex theses may require additional formatting
that may conflict with some of the requirements. We therefore \emph{highly
recommend} that you consult one of the FoGS staff for assistance and
an assessment of potential problems \emph{before} starting final draft.

While we have attemped to address most of the thesis formatting requirements
in these files, they do not constitute an official set of thesis requirements.
The official requirements are available at the following section of
the FoGS web site:

\begin{center}
\begin{tabular}{|l|}
\hline 
\url{http://www.grad.ubc.ca/students/thesis/}\tabularnewline
\hline
\end{tabular}
\par\end{center}

We recommend that you review these instructions carefully.
\end{document}

### 2014-04-14 paper/bpostlethwaite_Thesis.tex
thunk

@@
 Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
-The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and has enabled many insights into crustal structure and genesis.
+The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Ocean, and has enabled many insights into crustal structure and genesis.
 Recent summary studies utilizing Lithoprobe data include the analysis of {\it Eaton} (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement.

### 2014-04-14 paper/ubcThesisSubmission.tex
thunk

--- 2014-04-14 paper/bpostlethwaite_Thesis.tex
@@
-\documentclass[review]{elsarticle}
-\usepackage{graphicx,multirow,color,natbib,amssymb}
+%%
+%% This is file `ubcsample.tex',
+%% generated with the docstrip utility.
+%
+% The original source files were:
+%
+% ubcthesis.dtx  (with options: `ubcsampletex')
+%%
+%% This file was generated from the ubcthesis package.
+%% --------------------------------------------------------------
+%% NEED TO FIX: Envelope, ensure abs value --- add to figure caption
+%% Lith figure, tighten range.
+%% Rework final geochron conclusion. Data is weak for current claims
+%% figure -> Figure
+\documentclass[msc,oneside]{ubcthesis}
+
+\usepackage{graphicx,multirow,color,amssymb}
 \usepackage[margin=1in]{geometry}
-
-\makeatletter
-\renewcommand\@biblabel[1]{}
-\renewenvironment{thebibliography}[1]
-     {\section*{\refname}%
-      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
-      \list{}%
-           {\leftmargin0pt
-            \@openbib@code
-            \usecounter{enumiv}}%
-      \sloppy
-      \clubpenalty4000
-      \@clubpenalty \clubpenalty
-      \widowpenalty4000%
-      \sfcode`\.\@m}
-     {\def\@noitemerr
-       {\@latex@warning{Empty `thebibliography' environment}}%
-      \endlist}
-\makeatother
-
-\journal{Tectonophysics}
-
 \graphicspath{ {figures/} }
-
 \newcommand{\cc}[1]{\textcolor{red}{#1}}
 
+%******** afterpage ***************************
+% This package allows you to issue commands at the end of the current
+% page.  A good use for this is to use the command
+% \afterpage{\clearpage} right after a figure.  This will cause the
+% figure to be inserted on the page following the current one (or on
+% the current page if it will fit) but will not break the page in the
+% middle.
+\usepackage{afterpage}
+
+%******** float *********************************
+% This package allows you to customize the style of
+% "floats"---floating objects such as figures and tables.  In
+% addition, it allows you to define additional floating objects which
+% may be included in a list similar to that produces by \listoftables
+% and \listoffigures.  Common uses include introducing floats for
+% programs and other code bits in Compute Science and Chemical Schema.
+\usepackage{float}
+
+%******** tocloft *******************************
+% This package allows you to customize and define custom lists such
+% as a list of programs or Chemical Scheme.  Note: if you use the
+% subfigure package, you must specify that you do as an option here.
+% The title option uses the default formatting.  We do not use this
+% here as the default formatting is acceptable.  Use the float
+% package instead unless you need the extra formatting control
+% provided by tocloft.
+%\usepackage[subfigure, titles]{tocloft}
+
+%******** alltt *********************************
+% The alltt package allows you to include files and have them
+% formatted in a verbatim fashion.  This is useful for including
+% source code from an additional file.
+%\usepackage{alltt}
+
+%******** listings ******************************
+% The listings package may be used to include chunks of source code
+% and has facilities for pretty-printing many languages.
+%\usepackage{listings}
+
+%******** longtable *****************************
+% The longtable package allows you to define tables that span
+% multiple pages.
+% \usepackage{longtable}
+
+%******** graphics and graphicx *****************
+% This allows you to include encapsulated postscript files.  If you
+% don't have this, comment the \includegraphics{} line following the
+% comment "%includegraphics" later in this file.
+% \usepackage{graphicx}
+
+%******** subfigure *****************************
+% The subfigure package allows you to include multiple figures and
+% captions within a single figure environment.
+%\usepackage{subfigure}
+
+%******** here **********************************
+% The here package gives you more control over the placement of
+% figures and tables.  In particular, you can specify the placement
+% "H" which means "Put the figure here" rather than [h] which means
+% "I would suggest that you put the figure here if you think it looks
+% good."
+%\usepackage{here}
+
+%******** natbib ********************************
+% This is a very nice package for bibliographies.  It includes options
+% for sorting and compressing bibliographic entries.
+\usepackage[numbers,sort&compress]{natbib}
+
+%******** psfrag ******************************
+% This allows you to replace text in postscript pictures with formated
+% latex text.  This allows you to use math in graph labels
+% etc. Uncomment the psfrag lines following the "%psfrag" comment
+% later in this file if you don't have this package.  The replacements
+% will only be visible in the final postscript file: they will be
+% listed in the .dvi file but not performed.
+\usepackage{psfrag}
+
+%******** hyperref *****************************
+% Please read the manual:
+% http://www.tug.org/applications/hyperref/manual.html
+%
+% This adds hyperlinks to your document: with the right viewers (later
+% versions of xdvi, acrobat with pdftex, latex2html etc.) this will
+% make your equation, figure, citation references etc. hyperlinks so
+% that you can click on them.  Also, your table of contents will be
+% able to take you to the appropriate sections.  In the viewers that
+% support this, the links often appear with an underscore.  This
+% underscore will not appear in printed versions.
+%
+% Note: if you do not use the hypertex option, then the dvips driver
+% may be loaded by default.  This will cause the entries in the list
+% of figures and list of tables to be on a single line because dvips
+% does not deal with hyperlinks on broken lines properly.
+%
+% NOTE: HYPERREF is sensitive to the ORDER in which it is LOADED.
+% For example, it must be loaded AFTER natbib but BEFORE newly
+% defined float environments.  See the README file with the hyperref
+% for some help with this.  If you have some very obscure errors, try
+% first disabling hyperref.  If that fixes the problem, try various
+% orderings.
+%
+% Note also that there is a bug with versions before 2003/11/30
+% v6.74m that cause the float package to not function correctly.
+% Please ensure you have a current version of this package.  A
+% warning will be issued if you leave the date below but do not have
+% a current version installed.
+%
+% Some notes on options: depending on how you build your files, you
+% may need to choose the appropriate option (such as [pdftex]) for the
+% backend driver (see the hyperref manual for a complete list).  Also,
+% the default here is to make links from the page numbers in the table
+% of contents and lists of figures etc.  There are other options:
+% excluding the [linktocpage] option will make the entire text a
+% hyperref, but for some backends will prevent the text from wrapping
+% which can look terrible.  There is a [breaklinks=true] option that
+% will be set if the backend supports (dvipdfm for example supports
+% it but does not work with psfrag.)
+%
+% Finally, there are many options for choosing the colours of the
+% links.  These will be included by default in future versions but
+% you should probably consider changing some now for the electronic
+% version of your thesis.
+\usepackage[unicode=true,
+  linktocpage,
+  linkbordercolor={0.5 0.5 1},
+  citebordercolor={0.5 1 0.5},
+  linkcolor=blue]{hyperref}
+
+% If you would like to compile this sample thesis without the
+% hyperref package, then you will need to comment out the previous
+% \usepackage command and uncomment the following command which will
+% put the URL's in a typewriter font but not link them.
+%\newcommand\url[1]{\texttt{#1}}
+
+%******** setspace *******************************
+% The setspace package allows you to manually set the spacing of the
+% file.  UBC may require 1.5 spacing for microfilming of theses.  In
+% this case you may obtain this by including this package and issuing
+% one of the following commands:
+%\usepackage{setspace}
+%\singlespacing
+%\onehalfspacing
+%\doublespacing
+
+% These commands are optional.  The defaults are shown.  You only
+% need to include them if you need a different value
+\institution{The University Of British Columbia}
+
+% If you are at the Okanagan campus, then you should specify these
+% instead.
+%\faculty{The College of Graduate Studies}
+%\institutionaddress{Okanagan}
+\faculty{The Faculty of Graduate Studies}
+\institutionaddress{Vancouver}
+
+% You can issue as many of these as you have...
+\previousdegree{B.Sc., The University of British Columbia, 1999}
+\previousdegree{M.Sc., The University of British Columbia, 2001}
+\previousdegree{Ph.D., Massachusetts Institute of Technology, 2005}
+
+% You can override the option setting here.
+% \degreetitle{Jack of All Trades}
+
+% These commands are required.
+\title{Seismic velocities and composition of the Canadian crust}
+\author{Ben Postlethwaite}
+\copyrightyear{2014}
+\submitdate{\monthname\ \number\year} % The "\ " is required after
+                                      % \monthname to prevent the
+                                      % command from eating the space.
+\program{GeoPhysics}
+
+% These commands are presently not required for UBC theses as the
+% advisor's name and title are not presently required anywhere.
+%\advisor{Ariel R.~Zhitnitsky}
+%\advisortitle{Professor of Physics}
+
+% One might want to override the format of the section and chapter
+% numbers.  This shows you how to do it.  Note that the current
+% format is acceptable for submission to the FoGS: If you wish to modify
+% these, you should check with the FoGS explicity. prior to making
+% the modifications.
+\renewcommand\thepart         {\Roman{part}}
+\renewcommand\thechapter      {\arabic{chapter}}
+\renewcommand\thesection      {\thechapter.\arabic{section}}
+\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
+\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
+\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
+\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}
+
+\setcounter{tocdepth}{2}
+\setcounter{secnumdepth}{2}
+
+% Here is an example of a "Program" environment defined with the
+% "float" package.  The list of programs will be stored in the file
+% ubcsample.lop and the numbering will start with the chapter
+% number.  The style will be "ruled".
+\floatstyle{ruled}
+\newfloat{Program}{htbp}{lop}[chapter]
+
+% Here is the start of the document.
 \begin{document}
-\begin{frontmatter}
 
-\title{Seismic velocity estimation and the Canadian crust}
+%% This starts numbering in Roman numerals as required for the thesis
+%% style and is mandatory.
+\frontmatter
 
-\author[ubc]{B. Postlethwaite\corref{cor1}}
-\ead{bpostlet@eos.ubc.ca}
-\author[ubc]{M. G. Bostock}
-\ead{bostock@eos.ubc.ca}
-\author[ubc]{N. I. Christensen}
-\author[gsc]{D. B. Snyder}
+%%% The order of the following components should be preserved.  The order
+%%% listed here is the order currently required by FoGS:        \\
+%%% Title (Mandatory)                                           \\
+%%% Preface (Manditory if any collaborator contributions)       \\
+%%% Abstract (Mandatory)                                        \\
+%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
+%%% Acknowledgements (Optional)                                 \\
+%%% Dedication (Optional)                                       \\
 
-\cortext[cor2]{Corresponding author}
-
-\address[ubc]{Earth and Ocean Sciences, UBC, Vancouver BC, Canada}
-\address[gsc]{Geological Survey Canada, Natural Resources Canada, Ottawa ON, Canada}
-%% ------------------------------------------------------------------------ %%
-
-\begin{abstract}
+\maketitle                      %% Mandatory
+\begin{abstract}                %% Mandatory -  maximum 350 words
   It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
@@
 Analysis of results in conjunction with additional velocity estimates from active source seismic studies and a seismic property database affords improved constraints on bulk geological composition of the Canadian landmass that are used to evaluate competing models of crustal formation.
+
+  Ben Postlethwaite
+  post.ben.here@gmail.com
 \end{abstract}
 
-\end{frontmatter}
-%% -----------------------------------------------%%
-\section{Introduction}
+\chapter{Preface} % Manditory if any of the conditions are met
 
+You must include a preface if any part of your research was partly or
+wholly published in articles, was part of a collaboration, or required
+the approval of UBC Research Ethics Boards.
+
+The Preface must include the following:
+
+\begin{itemize}
+\item A statement indicating the relative contributions of all
+  collaborators and co-authors of publications (if any), emphasizing
+  details of your contribution, and stating the proportion of research
+  and writing conducted by you.
+\item A list of any publications arising from work presented in the
+  dissertation, and the chapter(s) in which the work is located.
+\item The name of the particular UBC Research Ethics Board, and the
+  Certificate Number(s) of the Ethics Certificate(s) obtained, if
+  ethics approval was required for the research.
+\end{itemize}
+
+%%% Sections and subsections etc. in the Preface should in general
+%%% not be listed in the table of contents, so use the starred form
+%%% of \section etc.
+\section*{Examples}
+Chapter~\ref{cha:apple_ref} is based on work conducted in UBC's Maple
+Syrup Laboratory by Dr. A.  Apple, Professor B. Boat, and Michael
+McNeil Forbes. I was responsible for tapping the trees in forests X
+and Z, conducted and supervised all boiling operations, and performed
+frequent quality control tests on the product.
+
+A version of chapter~\ref{cha:apple_ref} has been
+published~\cite{Apple:2010}. I conducted all the testing and wrote
+most of the manuscript. The section on ``Testing Implements'' was
+originally drafted by Boat, B.  Check the first pages of this
+chapter to see footnotes with similar information.
+
+Note that this preface must come before the table of contents.  Note
+also that this section ``Examples'' should not be listed in the table
+of contents, so we have used the starred form: \verb|\section*{Example}|.
+
+\tableofcontents                %% Mandatory
+\listoftables                   %% Mandatory if thesis has tables
+\listoffigures                  %% Mandatory if thesis has figures
+
+\chapter{Acknowledgements}      %% Optional
+This is the place to thank professional colleagues and people who have
+given you the most help during the course of your graduate work.
+
+% Now regular page numbering begins.
+\mainmatter
+
+% Parts are the largest structural units, but are optional.
+%\part{Thesis}
+
+% Chapters are the next main unit.
+\chapter{Introduction}
+
+% Sections are a sub-unit
+\section{Geology}
 The Canadian continental lithosphere is composed of at least fifteen large geological provinces representing crustal evolution from the early Archean to the present accretionary growth and orogenic deformation in the Western Cordillera.
@@
 The Interior Platform is largely Archean reworked during the Paleoproterozoic with bore-hole samples indicating amphibolite to granulite-grade Proterozoic metamorphism (Hammer et al., 2010).
-Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean to Paleoproterozoic crystalline basement.
+Extending west from the Interior platform, the Western Cordillera developed in a passive margin setting on and adjacent to late Archean and Paleoproterozoic crystalline basement.
 Between the middle of the Jurassic and Paleocene intense metamorphic, plutonic and uplifting events occurred as the result of Paleozoic and younger intraoceanic-arc accretionary events.
@@
 
-Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
-The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Ocean, and has enabled many insights into crustal structure and genesis.
-Recent summary studies utilizing Lithoprobe data include the analysis of {\it Eaton} (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement.
+\section{Previous Work}
+Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through {\sc Lithoprobe}, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
+The full suite of {\sc Lithoprobe} seismic profiles form a near-continuous transect from the Pacific to the Atlantic Ocean, and has enabled many insights into crustal structure and genesis.
+Recent summary studies utilizing {\sc Lithoprobe} data include the analysis of Eaton (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement.
 Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis whereas the presence of high-velocity, lower-crustal xenoliths in southern Alberta favours thick mafic underplating.
-Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large-scale eclogitization of the lower crust.
+Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lend support to the third hypothesis, that at least in some places, continental Moho forms by large-scale eclogitization of the lower crust.
 Finally, seismic imaging of subhorizontal reflections separated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional d\'ecollement.
@@
 
-In another recent summary study of Lithoprobe data, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust and note that Archean cratons and their boundaries contain structures similar to those found in Proterozoic and Phanerozoic orogens.
+In another recent summary study of {\sc Lithoprobe} data, Hammer and Clowes (2011) discuss broad structural characteristics of the Canadian crust and note that Archean cratons and their boundaries contain structures similar to those found in Proterozoic and Phanerozoic orogens.
 They also discuss the lack of evidence for significant crustal disruption from magmatic intrusion related to subduction, post-orogenic extension or orogenic collapse.
-The authors remark upon the striking uniformity in Moho depth that is imaged between 33 km and 43 km depth along 20,000 km of Lithoprobe transects, despite the diversity of overlying crust.
+The authors remark upon the striking uniformity in Moho depth that is imaged between 33 km and 43 km depth along 20,000 km of {\sc Lithoprobe} transects, despite the diversity of overlying crust.
 Both {\it Cook et al.} (2009) and {\it Hammer and Clowes} (2011) raise the possibility that, in some circumstances, the Moho and the base of seismic reflectivity do not represent the petrological crust-mantle boundary but rather eclogitized lower crust that is seismically indistinguishable from the upper mantle.
@@
 
-The Lith5.0 revised crustal model presented by {\it Perry et al.} (2002) was developed from crustal thickness estimates taken from Lithoprobe seismic profiles.
-This work updates the global Crust5.1 model (Mooney et al, 1998) and introduces broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick ($>$42.5 km) Phanerozoic crust in the south and west of the continental craton.
+The Lith5.0 revised crustal model presented by {\it Perry et al.} (2002) was developed from crustal thickness estimates taken from {\sc Lithoprobe} seismic profiles.
+This work updates the global Crust5.1 model (Mooney et al, 1998) and emphasizes broad crustal features of the Canadian crust such as relatively thin Archean Shield regions (32.5-40.0 km) as compared to the thick ($>$42.5 km) Phanerozoic crust in the south and west of the continental craton.
 
+\section{Summary}
 In this study, we examine the additional constraints that analyses of teleseismic receiver functions place upon Canadian bulk crustal properties when combined with available active source results.
@@
 Our data set comprises the majority of teleseismic data available for the Canadian landmass including both permanent and temporary deployments.
-Many of the temporary deployments are described by {\it Bostock et al.} (2010).
+Many of the temporary deployments are described by Bostock et al.
+(2010).
 Three previous studies employing similar methodologies but with more regional emphases deserve special mention: an examination of crustal thickness and $V_P/V_S$ variation in the Grenville Orogen (Eaton et al., 2005), a structural and compositional characterization of the Superior Province (Darbyshire et al.
@@
 Results from these studies will serve as benchmarks for our expanded effort.
+
 In the following sections we describe the data and methods employed to interrogate the Canadian crust.
@@
 
-%% -----------------------------------------------%%
-
-\section{Data and methods}
+\chapter{Data and Methods}
 
@@
   \includegraphics[width=\textwidth]{stationMap.pdf}
-  \caption{Map showing major Canadian geological provinces and seismic stations contributing to the data discussed in this study.}
+  \caption[Seismic broadband station locations]{Map showing major Canadian geological provinces and seismic stations contributing to the data discussed in this study.}
   \label{map:stationMap}
@@
 
-\subsection{Teleseismic Data Set}
-The primary data utilized in this study are computed from teleseismic {\it P} wave seismograms originating from more than 700 earthquake sources.
+\section{Teleseismic Data Set}
+The primary data utilized in this study are RF processed teleseismic {\it P} wave seismograms originating from more than 700 earthquake sources.
 These earthquakes occurred between 2000 and 2012 and were recorded on subsets of 343 broadband seismic stations distributed across Canada.
@@
 Quality control is performed by admitting only those seismograms with sufficiently high signal to noise ratios to permit visual observation of the dominant {\it P} wave energy and with sufficiently impulsive first arrivals to allow arrival onsets to be accurately measured.
-After selection and filtering more than 80,000 events are available for further processing.
+After selection and filtering more than 80,000 seismograms are available for further processing.
 Seismic stations contributing to the data utilized in this study are shown in figure \ref{map:stationMap}.
@@
 The resulting waveforms contain discrete pulses corresponding to {\it S} waves scattered from subsurface discontinuities including the base of the crust, or Moho.
-Confident identification of these peaks is required to estimate crustal parameters, and improved results can be obtained by transforming the displacement seismogram into their {\it P} and {\it S} contributions based on a 1-D model of propagation.
-This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels (Kennet, 1991).
-The direct {\it P} arrival of the signal on the resulting {\it P} wave component is used as an approximation to the source function as {\it P} component impulse response for teleseismic P approximates a delta function.
+Confident identification of these peaks is required to estimate crustal parameters, and improved results can be obtained by transforming the displacement seismogram into its {\it P} and {\it S} contribution based on a 1-D model of propagation.
+This procedure is accomplished by first rotating the N and E coordinates into radial and transverse directions and then performing a wave field decomposition using the radial and vertical channels (Kennett, 1991).
+The direct {\it P} arrival of the signal on the resulting {\it P} wave component is used as an approximation to the source function because the {\it P} component impulse response for teleseismic P approximates a delta function.
 This windowed source estimate is deconvolved from the {\it S} wave component computed from the wave field decomposition.
 
-We employ an $L_2$, multichannel, frequency domain approach to perform the deconvolution that has the advantage of computational efficiency and does not require {\it a priori} assumptions concerning the noise levels in the data.
+We employ an $L_2$, multichannel, frequency domain approach to perform the deconvolution that has the advantage of computational efficiency and does not require {\it a priori} assumptions concerning noise levels in the data.
 More specifically, we employ a simultaneous deconvolution of $N$ seismograms sharing a similar slowness to compute a single impulse response or receiver function $r(t)$.
@@
 \noindent where $F^{-1}$ is the inverse Fourier transform, $S_n$ represents the $n^{th}$ {\it S} wave seismogram, $P_n$ is the corresponding windowed {\it P} wave seismogram, $^*$ denotes complex conjugation and $\delta$ is the regularization parameter controlling the trade off between model smoothness and data misfit.
-The parameter $\delta$ which is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ is given by
+The parameter $\delta$ which is chosen programatically by minimizing the general cross validation function $GCV(\delta)$ (Golub et al., 1979) is given by
 
@@
 
-\noindent and $\omega$ is the frequency bin in the discrete Fourier transform. All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
+\noindent and $\omega_m$ is the $m^{th}$ discrete frequency. All resulting receiver functions, $r(t)$, are filtered between 0.04Hz and 3.0Hz.
 
-\subsection{Vp/Vs method} \label{section:VpVsMethod}
+\section{Vp/Vs method} \label{section:VpVsMethod}
 
@@
     \includegraphics[width=0.7\textwidth]{reflectedPhases.pdf}
-  \caption{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
+  \caption[Teleseismic reflection phase diagram]{Schematic diagram illustrating geometry of phases for the velocity contrast representing the Moho}
   \label{fig:reflectedPhases}
@@
 A well tested and widely published method for extracting the ratio of {\it P} velocity $V_P$ to {\it S} velocity $V_S$, hereafter referred to as $R$, and crustal thickness (or depth to Moho) $H$, is outlined by {\it Zhu and Kanamori} (2000), hereafter ZK.
-This method exploits the dependence of these parameters to the differential arrival times between the {\it S} wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct {\it P} wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
+This method exploits the dependence of these parameters on the differential arrival times between the {\it S} wave reflected phases $Ps$, $PpPs$, $PsPs$, $PpSs$ and $PsSs$ and the direct {\it P} wave arrival $Pp$ (Fig \ref{fig:reflectedPhases}).
 Note that $PsPs$ and $PpSs$ are kinematic analogues, such that the energy for these two phases arrive simultaneously for a 1-D model.
@@
 
-\begin{equation}
+\begin{equation} \label{eq:tpps}
 t_{Pps}(p_n) = \frac{H}{V_P} \left( \sqrt{ R^2 - V_P^2p_n^2} + \sqrt{1 - V_P^2p_n^2} \right)
@@
 
-\begin{equation}
+\begin{equation} \label{eq:tpss}
 t_{Pss}(p_n)= \frac{2H}{V_P} \sqrt{ R^2 - V_P^2p_n^2}
@@
 \noindent where $p_n$ is the slowness for the $n^{th}$ receiver function.
-Since strong reflected phases occur at sharp velocity contrasts, the Moho, the boundary targeted by ZK, tends to be well represented on most RF's.
+Since strong reflected phases occur at strong velocity contrasts, the Moho, the boundary targeted by ZK, tends to be well represented on most RF's.
 
@@
 In our implementation of the ZK approach, each station is assigned a $V_P$ value corresponding to the Crust 1.0 value for the corresponding $1^\circ$ containing cell.
-In the specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
+In specific cases where data are being compared to previously published results for quality control, $V_P$  values are chosen to match the values in the published study.
 
@@
 \ref{eq:stack}.
-Constructive interference where these bands intersect produce a maximum in the stacking function.
+Constructive interference where these bands intersect produces a maximum in the stacking function.
 The model, $R$ and $H$, that produces this maximum provides the best estimate for the bulk crustal parameters for a given seismic station.
@@
 
-\subsection{Additional data}
+\section{Additional data}
 
-We supplement receiver function estimates of crustal properties with data from controlled source experiments collected and compiled from GSC (Geological Survey of Canada) references by Walter Mooney (personal communication, 2012).
+We supplement receiver function estimates of crustal properties with data from controlled source experiments collected and compiled from Geological Survey of Canada (GSC) references by Walter Mooney (personal communication, 2012).
 The active source data provide $V_P$ and a few $V_S$ estimates for hundreds of locations across Canada, many along transects that cross major geological boundaries, faults and discontinuities.
@@
 The model includes estimates of $V_P$ and $V_S$ information with depth at $1^\circ$ resolution in latitude and longitude.
-In addition to employing local active source data, Crust 1.0 is also constrained by statistically aggregated data from geographically distant but geologically similar regions.
+In addition to employing local active source data Crust 1.0 is also constrained by statistically aggregated data from geographically distant but geologically similar regions.
 It therefore affords a somewhat independent dataset suited for comparison with the crustal estimates calculated in this study.
 
-%% -----------------------------------------------%%
+\chapter{Results}
 
-\section{Results}
+\section{Comparisons}
 
-\subsection{Comparisons}
-
-Three published studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
-The most recent study (Thompson et al., 2010) focused on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
+Three previous studies have provided parameter estimates for $R$ and $H$ for subsets of the seismic stations considered here.
+The most recent study (Thompson et al., 2010) focuses on the northern Canadian Shield and employs both the linear stacking approach used by ZK as well as a phase weighted method.
 A 2007 study of the Superior Province (Darbyshire et al, 2007) and a study of the Grenville Orogen (Eaton et al., 2006) utilize the ZK + semblance approach outlined in section \ref{section:VpVsMethod}.
@@
 study the linear stacking ZK approach is employed.
-Both correlation coefficient and mean difference are provided in (Table \ref{table:comparison}) as measures of similarity.
+Both correlation coefficient and mean difference are provided in Table \ref{table:comparison} as measures of similarity.
 
@@
   \end{tabular}
-  \caption{Comparison of $R$ and $H$ estimates with three published studies}
+  \caption[Multi-study comparison]{Comparison of $R$ and $H$ estimates with three published studies}
 \label{table:comparison}
@@
 Parameter estimates from the Superior Province (Darbyshire et al., 2007) show the lowest correlation in $R$ as well as the highest mean difference in both parameters.
-These differences are well over the averaged computed error, and can not be easily explained by deviation related to noise in the data.
+These differences are well over the averaged computed error, and cannot be easily explained.
 
-\subsection{Full Grid Search Analysis}
+\section{Full Grid Search Analysis}
 
@@
 
-\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended along the kinematic curve in the $H$ - $V_P$ space (Figure \ref{fig:kcurves} a).
-This trade-off contributes to the low correlation in $H$ between the two methods.
-Higher correlation in this quantity (0.96) can be restored by dividing out $V_P$ from $H$.
-The trade-off between the other parameters is lower as there is greater dependence between them such that a small change in one necessitates a large change in the other to maintain travel-time equality, which has the result of a well defined misfit surface (Figure \ref{fig:kcurves} b-f).
+\noindent As a consequence, contours of the misfit surface $s(H,R,V_P)$ are highly extended in $H$ - $V_P$ space (Figure \ref{fig:kcurves}a), contributing to the low correlation in $H$ between the two methods.
+In other words, the ratio $H/V_P$ is better constrained than either of the two parameters $H$, $V_P$ individually.
 
 For all but the cleanest stations, the trade-off between $H$ and $V_P$ renders recovery of $V_P$ unfeasible using the full grid approach.
-A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting less resolution in the reflected phases, illustrates this difficulty.
-RFs are arranged vertically, ordered by increasing slowness, and shown in figure \ref{fig:rfs}.
+A comparison between two stations, ULM, the station with the highest quality RFs, and DORN, a station exhibiting poorer resolution of the reflected phases $Pps$ and $Pss$, illustrates this difficulty.
+RFs are arranged vertically and ordered by increasing slowness, shown in figure \ref{fig:rfs}.
 Cross sections of $s(H,R,V_P)$ through each coordinate plane are displayed for ULM and DORN in figure \ref{fig:kcurves}.
@@
 Estimates computed from the full grid approach are roughly 0.25 km/s higher than the active source data.
-This difference is consistent across all four stations and is exemplified by a 99\% correlation, suggesting possible bias.
+This difference is consistent across all four stations and is exemplified by a 0.99 correlation coefficient, suggesting possible bias.
 However, the limited number of data means this bias is poorly constrained.
 
-There is less correlation between the crustal thickness data, though this is largely attributed to an approximate 10 km difference between the ARVN estimates.
-The reason for this discrepancy may be an anomalously low quality active source reading as both the ZK method and a published study (Thompson et al., 2010) provide an estimate of near 41 km for ARVN, much nearer the full grid value of 42 km.
-Despite lower correlation the apparent bias towards higher values computed from teleseismic data is also maintained for $H$.
-A positive bias in one parameter will tend to bias the other, owing to the relationship between thickness, velocity and times of a scattered phase.
-For a constant $t_{ps}$ one may compute how much $H$ must change for a 0.25 km/s bias in $V_P$ (Zhu and Kanamori, 2000).
-Using an average of the parameter estimates from the two cleanest stations, ULM and WHY, $\delta H$ will be 1.3 km, this fits well with the actual bias in the values.
+There is lower correlation between the crustal thickness estimates, though this is largely attributed to an approximate 10 km difference between the active and passive source values for station ARVN.
+This discrepancy may result from low signal to noise levels at this station although both the ZK method and a published study (Thompson et al., 2010) provide estimates of 41 km for ARVN, much nearer the full grid search value of 42 km.
+Despite lower correlation, the apparent bias towards higher values computed from teleseismic data is also maintained for $H$, and is not unexpected given the dependence of travel times in \ref{eq:tps}-\ref{eq:tpss} on $H/V_P$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{rfDORNULM.eps}
+  \caption[Stacked receiver functions]{Receiver functions stacked vertically for Dorn and ULM.
+ULM features stronger moveout curves for the reflected arrivals which produces higher certainties in resulting parameter estimates.
+}
+  \label{fig:rfs}
+\end{figure}
 
@@
   \includegraphics[width=\textwidth]{kinematicCurves.eps}
-  \caption{}
+  \caption[Misfit contours and kinematic curves]{2D slices through the 3D misfit surface of $s(H,R,V_P)$ for two Seismic stations, DORN and ULM.
+ULM is representative of clean data whereas DORN is representative of data containing lower SNR.
+Contours of the misfit surface are relatively highly extended along kinematic curves for the $V_P$ and $H$ pairings for both Dorn and ULM, a) and d).
+This shows the high degree of trade-off between the two parameters.
+}
   \label{fig:kcurves}
@@
   \centering
-  \includegraphics[width=\textwidth]{rfDORNULM.eps}
-  \caption{}
-  \label{fig:rfs}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width=\textwidth]{vpactive.eps}
-  \caption{Top, {\it P} wave velocity estimates computed from the full parameter search method compared to active source data and values from Crust 1.0.
-Bottom, crustal thickness estimates computed from the full parameter search method compared to active source data and values from Crust 1.0}
+  \caption[Comparison of {\it P} wave velocity estimates]{Top, {\it P} wave velocity estimates computed from the full grid search method compared to active source estimates and values from Crust 1.0.
+Bottom, crustal thickness estimates computed from the full grid search method compared to active source estimates and values from Crust 1.0.}
   \label{fig:vpactive}
@@
 
-\subsection{Crustal Discontinuities} \label{section:discontinuities}
+\section{Crustal Discontinuities} \label{section:discontinuities}
 
@@
   \includegraphics[width=\textwidth]{discPCA.eps}
-  \caption{Left, vertical station depth profiles.
-Right, the SVD singular vectors (principal components) of the station depth profile matrix.
+  \caption[Station depth profiles and SVD analysis]{Left, station depth profiles.
+Right, the singular vectors (principal components) of the station depth profile matrix.
 Each principal component represents a mode that captures some part of the total variance in the dataset.
@@
 We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method.
-The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:discPCA} (left) arranged by increasing crustal thickness.
+The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:discPCA} (left) arranged in order of increasing crustal thickness.
 By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components.
@@
 The first mode is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset.
-It features a maximum near 38.7 km that may be taken as an average Moho depth.
+It features a maximum near 38.7 km that may be taken as a measure of station average Moho depth.
 The second and third modes are dominated by a high amplitude shallow structures (e.g.
 sedimentary layers) within the top 10 km and Moho variability between 30 and 46 km .
-The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10 km to 20 km range.
-All 5 modes display markedly lower energy within a depth window roughly between 20 km and 30 km.
+The fourth and fifth modes account for a combined 11\% of the variance in the dataset, and contain energy at upper-mid crustal depths in the 10 km to 20 km range.
+All 5 modes display relatively little energy within a depth window roughly between 20 km and 30 km.
 
-\subsection{Regional Bulk Crustal Parameters}
+\section{Regional Bulk Crustal Parameters}
 
-We proceed to consider crustal thickness $H$ and velocity ratio, $R$, estimates for stations with assigned $R$ bootstrap errors below $\pm 0.06$ in a geographic context (figure \ref{map:stationMap}).
+We proceed to consider crustal thickness $H$ and velocity ratio $R$ estimates for stations with assigned $R$ bootstrap errors below $\pm 0.06$ in a geographic context (figure \ref{map:stationMap}).
 To compensate for the uneven distribution of seismic stations, weights are applied to both $R$ and $H$ estimates before averaging over a given region.
@@
 The Slave Province averages are similar to those for the Churchill province with a crustal thickness and seismic velocity ratio comparable to Crust 1.0.
-Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK + semblance $R$ values as well as crustal thickness $H$.
+Moving south and east from the Churchill Province through the Superior Province into the Grenville Province, we note an increase in the ZK $R$ values as well as crustal thickness $H$.
 The weighted $R$ average is 1.73 in the Churchill Province increasing slightly to 1.74 in the Superior and jumping to 1.78 in the Grenville.
@@
     \cline{2-7}
-    & \multicolumn{2}{ c }{ZK + semblance} & \multicolumn{3}{ c }{Crust 1.0} & \multicolumn{1}{ c }{Active Source} \\
+    & \multicolumn{2}{ c }{ZK} & \multicolumn{3}{ c }{Crust 1.0} & \multicolumn{1}{ c }{Active Source} \\
     \hline
@@
   \end{tabular}
-  \caption{Comparison of seismic velocity and crustal thickness estimates for Canada and selected geological provinces.}
+  \caption[Regional seismic parameter comparsion]{Comparison of seismic velocity and crustal thickness estimates for Canada and selected geological provinces.}
 \label{table:regionParameters}
@@
 
-%% -----------------------------------------------%%
-\section{Discussion and Conclusions}
-
+\chapter{Discussion}
+\section{Crustal reflectivity}
 Seismic refraction and reflection studies of continental crust have afforded valuable constraints on the processes governing crustal evolution.
 Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust (Conrad, 1925; Jeffreys, 1926).
-Evidence supporting this model was originally presented by {\it Conrad} (1925) which cites observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovi\v{c}i\'{c} in 1909 (Litak and Brown, 1989).
-Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence (Litak and Brown, 1989).
+Evidence supporting this model was originally presented by Conrad (1925) which cites observations of {\it P} arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovi\v{c}i\'{c} in 1909 (Litak and Brown, 1989).
+Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, lacking consistent evidence (Litak and Brown, 1989) for the Conrad Discontinuity.
 Technological improvements in reflection seismology allowed researchers to further argue for a complex heterogeneous crust and propose models wherein smooth random lateral inhomogeneities are superimposed on simple velocity gradient models to reproduce the range of seismic observations (Litak and Brown, 1989; Mereu and Ojo, 1981).
@@
 The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust (e.g.
-Ito, 1998; Mayer, 1996; Meissner et al., 2006).
+Ito, 1999; Mayer et al., 1997; Meissner et al., 2006).
 
@@
  \includegraphics[width=\textwidth]{discInterp.eps}
- \caption{Left, vertical station depth profiles interpolated to a common Moho. Right, normalized stack of depth profiles, grey line shows mean}
+ \caption[Normalized station depth profiles and SVD analysis]{Left, vertical station depth profiles interpolated to a common Moho. Right, normalized stack of depth profiles, grey line shows mean}
  \label{fig:discInterp}
@@
 There is a steady and marked decrease in energy over the bottom third of the profile that cannot be ascribed to sidelobe structure associated with the Moho pulse.
-This observation is ostensibly at odds with the {\it Mooney and Brocher} (1987) high reflectivity lower crustal model.
+This observation is ostensibly at odds with the Mooney and Brocher (1987) high reflectivity lower crustal model.
 Moreover, reflection profiles from {\sc Lithoprobe} transects (Hammer et al., 2010) frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced reflectivity within the lower crust.
-We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive.
+
+We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the two methods are sensitive.
 Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength (Bostock, 1999).
-For 1 Hz teleseismic waves, crustal layering with scale lengths of $\approx$ 800-2000 m should be resolvable.
-Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of $\approx$ 75-300 m.
-We infer then that the vertical scales of lower crustal reflectivity are better represented by scale lengths under $\approx$ 800 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.
+For 1 Hz teleseismic waves, crustal layering with scale lengths of $\sim$ 800-2000 m should be resolvable.
+Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of $\sim$ 75-300 m.
+We infer then that the vertical scales of lower crustal reflectivity are better represented by scale lengths under $\sim$ 800 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.
 Some authors, for example Mayer et al.
-(1996) in their study of the Rhine Graben, have employed the term Conrad Discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
-It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such signals, but the lower energy levels ($<5\%$) in this mode would indicate that Conrad Discontinuity so-defined is at best weakly represented within the Canadian crust.
-This finding is supported by crustal models characterized by a first order continuously increasing velocity profile such as the gradual transition from a more silicic uppercrust to a more mafic lower crust in shields and platforms observed by {\it Meissner} (1986).
-It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in lithoprobe data (Hammer et al., 2010).
-The increased transparency of the lower crust would appear to be better defined and relate to changes in rheology.
-Studies of crustal strength (Jackson, 2002), deformation (Burov and Watts, 2006; Meissner et al., 2006) and the observed $\sim$ 100 m lower crustal lamallae reported by Mooney and Brocher (1987) as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention; though see (Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999) for competing evidence.
+(1997) in their study of the Rhine Graben, have employed the term Conrad Discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
+It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure \ref{fig:disc}, are a manifestation of such signals, but that the lower energy levels ($<5\%$) in this mode would indicate that the Conrad Discontinuity so-defined is at best weakly represented within the Canadian crust.
+This finding is supported by crustal velocity models characterized by a near constant velocity gradient such as the gradual transition from a more silicic upper crust to a more mafic lower crust in shields and platforms observed by Meissner (1986).
+It is also supported by the lack of evidence for a consistent mid-crustal reflection boundary in {\sc lithoprobe} data (Hammer et al., 2010).
+The increased transparency of the lower crust would appear to relate to changes in rheology.
+Studies of crustal strength (Jackson, 2002), deformation (Burov and Watts, 2006; Meissner et al., 2006) and the inferred $\sim$ 100 m lower crustal lamallae reported by Mooney and Brocher (1987) as well as the widely noted reduction in seismicity of the lower crust *** ADD CITATION *** would support this contention; though see Deichmann (1992), Handy and Brun (2004) and Simpson (1999) for competing evidence.
 
+\section{Crustal Composition}
 Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
@@
 After the plate-tectonic revolution, Taylor (1967, 1977), proposed the `island arc' or `andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite.
-Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
+Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic, and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast.
 Hence this model cannot account for the intermediate composition of post-Archean crust.
 Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition (Christensen and Mooney, 1995; Rudnick and Fountain, 1995).
-Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of {\it Smithson et al.} (1981) although, whereas {\it Smithson et al.} utilizes mean crustal $V_P$, this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the Crust 1.0 model.
+Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et al.
+(1981) although, whereas these authors utilize mean crustal $V_P$, this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as represented in the Crust 1.0 model.
 
@@
   \includegraphics[width=\textwidth]{lithology.eps}
-  \caption{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
+  \caption[$V_P$ and $V_S$ estimates for lithologies and regions]{$V_P$ and $V_S$ estimates for common crust lithologies and for major Canadian-Shield provinces}
   \label{fig:lith}
@@
 In Figure \ref{fig:lith} we plot our area-weighted seismic velocity estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a graph of $V_P$ versus $V_S$.
-Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilations of laboratory velocity measurements by Christensen (1996) at 600 MPa and chemical analyses by Le Maitre (1976), all quoted at average crustal pressures of 600 MPa.
+Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilations of laboratory velocity measurements by Christensen (1996) at 600 MPa and chemical analyses by Le Maitre (1976), all quoted at an average crustal pressure of 600 MPa.
 Dashed lines delimit isocontours in $V_P/V_S$.
 The choice of common shield lithologies is motivated by observations along 30 km of paleo-depth represented in the Kapuskasing uplift of the Superior province (Percival and West, 1994).
-The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures comprised largely of tonalites and their metamorphosed equivalents.
+The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures composed largely of tonalites and their metamorphosed equivalents.
 Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons (Rudnick, 1995).
 
-Seismic velocity estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two lithologies: granite geniss and diorite.
-This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition e.g., (Rudnick and Fountain, 1995; Christensen and Mooney, 1995).
+Seismic velocity estimates for Canada and the Canadian Shield fall approximately midway between those of two lithologies: granite gneiss and diorite.
+This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition (e.g., Rudnick and Fountain, 1995; Christensen and Mooney, 1995).
 The Slave, Superior and Churchill provinces possess comparable seismic velocities implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition.
@@
 However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St.
-Lawrence Platform (where the geometric mean seismic ratio is 1.79).
+Lawrence Platform (where the mean seismic ratio is 1.79).
 Moreover, and more importantly, the average $V_P$ value for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition.
@@
   \end{tabular}
-  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Gao (2003) use total $Fe$ expressed as $FeO$.}
+  \caption[Bulk crustal composition comparison]{Bulk crustal composition comparison from selected sources. Rudnick and Gao (2003) use total $Fe$ expressed as $FeO$.}
 \label{table:composition}
@@
   \includegraphics[width=\textwidth]{composition.eps}
-  \caption{Comparisons of crustal composition from selected studies. All estimates have been normalized to the results computed in this study.}
+  \caption[Normalized crustal composition comparison]{Comparisons of crustal composition from selected studies following Rudnick and Gao (2003).
+All estimates have been normalized to the results computed in this study.}
   \label{fig:composition}
@@
 This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by {\it Le Maitre} (1976) to produce an estimate of bulk crustal composition.
-The resulting chemical breakdown is provided in table \ref{table:composition} along with selected previous estimates from global crustal studies.
+The resulting chemical breakdown is provided in Table \ref{table:composition} along with selected previous estimates from global crustal studies.
 Our analysis reveals a crust with marginally higher average $SiO_2$ composition and lower $MgO$ content than these previous studies.
@@
 Also included in this figure is our average crustal composition estimate for the Canadian Shield.
-Its values closely track the global crustal estimates of {\it Christensen and Mooney} (1995) and {\it Smithson} (1978).
+Its values closely track the global crustal estimates of Christensen and Mooney (1995) and Smithson (1978).
 These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss.
@@
 Our results are more consistent with those of Durrheim and Mooney (1991) who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
-The crustal thickness data presented in this study reveals no significant difference between Proterozoic crust (38.9 km) and Archean crust (38.0 km) and this difference is significantly smaller than the 5 km difference reported by Durrheim and Mooney.
+The crustal thickness data presented in this study reveals no significant difference between Proterozoic crust (38.9 km) and Archean crust (38.0 km) in contrast to the 5 km difference reported by Durrheim and Mooney.
 Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) versus the Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
 The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by {\it Durrheim and Mooney} (1991), has been persuasively augmented by recent work (Rudnick 1995; Rudnick and Gao, 2003) to now include a range of possible processes that may explain the data.
+
+\section{Conclusions}
 
@@
 
-%% ------------------------------------------------------------------------ %%
-
 \begin{thebibliography}{1}
-
-\bibitem{} Bostock, M. G. (1998), Mantle stratigraphy and the evolution of the Slave province, J. Geophys. Res., 103, 21183-21200.
 
@@
 
-\bibitem{} Bostock, M. G., M. R. Kumar (2010), Bias in seismic estimates of crustal properties, J. Geophys. Int., 182, 403-407.
-
-\bibitem{} Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere: jelly sandwich or crème brûlée?, GSA today, 16, 4.
+\bibitem{} Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere: jelly sandwich or cr\`{e}me br\^{u}l\`{e}e?, GSA today, 16, 4.
 
@@
 
+\bibitem {} Eaton, D. W. (2006), Multi-genetic origin of the continental Moho: insights from Lithoprobe, Terra Nova, 18, 34–43.
+
 \bibitem{} Eaton, D.
@@
 \bibitem{} Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
+
+\bibitem{} Hammer, P. T., R. M. Clowes (2011), The big picture: a lithospheric cross section of the North American continent, GSA Today, 21, 4-9.
 
@@
 
-\bibitem{} Hammer, P. T., R. M. Clowes (2011), The big picture: a lithospheric cross section of the North American continent, GSA Today, 21, 4-9.
-
 \bibitem{} Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.
 
-\bibitem{} Hoffman, P. F. (1998), United plates of America, the birth of a craton: Early Proterozoic Assembly and Growth of Laurentia, Ann. Rev. Earth Planet, 16, 543-603.
+\bibitem{} Hoffman, P. F. (1988), United plates of America, the birth of a craton: Early Proterozoic Assembly and Growth of Laurentia, Ann. Rev. Earth Planet, 16, 543-603.
 
@@
 
-\bibitem{} Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
+\bibitem{} Kennett, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.
 
@@
 Echtler, E.
-Lüschen, V.
+L\"{u}schen, V.
 Wehrle, B.
-Müller, K-P.
+M\"{u}ller, K-P.
 Bonjer, C.
@@
 
-\bibitem{} Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.
+\bititem{} Mooney, W. D., G. Laske, T. G. Masters (1998), CRUST 5.1: A global crustal model at $5\circ \times 5\circ$, J. Geophys. Res., 103(B1), 727–747
 
-\bibitem{} Mooney, W. D., T. M. Brocher (1987) Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.
+\bibitem{} Mooney, W. D., T. M. Brocher (1987), Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.
+
+\bibitem{} O'Neil, J., D. Francis, R. W. Carlson (2011), Implications of the Nuvvuagittuq Greenstone Belt for the Formation of Earth’s Early Crust, J. Petrology, 52, 985-1009.
+
+\bibitem{} Pakiser, L. C. R. Robinson (1966), Composition of the continental crust as estimated from seismic observations, American Geophysical Union, 10, 620-626.
 
@@
 \end{document}
-
-%% ------------------------------------------------------------------------ %%
+\endinput

### 2014-04-14 paper/ubcsample.tex
thunk

--- 2014-04-14 paper/bpostlethwaite_Thesis.tex
@@
-\documentclass[review]{elsarticle}
-\usepackage{graphicx,multirow,color,natbib,amssymb}
+%%
+%% This is file `ubcsample.tex',
+%% generated with the docstrip utility.
+%
+% The original source files were:
+%
+% ubcthesis.dtx  (with options: `ubcsampletex')
+%%
+%% This file was generated from the ubcthesis package.
+%% --------------------------------------------------------------
+%%
+%% Copyright (C) 2001
+%% Michael McNeil Forbes
+%% mforbes@alum.mit.edu
+%%
+\documentclass[msc,oneside]{ubcthesis}
+
+\usepackage{graphicx,multirow,color,amssymb}
 \usepackage[margin=1in]{geometry}
-
-\makeatletter
-\renewcommand\@biblabel[1]{}
-\renewenvironment{thebibliography}[1]
-     {\section*{\refname}%
-      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
-      \list{}%
-           {\leftmargin0pt
-            \@openbib@code
-            \usecounter{enumiv}}%
-      \sloppy
-      \clubpenalty4000
-      \@clubpenalty \clubpenalty
-      \widowpenalty4000%
-      \sfcode`\.\@m}
-     {\def\@noitemerr
-       {\@latex@warning{Empty `thebibliography' environment}}%
-      \endlist}
-\makeatother
-
-\journal{Tectonophysics}
-
 \graphicspath{ {figures/} }
-
 \newcommand{\cc}[1]{\textcolor{red}{#1}}
 
+%******** afterpage ***************************
+% This package allows you to issue commands at the end of the current
+% page.  A good use for this is to use the command
+% \afterpage{\clearpage} right after a figure.  This will cause the
+% figure to be inserted on the page following the current one (or on
+% the current page if it will fit) but will not break the page in the
+% middle.
+\usepackage{afterpage}
+
+%******** float *********************************
+% This package allows you to customize the style of
+% "floats"---floating objects such as figures and tables.  In
+% addition, it allows you to define additional floating objects which
+% may be included in a list similar to that produces by \listoftables
+% and \listoffigures.  Common uses include introducing floats for
+% programs and other code bits in Compute Science and Chemical Schema.
+\usepackage{float}
+
+%******** tocloft *******************************
+% This package allows you to customize and define custom lists such
+% as a list of programs or Chemical Scheme.  Note: if you use the
+% subfigure package, you must specify that you do as an option here.
+% The title option uses the default formatting.  We do not use this
+% here as the default formatting is acceptable.  Use the float
+% package instead unless you need the extra formatting control
+% provided by tocloft.
+%\usepackage[subfigure, titles]{tocloft}
+
+%******** alltt *********************************
+% The alltt package allows you to include files and have them
+% formatted in a verbatim fashion.  This is useful for including
+% source code from an additional file.
+%\usepackage{alltt}
+
+%******** listings ******************************
+% The listings package may be used to include chunks of source code
+% and has facilities for pretty-printing many languages.
+%\usepackage{listings}
+
+%******** longtable *****************************
+% The longtable package allows you to define tables that span
+% multiple pages.
+\usepackage{longtable}
+
+%******** graphics and graphicx *****************
+% This allows you to include encapsulated postscript files.  If you
+% don't have this, comment the \includegraphics{} line following the
+% comment "%includegraphics" later in this file.
+\usepackage{graphicx}
+
+%******** subfigure *****************************
+% The subfigure package allows you to include multiple figures and
+% captions within a single figure environment.
+%\usepackage{subfigure}
+
+%******** here **********************************
+% The here package gives you more control over the placement of
+% figures and tables.  In particular, you can specify the placement
+% "H" which means "Put the figure here" rather than [h] which means
+% "I would suggest that you put the figure here if you think it looks
+% good."
+%\usepackage{here}
+
+%******** pdflscape ********************************
+% This allows you to include landscape layout pages by using the
+% |landscape| environment.  The use of |pdflscape| is preferred over
+% the standard |lscape| package because it automatically rotates the
+% page in the pdf file for easier reading.  (Thanks to Joseph Shea
+% for pointing this out.)
+\usepackage{pdflscape}
+
+%******** natbib ********************************
+% This is a very nice package for bibliographies.  It includes options
+% for sorting and compressing bibliographic entries.
+\usepackage[numbers,sort&compress]{natbib}
+
+%******** psfrag ******************************
+% This allows you to replace text in postscript pictures with formated
+% latex text.  This allows you to use math in graph labels
+% etc. Uncomment the psfrag lines following the "%psfrag" comment
+% later in this file if you don't have this package.  The replacements
+% will only be visible in the final postscript file: they will be
+% listed in the .dvi file but not performed.
+\usepackage{psfrag}
+
+%******** hyperref *****************************
+% Please read the manual:
+% http://www.tug.org/applications/hyperref/manual.html
+%
+% This adds hyperlinks to your document: with the right viewers (later
+% versions of xdvi, acrobat with pdftex, latex2html etc.) this will
+% make your equation, figure, citation references etc. hyperlinks so
+% that you can click on them.  Also, your table of contents will be
+% able to take you to the appropriate sections.  In the viewers that
+% support this, the links often appear with an underscore.  This
+% underscore will not appear in printed versions.
+%
+% Note: if you do not use the hypertex option, then the dvips driver
+% may be loaded by default.  This will cause the entries in the list
+% of figures and list of tables to be on a single line because dvips
+% does not deal with hyperlinks on broken lines properly.
+%
+% NOTE: HYPERREF is sensitive to the ORDER in which it is LOADED.
+% For example, it must be loaded AFTER natbib but BEFORE newly
+% defined float environments.  See the README file with the hyperref
+% for some help with this.  If you have some very obscure errors, try
+% first disabling hyperref.  If that fixes the problem, try various
+% orderings.
+%
+% Note also that there is a bug with versions before 2003/11/30
+% v6.74m that cause the float package to not function correctly.
+% Please ensure you have a current version of this package.  A
+% warning will be issued if you leave the date below but do not have
+% a current version installed.
+%
+% Some notes on options: depending on how you build your files, you
+% may need to choose the appropriate option (such as [pdftex]) for the
+% backend driver (see the hyperref manual for a complete list).  Also,
+% the default here is to make links from the page numbers in the table
+% of contents and lists of figures etc.  There are other options:
+% excluding the [linktocpage] option will make the entire text a
+% hyperref, but for some backends will prevent the text from wrapping
+% which can look terrible.  There is a [breaklinks=true] option that
+% will be set if the backend supports (dvipdfm for example supports
+% it but does not work with psfrag.)
+%
+% Finally, there are many options for choosing the colours of the
+% links.  These will be included by default in future versions but
+% you should probably consider changing some now for the electronic
+% version of your thesis.
+\usepackage[unicode=true,
+  linktocpage,
+  linkbordercolor={0.5 0.5 1},
+  citebordercolor={0.5 1 0.5},
+  linkcolor=blue]{hyperref}
+
+% If you would like to compile this sample thesis without the
+% hyperref package, then you will need to comment out the previous
+% \usepackage command and uncomment the following command which will
+% put the URL's in a typewriter font but not link them.
+%\newcommand\url[1]{\texttt{#1}}
+
+%******** setspace *******************************
+% The setspace package allows you to manually set the spacing of the
+% file.  UBC may require 1.5 spacing for microfilming of theses.  In
+% this case you may obtain this by including this package and issuing
+% one of the following commands:
+%\usepackage{setspace}
+%\singlespacing
+%\onehalfspacing
+%\doublespacing
+
+% These commands are optional.  The defaults are shown.  You only
+% need to include them if you need a different value
+\institution{The University Of British Columbia}
+
+% If you are at the Okanagan campus, then you should specify these
+% instead.
+%\faculty{The College of Graduate Studies}
+%\institutionaddress{Okanagan}
+\faculty{The Faculty of Graduate Studies}
+\institutionaddress{Vancouver}
+
+% You can issue as many of these as you have...
+\previousdegree{B.Sc., The University of British Columbia, 1999}
+\previousdegree{M.Sc., The University of British Columbia, 2001}
+\previousdegree{Ph.D., Massachusetts Institute of Technology, 2005}
+
+% You can override the option setting here.
+% \degreetitle{Jack of All Trades}
+
+% These commands are required.
+\title{Seismic velocity estimation and the Canadian crust}
+\author{Ben Postlethwaite}
+\copyrightyear{2014}
+\submitdate{\monthname\ \number\year} % The "\ " is required after
+                                      % \monthname to prevent the
+                                      % command from eating the space.
+\program{GeoPhysics}
+
+% These commands are presently not required for UBC theses as the
+% advisor's name and title are not presently required anywhere.
+%\advisor{Ariel R.~Zhitnitsky}
+%\advisortitle{Professor of Physics}
+
+% One might want to override the format of the section and chapter
+% numbers.  This shows you how to do it.  Note that the current
+% format is acceptable for submission to the FoGS: If you wish to modify
+% these, you should check with the FoGS explicity. prior to making
+% the modifications.
+\renewcommand\thepart         {\Roman{part}}
+\renewcommand\thechapter      {\arabic{chapter}}
+\renewcommand\thesection      {\thechapter.\arabic{section}}
+\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
+\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
+\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
+\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}
+
+\setcounter{tocdepth}{2}
+\setcounter{secnumdepth}{2}
+
+% Here is an example of a "Program" environment defined with the
+% "float" package.  The list of programs will be stored in the file
+% ubcsample.lop and the numbering will start with the chapter
+% number.  The style will be "ruled".
+\floatstyle{ruled}
+\newfloat{Program}{htbp}{lop}[chapter]
+
+% Here is the start of the document.
 \begin{document}
-\begin{frontmatter}
 
-\title{Seismic velocity estimation and the Canadian crust}
+%% This starts numbering in Roman numerals as required for the thesis
+%% style and is mandatory.
+\frontmatter
 
-\author[ubc]{B. Postlethwaite\corref{cor1}}
-\ead{bpostlet@eos.ubc.ca}
-\author[ubc]{M. G. Bostock}
-\ead{bostock@eos.ubc.ca}
-\author[ubc]{N. I. Christensen}
-\author[gsc]{D. B. Snyder}
+%%% The order of the following components should be preserved.  The order
+%%% listed here is the order currently required by FoGS:        \\
+%%% Title (Mandatory)                                           \\
+%%% Preface (Manditory if any collaborator contributions)       \\
+%%% Abstract (Mandatory)                                        \\
+%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
+%%% Acknowledgements (Optional)                                 \\
+%%% Dedication (Optional)                                       \\
 
-\cortext[cor2]{Corresponding author}
-
-\address[ubc]{Earth and Ocean Sciences, UBC, Vancouver BC, Canada}
-\address[gsc]{Geological Survey Canada, Natural Resources Canada, Ottawa ON, Canada}
-%% ------------------------------------------------------------------------ %%
-
-\begin{abstract}
+\maketitle                      %% Mandatory
+\begin{abstract}                %% Mandatory -  maximum 350 words
   It has been suggested that processes driving crustal formation in the Archean and Proterozoic were dissimilar and produced crusts with unique bulk properties and average thicknesses.
@@
 
+\section{Previous Work}
 Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999).
@@
 
+\section{Summary}
 In this study, we examine the additional constraints that analyses of teleseismic receiver functions place upon Canadian bulk crustal properties when combined with available active source results.
@@
 
-\subsection{Teleseismic Data Set}
+\section{Teleseismic Data Set}
 The primary data utilized in this study are computed from teleseismic {\it P} wave seismograms originating from more than 700 earthquake sources.
@@
 
-\subsection{Vp/Vs method} \label{section:VpVsMethod}
+\section{Vp/Vs method} \label{section:VpVsMethod}
 
@@
 
-\subsection{Additional data}
+\section{Additional data}
 
@@
 
-%% -----------------------------------------------%%
+\chapter{Results}
 
-\section{Results}
-
-\subsection{Comparisons}
+\section{Comparisons}
 
@@
 
-\subsection{Full Grid Search Analysis}
+\section{Full Grid Search Analysis}
 
@@
 
-\subsection{Regional Bulk Crustal Parameters}
+\section{Regional Bulk Crustal Parameters}
 
@@
 
-%% -----------------------------------------------%%
-\section{Discussion and Conclusions}
-
+\chapter{Discussion}
+\section{Crustal reflectivity}
 Seismic refraction and reflection studies of continental crust have afforded valuable constraints on the processes governing crustal evolution.
@@
 
+\section{Crustal Composition}
 Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
@@
 
+\chapter{Conclusions}
+
 In summary, the teleseismic dataset employed in this study has provided an opportunity for comparison of crustal structure at regional scales and for characterization of the Canadian crust as a whole.
@@
 Finally, the dataset supports the hypothesis that processes driving crustal formation in the Archean and Proterozoic were dissimilar.
-
-%% ------------------------------------------------------------------------ %%
 
@@
 
+%% Use this to reset the appendix counter.  Note that the FoGS
+%% requires that the word ``Appendices'' appear in the table of
+%% contents either before each appendix lable or as a division
+%% denoting the start of the appendices.  We take the latter option
+%% here.  This is ensured by making the \texttt{appendicestoc} option
+%% a default option to the UBC thesis class.
+
+%%% If you only have one appendix, please uncomment the following line.
+% \renewcommand{\appendicesname}{Appendix}
+\appendix
+\chapter{First Appendix}
+Here you can have your appendices.  Note that if you only have a
+single appendix, you should issue
+\verb|\renewcommand{\appendicesname}{Appendix}| before calling
+\verb|\appendix| to display the singular ``Appendix'' rather than the
+default plural ``Appendices''.
+
+\chapter{Second Appendix}
+Here is the second appendix.
+
+%% This changes the headings and chapter titles (no numbers for
+%% example).
+\backmatter
+
+%% Indices come here if you have them.
+
+\chapter*{Additional Information}
+This chapter shows you how to include additional information in your
+thesis, the removal of which will not affect the submission.  Such
+material should be removed before the thesis is actually submitted.
+
+First, the chapter is unnumbered and not included in the Table of
+Contents.  Second, it is the last section of the thesis, so its
+removal will not alter any of the page numbering etc. for the previous
+sections.  Do not include any floats, however, as these will appear in
+the initial lists.
+
+The \texttt{ubcthesis} \LaTeX{} class has been designed to aid you in
+producing a thesis that conforms to the requirements of The
+University of British Columbia Faculty of Graduate Studies (FoGS).
+
+Proper use of this class and sample is highly recommended---and should
+produce a well formatted document that meets the FoGS requirement.
+Notwithstanding, complex theses may require additional formatting that
+may conflict with some of the requirements.  We therefore \emph{highly
+  recommend} that you consult one of the FoGS staff for assistance and
+an assessment of potential problems \emph{before} starting final
+draft.
+
+While we have attemped to address most of the thesis formatting
+requirements in these files, they do not constitute an official set of
+thesis requirements.  The official requirements are available at the
+following section of the FoGS web site:
+\begin{center}
+  \begin{tabular}{|l|}
+    \hline
+    \url{http://www.grad.ubc.ca/current-students/dissertation-thesis-preparation}\\
+    \hline
+  \end{tabular}
+\end{center}
+We recommend that you review these instructions carefully.
+
 \end{document}
-
-%% ------------------------------------------------------------------------ %%
+\endinput

### 2014-04-14 thesis-template/mitsample.tex
thunk

--- 2014-04-14 lyx/tex2lyx2texSample.tex
@@
-%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[english,msc]{ubcthesis}
-\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\usepackage{longtable}
-\usepackage{url}
-\usepackage{graphicx}
-\usepackage{esint}
-\usepackage[numbers]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
 %%
-%% This is file `ubcsample.tex',
+%% This is file `mitsample.tex',
 %% generated with the docstrip utility.
@@
 %
-% ubcthesis.dtx  (with options: `ubcsampletex')
+% ubcthesis.dtx  (with options: `mitsampletex')
 %% 
@@
 \NeedsTeXFormat{LaTeX2e}[1995/12/01]
-\ProvidesFile{ubcsample.tex}[2010/06/13 v1.66 ^^J
- University of British Columbia Sample Thesis]
-% This is the \documentclass[]{} command.  The manditory argument
-% specifies the "flavour" of thesis (ubcthesis for UBC).  The
-% optional arguments (in []) specify options that affect how the
-% thesis is displayed.  Please see the ubcthesis documentation for
-% details about the options.
+\ProvidesFile{mitsample.tex}[2012/04/07 v1.70 ^^J
+ Massachusetts Institute of Technology Sample Thesis]
 
+\documentclass[msc,10pt,oneside]{mitthesis}
 %
-% To compile this sample thesis, issue the following commands:
-% latex ubcsample
-% bibtex ubcsample
-% latex ubcsample
-% latex ubcsample
-% latex ubcsample
+% To compile issue the following commands:
+% latex mitsample
+% bibtex mitsample
+% latex mitsample
+% latex mitsample
+% latex mitsample
 %
 % To view use xdvi (on unix systems):
-% xdvi ubcsample.dvi
+% xdvi mitsample.dvi
 %
 % To make a postscript file, use dvips:
-% dvips -o ubcsample.ps ubcsample.dvi
+% dvips -o mitsample.ps mitsample.dvi
 %
 % To view the postscript file, use ghostview or gv (on unix systems):
-% gv ubcsample.ps
+% gv mitsample.ps
 %
@@
 %
-% The use of these packages is optional, but they provide various
-% tools for more flexible formating.  The sample thesis uses these,
-% but if you remove the example code, you should be able to exclude
-% these packages.  Only standard packages have been described here;
-% they should be installed with any complete LaTeX instalation, but
-% if not, you can find them at the Comprehensive TeX Archive Network
-% (CTAN): http://www.ctan.org/
+% The use of these packages is optional: they are standard now and
+% should be installed on your system, but if they are not, you might
+% have to comment out the appropriate lines to get this file to
+% compile.
 %
+%******** natbib ********************************
+% This is a very nice package for bibliographies.  It includes options
+% for sorting and compressing bibliographic entries.
+\usepackage[numbers,sort&compress]{natbib}
+
+%******** graphics and graphicx ******************************
+% This allows you to include encapsulated postscript files.  If you
+% don't have this, comment the \includegraphics{} line following the
+% comment "%includegraphics" later in this file.
+\usepackage{graphicx}
+
+%******** pdflscape ********************************
+% This allows you to include landscape layout pages by using the
+% |landscape| environment.  The use of |pdflscape| is preferred over
+% the standard |lscape| package because it automatically rotates the
+% page in the pdf file for easier reading.  (Thanks to Joseph Shea
+% for pointing this out.)
+\usepackage{pdflscape}
+
+%******** psfrag ******************************
+% This allows you to replace text in postscript pictures with formated
+% latex text.  This allows you to use math in graph labels
+% etc. Uncomment the psfrag lines following the "%psfrag" comment
+% later in this file if you don't have this package.  The replacements
+% will only be visible in the final postscript file: they will be
+% listed in the .dvi file but not performed.
+\usepackage{psfrag}
 
@@
 \usepackage{afterpage}
-
-%******** float *********************************
-% This package allows you to customize the style of
-% "floats"---floating objects such as figures and tables.  In
-% addition, it allows you to define additional floating objects which
-% may be included in a list similar to that produces by \listoftables
-% and \listoffigures.  Common uses include introducing floats for
-% programs and other code bits in Compute Science and Chemical Schema.
-\usepackage{float}
-
-%******** tocloft *******************************
-% This package allows you to customize and define custom lists such
-% as a list of programs or Chemical Scheme.  Note: if you use the
-% subfigure package, you must specify that you do as an option here.
-% The title option uses the default formatting.  We do not use this
-% here as the default formatting is acceptable.  Use the float
-% package instead unless you need the extra formatting control
-% provided by tocloft.
-%\usepackage[subfigure, titles]{tocloft}
-
-%******** alltt *********************************
-% The alltt package allows you to include files and have them
-% formatted in a verbatim fashion.  This is useful for including
-% source code from an additional file.
-%\usepackage{alltt}
-
-%******** listings ******************************
-% The listings package may be used to include chunks of source code
-% and has facilities for pretty-printing many languages.
-%\usepackage{listings}
-
-%******** longtable *****************************
-% The longtable package allows you to define tables that span
-% multiple pages.
-\usepackage{longtable}
-
-%******** graphics and graphicx *****************
-% This allows you to include encapsulated postscript files.  If you
-% don't have this, comment the \includegraphics{} line following the
-% comment "%includegraphics" later in this file.
-
-%******** subfigure *****************************
-% The subfigure package allows you to include multiple figures and
-% captions within a single figure environment.
-%\usepackage{subfigure}
-
-%******** here **********************************
-% The here package gives you more control over the placement of
-% figures and tables.  In particular, you can specify the placement
-% "H" which means "Put the figure here" rather than [h] which means
-% "I would suggest that you put the figure here if you think it looks
-% good."
-%\usepackage{here}
-
-%******** pdflscape ********************************
-% This allows you to include landscape layout pages by using the
-% |landscape| environment.  The use of |pdflscape| is preferred over
-% the standard |lscape| package because it automatically rotates the
-% page in the pdf file for easier reading.  (Thanks to Joseph Shea
-% for pointing this out.)
-\usepackage{pdflscape}
-
-%******** natbib ********************************
-% This is a very nice package for bibliographies.  It includes options
-% for sorting and compressing bibliographic entries.
-
-%******** psfrag ******************************
-% This allows you to replace text in postscript pictures with formated
-% latex text.  This allows you to use math in graph labels
-% etc. Uncomment the psfrag lines following the "%psfrag" comment
-% later in this file if you don't have this package.  The replacements
-% will only be visible in the final postscript file: they will be
-% listed in the .dvi file but not performed.
-\usepackage{psfrag}
 
@@
 
-%******** setspace *******************************
-% The setspace package allows you to manually set the spacing of the
-% file.  UBC may require 1.5 spacing for microfilming of theses.  In
-% this case you may obtain this by including this package and issuing
-% one of the following commands:
-%\usepackage{setspace}
-%\singlespacing
-%\onehalfspacing
-%\doublespacing
-
-% These commands are optional.  The defaults are shown.  You only
-% need to include them if you need a different value
-\institution{The University Of British Columbia}
-
-% If you are at the Okanagan campus, then you should specify these
-% instead.
-%\faculty{The College of Graduate Studies}
-%\institutionaddress{Okanagan}
-\faculty{The Faculty of Graduate Studies}
-\institutionaddress{Vancouver}
+% These commands are optional.  The defaults are shown.
+\institution{Massachusetts Institute of Technology}
+\institutionaddress{Cambridge}
+\program{Physics}
 
@@
 \previousdegree{M.Sc., The University of British Columbia, 2001}
-\previousdegree{Ph.D., Massachusetts Institute of Technology, 2005}
 
@@
 % These commands are required.
-\title{A Sample UBC Thesis}
+\title{A Sample Thesis}
 \subtitle{With a Subtitle}
@@
 \copyrightyear{2000}
-\submitdate{\today} % The "\ " is required after
-                                      % \monthname to prevent the
-                                      % command from eating the space.
-\program{Physics}
+\submitdate{June 2004}
 
-% These commands are presently not required for UBC theses as the
-% advisor's name and title are not presently required anywhere.
-%\advisor{Ariel R.~Zhitnitsky}
-%\advisortitle{Professor of Physics}
+% These commands are required by MIT.
+\advisor{Frank Wilczek}
+\advisortitle{Herman Feshbach Professor of Physics}
+\chairman{Thomas Greytak}{Professor and Associate Department Head for
+  Education}
+% One might want to override the format of the section and chapter
+% numbers.  This shows you how to do it.  Note that
+\renewcommand\thepart         {\Roman{part}}
+\renewcommand\thechapter      {\arabic{chapter}}
+\renewcommand\thesection      {\thechapter.\arabic{section}}
+\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
+\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
+\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
+\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}
 
-% One might want to override the format of the section and chapter
-% numbers.  This shows you how to do it.  Note that the current
-% format is acceptable for submission to the FoGS: If you wish to modify
-% these, you should check with the FoGS explicity. prior to making
-% the modifications.
-\renewcommand{\thepart}{\Roman{part}}
-\renewcommand{\thechapter}{\arabic{chapter}}
-\renewcommand{\thesection}{\thechapter.\arabic{section}}
-\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
-\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
-\renewcommand{\theparagraph}{\thesubsubsection.\arabic{paragraph}}
-\renewcommand{\thesubparagraph}{\theparagraph.\arabic{subparagraph}}
-
-% Here is an example of a "Program" environment defined with the
-% "float" package.  The list of programs will be stored in the file
-% ubcsample.lop and the numbering will start with the chapter
-% number.  The style will be "ruled".
-\floatstyle{ruled}
-\newfloat{Program}{htbp}{lop}[chapter]
+\setcounter{tocdepth}{2}
+\setcounter{secnumdepth}{2}
 
 % Here is the start of the document.
+\begin{document}
 
-\makeatother
+% Unlike the UBC thesis, page numbering for MIT theses should start
+% at 1 and continue.  Thus, there is no \frontmatter command issued
+% here as there was for the UBC thesis.
 
-\usepackage{babel}
-
-\begin{document}
-% This starts numbering in Roman numerals as required for the thesis
-% style.
-\frontmatter % Mandatory
-
-% The order of the following components should be preserved.  The order
-% listed here is the order currently required by FoGS.
-% Mandatory
-% Mandatory -  maximum 350 words
-
+\maketitle
+\authorizationform
 \begin{abstract}
-The \texttt{genthesis.cls} \LaTeX{} class file and accompanying documents,
-such as this sample thesis, are distributed in the hope that it will
-be useful but without any warranty (without even the implied warranty
-of fitness for a particular purpose). For a description of this file's
-purpose, and instructions on its use, see below.
+  The \texttt{genthesis.cls} \LaTeX{} class file and accompanying
+  documents, such as this sample thesis, are distributed in the hope
+  that it will be useful but without any warranty (without even the
+  implied warranty of fitness for a particular purpose).  For a
+  description of this file's purpose, and instructions on its use, see
+  below.
 
   These files are distributed under the GPL which should be included
-here in the future. Please let the author know of any changes or improvements
-that should be made.
+  here in the future.  Please let the author know of any changes or
+  improvements that should be made.
 
-Michael Forbes. mforbes@physics.ubc.ca 
+  Michael Forbes.
+  mforbes@alum.mit.edu
 \end{abstract}
-\tableofcontents{}% Mandatory
+
+\tableofcontents
 \listoftables
-% Mandatory if thesis has tables
 \listoffigures
-% Mandatory if thesis has figures
-\listof{Program}{List of Programs} % Optional
 % Any other lists should come here, i.e.
 % Abbreviation schemes, definitions, lists of formulae, list of
-% schemes, glossary, list of symbols etc.
+% schemes, etc.
+
+\chapter{Preface}
+These papers have been published earlier\ldots.
 
 \chapter{Acknowledgements}
+Thank you mother here.
 
-% Optional
-This is the place to thank professional colleagues and people who
-have given you the most help during the course of your graduate work.
+% Force a new page.
+\newpage
 
-\chapter{Dedication}
+% Any other unusual sections should come here between the
+% acknowledgements and the main body.
 
-% Optional
-The dedication is usually quite short, and is a personal rather than
-an academic recognition. The \emph{Dedication} does not have to be
-titled, but it must appear in the table of contents. If you want to
-skip the chapter title but still enter it into the Table of Contents,
-use this command \verb|\chapter[Dedication]{}|.
+% Suppress the running headers for this page only.
+\thispagestyle{plain}
+\chapter*{Disclaimer} % Unnumbered
+The \texttt{mitthesis} \LaTeX{} class and the accompanying sample files
+are \emph{unofficial} and are not supported by the Massachusetts
+Institute of Technology.  While I have attempted to make the style
+file and sample files conform to all of the requirements set forth by
+the library, you should always consult one of the library staff
+members for assistance with problems \emph{before} starting final
+draft.  You should be able to find the thesis requirements at one of
+the following sites:
+\begin{table}[h]
+  \begin{center}
+    \begin{tabular}{|l|}
+      \hline
+      \url{http://libraries.mit.edu/archives/thesis-specs/}\\
+      \url{http://libraries.mit.edu/archives/index.html}\\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:ubcurls}
+    Potential sources of information regarding thesis preparation at MIT.}
+\end{table}
 
-\chapter{Statement of Co-Authorship}
+% Force a new page.
+\newpage
 
-% Manditory if thesis co-written
-If any part of your thesis was co-written, you must include a Co-Authorship
-statement which gives details of your contribution to the following
-areas:
+% Suppress the running headers for this page only.
+\thispagestyle{plain}
+
+% Here we provide a short optional argument to \chapter[]{}.  This
+% optional argument will appear in the table of contents.  For long
+% titles, one should use this to give a single-line entry to the
+% table of contents.
+\chapter[Poem]{A Japanese Introduction}
+
+% Here is a quote:
+\begin{quote}
+  % It is centered
+  \begin{center}
+    This is a small poem,\\
+    a little poem, a Haiku,\\
+    to show you how to.\\
+    ---Michael Forbes.
+  \end{center}
+\end{quote}
+This small poem shows several features:
 \begin{itemize}
-\item identification and design of research program 
-\item performing the research 
-\item data analyses 
-\item manuscript preparation 
+\item The \verb|\newpage| command has been used to force a page break.
+\item The pagestyle has been set to suppress the headers using the
+  command \verb|\thispagestyle{plain}|.  Note that using
+  \verb|\pagestyle{plain}| would have affected all of the subsequent
+  pages.
+\item The \verb|\chapter[Poem]{A Japanese Introduction}| command has
+  been used with an optional argument to generate a title and to list
+  this ``chapter'' in the table of contents as ``Poem''.  If one did
+  not desire to have an entry in the table of contents, then one would
+  just use the starred command \verb|\chapter*{}|.  The use of an
+  optional argument is useful for long chapter and section titles that
+  take up too much space in the table of contents.
 \end{itemize}
-Note that this section is the last of the preliminary pages (with
-lowercase Roman numeral page numbers). It must be placed \emph{before}
-the \verb|\mainmatter| command. After that, Arabic numbered pages
-will begin.
 
-% Any other unusual prefactory material should come here before the
-% main body.
-
-% Now regular page numbering begins.
-\mainmatter
-
-% Parts are the largest structural units, but are optional.
-%\part{Thesis}
+% Parts are the largest units
+\part{Thesis}
 
 % Chapters are the next main unit.
-
 \chapter{This is a Chapter}
@@
 % Sections are a sub-unit
+\section{A Section}
+Here is a section with some text.  Equations look like this $y=x$.
 
-\section{A Section}
-
-Here is a section with some text. Equations look like this $y=x$.%
-\footnote{Here is a footnote.%
-}
-
-This is an example of a second paragraph in a section so you can see
-how much it is indented by.
+This is an example of a second paragraph in a section so you can
+see how much it is indented by.
 
 % Subsections follow
-
 \subsection{This is a Subsection}
-
-Here is an example of a citation: \citep{Forbes:2006ba}. The actual
-form of the citation is governed by the bibliographystyle. These citations
-are maintained in a BIB\TeX{} file \texttt{sample.bib}. You could
-type these directly into the file. For an example of the format to
-use look at the file \texttt{ubcsample.bbl} after you compile this
-file.%
-\footnote{Here is another footnote.%
-}
+Here is an example of a citation: \cite{Forbes:2006ba}.  The actual
+form of the citation is governed by the bibliographystyle.  These
+citations are maintained in a BIBTeX file \texttt{sample.bib}.  You
+could type these directly into the file.  For an example of the format
+to use look at the file \texttt{mitsample.bbl} after you compile this
+file.
 
@@
 \subsubsection{This is a Subsubsection}
-
-Here are some more citations \citep{LL3:1977,Peccei:1989,Turner:1999}.
+Here are some more citations \cite{LL3:1977,Peccei:1989,Turner:1999}.
 If you use the \texttt{natbib} package with the \verb+sort&compress+
 option, then the following citation will look the same as the first
-citation in this section: \citep{Turner:1999,Peccei:1989,LL3:1977}.
+citation in this section: \cite{Turner:1999,Peccei:1989,LL3:1977}.
 
-This is an example of a second paragraph in a subsubsection so you
-can see how much it is indented by.
+This is an example of a second paragraph in a subsubsection so you can
+see how much it is indented by.
 
 \paragraph{This is a Paragraph}
-
-Paragraphs and subparagraphs are the smallest units of text. There
-is no subsubsubsection etc.
+Paragraphs and subparagraphs are the smallest units of text.  There is
+no subsubsubsection etc.
 
 \subparagraph{This is a Subparagraph}
-
 This is the last level of organisation.  If you need more than this,
@@
 \begin{equation}
-\mathrm{f}(x)=\int_{-\infty}^{\int_{-\infty}^{x}e^{-\frac{y^{2}}{2}}\mathrm{d}{y}}e^{-z^{2}}\mathrm{d}z\end{equation}
+  \mathrm{f}(x)=\int_{-\infty}^{\int_{-\infty}^x
+    e^{-\frac{y^2}{2}}\mathrm{d}{y}}e^{-z^2}\mathrm{d}z
+\end{equation}
 
 In order to show you what a separate page would look like (i.e. without
-a chapter heading) I must type some more text. Thus I will babble
-a bit and keep babbling for at least one more page\ldots{} What you
+a chapter heading) I must type some more text.  Thus I will babble a
+bit and keep babbling for at least one more page\ldots  What you
 should notice is that the chapter titles appear substantially lower
-than the continuing text. Babble babble babble babble babble babble
+than the continuing text. Babble babble
 babble babble babble babble babble babble babble babble babble babble
@@
 babble babble babble babble babble babble babble babble babble babble
-babble babble babble babble babble.
+babble babble babble babble babble babble babble babble babble.
 
@@
 
-%
-\begin{table}[t]
- %optional [t, b or h];
- \begin{tabular}{|r||r@{.}l}
+\begin{table}[t]                 %optional [t, b or h];
+  \begin{tabular}{|r||r@{.}l|}
     \hline
-Phoenix  & \$960 & 35\tabularnewline
+    Phoenix & \$960&35\\
     \hline
-Calgary  & \$250 & 00\tabularnewline
+    Calgary & \$250&00\\
     \hline
-\end{tabular}\caption[Here is the caption for this wonderful table\ldots{}]{ \label{tab:Table1} Here is the caption for this wonderful table.
-It has not been centered and the positioning has been specified to
-be at the top of the page. Thus it appears above the babble rather
-than below where it is defined in the source file.}
-
+  \end{tabular}
+  \caption{
+    \label{tab:Table1}
+    Here is the caption for this wonderful table.Text of Caption}
 \end{table}
 
-% Force a new page: without this, the quote would appear on the
-% previous page.
-\newpage{}
-
-\section{Quote}
-
-% It is centered
-
-\begin{quote}
-\begin{center}
-This is a small poem,\\
- a little poem, a Haiku,\\
- to show you how to.\\
- ---Michael M$^{{\rm c}}$Neil Forbes. 
-\par\end{center}
-\end{quote}
-This small poem shows several features: 
-\begin{itemize}
-\item The use of the \verb|quote| and \verb|center| environments. 
-\item The \verb|\newpage| command has been used to force a page break.
-(Sections do not usually start on a new page.) 
-\item The pagestyle has been set to suppress the headers using the command
-\verb|\thispagestyle{plain}|. Note that using \verb|\pagestyle{plain}|
-would have affected all of the subsequent pages. 
-\end{itemize}
-
-\section{Programs}
-
-Here we give an example of a new float as defined using the \texttt{float}
-package. In the preamble we have use the commands \begin{verbatim}
-\floatstyle{ruled} \newfloat{Program}{htbp}{lop}{[}chapter{]}
-\end{verbatim} This creates a {}``Program'' environment that may
-be use for program fragments. A sample \texttt{python} program is
-show as Program~\ref{prog:fib}. (Note that Python places a fairly
-restrictive limit on recursion so trying to call this with a large
-$n$ before building up the cache is likely to fail unless you increase
-the recursion depth.) \begin{Program} \caption{\label{prog:fib} Python program that computes the $n^{{\rm th}}$
-Fibonacci number using memoization.}
-
-\begin{verbatim} def fib(n,cache={}): if n < 2: return 1 if n in
-cache: return cache{[}n{]} else: result = fib(n-1)+fib(n-2) cache{[}n{]}
-= result return result \end{verbatim} \end{Program} Instead of using
-a \texttt{verbatim} environment for your program chunks, you might
-like to \texttt{include} them within an \texttt{alltt} envrironment
-by including the \verb|\usepackage{alltt}| package (see page 187
-of the \LaTeX{} book). Another useful package is the \verb|\usepackage{listings}|
-which can pretty-print many different types of source code.
-
-% Force a new page
-\newpage{}
-
-% Here we provide a short optional argument to \chapter[]{}.  This
-% optional argument will appear in the table of contents.  For long
-% titles, one should use this to give a single-line entry to the
-% table of contents.
-
-\chapter[Another Chapter\ldots{}]{Another Chapter with a Very Long Chapter-name that will Probably
-Cause Problems}
-
+\chapter[Another Chapter\ldots]{Another Chapter with a Very Long
+  Chapter-name that will Probably Cause Problems}
 This chapter name is very long and does not display properly in the
 running headers or in the table of contents.  To deal with this, we
-provide a shorter version of the title as the optional argument to
-the \verb|\chapter[]{}| command.
-
-For example, this chapter's title and associated table of contents
-heading and running header was created with\\
- \verb|\chapter[Another Chapter\ldots]{Another Chapter with a Very Long|\\
- \verb|Chapter-name that will Probably Cause Problems}|.
-
-Note that, according to the thesis regulations, the heading included
-in the table of contents must be a truncation of the actual heading.
+provide a shorter version of the title as the optional argument to the
+\verb|\chapter[]{}| command.
 
 \section{Another Section}
-
-Another bunch of text to demonstrate what this file does. You might
-want a list for example:%
-\footnote{Here is a footnote in a different chapter. Footnotes should come after
-punctuation.%
-} 
+Another bunch of text to demonstrate what this file does.
+You might want a list for example:
 \begin{itemize}
@@
 
-\section*{An Unnumbered Section That is Not Included in the Table of Contents}
+\section*{An Unnumbered Section That is Not Included in the Table of
+  Contents}
+%% We would like to place the figure here, so we start with [h].
+%% Note that we have located the figure between paragraphs (rather,
+%% before one) so that it does not split up sentences.
+\begin{figure}[ht]
+  \begin{center}
+%% psfrag: comment the following line if not using the psfrag package
+    \psfrag{pie makes me happy!}{$\pi$ makes me happy!}
+%% includegraphics: comment the following if not using the graphicx package
+    \includegraphics[width=0.4\textwidth]{fig.eps}
+    \caption[Happy Face: figure example.]{\label{fig:happy} This is a
+      figure of a happy face with a \texttt{psfrag} replacement.  The
+      original figure (drawn in xfig and exported to a .eps file) has
+      the text ``pie makes me happy!''.  The \texttt{psfrag} package
+      replaces this with ``$\pi$ makes me happy!''.  Note that we have
+      used the optional argument for the caption command so that only
+      a short version of this caption occurs in the list of figures.}
+  \end{center}
+\end{figure}
+\afterpage{\clearpage}
+Here is an example of a figure environment.
+Perhaps I should say that the example of a figure can be seen in
+Figure~\ref{fig:happy}.  Figure placement can be tricky with \LaTeX\
+because figures and tables are treated as ``floats'': text can flow
+around them, but if there is not enough space, they will appear later.
+To prevent figures from going too far, the
+\verb|\afterpage{\clearpage}| command can be used.  This makes sure
+that the figure are typeset at the end of the page (possibly appear on
+their own on the following pages) and before any subsequent text.
 
-Here is an example of a figure environment. %
-\begin{figure}[ht]
- %psfrag: comment the following line if not using the psfrag package
-
-\centering{}\psfrag{pie makes me happy!}{$\pi$ makes me happy!}
-%includegraphics: comment the following if not using the graphicx package
- \includegraphics[width=0.4\textwidth]{fig} \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of a happy face with a \texttt{psfrag}
-replacement. The original figure (drawn in xfig and exported to a
-.eps file) has the text {}``pie makes me happy!''. The \texttt{psfrag}
-package replaces this with {}``$\pi$ makes me happy!''. Note that
-we have used the optional argument for the caption command so that
-only a short version of this caption occurs in the list of figures.}
-
-\end{figure}
-
-\afterpage{\clearpage{}} Perhaps I should say that the example
-of a figure can be seen in Figure~\ref{fig:happy}. Figure placement
-can be tricky with \LaTeX{}\ because figures and tables are treated
-as {}``floats'': text can flow around them, but if there is not
-enough space, they will appear later. To prevent figures from going
-too far, the \verb|\afterpage{\clearpage}| command can be used. This
-makes sure that the figure appears on the following page.
-
-The \verb|\clearpage| forces a page break so that the figure can
-be placed, but without the the \verb|\afterpage{}| command, the page
+The \verb|\clearpage| forces a page break so that the figure can be
+placed, but without the the \verb|\afterpage{}| command, the page
 would be broken too early (at the \verb|\clearpage| statement).  The
@@
 
-Figures can make a document more enjoyable as demonstrated by Figure~\ref{fig:happy}.
-
-\section{Tables}
-
-We have already included one table:~\ref{tab:Table1}. Another table
-is plopped right here. %
-\begin{table}[ht]
-
-\centering{}\begin{tabular}{|l||l|l||l|l|}
-\hline 
- & \multicolumn{2}{l|}{Singular} & \multicolumn{2}{l|}{Plural}\tabularnewline
-\hline 
- & English & \textbf{Gaeilge} & English & \textbf{Gaeilge}\tabularnewline
-\hline
-\hline 
-1st Person & at me & \textbf{agam} & at us & \textbf{againn}\tabularnewline
-2nd Person & at you & \textbf{agat} & at you & \textbf{agaibh}\tabularnewline
-3rd Person & at him & \textbf{aige} & at them & \textbf{acu}\tabularnewline
- & at her & \textbf{aici} &  & \tabularnewline
-\hline
-\end{tabular}\caption{ \label{tab:Table2} Another table.}
-
-\end{table}
-
-Well, actually, as with Figure~\ref{fig:happy}, the table does not
-necessarily appear right {}``here'' because tables are also {}``floats''.
-\LaTeX{} puts them where it can. Becuase of this, one should refer
-to floats by their labels rather than by their location. This example
-is demonstrated by Table~\ref{tab:Table2}. This one is pretty close,
-however.
-
-Another useful package is \verb|\usepackage{longtable}| which provides
-the \texttt{longtable} environment. This is nice because it allows
-tables to span multiple pages. Table~\ref{tab:longtable} has been
-formatted this way.
-
-\begin{center}
-\begin{longtable}{|l|l|l|}
-\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
-\tabularnewline
-\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
-\tabularnewline
-\hline 
-\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
-\endfirsthead
-\hline 
-\multicolumn{3}{c}{}\tabularnewline
-\hline 
-\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
-\hline
-\endhead
-\hline 
-\multicolumn{3}{|r|}{{Continued on next page}}\tabularnewline
-\hline
-\endfoot
-\hline 
-0  & (1, 11, 13725)  & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \tabularnewline
-274  & (1, 12, 10980)  & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \tabularnewline
-5490  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-8235  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-10980  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-13725  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-16470  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-19215  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-21960  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-24705  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-27450  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-30195  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-32940  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-35685  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-38430  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-41175  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-43920  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-46665  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-49410  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-52155  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-54900  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-57645  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-60390  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-63135  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-65880  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-68625  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-71370  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-74115  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-76860  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-79605  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-82350  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-85095  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-87840  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-90585  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-93330  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-96075  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-98820  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-101565  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-104310  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-107055  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-109800  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-112545  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-115290  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-118035  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-120780  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-123525  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-126270  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-129015  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-131760  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-134505  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-137250  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-139995  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-142740  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-145485  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-148230  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-150975  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-153720  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-156465  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-159210  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-161955  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-164700  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-\end{longtable}
-\par\end{center}
+Be careful when using the ``here'' placement option
+\verb|\begin{figure}[ht]| that you place the figure between paragraphs
+in your text, otherwise \LaTeX{} might actually insert it in the
+middle of a sentence (which does not look very good and is frowned
+upon by the editors!)
 
 \subsection*{An Unnumbered Subsection}
-
-Note that if you use subsections or further divisions under an unnumbered
-section, then you should make them unnumbered as well otherwise you
-will end up with zeros in the section numbering.
+Note that if you use subsections or further divisions under an
+unnumbered section, then you should make them unnumbered as well
+otherwise you will end up with zeros in the section numbering.
 
 \chapter{Landscape Mode}
-
-The landscape mode allows you to rotate a page through 90 degrees.
-It is generally not a good idea to make the chapter heading landscape,
+The landscape mode allows you to rotate a page through 90 degrees.  It
+is generally not a good idea to make the chapter heading landscape,
 but it can be useful for long tables etc.
 
-\begin{landscape} This text should appear rotated, allowing for formatting
-of very wide tables etc. Note that this might only work after you
-convert the \texttt{dvi} file to a postscript (\texttt{ps}) or \texttt{pdf}
-file using \texttt{dvips} or \texttt{dvipdf} etc. This features is
+\begin{landscape}
+  This text should appear rotated, allowing for formatting of very
+  wide tables etc.  Note that this might only work after you convert
+  the \texttt{dvi} file to a postscript (\texttt{ps}) or \texttt{pdf}
+  file using \texttt{dvips} or \texttt{dvipdf} etc.  This feature is
   provided by the \verb|lscape| and the \verb|pdflscape| packages.
   The latter is preferred if it works as it also rotates the pages in
-the pdf file for easier viewing. \end{landscape}
+  the pdf file for easier viewing.
+\end{landscape}
 
 % This file is setup to use a bibtex file sample.bib and uses the
-% plain style.  Other styles may be used depending on the conventions
-% of your field of study.
-% Note: the bibliography must come before the appendices.
-\bibliographystyle{plain} \bibliographystyle{plain}
+% plain style.  Note, the bibliography could come after the appendices.
+\bibliographystyle{plain}
 \bibliography{sample}
 
-% Use this to reset the appendix counter.  Note that the FoGS
-% requires that the word ``Appendices'' appear in the table of
-% contents either before each appendix lable or as a division
-% denoting the start of the appendices.  We take the latter option
-% here.  This is ensured by making the \texttt{appendicestoc} option
-% a default option to the UBC thesis class.
-
 % If you only have one appendix, please uncomment the following line.
-
 \appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
 \chapter{First Appendix}
-
-Here you can have your appendices. Note that if you only have a single
-appendix, you should issue \verb|\renewcommand{\appendicesname}{Appendix}|
-before calling \verb|\appendix| to display the singular {}``Appendix''
-rather than the default plural {}``Appendices'' some
+Here you can have your appendices.  Note that if you only have a
+single appendix, you should issue
+\verb|\renewcommand{\appendicesname}{Appendix}| before calling
+\verb|\appendix| to display the singular ``Appendix'' rather than the
+default plural ``Appendices''.
 
 \chapter{Second Appendix}
+Here is the second appendix.
 
-Here is the second appendix.
+%% This changes the headings and chapter titles (no numbers for
+%% example).
+\backmatter
 
@@
 
-\chapter*{Additional Information}
-
-This chapter shows you how to include additional information in your
-thesis, the removal of which will not affect the submission. Such
-material should be removed before the thesis is actually submitted.
-
-First, the chapter is unnumbered and not included in the Table of
-Contents. Second, it is the last section of the thesis, so its removal
-will not alter any of the page numbering etc. for the previous sections.
-Do not include any floats, however, as these will appear in the initial
-lists.
-
-The \texttt{ubcthesis} \LaTeX{} class has been designed to aid you
-in producing a thesis that conforms to the requirements of The University
-of British Columbia Faculty of Graduate Studies (FoGS).
-
-Proper use of this class and sample is highly recommended---and should
-produce a well formatted document that meets the FoGS requirement.
-Notwithstanding, complex theses may require additional formatting
-that may conflict with some of the requirements. We therefore \emph{highly
-recommend} that you consult one of the FoGS staff for assistance and
-an assessment of potential problems \emph{before} starting final draft.
-
-While we have attemped to address most of the thesis formatting requirements
-in these files, they do not constitute an official set of thesis requirements.
-The official requirements are available at the following section of
-the FoGS web site:
-
-\begin{center}
-\begin{tabular}{|l|}
-\hline 
-\url{http://www.grad.ubc.ca/students/thesis/}\tabularnewline
-\hline
-\end{tabular}
-\par\end{center}
-
-We recommend that you review these instructions carefully.
 \end{document}
+\endinput
+%%
+%% End of file `mitsample.tex'.

### 2014-04-14 thesis-template/ubcsample.tex
thunk

--- 2014-04-14 lyx/tex2lyx2texSample.tex
@@
-%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[english,msc]{ubcthesis}
-\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\usepackage{longtable}
-\usepackage{url}
-\usepackage{graphicx}
-\usepackage{esint}
-\usepackage[numbers]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
 %%
@@
 \NeedsTeXFormat{LaTeX2e}[1995/12/01]
-\ProvidesFile{ubcsample.tex}[2010/06/13 v1.66 ^^J
+\ProvidesFile{ubcsample.tex}[2012/04/07 v1.70 ^^J
  University of British Columbia Sample Thesis]
@@
 % details about the options.
-
+\documentclass[msc,oneside]{ubcthesis}
 %
@@
 % comment "%includegraphics" later in this file.
+\usepackage{graphicx}
 
@@
 % for sorting and compressing bibliographic entries.
+\usepackage[numbers,sort&compress]{natbib}
 
@@
 \copyrightyear{2000}
-\submitdate{\today} % The "\ " is required after
+\submitdate{\monthname\ \number\year} % The "\ " is required after
                                       % \monthname to prevent the
@@
 % the modifications.
-\renewcommand{\thepart}{\Roman{part}}
-\renewcommand{\thechapter}{\arabic{chapter}}
-\renewcommand{\thesection}{\thechapter.\arabic{section}}
-\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
-\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
-\renewcommand{\theparagraph}{\thesubsubsection.\arabic{paragraph}}
-\renewcommand{\thesubparagraph}{\theparagraph.\arabic{subparagraph}}
+\renewcommand\thepart         {\Roman{part}}
+\renewcommand\thechapter      {\arabic{chapter}}
+\renewcommand\thesection      {\thechapter.\arabic{section}}
+\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
+\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
+\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
+\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}
+
+\setcounter{tocdepth}{2}
+\setcounter{secnumdepth}{2}
 
@@
 % Here is the start of the document.
+\begin{document}
 
-\makeatother
+%% This starts numbering in Roman numerals as required for the thesis
+%% style and is mandatory.
+\frontmatter
 
-\usepackage{babel}
+%%% The order of the following components should be preserved.  The order
+%%% listed here is the order currently required by FoGS:        \\
+%%% Title (Mandatory)                                           \\
+%%% Preface (Manditory if any collaborator contributions)       \\
+%%% Abstract (Mandatory)                                        \\
+%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
+%%% Acknowledgements (Optional)                                 \\
+%%% Dedication (Optional)                                       \\
 
-\begin{document}
-% This starts numbering in Roman numerals as required for the thesis
-% style.
-\frontmatter % Mandatory
-
-% The order of the following components should be preserved.  The order
-% listed here is the order currently required by FoGS.
-% Mandatory
-% Mandatory -  maximum 350 words
-
-\begin{abstract}
-The \texttt{genthesis.cls} \LaTeX{} class file and accompanying documents,
-such as this sample thesis, are distributed in the hope that it will
-be useful but without any warranty (without even the implied warranty
-of fitness for a particular purpose). For a description of this file's
-purpose, and instructions on its use, see below.
+\maketitle                      %% Mandatory
+\begin{abstract}                %% Mandatory -  maximum 350 words
+  The \texttt{genthesis.cls} \LaTeX{} class file and accompanying
+  documents, such as this sample thesis, are distributed in the hope
+  that it will be useful but without any warranty (without even the
+  implied warranty of fitness for a particular purpose).  For a
+  description of this file's purpose, and instructions on its use, see
+  below.
 
   These files are distributed under the GPL which should be included
-here in the future. Please let the author know of any changes or improvements
-that should be made.
+  here in the future.  Please let the author know of any changes or
+  improvements that should be made.
 
-Michael Forbes. mforbes@physics.ubc.ca 
+  Michael Forbes.
+  mforbes@physics.ubc.ca
 \end{abstract}
-\tableofcontents{}% Mandatory
-\listoftables
-% Mandatory if thesis has tables
-\listoffigures
-% Mandatory if thesis has figures
-\listof{Program}{List of Programs} % Optional
-% Any other lists should come here, i.e.
-% Abbreviation schemes, definitions, lists of formulae, list of
-% schemes, glossary, list of symbols etc.
 
-\chapter{Acknowledgements}
+\chapter{Preface} % Manditory if any of the conditions are met
 
-% Optional
-This is the place to thank professional colleagues and people who
-have given you the most help during the course of your graduate work.
+You must include a preface if any part of your research was partly or
+wholly published in articles, was part of a collaboration, or required
+the approval of UBC Research Ethics Boards.
 
-\chapter{Dedication}
+The Preface must include the following:
 
-% Optional
+\begin{itemize}
+\item A statement indicating the relative contributions of all
+  collaborators and co-authors of publications (if any), emphasizing
+  details of your contribution, and stating the proportion of research
+  and writing conducted by you.
+\item A list of any publications arising from work presented in the
+  dissertation, and the chapter(s) in which the work is located.
+\item The name of the particular UBC Research Ethics Board, and the
+  Certificate Number(s) of the Ethics Certificate(s) obtained, if
+  ethics approval was required for the research.
+\end{itemize}
+
+%%% Sections and subsections etc. in the Preface should in general
+%%% not be listed in the table of contents, so use the starred form
+%%% of \section etc.
+\section*{Examples}
+Chapter~\ref{cha:apple_ref} is based on work conducted in UBC's Maple
+Syrup Laboratory by Dr. A.  Apple, Professor B. Boat, and Michael
+McNeil Forbes. I was responsible for tapping the trees in forests X
+and Z, conducted and supervised all boiling operations, and performed
+frequent quality control tests on the product.
+
+A version of chapter~\ref{cha:apple_ref} has been
+published~\cite{Apple:2010}. I conducted all the testing and wrote
+most of the manuscript. The section on ``Testing Implements'' was
+originally drafted by Boat, B.  Check the first pages of this
+chapter to see footnotes with similar information.
+
+Note that this preface must come before the table of contents.  Note
+also that this section ``Examples'' should not be listed in the table
+of contents, so we have used the starred form: \verb|\section*{Example}|.
+
+\tableofcontents                %% Mandatory
+\listoftables                   %% Mandatory if thesis has tables
+\listoffigures                  %% Mandatory if thesis has figures
+\listof{Program}{List of Programs} %% Optional
+%% Any other lists should come here, i.e.
+%% Abbreviation schemes, definitions, lists of formulae, list of
+%% schemes, glossary, list of symbols etc.
+
+\chapter{Acknowledgements}      %% Optional
+This is the place to thank professional colleagues and people who have
+given you the most help during the course of your graduate work.
+
+\chapter{Dedication} %% Optional
 The dedication is usually quite short, and is a personal rather than
@@
 
-\chapter{Statement of Co-Authorship}
-
-% Manditory if thesis co-written
-If any part of your thesis was co-written, you must include a Co-Authorship
-statement which gives details of your contribution to the following
-areas:
-\begin{itemize}
-\item identification and design of research program 
-\item performing the research 
-\item data analyses 
-\item manuscript preparation 
-\end{itemize}
 Note that this section is the last of the preliminary pages (with
-lowercase Roman numeral page numbers). It must be placed \emph{before}
-the \verb|\mainmatter| command. After that, Arabic numbered pages
-will begin.
+lowercase Roman numeral page numbers).  It must be placed
+\emph{before} the \verb|\mainmatter| command.  After that, Arabic
+numbered pages will begin.
 
@@
 % Chapters are the next main unit.
-
 \chapter{This is a Chapter}
@@
 % Sections are a sub-unit
+\section{A Section}
+Here is a section with some text.  Equations look like this
+$y=x$.\footnote{Here is a footnote.}
 
-\section{A Section}
-
-Here is a section with some text. Equations look like this $y=x$.%
-\footnote{Here is a footnote.%
-}
-
-This is an example of a second paragraph in a section so you can see
-how much it is indented by.
+This is an example of a second paragraph in a section so you can
+see how much it is indented by.
 
 % Subsections follow
-
 \subsection{This is a Subsection}
-
-Here is an example of a citation: \citep{Forbes:2006ba}. The actual
-form of the citation is governed by the bibliographystyle. These citations
-are maintained in a BIB\TeX{} file \texttt{sample.bib}. You could
-type these directly into the file. For an example of the format to
-use look at the file \texttt{ubcsample.bbl} after you compile this
-file.%
-\footnote{Here is another footnote.%
-}
+Here is an example of a citation: \cite{Forbes:2006ba}.  The actual
+form of the citation is governed by the bibliographystyle.  These
+citations are maintained in a BIBTeX file \texttt{sample.bib}.  You
+could type these directly into the file.  For an example of the format
+to use look at the file \texttt{ubcsample.bbl} after you compile this
+file.\footnote{Here is another footnote.}
 
@@
 
-%
-\begin{table}[t]
- %optional [t, b or h];
- \begin{tabular}{|r||r@{.}l}
+\begin{table}[t]                 % optional [t, b or h];
+  \begin{tabular}{|r||r@{.}l|}
     \hline
-Phoenix  & \$960 & 35\tabularnewline
+    Phoenix & \$960&35\\
     \hline
-Calgary  & \$250 & 00\tabularnewline
+    Calgary & \$250&00\\
     \hline
-\end{tabular}\caption[Here is the caption for this wonderful table\ldots{}]{ \label{tab:Table1} Here is the caption for this wonderful table.
-It has not been centered and the positioning has been specified to
-be at the top of the page. Thus it appears above the babble rather
-than below where it is defined in the source file.}
-
+  \end{tabular}
+  \caption[Here is the caption for this wonderful table\ldots]{
+    \label{tab:Table1}
+    Here is the caption for this wonderful table. It has not been
+    centered and the positioning has been specified to be at the top
+    of the page.  Thus it appears above the babble rather than below
+    where it is defined in the source file.}
 \end{table}
@@
 % previous page.
-\newpage{}
+\newpage
 
 \section{Quote}
-
+Here is a quote:
+\begin{quote}
   % It is centered
-
-\begin{quote}
   \begin{center}
@@
     to show you how to.\\
- ---Michael M$^{{\rm c}}$Neil Forbes. 
-\par\end{center}
+    ---Michael M$^{\rm c}$Neil Forbes.
+  \end{center}
 \end{quote}
+
 This small poem shows several features:
@@
 \item The use of the \verb|quote| and \verb|center| environments.
-\item The \verb|\newpage| command has been used to force a page break.
-(Sections do not usually start on a new page.) 
-\item The pagestyle has been set to suppress the headers using the command
-\verb|\thispagestyle{plain}|. Note that using \verb|\pagestyle{plain}|
-would have affected all of the subsequent pages. 
+\item The \verb|\newpage| command has been used to force a page
+  break.  (Sections do not usually start on a new page.)
+\item The pagestyle has been set to suppress the headers using the
+  command \verb|\thispagestyle{plain}|.  Note that using
+  \verb|\pagestyle{plain}| would have affected all of the subsequent
+  pages.
 \end{itemize}
-
 \section{Programs}
-
-Here we give an example of a new float as defined using the \texttt{float}
-package. In the preamble we have use the commands \begin{verbatim}
-\floatstyle{ruled} \newfloat{Program}{htbp}{lop}{[}chapter{]}
-\end{verbatim} This creates a {}``Program'' environment that may
-be use for program fragments. A sample \texttt{python} program is
-show as Program~\ref{prog:fib}. (Note that Python places a fairly
-restrictive limit on recursion so trying to call this with a large
-$n$ before building up the cache is likely to fail unless you increase
-the recursion depth.) \begin{Program} \caption{\label{prog:fib} Python program that computes the $n^{{\rm th}}$
-Fibonacci number using memoization.}
-
-\begin{verbatim} def fib(n,cache={}): if n < 2: return 1 if n in
-cache: return cache{[}n{]} else: result = fib(n-1)+fib(n-2) cache{[}n{]}
-= result return result \end{verbatim} \end{Program} Instead of using
-a \texttt{verbatim} environment for your program chunks, you might
-like to \texttt{include} them within an \texttt{alltt} envrironment
-by including the \verb|\usepackage{alltt}| package (see page 187
-of the \LaTeX{} book). Another useful package is the \verb|\usepackage{listings}|
-which can pretty-print many different types of source code.
+Here we give an example of a new float as defined using the
+\texttt{float} package.  In the preamble we have used the commands
+\begin{verbatim}
+\floatstyle{ruled}
+\newfloat{Program}{htbp}{lop}[chapter]
+\end{verbatim}
+This creates a ``Program'' environment that may be used for program
+fragments.  A sample \texttt{python} program is shown in
+Program~\ref{prog:fib}.  (Note that Python places a fairly restrictive
+limit on recursion so trying to call this with a large $n$ before
+building up the cache is likely to fail unless you increase the
+recursion depth.)
+\begin{Program}
+  \caption{\label{prog:fib} Python program that computes the $n^{\rm
+      th}$ Fibonacci number using memoization.}
+\begin{verbatim}
+def fib(n,_cache={}):
+    if n < 2:
+        return 1
+    if n in _cache:
+        return _cache[n]
+    else:
+        result = fib(n-1)+fib(n-2)
+        _cache[n] = result
+        return result
+\end{verbatim}
+\end{Program}
+Instead of using a \texttt{verbatim} environment for your program
+chunks, you might like to \texttt{include} them within an
+\texttt{alltt} envrironment by including the \verb|\usepackage{alltt}|
+package (see page 187 of the \LaTeX{} book).  Another useful package
+is the \verb|\usepackage{listings}| which can pretty-print many
+different types of source code.
 
 % Force a new page
-\newpage{}
+\newpage
 
-% Here we provide a short optional argument to \chapter[]{}.  This
-% optional argument will appear in the table of contents.  For long
-% titles, one should use this to give a single-line entry to the
-% table of contents.
-
-\chapter[Another Chapter\ldots{}]{Another Chapter with a Very Long Chapter-name that will Probably
-Cause Problems}
+%% Here we provide a short optional argument to \chapter[]{}.  This
+%% optional argument will appear in the table of contents.  For long
+%% titles, one should use this to give a single-line entry to the
+%% table of contents.
+\chapter[Another Chapter\ldots]{Another Chapter with a Very Long
+  Chapter-name that will Probably Cause Problems}
+\label{cha:apple_ref}
 
@@
 running headers or in the table of contents.  To deal with this, we
-provide a shorter version of the title as the optional argument to
-the \verb|\chapter[]{}| command.
+provide a shorter version of the title as the optional argument to the
+\verb|\chapter[]{}| command.
 
-For example, this chapter's title and associated table of contents
-heading and running header was created with\\
+For example, this chapter's title and associated table of contents heading and
+running header was created with\\
 \verb|\chapter[Another Chapter\ldots]{Another Chapter with a Very Long|\\
@@
 
+This Chapter was used as a demonstration in the Preface for how to
+attribute contribution from collaborators.  If there are any such
+contributions, details must be included in the Preface.  If you wish,
+you may additionally use a footnote such as this.\footnote{This
+  chapter is based on work conducted in UBC's Maple Syrup Laboratory
+  by Dr. A. Apple, Professor B. Boat, and C. Cat.}
+
 \section{Another Section}
-
-Another bunch of text to demonstrate what this file does. You might
-want a list for example:%
-\footnote{Here is a footnote in a different chapter. Footnotes should come after
-punctuation.%
-} 
+Another bunch of text to demonstrate what this file does.
+You might want a list for example:\footnote{Here is a footnote in a
+  different chapter.  Footnotes should come after punctuation.}
 \begin{itemize}
@@
 
-\section*{An Unnumbered Section That is Not Included in the Table of Contents}
+\section*{An Unnumbered Section That is Not Included in the Table of
+  Contents}
+\begin{figure}[ht]
+  \begin{center}
+%% psfrag: comment the following line if not using the psfrag package
+    \psfrag{pie makes me happy!}{$\pi$ makes me happy!}
+%% includegraphics: comment the following if not using the graphicx package
+    \includegraphics[width=0.4\textwidth]{fig.eps}
+    \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of
+      a happy face with a \texttt{psfrag} replacement.  The original figure
+      (drawn in xfig and exported to a .eps file) has the text ``pie makes me
+      happy!''.  The \texttt{psfrag} package replaces this with ``$\pi$ makes me
+      happy!''.  Note: the Makefile compiles the sample using pdf\LaTeX\ which
+      cannot use \texttt{psfrag} directly.  For some options that work with
+      pdf\LaTeX, please see this discussion:
+      \url{http://tex.stackexchange.com/questions/11839}.  For the caption, we
+      have used the optional argument for the caption command so that only a
+      short version of this caption occurs in the list of figures.}
+  \end{center}
+\end{figure}
+\afterpage{\clearpage}
+Here is an example of a figure environment.
+Perhaps I should say that the example of a figure can be seen in
+Figure~\ref{fig:happy}.  Figure placement can be tricky with \LaTeX\
+because figures and tables are treated as ``floats'': text can flow
+around them, but if there is not enough space, they will appear later.
+To prevent figures from going too far, the
+\verb|\afterpage{\clearpage}| command can be used.  This makes sure
+that the figure are typeset at the end of the page (possibly appear on
+their own on the following pages) and before any subsequent text.
 
-Here is an example of a figure environment. %
-\begin{figure}[ht]
- %psfrag: comment the following line if not using the psfrag package
-
-\centering{}\psfrag{pie makes me happy!}{$\pi$ makes me happy!}
-%includegraphics: comment the following if not using the graphicx package
- \includegraphics[width=0.4\textwidth]{fig} \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of a happy face with a \texttt{psfrag}
-replacement. The original figure (drawn in xfig and exported to a
-.eps file) has the text {}``pie makes me happy!''. The \texttt{psfrag}
-package replaces this with {}``$\pi$ makes me happy!''. Note that
-we have used the optional argument for the caption command so that
-only a short version of this caption occurs in the list of figures.}
-
-\end{figure}
-
-\afterpage{\clearpage{}} Perhaps I should say that the example
-of a figure can be seen in Figure~\ref{fig:happy}. Figure placement
-can be tricky with \LaTeX{}\ because figures and tables are treated
-as {}``floats'': text can flow around them, but if there is not
-enough space, they will appear later. To prevent figures from going
-too far, the \verb|\afterpage{\clearpage}| command can be used. This
-makes sure that the figure appears on the following page.
-
-The \verb|\clearpage| forces a page break so that the figure can
-be placed, but without the the \verb|\afterpage{}| command, the page
+The \verb|\clearpage| forces a page break so that the figure can be
+placed, but without the the \verb|\afterpage{}| command, the page
 would be broken too early (at the \verb|\clearpage| statement).  The
@@
 
-Figures can make a document more enjoyable as demonstrated by Figure~\ref{fig:happy}.
-
 \section{Tables}
-
 We have already included one table:~\ref{tab:Table1}.  Another table
-is plopped right here. %
+is plopped right here.
 \begin{table}[ht]
-
-\centering{}\begin{tabular}{|l||l|l||l|l|}
+  \begin{center}
+    \begin{tabular}{|l||l|l||l|l|}
       \hline
- & \multicolumn{2}{l|}{Singular} & \multicolumn{2}{l|}{Plural}\tabularnewline
+      &\multicolumn{2}{l|}{Singular}&\multicolumn{2}{l|}{Plural}\\
+      \cline{2-5}
+       &English&\textbf{Gaeilge}&English&\textbf{Gaeilge}\\
+      \hline\hline
+      1st Person&at me&\textbf{agam}&at us&\textbf{againn}\\
+      2nd Person&at you&\textbf{agat}&at you&\textbf{agaibh}\\
+      3rd Person&at him&\textbf{aige}&at them&\textbf{acu}\\
+       &at her&\textbf{aici}& & \\
       \hline
- & English & \textbf{Gaeilge} & English & \textbf{Gaeilge}\tabularnewline
-\hline
-\hline 
-1st Person & at me & \textbf{agam} & at us & \textbf{againn}\tabularnewline
-2nd Person & at you & \textbf{agat} & at you & \textbf{agaibh}\tabularnewline
-3rd Person & at him & \textbf{aige} & at them & \textbf{acu}\tabularnewline
- & at her & \textbf{aici} &  & \tabularnewline
-\hline
-\end{tabular}\caption{ \label{tab:Table2} Another table.}
-
+    \end{tabular}
+    \caption{
+      \label{tab:Table2}
+      Another table.}
+  \end{center}
 \end{table}
-
-Well, actually, as with Figure~\ref{fig:happy}, the table does not
-necessarily appear right {}``here'' because tables are also {}``floats''.
-\LaTeX{} puts them where it can. Becuase of this, one should refer
-to floats by their labels rather than by their location. This example
-is demonstrated by Table~\ref{tab:Table2}. This one is pretty close,
-however.
+Well, actually, as with Figures, tables do not
+necessarily appear right ``here'' because tables are also ``floats''.
+\LaTeX{} puts them where it can.  Because of this, one should refer to
+floats by their labels rather than by their location.  This example is
+demonstrated by Table~\ref{tab:Table2}.  This one is pretty close,
+however.  (Note: you should generally not put tables or figures in the
+middle of a paragraph.  This example is for demonstration purposes
+only.)
 
@@
 formatted this way.
-
 \begin{center}
   \begin{longtable}{|l|l|l|}
-\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
-\tabularnewline
-\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
-\tabularnewline
-\hline 
-\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
+    \caption{\label{tab:longtable}Feasible triples for
+      highly variable Grid}\\
+
+    \hline \multicolumn{1}{|c|}{\textbf{Time (s)}} &
+    \multicolumn{1}{c|}{\textbf{Triple chosen}} &
+    \multicolumn{1}{c|}{\textbf{Other feasible triples}} \\ \hline
     \endfirsthead
-\hline 
-\multicolumn{3}{c}{}\tabularnewline
-\hline 
-\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
-\hline
+
+    \multicolumn{3}{c}%
+    {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+    \hline \multicolumn{1}{|c|}{\textbf{Time (s)}} &
+    \multicolumn{1}{c|}{\textbf{Triple chosen}} &
+    \multicolumn{1}{c|}{\textbf{Other feasible triples}} \\ \hline
     \endhead
-\hline 
-\multicolumn{3}{|r|}{{Continued on next page}}\tabularnewline
-\hline
+
+    \hline \multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
     \endfoot
-\hline 
-0  & (1, 11, 13725)  & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \tabularnewline
-274  & (1, 12, 10980)  & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \tabularnewline
-5490  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-8235  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-10980  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-13725  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-16470  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-19215  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-21960  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-24705  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-27450  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-30195  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-32940  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-35685  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-38430  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-41175  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-43920  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-46665  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-49410  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-52155  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-54900  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-57645  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-60390  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-63135  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-65880  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-68625  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-71370  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-74115  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-76860  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-79605  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-82350  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-85095  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-87840  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-90585  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-93330  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-96075  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-98820  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-101565  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-104310  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-107055  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-109800  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-112545  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-115290  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-118035  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-120780  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-123525  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-126270  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-129015  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-131760  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-134505  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-137250  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-139995  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-142740  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-145485  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-148230  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
-150975  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-153720  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-156465  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-159210  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-161955  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
-164700  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
+
+    \hline \hline
+    \endlastfoot
+
+    0 & (1, 11, 13725) & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \\
+    274 & (1, 12, 10980) & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \\
+    5490 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    8235 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    10980 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    13725 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    16470 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    19215 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    21960 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    24705 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    27450 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    30195 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    32940 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    35685 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    38430 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    41175 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    43920 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    46665 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    49410 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    52155 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    54900 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    57645 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    60390 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    63135 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    65880 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    68625 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    71370 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    74115 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    76860 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    79605 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    82350 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    85095 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    87840 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    90585 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    93330 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    96075 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    98820 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    101565 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    104310 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    107055 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    109800 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    112545 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    115290 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    118035 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    120780 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    123525 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    126270 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    129015 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    131760 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    134505 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    137250 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    139995 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    142740 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    145485 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    148230 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
+    150975 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    153720 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    156465 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    159210 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    161955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
+    164700 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
 \end{longtable}
-\par\end{center}
+\end{center}
 
 \subsection*{An Unnumbered Subsection}
-
-Note that if you use subsections or further divisions under an unnumbered
-section, then you should make them unnumbered as well otherwise you
-will end up with zeros in the section numbering.
+Note that if you use subsections or further divisions under an
+unnumbered section, then you should make them unnumbered as well
+otherwise you will end up with zeros in the section numbering.
 
 \chapter{Landscape Mode}
-
-The landscape mode allows you to rotate a page through 90 degrees.
-It is generally not a good idea to make the chapter heading landscape,
+The landscape mode allows you to rotate a page through 90 degrees.  It
+is generally not a good idea to make the chapter heading landscape,
 but it can be useful for long tables etc.
 
-\begin{landscape} This text should appear rotated, allowing for formatting
-of very wide tables etc. Note that this might only work after you
-convert the \texttt{dvi} file to a postscript (\texttt{ps}) or \texttt{pdf}
-file using \texttt{dvips} or \texttt{dvipdf} etc. This features is
+\begin{landscape}
+  This text should appear rotated, allowing for formatting of very
+  wide tables etc.  Note that this might only work after you convert
+  the \texttt{dvi} file to a postscript (\texttt{ps}) or \texttt{pdf}
+  file using \texttt{dvips} or \texttt{dvipdf} etc.  This feature is
   provided by the \verb|lscape| and the \verb|pdflscape| packages.
   The latter is preferred if it works as it also rotates the pages in
-the pdf file for easier viewing. \end{landscape}
+  the pdf file for easier viewing.
+\end{landscape}
 
-% This file is setup to use a bibtex file sample.bib and uses the
-% plain style.  Other styles may be used depending on the conventions
-% of your field of study.
-% Note: the bibliography must come before the appendices.
-\bibliographystyle{plain} \bibliographystyle{plain}
+%% This file is setup to use a bibtex file sample.bib and uses the
+%% plain style.  Other styles may be used depending on the conventions
+%% of your field of study.
+%%
+%%% Note: the bibliography must come before the appendices.
+\bibliographystyle{plain}
 \bibliography{sample}
 
-% Use this to reset the appendix counter.  Note that the FoGS
-% requires that the word ``Appendices'' appear in the table of
-% contents either before each appendix lable or as a division
-% denoting the start of the appendices.  We take the latter option
-% here.  This is ensured by making the \texttt{appendicestoc} option
-% a default option to the UBC thesis class.
+%% Use this to reset the appendix counter.  Note that the FoGS
+%% requires that the word ``Appendices'' appear in the table of
+%% contents either before each appendix lable or as a division
+%% denoting the start of the appendices.  We take the latter option
+%% here.  This is ensured by making the \texttt{appendicestoc} option
+%% a default option to the UBC thesis class.
 
-% If you only have one appendix, please uncomment the following line.
-
+%%% If you only have one appendix, please uncomment the following line.
+% \renewcommand{\appendicesname}{Appendix}
 \appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
 \chapter{First Appendix}
-
-Here you can have your appendices. Note that if you only have a single
-appendix, you should issue \verb|\renewcommand{\appendicesname}{Appendix}|
-before calling \verb|\appendix| to display the singular {}``Appendix''
-rather than the default plural {}``Appendices'' some
+Here you can have your appendices.  Note that if you only have a
+single appendix, you should issue
+\verb|\renewcommand{\appendicesname}{Appendix}| before calling
+\verb|\appendix| to display the singular ``Appendix'' rather than the
+default plural ``Appendices''.
 
 \chapter{Second Appendix}
-
 Here is the second appendix.
 
-% Indices come here.
+%% This changes the headings and chapter titles (no numbers for
+%% example).
+\backmatter
+
+%% Indices come here if you have them.
 
 \chapter*{Additional Information}
-
 This chapter shows you how to include additional information in your
@@
 First, the chapter is unnumbered and not included in the Table of
-Contents. Second, it is the last section of the thesis, so its removal
-will not alter any of the page numbering etc. for the previous sections.
-Do not include any floats, however, as these will appear in the initial
-lists.
+Contents.  Second, it is the last section of the thesis, so its
+removal will not alter any of the page numbering etc. for the previous
+sections.  Do not include any floats, however, as these will appear in
+the initial lists.
 
-The \texttt{ubcthesis} \LaTeX{} class has been designed to aid you
-in producing a thesis that conforms to the requirements of The University
-of British Columbia Faculty of Graduate Studies (FoGS).
+The \texttt{ubcthesis} \LaTeX{} class has been designed to aid you in
+producing a thesis that conforms to the requirements of The
+University of British Columbia Faculty of Graduate Studies (FoGS).
 
@@
 produce a well formatted document that meets the FoGS requirement.
-Notwithstanding, complex theses may require additional formatting
-that may conflict with some of the requirements. We therefore \emph{highly
+Notwithstanding, complex theses may require additional formatting that
+may conflict with some of the requirements.  We therefore \emph{highly
   recommend} that you consult one of the FoGS staff for assistance and
-an assessment of potential problems \emph{before} starting final draft.
+an assessment of potential problems \emph{before} starting final
+draft.
 
-While we have attemped to address most of the thesis formatting requirements
-in these files, they do not constitute an official set of thesis requirements.
-The official requirements are available at the following section of
-the FoGS web site:
-
+While we have attemped to address most of the thesis formatting
+requirements in these files, they do not constitute an official set of
+thesis requirements.  The official requirements are available at the
+following section of the FoGS web site:
 \begin{center}
@@
     \hline
-\url{http://www.grad.ubc.ca/students/thesis/}\tabularnewline
+    \url{http://www.grad.ubc.ca/current-students/dissertation-thesis-preparation}\\
     \hline
   \end{tabular}
-\par\end{center}
+\end{center}
+We recommend that you review these instructions carefully.
 
-We recommend that you review these instructions carefully.
 \end{document}
+\endinput
+%%
+%% End of file `ubcsample.tex'.

### 2014-04-14 lyx/makeUBCthesisLyXfromTeX.py

@@ -1,0 +1,363 @@
+#!/usr/bin/python
+
+# This script automates the creation of a UBC Thesis Sample file in LyX,
+# using the LaTeX version. This means using tex2lyx to do the initial
+# conversion, and then fixing bugs, and then maybe making it look nicer
+# as well.
+# 2010 July 16: CPBL: first version
+
+# For safety, still call it ubcsamplelyx.lyx, so that exports from
+# this don't automatically get named the same thing as the original
+# .tex file...
+
+import os
+import re
+import errno
+
+'''
+# These are the commands I run: Michael
+rm -r convertToLyx
+export TEX2LYX=/Applications/LyX.app/Contents/MacOS/tex2lyx
+export LYX=/Applications/LyX.app/Contents/MacOS/lyx
+export DIFF=ediff
+python makeUBCthesisLyXfromTeX.py
+'''
+TEX2LYX = os.environ.get('TEX2LYX', 'tex2lyx')
+LYX = os.environ.get('LYX', 'lyx')
+DIFF = os.environ.get('DIFF', 'meld')
+PDFLATEX = os.environ.get('PDFLATEX', 'pdflatex')
+EPSTOPDF = os.environ.get('EPSTOPDF', 'epstopdf')
+
+UBCSAMPLE_TEX = '../../ubcsample.tex'
+
+def do_sys(cmd):
+    r"""Execute the command and print it."""
+    print(cmd)
+    os.system(cmd)
+
+def preprocess_tex(text):
+    r"""Return preprocessed .tex file to make it easier for the lyx
+    converter.
+
+    Parameters
+    ----------
+    text : String representing the original latex file.
+    """
+    # Remove preamble and ignore as this will end up in preambl
+    start = text.find(r"\begin{document}")
+    preamble = text[:start]
+
+    # Add comments package
+    ind = preamble.find(r"\usepackage")
+    preamble = "\n".join([preamble[:ind],
+                          r"\usepackage{comment}", "",
+                          preamble[ind:]])
+    
+    text = text[start:]
+
+    # Match solid blocks of comments that start as "%%% ".  We add a *
+    # so we can convert these to Notes later.
+    prog = re.compile(r"((^%%% .*\n)+)",
+                      flags=re.MULTILINE)
+    res = []
+    while True:
+        # For each block...
+        match = prog.search(text)
+        if match is None:
+            res.append(text)
+            break
[…]

### 2014-04-15 paper/ubcThesisSubmission.tex
thunk

@@
-%%
-%% This is file `ubcsample.tex',
-%% generated with the docstrip utility.
-%
-% The original source files were:
-%
-% ubcthesis.dtx  (with options: `ubcsampletex')
-%%
-%% This file was generated from the ubcthesis package.
-%% --------------------------------------------------------------
-%% NEED TO FIX: Envelope, ensure abs value --- add to figure caption
-%% Lith figure, tighten range.
-%% Rework final geochron conclusion. Data is weak for current claims
-%% figure -> Figure
 \documentclass[msc,oneside]{ubcthesis}
@@
 
-%******** afterpage ***************************
-% This package allows you to issue commands at the end of the current
-% page.  A good use for this is to use the command
-% \afterpage{\clearpage} right after a figure.  This will cause the
-% figure to be inserted on the page following the current one (or on
-% the current page if it will fit) but will not break the page in the
-% middle.
 \usepackage{afterpage}
 
-%******** float *********************************
-% This package allows you to customize the style of
-% "floats"---floating objects such as figures and tables.  In
-% addition, it allows you to define additional floating objects which
-% may be included in a list similar to that produces by \listoftables
-% and \listoffigures.  Common uses include introducing floats for
-% programs and other code bits in Compute Science and Chemical Schema.
 \usepackage{float}
 
-%******** tocloft *******************************
-% This package allows you to customize and define custom lists such
-% as a list of programs or Chemical Scheme.  Note: if you use the
-% subfigure package, you must specify that you do as an option here.
-% The title option uses the default formatting.  We do not use this
-% here as the default formatting is acceptable.  Use the float
-% package instead unless you need the extra formatting control
-% provided by tocloft.
-%\usepackage[subfigure, titles]{tocloft}
-
-%******** alltt *********************************
-% The alltt package allows you to include files and have them
-% formatted in a verbatim fashion.  This is useful for including
-% source code from an additional file.
-%\usepackage{alltt}
-
-%******** listings ******************************
-% The listings package may be used to include chunks of source code
-% and has facilities for pretty-printing many languages.
-%\usepackage{listings}
-
-%******** longtable *****************************
-% The longtable package allows you to define tables that span
-% multiple pages.
-% \usepackage{longtable}
-
-%******** graphics and graphicx *****************
-% This allows you to include encapsulated postscript files.  If you
-% don't have this, comment the \includegraphics{} line following the
-% comment "%includegraphics" later in this file.
-% \usepackage{graphicx}
-
-%******** subfigure *****************************
-% The subfigure package allows you to include multiple figures and
-% captions within a single figure environment.
-%\usepackage{subfigure}
-
-%******** here **********************************
-% The here package gives you more control over the placement of
-% figures and tables.  In particular, you can specify the placement
-% "H" which means "Put the figure here" rather than [h] which means
-% "I would suggest that you put the figure here if you think it looks
-% good."
-%\usepackage{here}
-
-%******** natbib ********************************
-% This is a very nice package for bibliographies.  It includes options
-% for sorting and compressing bibliographic entries.
 \usepackage[numbers,sort&compress]{natbib}
 
-%******** psfrag ******************************
-% This allows you to replace text in postscript pictures with formated
-% latex text.  This allows you to use math in graph labels
-% etc. Uncomment the psfrag lines following the "%psfrag" comment
-% later in this file if you don't have this package.  The replacements
-% will only be visible in the final postscript file: they will be
-% listed in the .dvi file but not performed.
-\usepackage{psfrag}
-
-%******** hyperref *****************************
-% Please read the manual:
-% http://www.tug.org/applications/hyperref/manual.html
-%
-% This adds hyperlinks to your document: with the right viewers (later
-% versions of xdvi, acrobat with pdftex, latex2html etc.) this will
-% make your equation, figure, citation references etc. hyperlinks so
-% that you can click on them.  Also, your table of contents will be
-% able to take you to the appropriate sections.  In the viewers that
-% support this, the links often appear with an underscore.  This
-% underscore will not appear in printed versions.
-%
-% Note: if you do not use the hypertex option, then the dvips driver
-% may be loaded by default.  This will cause the entries in the list
-% of figures and list of tables to be on a single line because dvips
-% does not deal with hyperlinks on broken lines properly.
-%
-% NOTE: HYPERREF is sensitive to the ORDER in which it is LOADED.
-% For example, it must be loaded AFTER natbib but BEFORE newly
-% defined float environments.  See the README file with the hyperref
-% for some help with this.  If you have some very obscure errors, try
-% first disabling hyperref.  If that fixes the problem, try various
-% orderings.
-%
-% Note also that there is a bug with versions before 2003/11/30
-% v6.74m that cause the float package to not function correctly.
-% Please ensure you have a current version of this package.  A
-% warning will be issued if you leave the date below but do not have
-% a current version installed.
-%
-% Some notes on options: depending on how you build your files, you
-% may need to choose the appropriate option (such as [pdftex]) for the
-% backend driver (see the hyperref manual for a complete list).  Also,
-% the default here is to make links from the page numbers in the table
-% of contents and lists of figures etc.  There are other options:
-% excluding the [linktocpage] option will make the entire text a
-% hyperref, but for some backends will prevent the text from wrapping
-% which can look terrible.  There is a [breaklinks=true] option that
-% will be set if the backend supports (dvipdfm for example supports
-% it but does not work with psfrag.)
-%
-% Finally, there are many options for choosing the colours of the
-% links.  These will be included by default in future versions but
-% you should probably consider changing some now for the electronic
 % version of your thesis.
@@
 
-% If you would like to compile this sample thesis without the
-% hyperref package, then you will need to comment out the previous
-% \usepackage command and uncomment the following command which will
-% put the URL's in a typewriter font but not link them.
-%\newcommand\url[1]{\texttt{#1}}
-
-%******** setspace *******************************
-% The setspace package allows you to manually set the spacing of the
-% file.  UBC may require 1.5 spacing for microfilming of theses.  In
-% this case you may obtain this by including this package and issuing
-% one of the following commands:
-%\usepackage{setspace}
-%\singlespacing
-%\onehalfspacing
-%\doublespacing
-
-% These commands are optional.  The defaults are shown.  You only
-% need to include them if you need a different value
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \institution{The University Of British Columbia}
 
-% If you are at the Okanagan campus, then you should specify these
-% instead.
-%\faculty{The College of Graduate Studies}
-%\institutionaddress{Okanagan}
 \faculty{The Faculty of Graduate Studies}
@@
 % You can issue as many of these as you have...
-\previousdegree{B.Sc., The University of British Columbia, 1999}
-\previousdegree{M.Sc., The University of British Columbia, 2001}
-\previousdegree{Ph.D., Massachusetts Institute of Technology, 2005}
-
-% You can override the option setting here.
-% \degreetitle{Jack of All Trades}
+\previousdegree{B.Sc., The University of British Columbia, 2011}
 
@@
                                       % command from eating the space.
-\program{GeoPhysics}
-
-% These commands are presently not required for UBC theses as the
-% advisor's name and title are not presently required anywhere.
-%\advisor{Ariel R.~Zhitnitsky}
-%\advisortitle{Professor of Physics}
+\program{Geophysics}
 
@@
 
-%% This starts numbering in Roman numerals as required for the thesis
-%% style and is mandatory.
 \frontmatter
-
-%%% The order of the following components should be preserved.  The order
-%%% listed here is the order currently required by FoGS:        \\
-%%% Title (Mandatory)                                           \\
-%%% Preface (Manditory if any collaborator contributions)       \\
-%%% Abstract (Mandatory)                                        \\
-%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
-%%% Acknowledgements (Optional)                                 \\
-%%% Dedication (Optional)                                       \\
 
@@
 Analysis of results in conjunction with additional velocity estimates from active source seismic studies and a seismic property database affords improved constraints on bulk geological composition of the Canadian landmass that are used to evaluate competing models of crustal formation.
-
-  Ben Postlethwaite
-  post.ben.here@gmail.com
 \end{abstract}
 
-\chapter{Preface} % Manditory if any of the conditions are met
+\chapter{Preface}
 
-You must include a preface if any part of your research was partly or
-wholly published in articles, was part of a collaboration, or required
-the approval of UBC Research Ethics Boards.
+Teleseismic data discussed in chapter \ref{cha:datamethods} section \ref{section:seismicdata} was processed using Python scripts written by myself and a set of shell scripts based off previous work by Michael Bostock.
+Deconvolution algorithms and coordinate transformation algorithms are likewise interpreted from previous work by Michael Bostock.
+Suggestions by David Snyder to investigate signals of the Conrad Discontinuity lead to the analysis of crustal discontinuities in chapter \ref{cha:results} and chapter \ref{cha:discussion}.
+Nik Christensen provided invaluable support during the discussion on crustal composition in chapter \ref{cha:discussion} section \ref{section:crustalcomp}.
+Specifically, Nik providing assistance in the interpretation and use of laboratory velocity measurements (Christensen, 1996) and rock chemical analyses (Le Maitre, 1976).
+His suggestions on crustal compositional literature and recommendations for the layout of figure \ref{fig:lith} were critical to this discussion.
+Michael Bostock provided advice and editorial recommendations in all chapters of this thesis.
 
-The Preface must include the following:
-
-\begin{itemize}
-\item A statement indicating the relative contributions of all
-  collaborators and co-authors of publications (if any), emphasizing
-  details of your contribution, and stating the proportion of research
-  and writing conducted by you.
-\item A list of any publications arising from work presented in the
-  dissertation, and the chapter(s) in which the work is located.
-\item The name of the particular UBC Research Ethics Board, and the
-  Certificate Number(s) of the Ethics Certificate(s) obtained, if
-  ethics approval was required for the research.
-\end{itemize}
-
-%%% Sections and subsections etc. in the Preface should in general
-%%% not be listed in the table of contents, so use the starred form
-%%% of \section etc.
-\section*{Examples}
-Chapter~\ref{cha:apple_ref} is based on work conducted in UBC's Maple
-Syrup Laboratory by Dr. A.  Apple, Professor B. Boat, and Michael
-McNeil Forbes. I was responsible for tapping the trees in forests X
-and Z, conducted and supervised all boiling operations, and performed
-frequent quality control tests on the product.
-
-A version of chapter~\ref{cha:apple_ref} has been
-published~\cite{Apple:2010}. I conducted all the testing and wrote
-most of the manuscript. The section on ``Testing Implements'' was
-originally drafted by Boat, B.  Check the first pages of this
-chapter to see footnotes with similar information.
-
-Note that this preface must come before the table of contents.  Note
-also that this section ``Examples'' should not be listed in the table
-of contents, so we have used the starred form: \verb|\section*{Example}|.
-
-\tableofcontents                %% Mandatory
-\listoftables                   %% Mandatory if thesis has tables
-\listoffigures                  %% Mandatory if thesis has figures
-
-\chapter{Acknowledgements}      %% Optional
-This is the place to thank professional colleagues and people who have
-given you the most help during the course of your graduate work.
+\tableofcontents
+\listoftables
+\listoffigures
 
@@
 % Chapters are the next main unit.
-\chapter{Introduction}
+\chapter{Introduction} \label{cha:intro}
 
@@
 In the northwest of the Superior Province on the eastern shores of the Hudson Bay lies the Nuvvuagittuq greenstone belt that recent geochronological work has dated at ~4.28 Ga making it the only known remnant of Hadean crust preserved on Earth (O'Neil et al., 2011).
+
 To the east lies the younger, 1.2-0.98 Ga, Grenville orogen, consisting largely of Superior and Southern Province rock reworked during Alpine-Himalayan like continental collision (Eaton et al., 2006).
@@
 Its western portion is the 4.0 Ga remnant of a larger Archean continent, whereas the eastern Slave forms a younger (2.6 Ga) paired accretionary prism and island arc (Kusky, 1989).
+
 To the west of the Superior is the extensive Interior Platform overlying the southern extensions of the Trans-Hudson Orogen, Hearne Domain and Alberta Orogen.
@@
 
-\chapter{Data and Methods}
+\chapter{Data and Methods} \label{cha:datamethods}
 
@@
 
-\section{Teleseismic Data Set}
+\section{Teleseismic Data Set} \label{section:seismicdata}
 The primary data utilized in this study are RF processed teleseismic {\it P} wave seismograms originating from more than 700 earthquake sources.
@@
 
-\chapter{Results}
+\chapter{Results} \label{cha:results}
 
@@
   \end{tabular}
-  \caption[Regional seismic parameter comparsion]{Comparison of seismic velocity and crustal thickness estimates for Canada and selected geological provinces.}
+  \caption[Regional seismic parameter comparison]{Comparison of seismic velocity and crustal thickness estimates for Canada and selected geological provinces.}
 \label{table:regionParameters}
@@
 
-\chapter{Discussion}
+\chapter{Discussion} \label{cha:discussion}
+
 \section{Crustal reflectivity}
@@
  \includegraphics[width=\textwidth]{discInterp.eps}
- \caption[Normalized station depth profiles and SVD analysis]{Left, vertical station depth profiles interpolated to a common Moho. Right, normalized stack of depth profiles, grey line shows mean}
+ \caption[Normalized station depth profiles and SVD analysis]{Left, vertical station depth profiles interpolated to a common Moho. Right, normalized stack of depth envelopes, grey line shows mean}
  \label{fig:discInterp}
@@
 (1997) in their study of the Rhine Graben, have employed the term Conrad Discontinuity to refer to an interface or transition between transparent upper crust and reflective lower crust.
-It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure \ref{fig:disc}, are a manifestation of such signals, but that the lower energy levels ($<5\%$) in this mode would indicate that the Conrad Discontinuity so-defined is at best weakly represented within the Canadian crust.
+It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure \ref{fig:discPCA}, are a manifestation of such signals, but that the lower energy levels ($<5\%$) in this mode would indicate that the Conrad Discontinuity so-defined is at best weakly represented within the Canadian crust.
 This finding is supported by crustal velocity models characterized by a near constant velocity gradient such as the gradual transition from a more silicic upper crust to a more mafic lower crust in shields and platforms observed by Meissner (1986).
@@
 The increased transparency of the lower crust would appear to relate to changes in rheology.
-Studies of crustal strength (Jackson, 2002), deformation (Burov and Watts, 2006; Meissner et al., 2006) and the inferred $\sim$ 100 m lower crustal lamallae reported by Mooney and Brocher (1987) as well as the widely noted reduction in seismicity of the lower crust *** ADD CITATION *** would support this contention; though see Deichmann (1992), Handy and Brun (2004) and Simpson (1999) for competing evidence.
+Studies of crustal strength (Jackson, 2002), deformation (Burov and Watts, 2006; Meissner et al., 2006) and the inferred $\sim$ 100 m lower crustal lamallae reported by Mooney and Brocher (1987) as well as the widely noted reduction in seismicity of the lower crust (Ito, 1990; Meissner and Wever, 1986) would support this contention; though see Deichmann (1992), Handy and Brun (2004) and Simpson (1999) for competing evidence.
 
-\section{Crustal Composition}
+\section{Crustal Composition} \label{section:crustalcomp}
 Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust.
@@
 
-Whereas Zandt and Ammon present evidence for increasing $V_P/V_S$ with crustal age, our estimates signify the opposite trend.
-Our results are more consistent with those of Durrheim and Mooney (1991) who argue that Proterozoic crust is thicker with higher {\it P}-velocity than Archean crust.
-The crustal thickness data presented in this study reveals no significant difference between Proterozoic crust (38.9 km) and Archean crust (38.0 km) in contrast to the 5 km difference reported by Durrheim and Mooney.
-Though the difference is not large, the higher seismic velocity ratio in the Proterozoic units (1.76) versus the Archean units (1.74) aligns with their argument for an evolution of crustal processes throughout geological time.
-The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by {\it Durrheim and Mooney} (1991), has been persuasively augmented by recent work (Rudnick 1995; Rudnick and Gao, 2003) to now include a range of possible processes that may explain the data.
+Whereas Zandt and Ammon present evidence for increasing $V_P/V_S$ with crustal age our estimates decrease in $V_P/V_S$ with crustal age.
+This decrease however is slight, from 1.76 in the Proterozoic units to 1.74 in the Archean units and signifies no trend in $V_P/V_S$ with crustal age.
+Similarly, whereas Durrheim and Mooney (1991) argue for thicker (>5km) Proterozoic crust with higher {\it P}-velocity than Archean crust, the crustal thickness data presented in this study reveals no significant difference between Proterozoic crust (38.9 km) and Archean crust (38.0 km).
+The hypothesis that Proterozoic crust is prone to basaltic underplating, thereby explaining higher velocities and thicker crust, suggested by {\it Durrheim and Mooney} (1991) does not seem to explain the data presented in this study.
+However, this hypothesis has been persuasively augmented by recent work (Rudnick 1995; Rudnick and Gao, 2003) to now include a range of possible processes that may better explain the data.
 
-\section{Conclusions}
+\chapter{Conclusions}
 
@@
 Constraints from $V_P/V_S$ estimates suggest that the composition of the Canadian Shield is comparable to the global continental crustal average whereas the Canadian crustal average is characterized by slightly lower $V_P$ and $V_S$ and higher $V_P/V_S$.
-Finally, the dataset supports the hypothesis that processes driving crustal formation in the Archean and Proterozoic were dissimilar.
 
@@
 
-\bibitem {} Eaton, D. W. (2006), Multi-genetic origin of the continental Moho: insights from Lithoprobe, Terra Nova, 18, 34–43.
+\bibitem{} Eaton, D. W. (2006), Multi-genetic origin of the continental Moho: insights from Lithoprobe, Terra Nova, 18, 34–43.
 
@@
 \bibitem{} Ito, K. (1999), Seismogenic layer, reflective lower crust, surface heat flow and large inland earthquakes, Tectonophysics, 306, 423-433.
+
+\bibitem{} Ito, K. (1990), Regional variations of the cutoff depth of seismicity in the crust and their relation to heat flow and large inland-earthquakes, J. Phys. Earth, 38, 223–250.
 
@@
 
+\bibitem{} Meissner, R., T. Wever (1986), Intracontinental seismicity, strength of crustal units, and the seismic signature of fault zones, Philos. Trans. R. Soc. London, 317, 45–61.
+
 \bibitem{} Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.
@@
 
-\bititem{} Mooney, W. D., G. Laske, T. G. Masters (1998), CRUST 5.1: A global crustal model at $5\circ \times 5\circ$, J. Geophys. Res., 103(B1), 727–747
+\bibitem{} Mooney, W. D., G. Laske, T. G. Masters (1998), CRUST 5.1: A global crustal model at 5$^\circ$ $\times$ 5$^\circ$, J. Geophys. Res., 103(B1), 727–747
 

### 2014-04-15 python/lithologyPlots.py

@@ -40,7 +40,7 @@ def project(ax, ay, bx, by, cx, cy):
 
 
 def addtext(ax, x, y, text, rotation):
-    ax.text(x-0.02, y, text,
+    ax.text(x, y, text,
             color = 'black',
             size = 'large',
             horizontalalignment = 'left',
@@ -54,7 +54,7 @@ liths = {
         "Vs": 3.583,
         "R": 1.732,
         "poisson": 0.250,
-        "x": -0.08,
+        "x": -0.0,
         "y": 0.0
         },
     "tonalite gneiss": {
@@ -62,7 +62,7 @@ liths = {
         "Vs": 3.606,
         "R": 1.747,
         "poisson": 0.257,
-        "x": 0.11,
+        "x": 0.0,
         "y": 0.
         },
     "mafic granulite": {
@@ -70,40 +70,40 @@ liths = {
         "Vs": 3.820,
         "R": 1.817,
         "poisson": 0.283,
-        "x": -0.08,
-        "y": 0.05
-        },
-    "mafic garnet granulite": {
-        "Vp": 7.249,
-        "Vs": 4.026,
-        "R": 1.801,
-        "poisson": 0.277,
-        "x": -0.15,
+        "x": -0.0,
         "y": 0.0
         },
-    "amphibolite": {
-        "Vp": 6.983,
-        "Vs": 3.959,
-        "R": 1.764,
-        "poisson": 0.263,
-        "x": 0.,
-        "y": 0.05
-        },
-    "hornblendite": {
-        "Vp": 7.261,
-        "Vs": 4.144,
-        "R": 1.752,
-        "poisson": 0.258,
-        "x": 0.,
-        "y": 0.05
-        },
+    # "mafic garnet granulite": {
+    #     "Vp": 7.249,
+    #     "Vs": 4.026,
+    #     "R": 1.801,
+    #     "poisson": 0.277,
+    #     "x": -0.15,
+    #     "y": 0.0
+    #     },
+    # "amphibolite": {
+    #     "Vp": 6.983,
+    #     "Vs": 3.959,
+    #     "R": 1.764,
+    #     "poisson": 0.263,
+    #     "x": 0.,
+    #     "y": 0.0
+    #     },
+    # "hornblendite": {
+    #     "Vp": 7.261,
+    #     "Vs": 4.144,
+    #     "R": 1.752,
+    #     "poisson": 0.258,
+    #     "x": 0.,
[…]
